### 2019-01-02 example programs/goEnrichment.GSTA4KO.2017-09-01.Rmd
add example programs

---
title: "GSTA4 KO analysis for Colin"
author: "Laura Saba"
date: "9/1/2017"
output: html_document
---

```{r}
rm(list=ls())
options(stringsAsFactors = FALSE)
library("GO.db")
library(dplyr)
wd <- "~/Documents/Shearn/GSTA4/"

orig <- read.table(file=paste(wd,"data/120d EtOH GSTA4 study Table for Laura Saba.txt",sep=""),sep="\t",quote="",header=TRUE,fill=TRUE)
orig <- orig[orig$Accession!="totals" & orig$Accession!=" " & orig$Accession!="",]

update <- read.table(file=paste(wd,"data/updateProteinID.2017-07-14.txt",sep=""),sep="\t",header=TRUE)
proteins <- merge(orig,update,by.x="Accession",by.y="orig",all=TRUE)
proteins$Accession[!is.na(proteins$new)] = proteins$new[!is.na(proteins$new)]

# GO data base
go_db <- read.table(file=paste(wd,"source/",sep=""),sep="\t",header=FALSE,comment.char="!",quote="")

library(biomaRt)
ensembl=useMart("ensembl")
ensembl = useDataset("mmusculus_gene_ensembl",mart=ensembl)

#convert protein ID to Uniprot symbols
protein_to_uniprot <- read.table(file=paste(wd,"data/protein_to_uniprot.2017-07-14.txt",sep=""),sep="\t",quote="",header=TRUE,fill=TRUE)

proteins <- merge(proteins,protein_to_uniprot,by.x="Accession",by.y="Entry.name",all.x=TRUE)

##problematic proteins
proteins <- proteins[!(proteins$Accession %in% c("PZP_MOUSE","SLF2_MOUSE")),]

##Convert to MGI Symbol
toMGI <- getBM(attributes=c('uniprot_gn', 'mgi_id',"mgi_symbol"), 
      filters = 'mgi_id', 
      values = go_db$V2, 
      mart = ensembl)
```

GO Enrichment

```{r}
uniprot_to_go <- merge(toMGI,go_db,by.x="mgi_id",by.y="V2")

Chow_GSTA4 <- unique(uniprot_to_go$uniprot_gn[uniprot_to_go$uniprot_gn %in% proteins$Entry[proteins$Chow.GSTA4...==1]])
Chow_SV <- unique(uniprot_to_go$uniprot_gn[uniprot_to_go$uniprot_gn %in% proteins$Entry[proteins$Chow.SV==1]])
EtOH_GSTA4 <- unique(uniprot_to_go$uniprot_gn[uniprot_to_go$uniprot_gn %in% proteins$Entry[proteins$EtOH.GSTA4...==1]])
EtOH_SV <- unique(uniprot_to_go$uniprot_gn[uniprot_to_go$uniprot_gn %in% proteins$Entry[proteins$EtOH.SV==1]])
Binge_GSTA4 <- unique(uniprot_to_go$uniprot_gn[uniprot_to_go$uniprot_gn %in% proteins$Entry[proteins$Binge.GSTA4...==1]])
Binge_SV <- unique(uniprot_to_go$uniprot_gn[uniprot_to_go$uniprot_gn %in% proteins$Entry[proteins$Binge.SV==1]])

pathEnrich <- function(x){
  all_GO <- unique(uniprot_to_go$uniprot_gn)
  sig_GO <- unique(x)

  all_GO_cnt <- uniprot_to_go[uniprot_to_go$uniprot_gn %in% all_GO,]
  sig_GO_cnt <- uniprot_to_go[uniprot_to_go$uniprot_gn %in% sig_GO,]

  all_GO_cnt <- all_GO_cnt %>%
    group_by(V5) %>%
    summarize(GO_cnt = length(uniprot_gn))

  sig_GO_cnt <- sig_GO_cnt %>%
    group_by(V5) %>%
    summarize(GO_sig = length(uniprot_gn))

  enrich_table <- merge(all_GO_cnt,sig_GO_cnt,by="V5",all=TRUE)
  enrich_table$GO_sig[is.na(enrich_table$GO_sig)] = 0

  enrich_table$numTested <- length(all_GO)
  enrich_table$numSig <- length(sig_GO)
  enrich_table <- enrich_table %>%
    mutate(expected = (numSig/numTested)*GO_cnt)

  #enrich_table <- merge(path_names_mmu,enrich_table,by.x="V1",by.y="pathway")

  enrich_table$enrich_p <- NA
  for(i in 1:nrow(enrich_table)){
    a = enrich_table[i,"GO_sig"]
    b = enrich_table[i,"GO_cnt"] - enrich_table[i,"GO_sig"]
    c = enrich_table[i,"numSig"] - enrich_table[i,"GO_sig"]
    d = enrich_table[i,"numTested"] - enrich_table[i,"numSig"] + enrich_table[i,"GO_sig"]
    enrich_table$enrich_p[i] = fisher.test(matrix(c(a,b,c,d),nr=2),alternative="greater")$p.value
  }

  enrich_table <- enrich_table[order(enrich_table$enrich_p),]
  enrich_table$fdr <- p.adjust(enrich_table$enrich_p,method="BH")
  return(enrich_table)
}

Chow_GSTA4_enrich <- pathEnrich(Chow_GSTA4)
Chow_SV_enrich <- pathEnrich(Chow_SV)
EtOH_GSTA4_enrich <- pathEnrich(EtOH_GSTA4)
EtOH_SV_enrich <- pathEnrich(EtOH_SV)
Binge_GSTA4_enrich <- pathEnrich(Binge_GSTA4)
Binge_SV_enrich <- pathEnrich(Binge_SV)

colnames(Chow_GSTA4_enrich) <- c("V1","GO_cnt","GO_sig_Chow_GSTA4","numTested","numSig_Chow_GSTA4","expected_Chow_GSTA4","enrich_p_Chow_GSTA4","fdr_Chow_GSTA4")
colnames(Chow_SV_enrich) <- c("V1","GO_cnt","GO_sig_Chow_SV","numTested","numSig_Chow_SV","expected_Chow_SV","enrich_p_Chow_SV","fdr_Chow_SV")

colnames(EtOH_GSTA4_enrich) <- c("V1","GO_cnt","GO_sig_EtOH_GSTA4","numTested","numSig_EtOH_GSTA4","expected_EtOH_GSTA4","enrich_p_EtOH_GSTA4","fdr_EtOH_GSTA4")
colnames(EtOH_SV_enrich) <- c("V1","GO_cnt","GO_sig_EtOH_SV","numTested","numSig_EtOH_SV","expected_EtOH_SV","enrich_p_EtOH_SV","fdr_EtOH_SV")

colnames(Binge_GSTA4_enrich) <- c("V1","GO_cnt","GO_sig_Binge_GSTA4","numTested","numSig_Binge_GSTA4","expected_Binge_GSTA4","enrich_p_Binge_GSTA4","fdr_Binge_GSTA4")
colnames(Binge_SV_enrich) <- c("V1","GO_cnt","GO_sig_Binge_SV","numTested","numSig_Binge_SV","expected_Binge_SV","enrich_p_Binge_SV","fdr_Binge_SV")

combined_enrich <- merge(Chow_GSTA4_enrich,Chow_SV_enrich,by=c("V1","GO_cnt","numTested"),all=TRUE)
combined_enrich <- merge(combined_enrich,EtOH_GSTA4_enrich,by=c("V1","GO_cnt","numTested"),all=TRUE)
combined_enrich <- merge(combined_enrich,EtOH_SV_enrich,by=c("V1","GO_cnt","numTested"),all=TRUE)
combined_enrich <- merge(combined_enrich,Binge_GSTA4_enrich,by=c("V1","GO_cnt","numTested"),all=TRUE)
combined_enrich <- merge(combined_enrich,Binge_SV_enrich,by=c("V1","GO_cnt","numTested"),all=TRUE)

threshold = 0.01
sum(combined_enrich$fdr_Chow_GSTA4<threshold,na.rm=TRUE)
sum(combined_enrich$fdr_Chow_SV<threshold,na.rm=TRUE)
sum(combined_enrich$fdr_EtOH_GSTA4<threshold,na.rm=TRUE)
sum(combined_enrich$fdr_EtOH_SV<threshold,na.rm=TRUE)
sum(combined_enrich$fdr_Binge_GSTA4<threshold,na.rm=TRUE)
sum(combined_enrich$fdr_Binge_SV<threshold,na.rm=TRUE)

combined_enrich$num_grps_sig <- rowSums(combined_enrich[,grep("fdr",colnames(combined_enrich))]<threshold)

of_interest <- combined_enrich[combined_enrich$num_grps_sig>0 & combined_enrich$num_grps_sig<6,]

groups <- paste(rep(c("Chow","EtOH","Binge"),each=2),rep(c("GSTA4","SV"),3),sep="_")
of_interest[,c("V1","V2",paste("fdr_",groups,sep=""))]

combined_enrich$sig_pattern <- apply(combined_enrich[,grep("fdr",colnames(combined_enrich))],1,function(a) paste(as.numeric(a<threshold),collapse=""))
table(rowSums(combined_enrich[,grep("enrich_p",colnames(combined_enrich))]<threshold))
```

##Differential Enrichment

```{r}
diffEnrich <- function(a,b,c,d,OR){
  y = fisher.test(matrix(c(a,b,c,d),nr=2))
  if(OR) z = y$estimate
  if(!OR) z = y$p.value
  return(z)
}

#diff enrich EtOH effect - WT Chow vs WT EtOH

combined_enrich$EtOH_Effect_p = apply(combined_enrich[,c("GO_sig_Chow_SV","GO_sig_EtOH_SV","numSig_Chow_SV","numSig_EtOH_SV")],1,function(a) diffEnrich(a[1],a[2],a[3],a[4],OR=FALSE))

combined_enrich$EtOH_Effect_OddsRatio = apply(combined_enrich[,c("GO_sig_Chow_SV","GO_sig_EtOH_SV","numSig_Chow_SV","numSig_EtOH_SV")],1,function(a) diffEnrich(a[1],a[2],a[3],a[4],OR=TRUE))

#diff enrich EtOH effect in GSTA4 - GSTA4 Chow vs GSTA4 EtOH

combined_enrich$EtOHinGSTA4_Effect_p = apply(combined_enrich[,c("GO_sig_Chow_GSTA4","GO_sig_EtOH_GSTA4","numSig_Chow_GSTA4","numSig_EtOH_GSTA4")],1,function(a) diffEnrich(a[1],a[2],a[3],a[4],OR=FALSE))

combined_enrich$EtOHinGSTA4_Effect_OddsRatio = apply(combined_enrich[,c("GO_sig_Chow_GSTA4","GO_sig_EtOH_GSTA4","numSig_Chow_GSTA4","numSig_EtOH_GSTA4")],1,function(a) diffEnrich(a[1],a[2],a[3],a[4],OR=TRUE))

#diff enrich Binge effect - WT Binge vs WT EtOH

combined_enrich$Binge_Effect_p = apply(combined_enrich[,c("GO_sig_EtOH_SV","GO_sig_Binge_SV","numSig_EtOH_SV","numSig_Binge_SV")],1,function(a) diffEnrich(a[1],a[2],a[3],a[4],OR=FALSE))

combined_enrich$Binge_Effect_OddsRatio = apply(combined_enrich[,c("GO_sig_EtOH_SV","GO_sig_Binge_SV","numSig_EtOH_SV","numSig_Binge_SV")],1,function(a) diffEnrich(a[1],a[2],a[3],a[4],OR=TRUE))

#diff enrich Binge effect in GSTA4 - GSTA4 Binge vs GSTA4 EtOH

combined_enrich$BingeInGSTA4_Effect_p = apply(combined_enrich[,c("GO_sig_EtOH_GSTA4","GO_sig_Binge_GSTA4","numSig_EtOH_GSTA4","numSig_Binge_GSTA4")],1,function(a) diffEnrich(a[1],a[2],a[3],a[4],OR=FALSE))

combined_enrich$BingeInGSTA4_Effect_OddsRatio = apply(combined_enrich[,c("GO_sig_EtOH_GSTA4","GO_sig_Binge_GSTA4","numSig_EtOH_GSTA4","numSig_Binge_GSTA4")],1,function(a) diffEnrich(a[1],a[2],a[3],a[4],OR=TRUE))

#diff enrich Genotype Effect - WT Chow vs GSTA4 Chow

combined_enrich$Genotype_Effect_p = apply(combined_enrich[,c("GO_sig_Chow_SV","GO_sig_Chow_GSTA4","numSig_Chow_SV","numSig_Chow_GSTA4")],1,function(a) diffEnrich(a[1],a[2],a[3],a[4],OR=FALSE))

combined_enrich$Genotype_Effect_OddsRatio = apply(combined_enrich[,c("GO_sig_Chow_SV","GO_sig_Chow_GSTA4","numSig_Chow_SV","numSig_Chow_GSTA4")],1,function(a) diffEnrich(a[1],a[2],a[3],a[4],OR=TRUE))

#diff enrich Genotype Effect With Ethanol - WT EtOH vs GSTA4 EtOH

combined_enrich$GenoInEtOH_Effect_p = apply(combined_enrich[,c("GO_sig_EtOH_SV","GO_sig_EtOH_GSTA4","numSig_EtOH_SV","numSig_EtOH_GSTA4")],1,function(a) diffEnrich(a[1],a[2],a[3],a[4],OR=FALSE))

combined_enrich$GenoInEtOH_Effect_OddsRatio = apply(combined_enrich[,c("GO_sig_EtOH_SV","GO_sig_EtOH_GSTA4","numSig_EtOH_SV","numSig_EtOH_GSTA4")],1,function(a) diffEnrich(a[1],a[2],a[3],a[4],OR=TRUE))

combined_enrich[combined_enrich$GenoInEtOH_Effect_p<0.05,]

#diff enrich Genotype Effect With Binge - WT Binge vs GSTA4 Binge

combined_enrich$GenoInBinge_Effect_p = apply(combined_enrich[,c("GO_sig_Binge_SV","GO_sig_Binge_GSTA4","numSig_Binge_SV","numSig_Binge_GSTA4")],1,function(a) diffEnrich(a[1],a[2],a[3],a[4],OR=FALSE))

combined_enrich$GenoInBinge_Effect_OddsRatio = apply(combined_enrich[,c("GO_sig_Binge_SV","GO_sig_Binge_GSTA4","numSig_Binge_SV","numSig_Binge_GSTA4")],1,function(a) diffEnrich(a[1],a[2],a[3],a[4],OR=TRUE))

table(rowSums(combined_enrich[,grep("_Effect_p",colnames(combined_enrich),fixed=TRUE)]<0.05))

### add FDR  ##
combined_enrich$GenoInBinge_Effect_fdr = p.adjust(combined_enrich$GenoInBinge_Effect_p,method="BH")
combined_enrich$GenoInEtOH_Effect_fdr = p.adjust(combined_enrich$GenoInEtOH_Effect_p,method="BH")
combined_enrich$Genotype_Effect_fdr = p.adjust(combined_enrich$Genotype_Effect_p,method="BH")
combined_enrich$BingeInGSTA4_Effect_fdr = p.adjust(combined_enrich$BingeInGSTA4_Effect_p,method="BH")
combined_enrich$Binge_Effect_fdr = p.adjust(combined_enrich$Binge_Effect_p,method="BH")
combined_enrich$EtOHinGSTA4_Effect_fdr = p.adjust(combined_enrich$EtOHinGSTA4_Effect_p,method="BH")
combined_enrich$EtOH_Effect_fdr = p.adjust(combined_enrich$EtOH_Effect_p,method="BH")

tmp_results <- combined_enrich[rowSums(combined_enrich[,grep("_Effect_p",colnames(combined_enrich),fixed=TRUE)]<0.05)>0,]

xx<-as.list(GOTERM)
y <- xx["GO:0002818"]]

anno <- do.call('rbind',lapply(xx[tmp_results$V1],function(y) data.frame(GO_ID = GOID(y),
           Term = Term(y),
           Onto = Ontology(y),
           Defin = Definition(y))))

tmp_results <- merge(anno,tmp_results,by.x="GO_ID",by.y="V1")
write.table(tmp_results,file="~/Desktop/tmp_results_2017-09-07.txt",sep="\t",row.names = FALSE,col.names = TRUE)
```

### 2019-01-02 example programs/keggEnrichment.GSTA4KO.2017-08-30.Rmd
add example programs

---
title: "GSTA4 KO analysis for Colin"
author: "Laura Saba"
date: "8/30/2017"
output: html_document
---

```{r}
rm(list=ls())
options(stringsAsFactors = FALSE)
library(dplyr)
wd <- "~/Documents/Shearn/GSTA4/"

orig <- read.table(file=paste(wd,"data/120d EtOH GSTA4 study Table for Laura Saba.txt",sep=""),sep="\t",quote="",header=TRUE,fill=TRUE)
orig <- orig[orig$Accession!="totals" & orig$Accession!=" " & orig$Accession!="",]

update <- read.table(file=paste(wd,"data/updateProteinID.2017-07-14.txt",sep=""),sep="\t",header=TRUE)
proteins <- merge(orig,update,by.x="Accession",by.y="orig",all=TRUE)
proteins$Accession[!is.na(proteins$new)] = proteins$new[!is.na(proteins$new)]

#convert protein ID to Uniprot symbols
protein_to_uniprot <- read.table(file=paste(wd,"data/protein_to_uniprot.2017-07-14.txt",sep=""),sep="\t",quote="",header=TRUE,fill=TRUE)

proteins <- merge(proteins,protein_to_uniprot,by.x="Accession",by.y="Entry.name",all.x=TRUE)

##problematic proteins
proteins <- proteins[!(proteins$Accession %in% c("PZP_MOUSE","SLF2_MOUSE")),]
```

KEGG Enrichment
```{r,eval=FALSE}
uniprot_to_kegg <- read.table(file="http://rest.kegg.jp/conv/mmu/uniprot",fill=TRUE,sep="\t",quote="")
write.table(uniprot_to_kegg,file=paste(wd,"data/uniprot_to_kegg",Sys.Date(),".txt",sep=""),sep="\t",row.names=FALSE,col.names=FALSE,quote=FALSE)

kegg_to_pathway <- read.table(file="http://rest.kegg.jp/link/pathway/mmu",fill=TRUE,sep="\t",quote="")
write.table(kegg_to_pathway,file=paste(wd,"data/kegg_to_pathway",Sys.Date(),".txt",sep=""),sep="\t",row.names=FALSE,col.names=FALSE,quote=FALSE)

path_names_mmu <- read.table(file="http://rest.kegg.jp/list/pathway/mmu",fill=TRUE,sep="\t",quote="")
write.table(path_names_mmu,file=paste(wd,"data/path_names_mmu",Sys.Date(),".txt",sep=""),sep="\t",row.names=FALSE,col.names=FALSE,quote=FALSE)
```

```{r}
uniprot_to_kegg <- read.table(file=paste(wd,"data/uniprot_to_kegg2017-07-14.txt",sep=""),sep="\t",header=FALSE)
colnames(uniprot_to_kegg) <- c("uniprot_id","gene")
uniprot_to_kegg$Entry <- gsub("up:","",uniprot_to_kegg$uniprot_id)

kegg_to_pathway <- read.table(file=paste(wd,"data/kegg_to_pathway2017-07-14.txt",sep=""),sep="\t",header =FALSE)
colnames(kegg_to_pathway) <- c("gene","pathway")

uniprot_to_pathway <- merge(uniprot_to_kegg,kegg_to_pathway,by="gene")
uniprot_to_pathway <- uniprot_to_pathway[,c("Entry","pathway")]
uniprot_to_pathway <- uniprot_to_pathway[!duplicated(uniprot_to_pathway),]

path_names_mmu <- read.table(file=paste(wd,"data/path_names_mmu2017-07-14.txt",sep=""),sep="\t",header=FALSE)

Chow_GSTA4 <- unique(uniprot_to_pathway$Entry[uniprot_to_pathway$Entry %in% proteins$Entry[proteins$Chow.GSTA4...==1]])
Chow_SV <- unique(uniprot_to_pathway$Entry[uniprot_to_pathway$Entry %in% proteins$Entry[proteins$Chow.SV==1]])
EtOH_GSTA4 <- unique(uniprot_to_pathway$Entry[uniprot_to_pathway$Entry %in% proteins$Entry[proteins$EtOH.GSTA4...==1]])
EtOH_SV <- unique(uniprot_to_pathway$Entry[uniprot_to_pathway$Entry %in% proteins$Entry[proteins$EtOH.SV==1]])
Binge_GSTA4 <- unique(uniprot_to_pathway$Entry[uniprot_to_pathway$Entry %in% proteins$Entry[proteins$Binge.GSTA4...==1]])
Binge_SV <- unique(uniprot_to_pathway$Entry[uniprot_to_pathway$Entry %in% proteins$Entry[proteins$Binge.SV==1]])

pathEnrich <- function(x){
  all_KEGG <- unique(uniprot_to_pathway$Entry)
  sig_KEGG <- unique(x)

  all_KEGG_cnt <- uniprot_to_pathway[uniprot_to_pathway$Entry %in% all_KEGG,]
  sig_KEGG_cnt <- uniprot_to_pathway[uniprot_to_pathway$Entry %in% sig_KEGG,]

  all_KEGG_cnt <- all_KEGG_cnt %>%
    group_by(pathway) %>%
    summarize(KEGG_cnt = length(Entry))

  sig_KEGG_cnt <- sig_KEGG_cnt %>%
    group_by(pathway) %>%
    summarize(KEGG_sig = length(Entry))

  enrich_table <- merge(all_KEGG_cnt,sig_KEGG_cnt,by="pathway",all=TRUE)
  enrich_table$KEGG_sig[is.na(enrich_table$KEGG_sig)] = 0

  enrich_table$numTested <- length(all_KEGG)
  enrich_table$numSig <- length(sig_KEGG)
  enrich_table <- enrich_table %>%
    mutate(expected = (numSig/numTested)*KEGG_cnt)

  enrich_table <- merge(path_names_mmu,enrich_table,by.x="V1",by.y="pathway")

  enrich_table$enrich_p <- NA
  for(i in 1:nrow(enrich_table)){
    a = enrich_table[i,"KEGG_sig"]
    b = enrich_table[i,"KEGG_cnt"] - enrich_table[i,"KEGG_sig"]
    c = enrich_table[i,"numSig"] - enrich_table[i,"KEGG_sig"]
    d = enrich_table[i,"numTested"] - enrich_table[i,"numSig"] + enrich_table[i,"KEGG_sig"]
    enrich_table$enrich_p[i] = fisher.test(matrix(c(a,b,c,d),nr=2),alternative="greater")$p.value
  }

  enrich_table <- enrich_table[order(enrich_table$enrich_p),]
  enrich_table$fdr <- p.adjust(enrich_table$enrich_p,method="BH")
  return(enrich_table)
}

Chow_GSTA4_enrich <- pathEnrich(Chow_GSTA4)
Chow_SV_enrich <- pathEnrich(Chow_SV)
EtOH_GSTA4_enrich <- pathEnrich(EtOH_GSTA4)
EtOH_SV_enrich <- pathEnrich(EtOH_SV)
Binge_GSTA4_enrich <- pathEnrich(Binge_GSTA4)
Binge_SV_enrich <- pathEnrich(Binge_SV)

colnames(Chow_GSTA4_enrich) <- c("V1","V2","KEGG_cnt","KEGG_sig_Chow_GSTA4","numTested","numSig_Chow_GSTA4","expected_Chow_GSTA4","enrich_p_Chow_GSTA4","fdr_Chow_GSTA4")
colnames(Chow_SV_enrich) <- c("V1","V2","KEGG_cnt","KEGG_sig_Chow_SV","numTested","numSig_Chow_SV","expected_Chow_SV","enrich_p_Chow_SV","fdr_Chow_SV")

colnames(EtOH_GSTA4_enrich) <- c("V1","V2","KEGG_cnt","KEGG_sig_EtOH_GSTA4","numTested","numSig_EtOH_GSTA4","expected_EtOH_GSTA4","enrich_p_EtOH_GSTA4","fdr_EtOH_GSTA4")
colnames(EtOH_SV_enrich) <- c("V1","V2","KEGG_cnt","KEGG_sig_EtOH_SV","numTested","numSig_EtOH_SV","expected_EtOH_SV","enrich_p_EtOH_SV","fdr_EtOH_SV")

colnames(Binge_GSTA4_enrich) <- c("V1","V2","KEGG_cnt","KEGG_sig_Binge_GSTA4","numTested","numSig_Binge_GSTA4","expected_Binge_GSTA4","enrich_p_Binge_GSTA4","fdr_Binge_GSTA4")
colnames(Binge_SV_enrich) <- c("V1","V2","KEGG_cnt","KEGG_sig_Binge_SV","numTested","numSig_Binge_SV","expected_Binge_SV","enrich_p_Binge_SV","fdr_Binge_SV")

combined_enrich <- merge(Chow_GSTA4_enrich,Chow_SV_enrich,by=c("V1","V2","KEGG_cnt","numTested"))
combined_enrich <- merge(combined_enrich,EtOH_GSTA4_enrich,by=c("V1","V2","KEGG_cnt","numTested"))
combined_enrich <- merge(combined_enrich,EtOH_SV_enrich,by=c("V1","V2","KEGG_cnt","numTested"))
combined_enrich <- merge(combined_enrich,Binge_GSTA4_enrich,by=c("V1","V2","KEGG_cnt","numTested"))
combined_enrich <- merge(combined_enrich,Binge_SV_enrich,by=c("V1","V2","KEGG_cnt","numTested"))

threshold = 0.01
sum(combined_enrich$fdr_Chow_GSTA4<threshold)
sum(combined_enrich$fdr_Chow_SV<threshold)
sum(combined_enrich$fdr_EtOH_GSTA4<threshold)
sum(combined_enrich$fdr_EtOH_SV<threshold)
sum(combined_enrich$fdr_Binge_GSTA4<threshold)
sum(combined_enrich$fdr_Binge_SV<threshold)

combined_enrich$num_grps_sig <- rowSums(combined_enrich[,grep("fdr",colnames(combined_enrich))]<threshold)

of_interest <- combined_enrich[combined_enrich$num_grps_sig>0 & combined_enrich$num_grps_sig<6,]

groups <- paste(rep(c("Chow","EtOH","Binge"),each=2),rep(c("GSTA4","SV"),3),sep="_")
of_interest[,c("V1","V2",paste("fdr_",groups,sep=""))]

combined_enrich$sig_pattern <- apply(combined_enrich[,grep("fdr",colnames(combined_enrich))],1,function(a) paste(as.numeric(a<threshold),collapse=""))
table(rowSums(combined_enrich[,grep("enrich_p",colnames(combined_enrich))]<threshold))
```

##Differential Enrichment

```{r}
diffEnrich <- function(a,b,c,d,OR){
  y = fisher.test(matrix(c(a,b,c,d),nr=2))
  if(OR) z = y$estimate
  if(!OR) z = y$p.value
  return(z)
}

#diff enrich EtOH effect - WT Chow vs WT EtOH

combined_enrich$EtOH_Effect_p = apply(combined_enrich[,c("KEGG_sig_Chow_SV","KEGG_sig_EtOH_SV","numSig_Chow_SV","numSig_EtOH_SV")],1,function(a) diffEnrich(a[1],a[2],a[3],a[4],OR=FALSE))

combined_enrich$EtOH_Effect_OddsRatio = apply(combined_enrich[,c("KEGG_sig_Chow_SV","KEGG_sig_EtOH_SV","numSig_Chow_SV","numSig_EtOH_SV")],1,function(a) diffEnrich(a[1],a[2],a[3],a[4],OR=TRUE))

#diff enrich EtOH effect in GSTA4 - GSTA4 Chow vs GSTA4 EtOH

combined_enrich$EtOHinGSTA4_Effect_p = apply(combined_enrich[,c("KEGG_sig_Chow_GSTA4","KEGG_sig_EtOH_GSTA4","numSig_Chow_GSTA4","numSig_EtOH_GSTA4")],1,function(a) diffEnrich(a[1],a[2],a[3],a[4],OR=FALSE))

combined_enrich$EtOHinGSTA4_Effect_OddsRatio = apply(combined_enrich[,c("KEGG_sig_Chow_GSTA4","KEGG_sig_EtOH_GSTA4","numSig_Chow_GSTA4","numSig_EtOH_GSTA4")],1,function(a) diffEnrich(a[1],a[2],a[3],a[4],OR=TRUE))

#diff enrich Binge effect - WT Binge vs WT EtOH

combined_enrich$Binge_Effect_p = apply(combined_enrich[,c("KEGG_sig_EtOH_SV","KEGG_sig_Binge_SV","numSig_EtOH_SV","numSig_Binge_SV")],1,function(a) diffEnrich(a[1],a[2],a[3],a[4],OR=FALSE))

combined_enrich$Binge_Effect_OddsRatio = apply(combined_enrich[,c("KEGG_sig_EtOH_SV","KEGG_sig_Binge_SV","numSig_EtOH_SV","numSig_Binge_SV")],1,function(a) diffEnrich(a[1],a[2],a[3],a[4],OR=TRUE))

#diff enrich Binge effect in GSTA4 - GSTA4 Binge vs GSTA4 EtOH

combined_enrich$BingeInGSTA4_Effect_p = apply(combined_enrich[,c("KEGG_sig_EtOH_GSTA4","KEGG_sig_Binge_GSTA4","numSig_EtOH_GSTA4","numSig_Binge_GSTA4")],1,function(a) diffEnrich(a[1],a[2],a[3],a[4],OR=FALSE))

combined_enrich$BingeInGSTA4_Effect_OddsRatio = apply(combined_enrich[,c("KEGG_sig_EtOH_GSTA4","KEGG_sig_Binge_GSTA4","numSig_EtOH_GSTA4","numSig_Binge_GSTA4")],1,function(a) diffEnrich(a[1],a[2],a[3],a[4],OR=TRUE))

#diff enrich Genotype Effect - WT Chow vs GSTA4 Chow

combined_enrich$Genotype_Effect_p = apply(combined_enrich[,c("KEGG_sig_Chow_SV","KEGG_sig_Chow_GSTA4","numSig_Chow_SV","numSig_Chow_GSTA4")],1,function(a) diffEnrich(a[1],a[2],a[3],a[4],OR=FALSE))

combined_enrich$Genotype_Effect_OddsRatio = apply(combined_enrich[,c("KEGG_sig_Chow_SV","KEGG_sig_Chow_GSTA4","numSig_Chow_SV","numSig_Chow_GSTA4")],1,function(a) diffEnrich(a[1],a[2],a[3],a[4],OR=TRUE))

#diff enrich Genotype Effect With Ethanol - WT EtOH vs GSTA4 EtOH

combined_enrich$GenoInEtOH_Effect_p = apply(combined_enrich[,c("KEGG_sig_EtOH_SV","KEGG_sig_EtOH_GSTA4","numSig_EtOH_SV","numSig_EtOH_GSTA4")],1,function(a) diffEnrich(a[1],a[2],a[3],a[4],OR=FALSE))

combined_enrich$GenoInEtOH_Effect_OddsRatio = apply(combined_enrich[,c("KEGG_sig_EtOH_SV","KEGG_sig_EtOH_GSTA4","numSig_EtOH_SV","numSig_EtOH_GSTA4")],1,function(a) diffEnrich(a[1],a[2],a[3],a[4],OR=TRUE))

combined_enrich[combined_enrich$GenoInEtOH_Effect_p<0.05,]

#diff enrich Genotype Effect With Binge - WT Binge vs GSTA4 Binge

combined_enrich$GenoInBinge_Effect_p = apply(combined_enrich[,c("KEGG_sig_Binge_SV","KEGG_sig_Binge_GSTA4","numSig_Binge_SV","numSig_Binge_GSTA4")],1,function(a) diffEnrich(a[1],a[2],a[3],a[4],OR=FALSE))

combined_enrich$GenoInBinge_Effect_OddsRatio = apply(combined_enrich[,c("KEGG_sig_Binge_SV","KEGG_sig_Binge_GSTA4","numSig_Binge_SV","numSig_Binge_GSTA4")],1,function(a) diffEnrich(a[1],a[2],a[3],a[4],OR=TRUE))

table(rowSums(combined_enrich[,grep("_Effect_p",colnames(combined_enrich),fixed=TRUE)]<0.05))

combined_enrich[combined_enrich$EtOHinGSTA4_Effect_p<0.05,]

tmp_results <- combined_enrich[rowSums(combined_enrich[,grep("_Effect_p",colnames(combined_enrich),fixed=TRUE)]<0.20)>0,]

write.table(tmp_results,file="~/Desktop/tmp_results_2017-08-30.txt",sep="\t",row.names = FALSE,col.names = TRUE)
```

Proteins from Ribosome KEGG Pathway
```{r}

ribosome_genes <- kegg_to_pathway[grep("mmu03010",kegg_to_pathway$pathway),]
ribosome_proteins <- uniprot_to_kegg[uniprot_to_kegg$gene %in% ribosome_genes$gene,]
ribosome_withAnno <- proteins[proteins$Entry %in% ribosome_proteins$Entry,]

write.table(ribosome_withAnno,file=paste(wd,"results/KEGGpathway_Ribosome_mmu03010.2017-09-08.txt",sep=""),sep="\t",row.names = FALSE,col.names = TRUE)
```

### 2019-01-09 README.Rmd
updated Description and knit README.rmd

---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```
# diffEnrich

The goal of diffEnrich is Given a List of Gene Symbols, Performs Differential 
Enrichment Ananlysis. # diffEnrich

development of methods and software for differential enrichment analysis

Papers from Saba Lab where differentially enrichment has been used:
1. [Shearn et al 2018 - ACER](https://onlinelibrary.wiley.com/doi/full/10.1111/acer.13766)

## Installation

You can install the released version of diffEnrich from [CRAN](https://CRAN.R-project.org) with:

``` r
install.packages("diffEnrich")
```

## Example

This is a basic example which shows you how to solve a common problem:

```{r example}
## basic example code
```

What is special about using `README.Rmd` instead of just `README.md`? You can include R chunks like so:

```{r cars}
summary(cars)
```

You'll still need to render `README.Rmd` regularly, to keep `README.md` up-to-date.

You can also embed plots, for example:

```{r pressure, echo = FALSE}
plot(pressure)
```

In that case, don't forget to commit and push the resulting figure files, so they display on GitHub!

### 2019-04-26 R/get_kegg.R
Added code for user to provide any path for file export

@@ -26,47 +26,124 @@
 #'
 get_kegg <- function(species, path = NULL){
   options(stringsAsFactors = F)
-  # Define user's base file path
-  base_path <- here::here()
-  flist <- list.files(base_path)
+  ## Argument checks
   if(missing(species)){stop("Must choose one of the 3 species options: human, mouse, rat")}
   if(missing(species) | !(species %in% c('hsa','mmu','rno'))){stop("Must choose one of the 3 species options: human: use 'hsa', mouse: use 'mmu', rat: use rno")}
-  else {
-    # Define base api path and define list of operations/arguments
-    api_base <- "http://rest.kegg.jp/"
-    op <- list("info"="info", "list"="list", "find"="find", "get"="get",
-               "conv"="conv", "link"="link", "ddi"="ddi")
-    db <- list("pathway"="pathway")
-    org <- list("hsa"="hsa", "mmu"="mmu", "rno"="rno")
-    # Build api paths for
-    # 1) ncbi to kegg
-    ncbi_to_kegg_path <- paste0(api_base, op[["conv"]], "/",
-                                org[[species]], "/", "ncbi-geneid")
-    # 2) kegg to pathway
-    kegg_to_pathway_path <- paste0(api_base, op[["link"]], "/", db[["pathway"]],
-                                   "/", org[[species]])
-    # 3) pathway to species
-    pathway_to_species_path <- paste0(api_base, op[["list"]], "/",
-                                      db[["pathway"]],
-                                      "/", org[[species]])
-    # 4) pathway to kegg release
-    pathway_to_kegg_release <- paste0(api_base, op[["info"]], "/",
-                                      "kegg")
-    ## api pull
-    kegg_release <- utils::read.table(file = pathway_to_kegg_release,
-                                      fill = TRUE,
-                                      sep = "\t",
-                                      quote = "")[2, 1]
-    kegg_release <- stringr::str_extract(kegg_release, ".{0,0}Release.{0,30}")
-    kegg_release <- gsub(",", "", kegg_release, fixed = T)
-    kegg_release <- gsub(" ", "_", kegg_release, fixed = T)
-    kegg_release <- gsub("/", "_", kegg_release, fixed = T)
-    find_files <- c(paste("ncbi_to_kegg",Sys.Date(), kegg_release, ".txt",sep=""),
-                    paste("kegg_to_pathway",Sys.Date(), kegg_release, ".txt",sep=""),
[…]

### 2019-04-26 R/get_kegg.R
updated get_kegg documentation

@@ -5,6 +5,8 @@
 #' written to the working directory.
 #'
 #' @param species character. The species to use in kegg data pull
+#' @param path character. A character string describing the path to write out KEGG
+#' API data sets. If not provided, defaults to current working directory.
 #'
 #' @return kegg_out: A named list of the data pulled from kegg api when the
 #' function was run. This may be different if the function is run at
@@ -21,7 +23,7 @@
 #' @examples
 #' kegg <- get_kegg(species = "rno")
 #' \dontrun{
-#' kegg <- get_kegg(species = "mmu")
+#' kegg <- get_kegg(species = "mmu", path = "usr/data/out/")
 #' }
 #'
 get_kegg <- function(species, path = NULL){

### 2019-04-26 R/pathEnrich.R
changed KEGG_sig to KEGG_in_list

@@ -64,11 +64,11 @@ pathEnrich <- function(gk_obj, gene_list){
   sig_KEGG_cnt <- ncbi_to_pathway %>%
     dplyr::filter(.data$Entry %in% sig_KEGG) %>%
     dplyr::group_by(.data$pathway) %>%
-    dplyr::summarize(KEGG_sig = length(.data$Entry))
+    dplyr::summarize(KEGG_in_list = length(.data$Entry))
 
   ## Set up enrichment table
   enrich_table <- merge(all_KEGG_cnt, sig_KEGG_cnt, by = "pathway", all = TRUE)
-  enrich_table$KEGG_sig[is.na(enrich_table$KEGG_sig)] = 0
+  enrich_table$KEGG_in_list[is.na(enrich_table$KEGG_in_list)] = 0
   enrich_table <- enrich_table %>%
     dplyr::mutate(numTested = length(all_KEGG),
                   numSig = length(sig_KEGG),
@@ -78,10 +78,10 @@ pathEnrich <- function(gk_obj, gene_list){
   ## Perform Fisher's test
   enrich_table$enrich_p <- NA
   for(i in 1:nrow(enrich_table)){
-    a = enrich_table[i, "KEGG_sig"]
-    b = enrich_table[i, "KEGG_cnt"] - enrich_table[i, "KEGG_sig"]
-    c = enrich_table[i, "numSig"] - enrich_table[i, "KEGG_sig"]
-    d = enrich_table[i, "numTested"] - enrich_table[i, "numSig"] + enrich_table[i, "KEGG_sig"]
+    a = enrich_table[i, "KEGG_in_list"]
+    b = enrich_table[i, "KEGG_cnt"] - enrich_table[i, "KEGG_in_list"]
+    c = enrich_table[i, "numSig"] - enrich_table[i, "KEGG_in_list"]
+    d = enrich_table[i, "numTested"] - enrich_table[i, "numSig"] + enrich_table[i, "KEGG_in_list"]
     enrich_table$enrich_p[i] = stats::fisher.test(matrix(c(a,b,c,d), nrow = 2), alternative = "greater")$p.value
   }
 

### 2019-04-26 R/pathEnrich.R
Fixed numSig bug

@@ -55,7 +55,7 @@ pathEnrich <- function(gk_obj, gene_list){
   pathway_to_species <- gk_obj[["pathway_to_species"]]
 
   ## set up enrichment
-  all_KEGG <- unique(ncbi_to_kegg$Entry)
+  all_KEGG <- unique(ncbi_to_pathway$Entry)
   sig_KEGG <- unique(gene_list)
   all_KEGG_cnt <- ncbi_to_pathway %>%
     dplyr::filter(.data$Entry %in% all_KEGG) %>%
@@ -71,7 +71,7 @@ pathEnrich <- function(gk_obj, gene_list){
   enrich_table$KEGG_in_list[is.na(enrich_table$KEGG_in_list)] = 0
   enrich_table <- enrich_table %>%
     dplyr::mutate(numTested = length(all_KEGG),
-                  numSig = length(sig_KEGG),
+                  numSig = length(all_KEGG[which(all_KEGG %in% sig_KEGG)]),
                   expected = (.data$numSig/.data$numTested)*.data$KEGG_cnt)
   enrich_table <- merge(pathway_to_species, enrich_table, by.x = "V1", by.y = "pathway")
 

### 2019-04-26 R/pathEnrich.R
updated pathEnrich.rd. Also cleaned up imports.

@@ -25,7 +25,6 @@
 #' @importFrom rlang .data
 #' @import dplyr
 #' @import utils
-#' @import org.Rn.eg.db
 #' @examples
 #'
 #' sig_pe <- pathEnrich(gk_obj = kegg, gene_list = geneLists$sigGenes)

### 2019-04-26 README.Rmd
Added get_kegg example to README

@@
 
-The goal of diffEnrich is Given a List of Gene Symbols, Performs Differential 
-Enrichment Ananlysis. # diffEnrich
+The goal of diffEnrich is simple. Given a list of gene symbols, *diffEnrich*  will
+perform differential enrichment ananlysis using the Kyoto Encyclopedia of Genes
+and Genomes (KEGG) REST API. This package provides a number of functions that are 
+intended to be used in a pipeline (See Figure N). Breifly, the workflow will
+download and clean species specific ENTREZ gene IDs and map them to their respective
+KEGG pathways by accessing KEGG's REST API. This way the user will always have 
+the most up-to-date pathway data from KEGG. Next, the user will identify significantly
+enriched pathways from two different gene sets, and finally, the user will identify 
+pathways that are differentially expressed between the two gene sets. In addition to 
+the analysis pipeline, this package also provides functions for data visualizations.
 
@@
 
-This is a basic example which shows you how to solve a common problem:
+First we will use the \code{get_kegg} function access the KEGG REST API and download the data sets required to perform 
+our downstream analysis. This function takes two arguments. The first, \code{species} is required and is the species of 
+interest. Currently, *diffEnrich* supports three species, and the argument is a character string using the KEGG
+code [REF](https://www.pnas.org/content/suppl/2008/09/11/0806162105.DCSupplemental/ST1_PDF.pdf): 
+Homo sapiens (human), use 'hsa'; Mus musculus (mouse), use 'mmu', and Rattus norvegicus (rat), use 'rno'. The second,
+\code{path} is also passed as a character string, and is the path of the directory in which the user would like to
+write the data sets downloaded from the KEGG REST API. If the user does not provide a path, the data sets will 
+be automatically written to the current working directory using the \code{here::here()} functionality. These data 
+sets will be tab delimited files with a name describing the data, and for reproducibility the date they were generated
+and the version of KEGG when the API was accessed. In addition to these flat files, \code{get_kegg} will also
+create a named list with the three relavent KEGG data sets. The names of this list will describe the the data set.
 
-```{r example}
-## basic example code
+```{r get_kegg_exp1}
+suppressMessages(library(diffEnrich))
+
+## run get_kegg() using rat
+kegg_rno <- get_kegg('rno')
 ```
 
-What is special about using `README.Rmd` instead of just `README.md`? You can include R chunks like so:
+Here are the files:
 
-```{r cars}
-summary(cars)
+    kegg_to_pathway2019-04-26Release_90.0+_04-26_Apr_19.txt
+    ncbi_to_kegg2019-04-26Release_90.0+_04-26_Apr_19.txt
+    pathway_to_species2019-04-26Release_90.0+_04-26_Apr_19.txt
+
+**Note:** Because it is assumed that the user might want to use the data sets generated by \code{get_kegg}, it is careful not to
+overwrite data sets with exact names. \code{get_kegg} checks the path provided for data sets generated 'same-day/same-version', 
+and if it finds even one of the three, it will not re-write any of the data sets. It will still however, let the user know
+it is not writing out new data sets and still generate the named list object. Users can of course generate 'same-day/same-version'
+data sets in different directories if they so choose.
+
+```{r get_kegg_exp2}
+## run get_kegg() using rat
+kegg_rno <- get_kegg('rno')
 ```
-
-You'll still need to render `README.Rmd` regularly, to keep `README.md` up-to-date.
-
-You can also embed plots, for example:
-
-```{r pressure, echo = FALSE}
-plot(pressure)
-```
-
-In that case, don't forget to commit and push the resulting figure files, so they display on GitHub!

### 2019-06-24 R/combineEnrich.R
Use logical to numeric to summarize sig_pattern

@@ -40,10 +40,14 @@ combineEnrich <- function(sig_pe, bkg_pe, threshold = 0.05, range){
 
   ## Get sum of significantly enriched genes based on fdr cutoff
   combined_enrich$num_groups_sig <- rowSums(combined_enrich[, grep("fdr", colnames(combined_enrich))] < threshold)
+
   ## Get group range of interest and subset data to only include pathways and fdr values
   of_interest <- combined_enrich[combined_enrich$num_groups_sig > range[1] & combined_enrich$num_groups_sig < range[2], ]
   groups <- c("sig", "bkg")
   of_interest_subset <- of_interest[, c("V1", "V2", paste("fdr", groups, sep = "_"))]
+
+  ## Use logical to numeric to summarize sig_pattern
+  combined_enrich$sig_pattern <- apply(combined_enrich[, grep("fdr", colnames(combined_enrich))], 1, function(a) paste(as.numeric(a<threshold), collapse = ""))
 
 
 }

### 2019-06-24 R/combineEnrich.R
generate table of sig

@@ -48,6 +48,7 @@ combineEnrich <- function(sig_pe, bkg_pe, threshold = 0.05, range){
 
   ## Use logical to numeric to summarize sig_pattern
   combined_enrich$sig_pattern <- apply(combined_enrich[, grep("fdr", colnames(combined_enrich))], 1, function(a) paste(as.numeric(a<threshold), collapse = ""))
+  table(rowSums(combined_enrich[, grep("enrich_p", colnames(combined_enrich))]<threshold))
 
 
 }

### 2019-06-24 R/combineEnrich.R
made list.out for combineEnrich

@@ -48,7 +48,11 @@ combineEnrich <- function(sig_pe, bkg_pe, threshold = 0.05, range){
 
   ## Use logical to numeric to summarize sig_pattern
   combined_enrich$sig_pattern <- apply(combined_enrich[, grep("fdr", colnames(combined_enrich))], 1, function(a) paste(as.numeric(a<threshold), collapse = ""))
-  table(rowSums(combined_enrich[, grep("enrich_p", colnames(combined_enrich))]<threshold))
+  sig_sum <- as.data.frame(table(rowSums(combined_enrich[, grep("enrich_p", colnames(combined_enrich))]<threshold)))
 
-
+  ## Put out objects in list *THIS WILL NEED TO BE UPDATED*
+  list.out <- list("combined_enrich" = combined_enrich,
+                   "of_interest" = of_interest_subset,
+                   "sig_sum" = sig_sum)
+  return(list.out)
 }

### 2019-06-24 R/combineEnrich.R
Added argument check for range parameter

@@ -30,6 +30,7 @@ combineEnrich <- function(sig_pe, bkg_pe, threshold = 0.05, range){
   if(missing(sig_pe)){stop("Argument missing: sig_pe")}
   if(missing(bkg_pe)){stop("Argument missing: bkg_pe")}
   if(missing(threshold)){stop("Argument missing: threshold. Please provide an FDR cutoff for gene inclusion in summary table")}
+  if(missing(range)){stop("Argument missing: range. Please provide a range of FDR cutoffs to filter of_interest pathways")}
 
   ## Merge results from first enrichment
   combined_enrich <- merge(sig_pe, bkg_pe, by = c("V1", "V2", "KEGG_cnt", "numTested"))

### 2019-06-24 R/combineEnrich.R
Added range param to roxygen skeleton

@@ -14,6 +14,8 @@
 #' generated from \code{\link{pathEnrich}}. See example for \code{\link{pathEnrich}}.
 #' @param threshold numeric. FDR cutoff for significant genes included in summary table.
 #' (Default = 0.05).
+#' @param range numeric vector. Vector containing the smallest and largest number of
+#' significant groups for use in filtering pathways of interest (e.g. \code{c(0,6)}.
 #'
 #' @return combined_enrich: An object of class data.frame that is the result of merging
 #' \code{sig_pe} and \code{bkg_pe}, using the default joining implemented in the base

### 2019-06-24 R/combineEnrich.R
Added example to roxygen skeleton

@@ -27,6 +27,8 @@
 #' @export
 #'
 #' @examples
+#' combined_enrich <- combineEnrich(sig_pe, bkg_pe, threshold = 0.05, range = c(0, 6))
+#'
 combineEnrich <- function(sig_pe, bkg_pe, threshold = 0.05, range){
   ## argument check
   if(missing(sig_pe)){stop("Argument missing: sig_pe")}

### 2019-06-24 R/combineEnrich.R
Updated example for combineEnrich

@@ -27,6 +27,10 @@
 #' @export
 #'
 #' @examples
+#' ## Generate individual enrichment results
+#' sig_pe <- pathEnrich(gk_obj = kegg, gene_list = geneLists$sigGenes)
+#' bkg_pe <- pathEnrich(gk_obj = kegg, gene_list = geneLists$background)
+#' ## Combine
 #' combined_enrich <- combineEnrich(sig_pe, bkg_pe, threshold = 0.05, range = c(0, 6))
 #'
 combineEnrich <- function(sig_pe, bkg_pe, threshold = 0.05, range){

### 2019-06-24 README.Rmd
fixed README.Rmd so code is ital

@@
 
-First we will use the \code{get_kegg} function access the KEGG REST API and download the data sets required to perform 
-our downstream analysis. This function takes two arguments. The first, \code{species} is required and is the species of 
+First we will use the *get_kegg* function access the KEGG REST API and download the data sets required to perform 
+our downstream analysis. This function takes two arguments. The first, 'species' is required and is the species of 
 interest. Currently, *diffEnrich* supports three species, and the argument is a character string using the KEGG
@@
 Homo sapiens (human), use 'hsa'; Mus musculus (mouse), use 'mmu', and Rattus norvegicus (rat), use 'rno'. The second,
-\code{path} is also passed as a character string, and is the path of the directory in which the user would like to
+'path' is also passed as a character string, and is the path of the directory in which the user would like to
 write the data sets downloaded from the KEGG REST API. If the user does not provide a path, the data sets will 
-be automatically written to the current working directory using the \code{here::here()} functionality. These data 
-sets will be tab delimited files with a name describing the data, and for reproducibility the date they were generated
-and the version of KEGG when the API was accessed. In addition to these flat files, \code{get_kegg} will also
+be automatically written to the current working directory using the *here::here()* functionality. These data 
+sets will be tab delimited files with a name describing the data, and for reproducibility, the date they were generated
+and the version of KEGG when the API was accessed. In addition to these flat files, *get_kegg* will also
 create a named list with the three relavent KEGG data sets. The names of this list will describe the the data set.
@@
 
-**Note:** Because it is assumed that the user might want to use the data sets generated by \code{get_kegg}, it is careful not to
-overwrite data sets with exact names. \code{get_kegg} checks the path provided for data sets generated 'same-day/same-version', 
+**Note:** Because it is assumed that the user might want to use the data sets generated by *get_kegg*, it is careful not to
+overwrite data sets with exact names. *get_kegg* checks the path provided for data sets generated 'same-day/same-version', 
 and if it finds even one of the three, it will not re-write any of the data sets. It will still however, let the user know

### 2019-06-24 README.Rmd
Started step 2 of README

@@
 
+### Step 1: Collect and clean pathways from KEGG API
 First we will use the *get_kegg* function access the KEGG REST API and download the data sets required to perform 
@@
 ```
+
+### Step 2: Perform individual enrichment analysis
+
+In this step we will use the *pathEnrich* function to identify KEGG pathways that are enriched based on a list of genes we
+are interested in and based on a list of background genes. This function may not always use the complete list of genes provided 
+by the user. Specifically, it will only use the genes from the list provided that are also in
+the most current species list pulled from the KEGG REST API using *get_kegg*, or from the older KEGG data loaded by the user from a 
+previous *get_kegg* call. The *pathEnrich* function should be run at least twice, once for the genes of interest and once for the background.
+Each *pathEnrich* call generates a dataframe summarizing the results of a traditional pathway enrichment analysis in which a Fisher's
+Exact test is used to identify which KEGG pathways are enriched by the user's list of interesting genes with repect to background
+enrichment.
+
+```{r pathEnrich}
+# run pathEnrich using kegg_rno
+## Genes of interest
+sig_pe <- pathEnrich(gk_obj = kegg, gene_list = geneLists$sigGenes)
+## Background
+bkg_pe <- pathEnrich(gk_obj = kegg, gene_list = geneLists$background)
+head(sig_pe)
+```
+
+*pathEnrich* generates a dataframe with 9 columns described below.

### 2019-06-24 README.Rmd
Added kableExtra table for pathEnrich in README

@@
 bkg_pe <- pathEnrich(gk_obj = kegg, gene_list = geneLists$background)
-head(sig_pe)
+knitr::kable(head(sig_pe),
+             caption = "Table 1. Head of sig_pe dataframe generated using pathEnrich") %>%
+  kableExtra::kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = T)
 ```

### 2019-06-24 README.Rmd
removed rownames from pathEnrich table in README

@@
 knitr::kable(head(sig_pe),
-             caption = "Table 1. Head of sig_pe dataframe generated using pathEnrich") %>%
+             caption = "Table 1. Head of sig_pe dataframe generated using pathEnrich", row.names = FALSE) %>%
   kableExtra::kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = T)

### 2019-06-24 README.Rmd
Added Table 2 to README

@@
 output: github_document
+always_allow_html: yes
 ---
@@
 
-```{r pathEnrich}
+```{r pathEnrich, warning=FALSE}
 # run pathEnrich using kegg_rno
@@
 *pathEnrich* generates a dataframe with 9 columns described below.
+
+```{r pathEnrich_description, echo=FALSE}
+# get column names
+names <- colnames(sig_pe)
+# Define column names
+description <- c("KEGG ID",
+                 "KEGG pathway description",
+                 "PLACE HOLDER",
+                 "PLACE HOLDER",
+                 "Total number of pathways in KEGG release",
+                 "PLACE HOLDER",
+                 "PLACE HOLDER",
+                 "Unadjusted p-value from Fisher's Exact test",
+                 "False Discovery Rate (FDR)")
+# Make dataframe
+df1 <- data.frame(names, description)
+colnames(df1) <- c("Column Names", "Column Description")
+
+knitr::kable(df1,
+             caption = "Table 2. Description of columns is dataframe generated by pathEnrich", row.names = FALSE) %>%
+  kableExtra::kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = F)
+```

### 2019-07-01 R/pathEnrich.R
Add menaingful column names for the C1 and C2 in pathEnrich

@@ -87,5 +87,7 @@ pathEnrich <- function(gk_obj, gene_list){
   ## clean and return
   enrich_table <- enrich_table[order(enrich_table$enrich_p), ]
   enrich_table$fdr <- stats::p.adjust(enrich_table$enrich_p, method = 'BH')
+  ## Add menaingful column names for the C1 and C2
+  colnames(enrich_table)[c(1,2)] <- c("KEGG_ID", "KEGG_description")
   return(enrich_table)
 }

### 2019-07-01 R/diffEnrich.R
updated combineEnrich now that column names changed in pathEnrich out

@@ -72,7 +72,7 @@ diffEnrich <- function(sig_pe, bkg_pe){
   if(missing(bkg_pe)){stop("Argument missing: bkg_pe")}
 
   ## Merge results from first enrichment
-  combined_enrich <- merge(sig_pe, bkg_pe, by = c("V1", "V2", "KEGG_cnt", "numTested"))
+  combined_enrich <- merge(sig_pe, bkg_pe, by = c("KEGG_ID", "KEGG_description", "KEGG_cnt", "numTested"))
   colnames(combined_enrich) <- gsub(".x", "_sig", colnames(combined_enrich), fixed = TRUE)
   colnames(combined_enrich) <- gsub(".y", "_bkg", colnames(combined_enrich), fixed = TRUE)
 

### 2019-07-01 README.Rmd
Added Formula 1 to README

@@
 
-In this step we will use the *pathEnrich* function to identify KEGG pathways that are enriched based on a list of genes we
+In this step we will use the *pathEnrich* function to identify KEGG pathways that are enriched (i.e. over-represented) based on a list of genes we
 are interested in and based on a list of background genes. This function may not always use the complete list of genes provided 
@@
 previous *get_kegg* call. The *pathEnrich* function should be run at least twice, once for the genes of interest and once for the background.
-Each *pathEnrich* call generates a dataframe summarizing the results of a traditional pathway enrichment analysis in which a Fisher's
+Each *pathEnrich* call generates a dataframe summarizing the results of an enrichment analysis in which a Fisher's
 Exact test is used to identify which KEGG pathways are enriched by the user's list of interesting genes with repect to background
-enrichment.
+enrichment, and a p-value is calculated using a hypergeometric distribution (Formula 1). P-values are adjusted for multiple comparisons by controlling the False Discovery Rate (FDR) at 0.05.
+
+$Formula 1:$
+
+  $$P = 1-\sum_{i=0}^{k-1} \frac{\binom{m}{i}\binom{N-M}{n-i}}{\binom{N}{n}}$$
 

### 2019-07-01 README.Rmd
changed formula formatting

@@
 
-$Formula 1:$
+Formula 1:
 
-  $$P = 1-\sum_{i=0}^{k-1} \frac{\binom{m}{i}\binom{N-M}{n-i}}{\binom{N}{n}}$$
+$$ P = 1-\sum_{i=0}^{k-1} \frac{ \binom{m} {i} \binom{N-M} {n-i} } {\binom{N} {n} } $$
+
+where N is the total number of genes in the background, M is the number of genes within that background that also in the gene set of interest, n is the size of the list of genes of interest and k is the number of genes within that are in the background.
+The background by default is all the genes that have annotation.
 

### 2019-07-01 README.Rmd
Fixed github_document equation rendering
Added diffEnrich step to README

@@
 ---
-output: github_document
+output: 
+  github_document:
+    pandoc_args: --webtex
 always_allow_html: yes
@@
 ```
+
+### Step 3: Identify differentially enriched KEGG pathways
+
+The diffEnrich function will merge the results from the *pathEnrich* calls generated above. Specifically, the dataframe
+'sig_pe' and the dataframe 'bkg_pe' will be merged using the base *merge* functionality and be merged by the following columns:
+'c("KEGG_ID", "KEGG_description", "KEGG_cnt", "numTested")'. This merged data set will then be used to perform differential 
+enrichment usng the same method and p-value calculation as described above.
+
+```{r diffEnrich}
+## Perform differential enrichment
+diff_enrich <- diffEnrich(sig_pe = sig_pe, bkg_pe = bkg_pe)
+
+knitr::kable(head(diff_enrich),
+             caption = "Table 2. Description of columns is dataframe generated by pathEnrich", row.names = FALSE) %>%
+  kableExtra::kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = F)
+```
+
+The result of the *diffEnrich* call is the merged dataframe with the estimated odds ration generated by the Fisher's Exact test 
+and the associated p-value.

### 2019-07-01 README.Rmd
Spell Check and knit draft of README for github

--- 2019-07-01 README.Rmd
@@
 The goal of diffEnrich is simple. Given a list of gene symbols, *diffEnrich*  will
-perform differential enrichment ananlysis using the Kyoto Encyclopedia of Genes
+perform differential enrichment analysis using the Kyoto Encyclopedia of Genes
 and Genomes (KEGG) REST API. This package provides a number of functions that are 
-intended to be used in a pipeline (See Figure N). Breifly, the workflow will
+intended to be used in a pipeline (See Figure N). Briefly, the workflow will
 download and clean species specific ENTREZ gene IDs and map them to their respective
@@
 and the version of KEGG when the API was accessed. In addition to these flat files, *get_kegg* will also
-create a named list with the three relavent KEGG data sets. The names of this list will describe the the data set.
+create a named list with the three relevant KEGG data sets. The names of this list will describe the the data set.
 
@@
 previous *get_kegg* call. The *pathEnrich* function should be run at least twice, once for the genes of interest and once for the background.
-Each *pathEnrich* call generates a dataframe summarizing the results of an enrichment analysis in which a Fisher's
-Exact test is used to identify which KEGG pathways are enriched by the user's list of interesting genes with repect to background
+Each *pathEnrich* call generates a data frame summarizing the results of an enrichment analysis in which a Fisher's
+Exact test is used to identify which KEGG pathways are enriched by the user's list of interesting genes with respect to background
 enrichment, and a p-value is calculated using a hypergeometric distribution (Formula 1). P-values are adjusted for multiple comparisons by controlling the False Discovery Rate (FDR) at 0.05.
@@
 
-*pathEnrich* generates a dataframe with 9 columns described below.
+*pathEnrich* generates a data frame with 9 columns described below.
 
@@
 ```
+
+### Step 3: Identify differentially enriched KEGG pathways
+
+The diffEnrich function will merge the results from the *pathEnrich* calls generated above. Specifically, the data frame
+'sig_pe' and the data frame 'bkg_pe' will be merged using the base *merge* functionality and be merged by the following columns:
+'c("KEGG_ID", "KEGG_description", "KEGG_cnt", "numTested")'. This merged data set will then be used to perform differential 
+enrichment using the same method and p-value calculation as described above.
+
+```{r diffEnrich}
+## Perform differential enrichment
+diff_enrich <- diffEnrich(sig_pe = sig_pe, bkg_pe = bkg_pe)
+
+knitr::kable(head(diff_enrich),
+             caption = "Table 2. Description of columns is dataframe generated by pathEnrich", row.names = FALSE) %>%
+  kableExtra::kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = F)
+```
+
+The result of the *diffEnrich* call is the merged data frame with the estimated odds ration generated by the Fisher's Exact test 
+and the associated p-value.

### 2019-07-01 R/get_kegg.R
fixed message when user provides path

@@ -180,7 +180,7 @@ get_kegg <- function(species, path = NULL){
                                               quote = "")
 
       message("3 data sets will be written as tab delimited text files")
-      message("File location: ", here::here())
+      message("File location: ", path)
       message("Kegg Release: ", kegg_release)
       # write files
       utils::write.table(ncbi_to_kegg,

### 2019-07-01 R/pathEnrich.R
no change

@@ -87,6 +87,7 @@ pathEnrich <- function(gk_obj, gene_list){
   ## clean and return
   enrich_table <- enrich_table[order(enrich_table$enrich_p), ]
   enrich_table$fdr <- stats::p.adjust(enrich_table$enrich_p, method = 'BH')
+
   ## Add menaingful column names for the C1 and C2
   colnames(enrich_table)[c(1,2)] <- c("KEGG_ID", "KEGG_description")
   return(enrich_table)

### 2019-07-01 R/pathEnrich.R
Removed "path:" from KEGG ID

@@ -87,7 +87,7 @@ pathEnrich <- function(gk_obj, gene_list){
   ## clean and return
   enrich_table <- enrich_table[order(enrich_table$enrich_p), ]
   enrich_table$fdr <- stats::p.adjust(enrich_table$enrich_p, method = 'BH')
-
+  enrich_table$V1 <- gsub("path:", "", enrich_table$V1, fixed = TRUE)
   ## Add menaingful column names for the C1 and C2
   colnames(enrich_table)[c(1,2)] <- c("KEGG_ID", "KEGG_description")
   return(enrich_table)

### 2019-07-01 data-raw/create_sample_data.R
Changed names in data file to list1 and list2

@@ -21,14 +21,14 @@ backgroundENTZ <- mapIds(org.Rn.eg.db, backgroundList, 'ENTREZID', 'SYMBOL')
 backgroundENTZ <- base::unname(backgroundENTZ)
 sigGeneENTZ <- mapIds(org.Rn.eg.db, sigGenesList, 'ENTREZID', 'SYMBOL')
 sigGeneENTZ <- base::unname(sigGeneENTZ)
-geneLists <- list(background = backgroundENTZ,
-                  sigGenes = sigGeneENTZ)
+geneLists <- list(list1 = backgroundENTZ,
+                  list2 = sigGeneENTZ)
 
 ## save and compress
-usethis::use_data(geneLists, compress = "xz")
+usethis::use_data(geneLists, compress = "xz", overwrite = TRUE)
 
 ## generate kegg data
-kegg <- get_kegg('rat')
+kegg <- get_kegg('rno')
 
 ## save and compress
-usethis::use_data(kegg, compress = "xz")
+usethis::use_data(kegg, compress = "xz", overwrite = TRUE)

### 2019-07-01 R/pathEnrich.R
Changed colnames to pathErich res

@@ -9,7 +9,7 @@
 #' defined in \code{get_kegg}. If the user wishes to use an older version of data
 #' generated by \code{get_kegg}, they must first load that data and put it in a named
 #' list that matches the names given in the list generated by \code{get_kegg}.
-#' @param gene_list Vector. Vector of NCBI (ENTREZ) geneIDs
+#' @param gene_list Vector. Vector of NCBI (ENTREZ) geneIDs.
 #'
 #' @return enrich_table: An object of class data.frame that summarizes the results
 #' of the pathway analysis.
@@ -27,9 +27,9 @@
 #' @import utils
 #' @examples
 #'
-#' sig_pe <- pathEnrich(gk_obj = kegg, gene_list = geneLists$sigGenes)
+#' list1_pe <- pathEnrich(gk_obj = kegg, gene_list = geneLists$list1)
 #' \dontrun{
-#' bkg_pe <- pathEnrich(gk_obj = kegg, gene_list = geneLists$background)
+#' list2_pe <- pathEnrich(gk_obj = kegg, gene_list = geneLists$list2)
 #' }
 #'
 pathEnrich <- function(gk_obj, gene_list){
@@ -89,6 +89,8 @@ pathEnrich <- function(gk_obj, gene_list){
   enrich_table$fdr <- stats::p.adjust(enrich_table$enrich_p, method = 'BH')
   enrich_table$V1 <- gsub("path:", "", enrich_table$V1, fixed = TRUE)
   ## Add menaingful column names for the C1 and C2
-  colnames(enrich_table)[c(1,2)] <- c("KEGG_ID", "KEGG_description")
+  colnames(enrich_table)[c(1,2)] <- c("KEGG_ID", "KEGG_description", "KEGG_cnt", "KEGG_in_list",
+                                      "num_in_background", "num_in_list", "expected", "enrich_p",
+                                      "fdr")
   return(enrich_table)
 }

### 2019-07-01 R/diffEnrich.R
Carried over names changes of list1 and list2 to diffEnrich

@@ -8,9 +8,9 @@
 #' pathways with their associated estimated odds ratio, unadjusted p-value, and fdr adjusted
 #' p-value.
 #'
-#' @param sig_pe data.frame. Dataframe of enrichment results for genes of interest
+#' @param list1_pe data.frame. Dataframe of enrichment results for genes of interest
 #' generated from \code{\link{pathEnrich}}. See example for \code{\link{pathEnrich}}.
-#' @param  bkg_pe data.frame. Dataframe of enrichment results for background genes
+#' @param list2_pe data.frame. Dataframe of enrichment results for background genes
 #' generated from \code{\link{pathEnrich}}. See example for \code{\link{pathEnrich}}.
 #'
 #' @return data.frame. Dataframe generated from merging pathEnrich dataframes with the following added columns:
@@ -22,15 +22,15 @@
 #'
 #' @examples
 #' ## Generate individual enrichment reults
-#' sig_pe <- pathEnrich(gk_obj = kegg, gene_list = geneLists$sigGenes)
-#' bkg_pe <- pathEnrich(gk_obj = kegg, gene_list = geneLists$background)
+#' list1_pe <- pathEnrich(gk_obj = kegg, gene_list = geneLists$sigGenes)
+#' list2_pe <- pathEnrich(gk_obj = kegg, gene_list = geneLists$background)
 #'
 #' ## Perform differential enrichment
-#' dif_enrich <- diffEnrich(sig_pe = sig_pe, bkg_pe = bkg_pe)
+#' dif_enrich <- diffEnrich(list1_pe = list1_pe, list2_pe = list2_pe)
 #'
-diffEnrich <- function(sig_pe, bkg_pe){
+diffEnrich <- function(list1_pe, list2_pe){
   ## Call .combineEnrich helper function
-  ce <- .combineEnrich(sig_pe = sig_pe, bkg_pe = bkg_pe)
+  ce <- .combineEnrich(list1_pe = list1_pe, list2_pe = list2_pe)
 
   ## Build diffEnrich Fisher's Exact function
   de <- function(a,b,c,d){
@@ -56,25 +56,25 @@ diffEnrich <- function(sig_pe, bkg_pe){
 #' in both objects (\code{by = c("V1", "V2", "KEGG_cnt", "numTested")}). This merged dataframe
 #' will be used as the input for the differential enrichment function.
 #'
-#' @param sig_pe data.frame. Dataframe of enrichment results for genes of interest
[…]

### 2019-07-01 R/diffEnrich.R
colname changes to res generation in diffEnrich

@@ -41,7 +41,7 @@ diffEnrich <- function(list1_pe, list2_pe){
     return(out.de)
   }
   ## perform differential enrichment
-  res <- cbind(ce, do.call('rbind', apply(ce[, c("KEGG_in_list_bkg", "KEGG_in_list_sig", "numSig_bkg", "numSig_sig")], 1,
+  res <- cbind(ce, do.call('rbind', apply(ce[, c("KEGG_in_list_list2", "KEGG_in_list_list1", "numSig_list2", "numSig_list1")], 1,
                function(a){ de(a[1], a[2], a[3], a[4])})))
   return(res)
 }

### 2019-07-01 R/diffEnrich.R
update rownames in diffEnrich output

@@ -22,8 +22,8 @@
 #'
 #' @examples
 #' ## Generate individual enrichment reults
-#' list1_pe <- pathEnrich(gk_obj = kegg, gene_list = geneLists$sigGenes)
-#' list2_pe <- pathEnrich(gk_obj = kegg, gene_list = geneLists$background)
+#' list1_pe <- pathEnrich(gk_obj = kegg, gene_list = geneLists$list1)
+#' list2_pe <- pathEnrich(gk_obj = kegg, gene_list = geneLists$list2)
 #'
 #' ## Perform differential enrichment
 #' dif_enrich <- diffEnrich(list1_pe = list1_pe, list2_pe = list2_pe)
@@ -43,6 +43,8 @@ diffEnrich <- function(list1_pe, list2_pe){
   ## perform differential enrichment
   res <- cbind(ce, do.call('rbind', apply(ce[, c("KEGG_in_list_list2", "KEGG_in_list_list1", "numSig_list2", "numSig_list1")], 1,
                function(a){ de(a[1], a[2], a[3], a[4])})))
+  ## update rownames
+  rownames(res) <- res$KEGG_ID
   return(res)
 }
 

### 2019-07-01 R/diffEnrich.R
Added multiple testing option to diffEnrich

@@ -12,6 +12,9 @@
 #' generated from \code{\link{pathEnrich}}. See example for \code{\link{pathEnrich}}.
 #' @param list2_pe data.frame. Dataframe of enrichment results for background genes
 #' generated from \code{\link{pathEnrich}}. See example for \code{\link{pathEnrich}}.
+#' @param method character. Character string telling \code{diffEnrich} which method to
+#' use for multiple testing correction. Available methods are thos provided by
+#' \code{\link{p.adjust}}, and the default is "BH", or False Discovery Rate (FDR).
 #'
 #' @return data.frame. Dataframe generated from merging pathEnrich dataframes with the following added columns:
 #'                     Estimate: Estimated odds ration calculated from Fisher's Exact test
@@ -26,9 +29,9 @@
 #' list2_pe <- pathEnrich(gk_obj = kegg, gene_list = geneLists$list2)
 #'
 #' ## Perform differential enrichment
-#' dif_enrich <- diffEnrich(list1_pe = list1_pe, list2_pe = list2_pe)
+#' dif_enrich <- diffEnrich(list1_pe = list1_pe, list2_pe = list2_pe, method = 'none')
 #'
-diffEnrich <- function(list1_pe, list2_pe){
+diffEnrich <- function(list1_pe, list2_pe, method = 'BH'){
   ## Call .combineEnrich helper function
   ce <- .combineEnrich(list1_pe = list1_pe, list2_pe = list2_pe)
 
@@ -43,6 +46,7 @@ diffEnrich <- function(list1_pe, list2_pe){
   ## perform differential enrichment
   res <- cbind(ce, do.call('rbind', apply(ce[, c("KEGG_in_list_list2", "KEGG_in_list_list1", "numSig_list2", "numSig_list1")], 1,
                function(a){ de(a[1], a[2], a[3], a[4])})))
+  res$adjusted_p <- stats::p.adjust(res$pv, method = method)
   ## update rownames
   rownames(res) <- res$KEGG_ID
   return(res)

### 2019-07-01 R/pathEnrich.R
added variable description table to roxygen skeleton for pathEnrich

@@ -12,7 +12,19 @@
 #' @param gene_list Vector. Vector of NCBI (ENTREZ) geneIDs.
 #'
 #' @return enrich_table: An object of class data.frame that summarizes the results
-#' of the pathway analysis.
+#' of the pathway analysis and contains the following variables:
+#'
+#' \describe{
+#'   \item{KEGG_ID}{KEGG ID}
+#'   \item{KEGG_description}{Description og KEGG pathway}
+#'   \item{KEGG_cnt}{Number of genes in KEGG pathway}
+#'   \item{KEGG_in_list}{Number of gene in KEGG pathway AND in user provided gene list}
+#'   \item{num_in_background}{Number of genes in }
+#'   \item{num_in_list}{Number of genes in }
+#'   \item{expected}{Estimated odd's ratio from Fisher's Exact test}
+#'   \item{enrich_p}{Unadjusted p-value}
+#'   \item{fdr}{Adjusted p-value. False Discovery Rate (FDR), using the Benjamini & Hochberg (1995)}
+#' }
 #'
 #' @details This function may not always use the complete list of genes provided by the user.
 #' Specifically, it will only use the genes from the list provided that are also in
@@ -89,7 +101,7 @@ pathEnrich <- function(gk_obj, gene_list){
   enrich_table$fdr <- stats::p.adjust(enrich_table$enrich_p, method = 'BH')
   enrich_table$V1 <- gsub("path:", "", enrich_table$V1, fixed = TRUE)
   ## Add menaingful column names for the C1 and C2
-  colnames(enrich_table)[c(1,2)] <- c("KEGG_ID", "KEGG_description", "KEGG_cnt", "KEGG_in_list",
+  colnames(enrich_table) <- c("KEGG_ID", "KEGG_description", "KEGG_cnt", "KEGG_in_list",
                                       "num_in_background", "num_in_list", "expected", "enrich_p",
                                       "fdr")
   return(enrich_table)

### 2019-07-01 data-raw/create_sample_data.R
updated data packed in package

@@ -21,8 +21,8 @@ backgroundENTZ <- mapIds(org.Rn.eg.db, backgroundList, 'ENTREZID', 'SYMBOL')
 backgroundENTZ <- base::unname(backgroundENTZ)
 sigGeneENTZ <- mapIds(org.Rn.eg.db, sigGenesList, 'ENTREZID', 'SYMBOL')
 sigGeneENTZ <- base::unname(sigGeneENTZ)
-geneLists <- list(list1 = backgroundENTZ,
-                  list2 = sigGeneENTZ)
+geneLists <- list(list1 = sigGeneENTZ,
+                  list2 = backgroundENTZ)
 
 ## save and compress
 usethis::use_data(geneLists, compress = "xz", overwrite = TRUE)

### 2019-07-01 R/pathEnrich.R
Added colnames from Laura to pathEnrich res

@@ -101,8 +101,8 @@ pathEnrich <- function(gk_obj, gene_list){
   enrich_table$fdr <- stats::p.adjust(enrich_table$enrich_p, method = 'BH')
   enrich_table$V1 <- gsub("path:", "", enrich_table$V1, fixed = TRUE)
   ## Add menaingful column names for the C1 and C2
-  colnames(enrich_table) <- c("KEGG_ID", "KEGG_description", "KEGG_cnt", "KEGG_in_list",
-                                      "num_in_background", "num_in_list", "expected", "enrich_p",
+  colnames(enrich_table) <- c("KEGG_PATHWAY_ID", "KEGG_PATHWAY_description", "KEGG_PATHWAY_cnt", "KEGG_PATHWAY_in_list",
+                                      "KEGG_DATABASE_cnt", "KEG_DATASE_in_list", "expected", "enrich_p",
                                       "fdr")
   return(enrich_table)
 }

### 2019-07-01 R/pathEnrich.R
Added column header descriptions to pathEnrich documentation

@@ -15,15 +15,15 @@
 #' of the pathway analysis and contains the following variables:
 #'
 #' \describe{
-#'   \item{KEGG_ID}{KEGG ID}
-#'   \item{KEGG_description}{Description og KEGG pathway}
-#'   \item{KEGG_cnt}{Number of genes in KEGG pathway}
-#'   \item{KEGG_in_list}{Number of gene in KEGG pathway AND in user provided gene list}
-#'   \item{num_in_background}{Number of genes in }
-#'   \item{num_in_list}{Number of genes in }
-#'   \item{expected}{Estimated odd's ratio from Fisher's Exact test}
-#'   \item{enrich_p}{Unadjusted p-value}
-#'   \item{fdr}{Adjusted p-value. False Discovery Rate (FDR), using the Benjamini & Hochberg (1995)}
+#'   \item{KEGG_PATHWAY_ID}{KEGG Pathway Identifier}
+#'   \item{KEGG_PATHWAY_description}{Description of KEGG Pathway (provided by KEGG)}
+#'   \item{KEGG_PATHWAY_cnt}{Number of Genes in KEGG Pathway}
+#'   \item{KEGG_PATHWAY_in_list}{Number of Genes from gene list in KEGG Pathway}
+#'   \item{KEGG_DATABASE_cnt}{Number of Genes in KEGG Database}
+#'   \item{KEG_DATASE_in_list}{Number of Genes from gene list in KEGG Database}
+#'   \item{expected}{Expected number of genes from list to be in KEGG pathway by chance (i.e., not enriched)}
+#'   \item{enrich_p}{P-value for enrichment of list genes related to KEGG pathway}
+#'   \item{fdr}{False Discovery Rate (Benjamini and Hochberg) to account for multiple testing across KEGG pathways}
 #' }
 #'
 #' @details This function may not always use the complete list of genes provided by the user.

### 2019-07-01 R/diffEnrich.R
Added colname descriptions to diffEnrich documentation

@@ -16,9 +16,28 @@
 #' use for multiple testing correction. Available methods are thos provided by
 #' \code{\link{p.adjust}}, and the default is "BH", or False Discovery Rate (FDR).
 #'
-#' @return data.frame. Dataframe generated from merging pathEnrich dataframes with the following added columns:
-#'                     Estimate: Estimated odds ration calculated from Fisher's Exact test
-#'                     P_value: Unadjusted p_value from Fisher's Exact test
+#' @return res: An object of class data.frame that summarizes the results
+#' of the differential enrichment and contains the following variables:
+#'
+#' \describe{
+#'   \item{KEGG_PATHWAY_ID}{KEGG Pathway Identifier}
+#'   \item{KEGG_PATHWAY_description}{Description of KEGG Pathway (provided by KEGG)}
+#'   \item{KEGG_PATHWAY_cnt}{Number of Genes in KEGG Pathway}
+#'   \item{KEGG_DATABASE_cnt}{Number of Genes in KEGG Database}
+#'   \item{KEGG_PATHWAY_in_list1}{Number of Genes from gene list 1 in KEGG Pathway}
+#'   \item{KEGG_DATABASE_in_list1}{Number of Genes from gene list 1 in KEGG Database}
+#'   \item{expected_list1}{Expected number of genes from list 1 to be in KEGG pathway by chance (i.e., not enriched)}
+#'   \item{enrich_p_list1}{P-value for enrichment of list 1 genes related to KEGG pathway}
+#'   \item{fdr_list1}{False Discovery Rate (Benjamini and Hochberg) of enrich_p_list1 to account for multiple testing across KEGG pathways}
+#'   \item{KEGG_PATHWAY_in_list2}{Number of Genes from gene list 2 in KEGG Pathway}
+#'   \item{KEGG_DATABASE_in_list2}{Number of Genes from gene list 2 in KEGG Database}
+#'   \item{expected_list2}{Expected number of genes from list 2 to be in KEGG pathway by chance (i.e., not enriched)}
+#'   \item{enrich_p_list2}{P-value for enrichment of list 2 genes related to KEGG pathway}
+#'   \item{fdr_list2}{False Discovery Rate (Benjamini and Hochberg) of enrich_p_list2 to account for multiple testing across KEGG pathways}
[…]

### 2019-07-01 R/pathEnrich.R

@@ -20,7 +20,7 @@
 #'   \item{KEGG_PATHWAY_cnt}{Number of Genes in KEGG Pathway}
 #'   \item{KEGG_PATHWAY_in_list}{Number of Genes from gene list in KEGG Pathway}
 #'   \item{KEGG_DATABASE_cnt}{Number of Genes in KEGG Database}
-#'   \item{KEG_DATASE_in_list}{Number of Genes from gene list in KEGG Database}
+#'   \item{KEGG_DATABASE_in_list}{Number of Genes from gene list in KEGG Database}
 #'   \item{expected}{Expected number of genes from list to be in KEGG pathway by chance (i.e., not enriched)}
 #'   \item{enrich_p}{P-value for enrichment of list genes related to KEGG pathway}
 #'   \item{fdr}{False Discovery Rate (Benjamini and Hochberg) to account for multiple testing across KEGG pathways}

### 2019-07-01 README.Rmd
Updated README

@@
 # run pathEnrich using kegg_rno
-## Genes of interest
-sig_pe <- pathEnrich(gk_obj = kegg, gene_list = geneLists$sigGenes)
-## Background
-bkg_pe <- pathEnrich(gk_obj = kegg, gene_list = geneLists$background)
-knitr::kable(head(sig_pe),
-             caption = "Table 1. Head of sig_pe dataframe generated using pathEnrich", row.names = FALSE) %>%
+## List 1
+list1_pe <- pathEnrich(gk_obj = kegg, gene_list = geneLists$list1)
+## list2
+list2_pe <- pathEnrich(gk_obj = kegg, gene_list = geneLists$list2)
+```
+
+```{r pathEnrich_kable, warning=FALSE, echo=FALSE}
+knitr::kable(head(list1_pe),
+             caption = "Table 1. Head of list1_pe data frame generated using pathEnrich", row.names = FALSE) %>%
   kableExtra::kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = T)
@@
 # get column names
-names <- colnames(sig_pe)
+names <- colnames(list1_pe)
 # Define column names
-description <- c("KEGG ID",
-                 "KEGG pathway description",
-                 "PLACE HOLDER",
-                 "PLACE HOLDER",
-                 "Total number of pathways in KEGG release",
-                 "PLACE HOLDER",
-                 "PLACE HOLDER",
-                 "Unadjusted p-value from Fisher's Exact test",
-                 "False Discovery Rate (FDR)")
+description <- c("KEGG Pathway Identifier",
+                 "Description of KEGG Pathway (provided by KEGG)",
+                 "Number of Genes in KEGG Pathway",
+                 "Number of Genes from gene list in KEGG Pathway",
+                 "Number of Genes in KEGG Database",
+                 "Number of Genes from gene list in KEGG Database",
+                 "Expected number of genes from list to be in KEGG pathway by chance (i.e., not enriched)",
+                 "P-value for enrichment of list genes related to KEGG pathway",
+                 "False Discovery Rate (Benjamini and Hochberg) to account for multiple testing across KEGG pathways")
 # Make dataframe
@@
 colnames(df1) <- c("Column Names", "Column Description")
+```
 
+```{r pathEnrich_vars_kable, warning=FALSE, echo=FALSE}
 knitr::kable(df1,
@@
 The diffEnrich function will merge the results from the *pathEnrich* calls generated above. Specifically, the data frame
-'sig_pe' and the data frame 'bkg_pe' will be merged using the base *merge* functionality and be merged by the following columns:
-'c("KEGG_ID", "KEGG_description", "KEGG_cnt", "numTested")'. This merged data set will then be used to perform differential 
-enrichment using the same method and p-value calculation as described above.
+'list1_pe' and the data frame 'list2_pe' will be merged using the base *merge* functionality and be merged by the following columns:
+'c("KEGG_PATHWAY_ID", "KEGG_PATHWAY_description", "KEGG_PATHWAY_cnt", "KEGG_DATABASE_cnt")'.
+This merged data set will then be used to perform differential enrichment using the same method and p-value calculation as described above.
+Users do have the option of choosing a method
+for multiple testing adjustment.
 
@@
 ## Perform differential enrichment
-diff_enrich <- diffEnrich(sig_pe = sig_pe, bkg_pe = bkg_pe)
+diff_enrich <- diffEnrich(list1_pe = list1_pe, list2_pe = list2_pe, method = 'none')
+```
 
+```{r diffEnrich_kable, warning=FALSE, echo=FALSE}
 knitr::kable(head(diff_enrich),
-             caption = "Table 2. Description of columns is dataframe generated by pathEnrich", row.names = FALSE) %>%
+             caption = "Table 3. Description of columns is data frame generated by diffEnrich", row.names = FALSE) %>%
   kableExtra::kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = F)
@@
 The result of the *diffEnrich* call is the merged data frame with the estimated odds ration generated by the Fisher's Exact test 
-and the associated p-value.
+and the associated p-value. Users can choose from those supported by *stats::p.adjust*, and the defualt is the False Discovery Rate (Benjamini and Hochberg).

### 2019-07-01 R/get_kegg.R
added describe section to get_kegg roxygen skeleton

@@ -2,7 +2,7 @@
 #' @description This function connects to the KEGG API, downloads, and cleans
 #' ncbi gene ID data, KEGG pathway descriptions, and species specific data.
 #' Currently, this function supports Human, Mouse, and Rat. Files will be
-#' written to the working directory.
+#' written to the working directory unless otherwise specified by the user.
 #'
 #' @param species character. The species to use in kegg data pull
 #' @param path character. A character string describing the path to write out KEGG
@@ -12,9 +12,13 @@
 #' function was run. This may be different if the function is run at
 #' different times. For reproducible results, use text files generated
 #' by function that include the date they were pulled.
-#' kegg_out: "ncbi_to_kegg" = ncbi_to_kegg dataframe,
-#'           "kegg_to_pathway" = kegg_to_pathway dataframe,
-#'           "pathway_to_species" = pathway_to_species dataframe
+#'
+#' \describe{
+#'   \item{ncbi_to_kegg}{ncbi_to_kegg mappings as class data.frame}
+#'   \item{kegg_to_pathway}{kegg_to_pathway mappings as class data.frame}
+#'   \item{pathway_to_species}{pathway_to_species mappings as class data.frame}
+#'   }
+#'
 #' @export
 #' @importFrom here here
 #' @import utils

### 2019-07-05 R/get_kegg.R
Wrote helper function called .api_pull. Updated get_kegg

@@ -1,5 +1,5 @@
 #' @title get_kegg
-#' @description This function connects to the KEGG API, downloads, and cleans
+#' @description This function calls an internal helper function that connects to the KEGG API, downloads, and cleans
 #' ncbi gene ID data, KEGG pathway descriptions, and species specific data.
 #' Currently, this function supports Human, Mouse, and Rat. Files will be
 #' written to the working directory unless otherwise specified by the user.
@@ -21,8 +21,6 @@
 #'
 #' @export
 #' @importFrom here here
-#' @import utils
-#' @importFrom stringr str_extract
 #'
 #' @examples
 #' kegg <- get_kegg(species = "rno")
@@ -30,7 +28,47 @@
 #' kegg <- get_kegg(species = "mmu", path = "usr/data/out/")
 #' }
 #'
-get_kegg <- function(species, path = NULL){
+
+
+test_kegg <- function(species, path = NULL){
+  ## API pull if path = NULL
+  if(is.null(path)){
+    wkd <- here::here()
+    res <- .api_pull(species, path = wkd)
+  } else {
+    # API pull with user's path
+    res <- .api_pull(species, path = path)
+  }
+  return(res)
+}
+
+
+
+
+
+
+
+
+#' @title .api_pull
+#' @description This function connects to the KEGG API, downloads, and cleans
+#' ncbi gene ID data, KEGG pathway descriptions, and species specific data.
+#' Currently, this function supports Human, Mouse, and Rat. Files will be
+#' written to the working directory unless otherwise specified by the user.
+#'
+#' @param species character. The species to use in kegg data pull
+#' @param path character. A character string describing the path to write out KEGG
+#' API data sets. If not provided, defaults to current working directory.
+#'
+#' @return kegg_out: A named list of the data pulled from kegg api when the
+#' function was run. This may be different if the function is run at
+#' different times. For reproducible results, use text files generated
+#' by function that include the date they were pulled.
+#'
+#' @importFrom here here
+#' @import utils
+#' @importFrom stringr str_extract
+#'
+.api_pull <- function(species, path = path){
   options(stringsAsFactors = F)
   ## Argument checks
   if(missing(species)){stop("Must choose one of the 3 species options: human, mouse, rat")}
@@ -71,144 +109,69 @@ get_kegg <- function(species, path = NULL){
                   paste("kegg_to_pathway",Sys.Date(), kegg_release, ".txt",sep=""),
[…]

### 2019-07-05 R/get_kegg.R
Updated get_kegg documentation files

@@ -30,7 +30,7 @@
 #'
 
 
-test_kegg <- function(species, path = NULL){
+get_kegg <- function(species, path = NULL){
   ## API pull if path = NULL
   if(is.null(path)){
     wkd <- here::here()

### 2019-07-05 R/get_kegg.R
Cleaned up get_kegg.R

@@ -28,8 +28,6 @@
 #' kegg <- get_kegg(species = "mmu", path = "usr/data/out/")
 #' }
 #'
-
-
 get_kegg <- function(species, path = NULL){
   ## API pull if path = NULL
   if(is.null(path)){
@@ -41,11 +39,6 @@ get_kegg <- function(species, path = NULL){
   }
   return(res)
 }
-
-
-
-
-
 
 
 

### 2019-07-15 R/pathEnrich.R
updated calculation of d

@@ -92,7 +92,7 @@ pathEnrich <- function(gk_obj, gene_list){
     a = enrich_table[i, "KEGG_in_list"]
     b = enrich_table[i, "KEGG_cnt"] - enrich_table[i, "KEGG_in_list"]
     c = enrich_table[i, "numSig"] - enrich_table[i, "KEGG_in_list"]
-    d = enrich_table[i, "numTested"] - enrich_table[i, "numSig"] + enrich_table[i, "KEGG_in_list"]
+    d = enrich_table[i, "numTested"] - a - b - c
     enrich_table$enrich_p[i] = stats::fisher.test(matrix(c(a,b,c,d), nrow = 2), alternative = "greater")$p.value
   }
 

### 2019-07-15 R/pathEnrich.R
Spelling fix in results table

@@ -102,7 +102,7 @@ pathEnrich <- function(gk_obj, gene_list){
   enrich_table$V1 <- gsub("path:", "", enrich_table$V1, fixed = TRUE)
   ## Add menaingful column names for the C1 and C2
   colnames(enrich_table) <- c("KEGG_PATHWAY_ID", "KEGG_PATHWAY_description", "KEGG_PATHWAY_cnt", "KEGG_PATHWAY_in_list",
-                                      "KEGG_DATABASE_cnt", "KEG_DATASE_in_list", "expected", "enrich_p",
+                                      "KEGG_DATABASE_cnt", "KEG_DATABASE_in_list", "expected", "enrich_p",
                                       "fdr")
   return(enrich_table)
 }

### 2019-07-15 R/pathEnrich.R
Added code to calculate fold enrichment in pathEnrich

@@ -100,9 +100,12 @@ pathEnrich <- function(gk_obj, gene_list){
   enrich_table <- enrich_table[order(enrich_table$enrich_p), ]
   enrich_table$fdr <- stats::p.adjust(enrich_table$enrich_p, method = 'BH')
   enrich_table$V1 <- gsub("path:", "", enrich_table$V1, fixed = TRUE)
+
+  ## Calulate fold enrichment column
+  enrich_table$fold_enrichment <- enrich_table$KEGG_in_list/enrich_table$expected
   ## Add menaingful column names for the C1 and C2
   colnames(enrich_table) <- c("KEGG_PATHWAY_ID", "KEGG_PATHWAY_description", "KEGG_PATHWAY_cnt", "KEGG_PATHWAY_in_list",
                                       "KEGG_DATABASE_cnt", "KEG_DATABASE_in_list", "expected", "enrich_p",
-                                      "fdr")
+                                      "fdr", "fold_enrichment")
   return(enrich_table)
 }

### 2019-07-15 R/diffEnrich.R
updated diffEnrich and .combineEnrich after adding fold enrichment variables

@@ -69,8 +69,8 @@ diffEnrich <- function(list1_pe, list2_pe, method = 'BH'){
   res$adjusted_p <- stats::p.adjust(res$pv, method = method)
   colnames(res) <- c("KEGG_PATHWAY_ID", "KEGG_PATHWAY_description", "KEGG_PATHWAY_cnt", "KEGG_DATABASE_cnt",
                      "KEGG_PATHWAY_in_list1", "KEGG_DATABASE_in_list1", "expected_list1", "enrich_p_list1",
-                     "fdr_list1", "KEGG_PATHWAY_in_list2", "KEGG_DATABASE_in_list2", "expected_list2",
-                     "enrich_p_list2", "fdr_list2", "odd_ratio", "diff_enrich_p", "diff_enrich_adjusted")
+                     "fdr_list1", "fold_enrichment_list1", "KEGG_PATHWAY_in_list2", "KEGG_DATABASE_in_list2", "expected_list2",
+                     "enrich_p_list2", "fdr_list2", "fold_enrichment_list2", "odd_ratio", "diff_enrich_p", "diff_enrich_adjusted")
   ## update rownames
   rownames(res) <- res$KEGG_PATHWAY_ID
   return(res)
@@ -107,8 +107,8 @@ diffEnrich <- function(list1_pe, list2_pe, method = 'BH'){
   colnames(combined_enrich) <- gsub(".y", "_list2", colnames(combined_enrich), fixed = TRUE)
   colnames(combined_enrich) <- c("KEGG_PATHWAY_ID", "KEGG_PATHWAY_description", "KEGG_PATHWAY_cnt", "KEGG_DATABASE_cnt",
                                  "KEGG_PATHWAY_in_list1", "KEGG_DATABASE_in_list1", "expected_list1", "enrich_p_list1",
-                                 "fdr_list1", "KEGG_PATHWAY_in_list2", "KEGG_DATABASE_in_list2", "expected_list2",
-                                 "enrich_p_list2", "fdr_list2")
+                                 "fdr_list1", "fold_enrichment_list1", "KEGG_PATHWAY_in_list2", "KEGG_DATABASE_in_list2", "expected_list2",
+                                 "enrich_p_list2", "fdr_list2", "fold_enrichment_list2")
 
   out <- combined_enrich
   return(out)

### 2019-07-15 R/diffEnrich.R
Updated documentation for pathEnrich and diffEnrich after adding fold enrichment variables

@@ -29,11 +29,13 @@
 #'   \item{expected_list1}{Expected number of genes from list 1 to be in KEGG pathway by chance (i.e., not enriched)}
 #'   \item{enrich_p_list1}{P-value for enrichment of list 1 genes related to KEGG pathway}
 #'   \item{fdr_list1}{False Discovery Rate (Benjamini and Hochberg) of enrich_p_list1 to account for multiple testing across KEGG pathways}
+#'   \item{fold_enrichment_list1}{KEGG_PATHWAY_in_list1/expected_list1}
 #'   \item{KEGG_PATHWAY_in_list2}{Number of Genes from gene list 2 in KEGG Pathway}
 #'   \item{KEGG_DATABASE_in_list2}{Number of Genes from gene list 2 in KEGG Database}
 #'   \item{expected_list2}{Expected number of genes from list 2 to be in KEGG pathway by chance (i.e., not enriched)}
 #'   \item{enrich_p_list2}{P-value for enrichment of list 2 genes related to KEGG pathway}
 #'   \item{fdr_list2}{False Discovery Rate (Benjamini and Hochberg) of enrich_p_list2 to account for multiple testing across KEGG pathways}
+#'   \item{fold_enrichment_list2}{KEGG_PATHWAY_in_list2/expected_list2}
 #'   \item{odd_ratio}{Odds of a gene from list 2 being from this KEGG pathway / Odds of a gene from list 1 being from this KEGG pathway}
 #'   \item{diff_enrich_p}{P-value for differential enrichment of this KEGG pathway between list 1 and list 2}
 #'   \item{diff_enrich_adjusted}{Multiple testing adjustment of diff_enrich_p (default = False Discovery Rate (Benjamini and Hochberg))}

### 2019-07-15 R/pathEnrich.R

@@ -24,6 +24,7 @@
 #'   \item{expected}{Expected number of genes from list to be in KEGG pathway by chance (i.e., not enriched)}
 #'   \item{enrich_p}{P-value for enrichment of list genes related to KEGG pathway}
 #'   \item{fdr}{False Discovery Rate (Benjamini and Hochberg) to account for multiple testing across KEGG pathways}
+#'   \item{fold_enrichment}{KEGG_PATHWAY_in_list/expected}
 #' }
 #'
 #' @details This function may not always use the complete list of genes provided by the user.

### 2019-07-15 README.Rmd
kegg release updated in README
Updated pathEnrich so reknit README
Updated README table of results variable descriptions

@@
                  "P-value for enrichment of list genes related to KEGG pathway",
-                 "False Discovery Rate (Benjamini and Hochberg) to account for multiple testing across KEGG pathways")
+                 "False Discovery Rate (Benjamini and Hochberg) to account for multiple testing across KEGG pathways",
+                 "KEGG_PATHWAY_in_list/expected")
 # Make dataframe

### 2019-07-29 R/diffEnrich.R
added code to re-order results table based on adjusted p-value

@@ -43,6 +43,7 @@
 #'
 #' @export
 #' @importFrom  stats fisher.test
+#' @import dplyr
 #'
 #' @examples
 #' ## Generate individual enrichment reults
@@ -73,6 +74,12 @@ diffEnrich <- function(list1_pe, list2_pe, method = 'BH'){
                      "KEGG_PATHWAY_in_list1", "KEGG_DATABASE_in_list1", "expected_list1", "enrich_p_list1",
                      "fdr_list1", "fold_enrichment_list1", "KEGG_PATHWAY_in_list2", "KEGG_DATABASE_in_list2", "expected_list2",
                      "enrich_p_list2", "fdr_list2", "fold_enrichment_list2", "odd_ratio", "diff_enrich_p", "diff_enrich_adjusted")
+
+  ## re-order table based on adjusted p-value
+  library(dplyr)
+  res <- res %>%
+    arrange(diff_enrich_adjusted)
+
   ## update rownames
   rownames(res) <- res$KEGG_PATHWAY_ID
   return(res)

### 2019-08-22 R/plotFoldEnrichment.R
added hack version of plot

@@ -66,31 +66,43 @@ plotFoldEnrichment <- function(de_res, pval, N){
   ###########################################################
   # Generate plot
   ###########################################################
-   ## TEST 1
-  p <- ggplot(bardat, aes(x=reorder(KEGG_PATHWAY_description, -pvals), y=value)) +
-    geom_bar(stat="identity", aes(fill=variable, alpha = pvals), position="dodge") +
-    ylim(0, max(bardat$value) + 0.6) + xlab("") +
+  g <- ggplot(bardat, aes(x=reorder(KEGG_PATHWAY_description, -pvals), y=value)) +
+    geom_bar(stat="identity", aes(col=variable, group=variable, fill=pvals), position="dodge") +
+    ylim(0, max(d$value) + 0.6) + xlab("") +
     coord_flip() +
     scale_fill_brewer(palette = "Set1",
                       name="",
                       breaks=c("fold_enrichment_list1", "fold_enrichment_list2"),
                       labels=c("Fold Enrichment in \nlist 1\n", "Fold enrichment in \nlist 2\n")) +
-    scale_alpha(trans = "log10") +
+    scale_fill_continuous(trans = 'log10') +
     geom_text(data=subset(df.ss, variable %in% c("diff_enrich_adjusted")),
               aes(x = KEGG_PATHWAY_description, y = (max(bardat$value) + 0.3), label = round(value, 4))) +
     labs(alpha = "List specific p-value")
-  ## TEST 2
-  # p2 <- ggplot(bardat, aes(x=reorder(KEGG_PATHWAY_description, -pvals), y=value)) +
-  #   geom_bar(stat="identity", aes(group=variable, fill=pvals), position="dodge") +
-  #   ylim(0, max(bardat$value) + 0.6) + xlab("") +
-  #   coord_flip() +
-  #   scale_fill_brewer(palette = "Set1",
-  #                     name="",
-  #                     breaks=c("fold_enrichment_list1", "fold_enrichment_list2"),
-  #                     labels=c("Fold Enrichment in \nlist 1\n", "Fold enrichment in \nlist 2\n")) +
-  #   scale_fill_continuous(trans = 'log10') +
-  #   geom_text(data=subset(df.ss, variable %in% c("diff_enrich_adjusted")),
-  #             aes(x = KEGG_PATHWAY_description, y = (max(bardat$value) + 0.3), label = round(value, 4))) +
[…]

### 2019-08-22 R/plotFoldEnrichment.R
finally have a working prototype

@@ -66,7 +66,8 @@ plotFoldEnrichment <- function(de_res, pval, N){
   ###########################################################
   # Generate plot
   ###########################################################
-  g <- ggplot(bardat, aes(x=reorder(KEGG_PATHWAY_description, -pvals), y=value)) +
+    library(ggnewscale)
+    g <- ggplot(bardat, aes(x=reorder(KEGG_PATHWAY_description, -pvals), y=value)) +
     geom_bar(stat="identity", aes(col=variable, group=variable, fill=pvals), position="dodge") +
     ylim(0, max(d$value) + 0.6) + xlab("") +
     coord_flip() +
@@ -91,8 +92,14 @@ plotFoldEnrichment <- function(de_res, pval, N){
   ld$descr <- bardat$KEGG_PATHWAY_description[matches]
   ld$vars <- bardat$variable[matches]
 
+  ## Merge ld with df.ss
+  test <- merge(ld, df.ss, by.x = "descr", by.y = "KEGG_PATHWAY_description")
+  test <- subset(test, variable %in% c("diff_enrich_adjusted")) %>%
+    filter(!duplicated(value))
+
   ggplot(mapping = aes(xmin = xmin, xmax = xmax, ymin = ymin, ymax = ymax)) +
     geom_rect(data = ld[ld$vars == "fold_enrichment_list1", ], aes(fill = pvals)) +
+    ylim(0, max(d$value) + 1.0) + xlab("") +
     scale_fill_gradient(low = "red", high = "transparent",
                         limits = c(min(ld$pvals), 0),
                         name = "P-values List 1") +
@@ -103,6 +110,8 @@ plotFoldEnrichment <- function(de_res, pval, N){
                         name = "P-values List 2") +
     scale_x_continuous(breaks = seq_along(unique(ld$descr)),
                        labels = unique(ld$descr)) +
-    coord_flip()
+    coord_flip() +
+    geom_text(data=test,
+              aes(x = 1:5, y = (max(bardat$value) + 0.3), label = round(value, 4)))
   return(p)
 }

### 2019-08-22 R/plotFoldEnrichment.R
Added comments to plotting code

@@ -67,6 +67,7 @@ plotFoldEnrichment <- function(de_res, pval, N){
   # Generate plot
   ###########################################################
     library(ggnewscale)
+  # First, we'll make a plot and save it as a variable
     g <- ggplot(bardat, aes(x=reorder(KEGG_PATHWAY_description, -pvals), y=value)) +
     geom_bar(stat="identity", aes(col=variable, group=variable, fill=pvals), position="dodge") +
     ylim(0, max(d$value) + 0.6) + xlab("") +
@@ -80,9 +81,9 @@ plotFoldEnrichment <- function(de_res, pval, N){
               aes(x = KEGG_PATHWAY_description, y = (max(bardat$value) + 0.3), label = round(value, 4))) +
     labs(alpha = "List specific p-value")
 
+  # Next, we'll take the coordinates of this layers data and match them back to the original data.
   ld <- layer_data(g)
   ld <- ld[, c("xmin", "xmax", "ymin", "ymax")]
-
 
   # Match back to original data
   matches <- match(ld$ymax, bardat$value)
@@ -93,11 +94,12 @@ plotFoldEnrichment <- function(de_res, pval, N){
   ld$vars <- bardat$variable[matches]
 
   ## Merge ld with df.ss
-  test <- merge(ld, df.ss, by.x = "descr", by.y = "KEGG_PATHWAY_description")
-  test <- subset(test, variable %in% c("diff_enrich_adjusted")) %>%
+  df_ptext <- merge(ld, df.ss, by.x = "descr", by.y = "KEGG_PATHWAY_description")
+  df_ptext <- subset(df_ptext, variable %in% c("diff_enrich_adjusted")) %>%
     filter(!duplicated(value))
 
-  ggplot(mapping = aes(xmin = xmin, xmax = xmax, ymin = ymin, ymax = ymax)) +
+  ## Generate fineal plot
+  p <- ggplot(mapping = aes(xmin = xmin, xmax = xmax, ymin = ymin, ymax = ymax)) +
     geom_rect(data = ld[ld$vars == "fold_enrichment_list1", ], aes(fill = pvals)) +
     ylim(0, max(d$value) + 1.0) + xlab("") +
     scale_fill_gradient(low = "red", high = "transparent",
@@ -111,7 +113,7 @@ plotFoldEnrichment <- function(de_res, pval, N){
     scale_x_continuous(breaks = seq_along(unique(ld$descr)),
[…]

### 2019-08-22 R/plotFoldEnrichment.R
Added axis labels

@@ -98,10 +98,10 @@ plotFoldEnrichment <- function(de_res, pval, N){
   df_ptext <- subset(df_ptext, variable %in% c("diff_enrich_adjusted")) %>%
     filter(!duplicated(value))
 
-  ## Generate fineal plot
+  ## Generate finale plot
   p <- ggplot(mapping = aes(xmin = xmin, xmax = xmax, ymin = ymin, ymax = ymax)) +
     geom_rect(data = ld[ld$vars == "fold_enrichment_list1", ], aes(fill = pvals)) +
-    ylim(0, max(d$value) + 1.0) + xlab("") +
+    ylim(0, max(d$value) + 1.0) + xlab("") + ylab("Fold Enrichment") +
     scale_fill_gradient(low = "red", high = "transparent",
                         limits = c(min(ld$pvals), 0),
                         name = "P-values List 1") +

### 2019-08-22 R/plotFoldEnrichment.R
Updated legend titles

@@ -104,12 +104,12 @@ plotFoldEnrichment <- function(de_res, pval, N){
     ylim(0, max(d$value) + 1.0) + xlab("") + ylab("Fold Enrichment") +
     scale_fill_gradient(low = "red", high = "transparent",
                         limits = c(min(ld$pvals), 0),
-                        name = "P-values List 1") +
+                        name = "Log10 P-values List 1") +
     new_scale_fill() +
     geom_rect(data = ld[ld$vars == "fold_enrichment_list2", ], aes(fill = pvals)) +
     scale_fill_gradient(low =  "blue", high = "transparent",
                         limits = c(min(ld$pvals), 0),
-                        name = "P-values List 2") +
+                        name = "Log10 P-values List 2") +
     scale_x_continuous(breaks = seq_along(unique(ld$descr)),
                        labels = unique(ld$descr)) +
     coord_flip() +

### 2019-08-22 R/plotFoldEnrichment.R
Changed aes x = 1:5 to 1:N

@@ -114,6 +114,6 @@ plotFoldEnrichment <- function(de_res, pval, N){
                        labels = unique(ld$descr)) +
     coord_flip() +
     geom_text(data=df_ptext,
-              aes(x = 1:5, y = (max(bardat$value) + 0.3), label = round(value, 4)))
+              aes(x = 1:N, y = (max(bardat$value) + 0.3), label = round(value, 4)))
   return(p)
 }

### 2019-08-22 R/plotFoldEnrichment.R
Formatted p-values in legend

@@ -89,7 +89,7 @@ plotFoldEnrichment <- function(de_res, pval, N){
   matches <- match(ld$ymax, bardat$value)
 
   # Supplement with original data
-  ld$pvals <- log10((bardat$pvals[matches]))
+  ld$pvals <- log10(bardat$pvals[matches])
   ld$descr <- bardat$KEGG_PATHWAY_description[matches]
   ld$vars <- bardat$variable[matches]
 
@@ -103,12 +103,18 @@ plotFoldEnrichment <- function(de_res, pval, N){
     geom_rect(data = ld[ld$vars == "fold_enrichment_list1", ], aes(fill = pvals)) +
     ylim(0, max(d$value) + 1.0) + xlab("") + ylab("Fold Enrichment") +
     scale_fill_gradient(low = "red", high = "transparent",
+                        #trans = 'log10',
                         limits = c(min(ld$pvals), 0),
+                        breaks = as.numeric(summary(ld$pvals))[c(1,2,3,5,6)],
+                        labels = as.character(formatC(as.numeric(summary(bardat$pvals))[c(1,2,3,5,6)], format = "e", digits = 2)),
                         name = "Log10 P-values List 1") +
     new_scale_fill() +
     geom_rect(data = ld[ld$vars == "fold_enrichment_list2", ], aes(fill = pvals)) +
     scale_fill_gradient(low =  "blue", high = "transparent",
+                        #trans = 'log10',
                         limits = c(min(ld$pvals), 0),
+                        breaks = as.numeric(summary(ld$pvals))[c(1,2,3,5,6)],
+                        labels = as.character(formatC(as.numeric(summary(bardat$pvals))[c(1,2,3,5,6)], format = 'e', digits = 2)),
                         name = "Log10 P-values List 2") +
     scale_x_continuous(breaks = seq_along(unique(ld$descr)),
                        labels = unique(ld$descr)) +

### 2019-08-22 R/plotFoldEnrichment.R
updated legend titles

@@ -107,7 +107,7 @@ plotFoldEnrichment <- function(de_res, pval, N){
                         limits = c(min(ld$pvals), 0),
                         breaks = as.numeric(summary(ld$pvals))[c(1,2,3,5,6)],
                         labels = as.character(formatC(as.numeric(summary(bardat$pvals))[c(1,2,3,5,6)], format = "e", digits = 2)),
-                        name = "Log10 P-values List 1") +
+                        name = "P-values List 1") +
     new_scale_fill() +
     geom_rect(data = ld[ld$vars == "fold_enrichment_list2", ], aes(fill = pvals)) +
     scale_fill_gradient(low =  "blue", high = "transparent",
@@ -115,7 +115,7 @@ plotFoldEnrichment <- function(de_res, pval, N){
                         limits = c(min(ld$pvals), 0),
                         breaks = as.numeric(summary(ld$pvals))[c(1,2,3,5,6)],
                         labels = as.character(formatC(as.numeric(summary(bardat$pvals))[c(1,2,3,5,6)], format = 'e', digits = 2)),
-                        name = "Log10 P-values List 2") +
+                        name = "P-values List 2") +
     scale_x_continuous(breaks = seq_along(unique(ld$descr)),
                        labels = unique(ld$descr)) +
     coord_flip() +

### 2019-08-22 R/plotFoldEnrichment.R
changed ggplot to theme_bw

@@ -14,6 +14,7 @@
 #'
 #' @import dplyr
 #'         ggplot2
+#'         ggnewscale
 #' @importFrom reshape2 melt
 #' @export
 #'
@@ -55,10 +56,10 @@ plotFoldEnrichment <- function(de_res, pval, N){
                          "diff_enrich_adjusted"))
 
   ## get vector of pvals
-  pvals <- dplyr::subset(df.ss, variable %in% c("enrich_p_list1", "enrich_p_list2"))
+  pvals <- subset(df.ss, variable %in% c("enrich_p_list1", "enrich_p_list2"))
 
   ## Generate data set to be used for plotting
-  bardat <- dplyr::subset(df.ss, variable %in% c("fold_enrichment_list1", "fold_enrichment_list2")) %>%
+  bardat <- subset(df.ss, variable %in% c("fold_enrichment_list1", "fold_enrichment_list2")) %>%
     dplyr::mutate(alpha = log10(pvals$value),
            pvals = pvals$value) %>%
     dplyr::arrange(pvals)
@@ -82,7 +83,7 @@ plotFoldEnrichment <- function(de_res, pval, N){
     labs(alpha = "List specific p-value")
 
   # Next, we'll take the coordinates of this layers data and match them back to the original data.
-  ld <- layer_data(g)
+  ld <- ggplot2::layer_data(g)
   ld <- ld[, c("xmin", "xmax", "ymin", "ymax")]
 
   # Match back to original data
@@ -96,7 +97,7 @@ plotFoldEnrichment <- function(de_res, pval, N){
   ## Merge ld with df.ss
   df_ptext <- merge(ld, df.ss, by.x = "descr", by.y = "KEGG_PATHWAY_description")
   df_ptext <- subset(df_ptext, variable %in% c("diff_enrich_adjusted")) %>%
-    filter(!duplicated(value))
+    dplyr::filter(!duplicated(value))
 
   ## Generate finale plot
   p <- ggplot(mapping = aes(xmin = xmin, xmax = xmax, ymin = ymin, ymax = ymax)) +
@@ -108,7 +109,7 @@ plotFoldEnrichment <- function(de_res, pval, N){
                         breaks = as.numeric(summary(ld$pvals))[c(1,2,3,5,6)],
                         labels = as.character(formatC(as.numeric(summary(bardat$pvals))[c(1,2,3,5,6)], format = "e", digits = 2)),
                         name = "P-values List 1") +
-    new_scale_fill() +
[…]

### 2019-08-22 R/plotFoldEnrichment.R
Changed colors in plot

@@ -103,7 +103,7 @@ plotFoldEnrichment <- function(de_res, pval, N){
   p <- ggplot(mapping = aes(xmin = xmin, xmax = xmax, ymin = ymin, ymax = ymax)) +
     geom_rect(data = ld[ld$vars == "fold_enrichment_list1", ], aes(fill = pvals)) +
     ylim(0, max(d$value) + 1.0) + xlab("") + ylab("Fold Enrichment") +
-    scale_fill_gradient(low = "red", high = "transparent",
+    scale_fill_gradient(low = "darkred", high = "transparent",
                         #trans = 'log10',
                         limits = c(min(ld$pvals), 0),
                         breaks = as.numeric(summary(ld$pvals))[c(1,2,3,5,6)],
@@ -111,7 +111,7 @@ plotFoldEnrichment <- function(de_res, pval, N){
                         name = "P-values List 1") +
     ggnewscale::new_scale_fill() +
     geom_rect(data = ld[ld$vars == "fold_enrichment_list2", ], aes(fill = pvals)) +
-    scale_fill_gradient(low =  "blue", high = "transparent",
+    scale_fill_gradient(low =  "navy", high = "transparent",
                         #trans = 'log10',
                         limits = c(min(ld$pvals), 0),
                         breaks = as.numeric(summary(ld$pvals))[c(1,2,3,5,6)],

### 2019-08-22 README.Rmd
Re-knit README to reflect new ordering of diffEnrich results table
Bug fixes and knit Rmd with new plotFoldEnrichment function

@@
 The result of the *diffEnrich* call is the merged data frame with the estimated odds ration generated by the Fisher's Exact test 
-and the associated p-value. Users can choose from those supported by *stats::p.adjust*, and the defualt is the False Discovery Rate (Benjamini and Hochberg).
+and the associated p-value. Users can choose from those supported by *stats::p.adjust*, and the default is the False Discovery Rate (Benjamini and Hochberg).
+
+### Step 4: Plot fold enrichment
+
+```{r plotFoldEnrichment}
+## Plot fold enrichment
+plotFoldEnrichment(de_res = diff_enrich, pval = 0.05, N = 5)
+```

### 2019-08-22 R/plotFoldEnrichment.R

@@ -10,7 +10,11 @@
 #' @param pval Numeric. Threshold for filtering pathways based on adjusted pvalue in de_res
 #' @param N Numeric. Number of top pathways to plot after filtering based on pval
 #'
-#' @return
+#' @return ggplot object. If the user calls \code{\link{plotFoldEnrich}} and
+#' assigns it to an object (see example) then no plot will print in viewer,
+#' but if \code{\link{plotFoldEnrich}} is called without being assigned to an
+#' object the plot will print to the viewer. Users can edit the ggplot object
+#' as they would any other ggplot object (e.g. add title, theme, etc.).
 #'
 #' @import dplyr
 #'         ggplot2
@@ -71,7 +75,7 @@ plotFoldEnrichment <- function(de_res, pval, N){
   # First, we'll make a plot and save it as a variable
     g <- ggplot(bardat, aes(x=reorder(KEGG_PATHWAY_description, -pvals), y=value)) +
     geom_bar(stat="identity", aes(col=variable, group=variable, fill=pvals), position="dodge") +
-    ylim(0, max(d$value) + 0.6) + xlab("") +
+    ylim(0, max(bardat$value) + 0.6) + xlab("") +
     coord_flip() +
     scale_fill_brewer(palette = "Set1",
                       name="",
@@ -102,7 +106,7 @@ plotFoldEnrichment <- function(de_res, pval, N){
   ## Generate finale plot
   p <- ggplot(mapping = aes(xmin = xmin, xmax = xmax, ymin = ymin, ymax = ymax)) +
     geom_rect(data = ld[ld$vars == "fold_enrichment_list1", ], aes(fill = pvals)) +
-    ylim(0, max(d$value) + 1.0) + xlab("") + ylab("Fold Enrichment") +
+    ylim(0, max(bardat$value) + 1.0) + xlab("") + ylab("Fold Enrichment") +
     scale_fill_gradient(low = "darkred", high = "transparent",
                         #trans = 'log10',
                         limits = c(min(ld$pvals), 0),

### 2019-08-22 R/plotFoldEnrichment.R
Added details section to plotFoldEnrichment

@@ -16,6 +16,17 @@
 #' object the plot will print to the viewer. Users can edit the ggplot object
 #' as they would any other ggplot object (e.g. add title, theme, etc.).
 #'
+#' @details This function generates a grouped bar plot using ggplot2 and the
+#' ggnewscale package. KEGG pathways are plotted on the y-axis and fold
+#' enrichment is plotted on the x-axis. each KEGG pathway has a bar plotting
+#' its fold enrichment in list 1 (red) and its fold enrichment in list 2 (blue).
+#' The transparency of the bars correspond to the adjusted p-value for the
+#' pathway's enrichment in the given list. The p-value presented as text to the
+#' right of each pair of bars is the adjusted p-value associated with the
+#' differential enrichment of the pathway between the two lists, and the pathways
+#' are ordered from top to bottom by this p-value (i.e. smallest p-value on top
+#' of plot, and largest p-value on bottom of plot).
+#'
 #' @import dplyr
 #'         ggplot2
 #'         ggnewscale

### 2019-08-22 R/diffEnrich.R
defined global variables in dplyr and ggplot functions

@@ -43,6 +43,7 @@
 #'
 #' @export
 #' @importFrom  stats fisher.test
+#' @importFrom  rlang .data
 #' @import dplyr
 #'
 #' @examples
@@ -78,7 +79,7 @@ diffEnrich <- function(list1_pe, list2_pe, method = 'BH'){
   ## re-order table based on adjusted p-value
   library(dplyr)
   res <- res %>%
-    arrange(diff_enrich_adjusted)
+    arrange(.data$diff_enrich_adjusted)
 
   ## update rownames
   rownames(res) <- res$KEGG_PATHWAY_ID

### 2019-08-22 R/plotFoldEnrichment.R

@@ -10,9 +10,9 @@
 #' @param pval Numeric. Threshold for filtering pathways based on adjusted pvalue in de_res
 #' @param N Numeric. Number of top pathways to plot after filtering based on pval
 #'
-#' @return ggplot object. If the user calls \code{\link{plotFoldEnrich}} and
+#' @return ggplot object. If the user calls \code{plotFoldEnrich} and
 #' assigns it to an object (see example) then no plot will print in viewer,
-#' but if \code{\link{plotFoldEnrich}} is called without being assigned to an
+#' but if \code{plotFoldEnrich} is called without being assigned to an
 #' object the plot will print to the viewer. Users can edit the ggplot object
 #' as they would any other ggplot object (e.g. add title, theme, etc.).
 #'
@@ -31,6 +31,8 @@
 #'         ggplot2
 #'         ggnewscale
 #' @importFrom reshape2 melt
+#' @importFrom stats reorder
+#' @importFrom rlang .data
 #' @export
 #'
 #' @examples
@@ -38,8 +40,8 @@
 #' plot <- plotFoldEnrichment(de_res = diffEnrich, pval = 0.05, N = 5)
 #' }
 plotFoldEnrichment <- function(de_res, pval, N){
-  library(dplyr)
-  library(ggplot2)
+  # library(dplyr)
+  # library(ggplot2)
   ## Check arguments
   if(missing(de_res)){stop("Argument missing: de_res")}
   if(missing(pval)){stop("Argument missing: pval - please provide a threshold")}
@@ -52,13 +54,13 @@ plotFoldEnrichment <- function(de_res, pval, N){
 
   ## Strip extra columns from de_res and filter based on pval. Then sort by pval.
   df <- de_res %>%
-    dplyr::select(KEGG_PATHWAY_ID, KEGG_PATHWAY_description,
-           fold_enrichment_list1, fold_enrichment_list2,
-           enrich_p_list1, enrich_p_list2,
-           odd_ratio, diff_enrich_adjusted) %>%
-    dplyr::mutate(KEGG_PATHWAY_description = sapply(strsplit(KEGG_PATHWAY_description, split = " - "), function(x) x[1])) %>%
-    dplyr::arrange(diff_enrich_adjusted) %>%
-    dplyr::filter(diff_enrich_adjusted < pval) %>%
+    dplyr::select(.data$KEGG_PATHWAY_ID, .data$KEGG_PATHWAY_description,
[…]

### 2019-08-22 R/plotFoldEnrichment.R
bug fixes in plotFoldEnrichment after defing global variables

@@ -95,9 +95,8 @@ plotFoldEnrichment <- function(de_res, pval, N){
                       breaks=c("fold_enrichment_list1", "fold_enrichment_list2"),
                       labels=c("Fold Enrichment in \nlist 1\n", "Fold enrichment in \nlist 2\n")) +
     scale_fill_continuous(trans = 'log10') +
-    geom_text(data=subset(df.ss, .data$variable %in% c("diff_enrich_adjusted")),
-              aes(x = .data$KEGG_PATHWAY_description, y = (max(bardat$value) + 0.3), label = round(.data$value, 4))) +
-    labs(alpha = "List specific p-value")
+    geom_text(data=subset(df.ss, df.ss$variable %in% c("diff_enrich_adjusted")),
+              aes(x = .data$KEGG_PATHWAY_description, y = (max(bardat$value) + 0.3), label = round(.data$value, 4)))
 
   # Next, we'll take the coordinates of this layers data and match them back to the original data.
   ld <- ggplot2::layer_data(g)
@@ -113,7 +112,7 @@ plotFoldEnrichment <- function(de_res, pval, N){
 
   ## Merge ld with df.ss
   df_ptext <- merge(ld, df.ss, by.x = "descr", by.y = "KEGG_PATHWAY_description")
-  df_ptext <- subset(df_ptext, .data$variable %in% c("diff_enrich_adjusted")) %>%
+  df_ptext <- subset(df_ptext, df_ptext$variable %in% c("diff_enrich_adjusted")) %>%
     dplyr::filter(!duplicated(.data$value))
 
   ## Generate finale plot

### 2019-08-22 R/diffEnrich.R
Fixes due to R CMD Check

@@ -77,7 +77,7 @@ diffEnrich <- function(list1_pe, list2_pe, method = 'BH'){
                      "enrich_p_list2", "fdr_list2", "fold_enrichment_list2", "odd_ratio", "diff_enrich_p", "diff_enrich_adjusted")
 
   ## re-order table based on adjusted p-value
-  library(dplyr)
+  # library(dplyr)
   res <- res %>%
     arrange(.data$diff_enrich_adjusted)
 

### 2019-08-22 R/plotFoldEnrichment.R

@@ -4,7 +4,7 @@
 #' pathways stratified by their enrichment in list 1 or list 2. Users can plot
 #' all pathways based on the adjusted p-value threshold used in \code{\link{diffEnrich}}
 #' and the top N pathways sorted by the adjusted p-value threshold used in \code{\link{diffEnrich}}.
-#' \code{\link{plotFoldEnrich}} returns a ggplot2 object so users can add additional
+#' \code{plotFoldEnrich} returns a ggplot2 object so users can add additional
 #' customizations.
 #' @param de_res Dataframe. Generated using \code{\link{diffEnrich}}
 #' @param pval Numeric. Threshold for filtering pathways based on adjusted pvalue in de_res
@@ -68,7 +68,7 @@ plotFoldEnrichment <- function(de_res, pval, N){
 
   ## Clean up melted data frame
   df.ss <- df.melt %>%
-    dplyr::filter(variable %in% c("fold_enrichment_list1", "fold_enrichment_list2",
+    dplyr::filter(.data$variable %in% c("fold_enrichment_list1", "fold_enrichment_list2",
                          "enrich_p_list1", "enrich_p_list2",
                          "diff_enrich_adjusted"))
 

### 2019-08-22 R/plotFoldEnrichment.R
Updated Imports from Depends

@@ -29,10 +29,10 @@
 #'
 #' @import dplyr
 #'         ggplot2
-#'         ggnewscale
 #' @importFrom reshape2 melt
 #' @importFrom stats reorder
 #' @importFrom rlang .data
+#' @importFrom ggnewscale new_scale_fill
 #' @export
 #'
 #' @examples
@@ -73,10 +73,10 @@ plotFoldEnrichment <- function(de_res, pval, N){
                          "diff_enrich_adjusted"))
 
   ## get vector of pvals
-  pvals <- subset(df.ss, variable %in% c("enrich_p_list1", "enrich_p_list2"))
+  pvals <- subset(df.ss, df.ss$variable %in% c("enrich_p_list1", "enrich_p_list2"))
 
   ## Generate data set to be used for plotting
-  bardat <- subset(df.ss, variable %in% c("fold_enrichment_list1", "fold_enrichment_list2")) %>%
+  bardat <- subset(df.ss, df.ss$variable %in% c("fold_enrichment_list1", "fold_enrichment_list2")) %>%
     dplyr::mutate(alpha = log10(pvals$value),
            pvals = pvals$value) %>%
     dplyr::arrange(.data$pvals)

### 2019-08-23 README.Rmd
Added text to step 4 in README and vignette

@@
 
+This function generates a grouped bar plot using ggplot2 and the
+ggnewscale package. KEGG pathways are plotted on the y-axis and fold
+enrichment is plotted on the x-axis. each KEGG pathway has a bar plotting
+its fold enrichment in list 1 (red) and its fold enrichment in list 2 (blue).
+The transparency of the bars correspond to the adjusted p-value for the
+pathway's enrichment in the given list. The p-value presented as text to the
+right of each pair of bars is the adjusted p-value associated with the
+differential enrichment of the pathway between the two lists, and the pathways
+are ordered from top to bottom by this p-value (i.e. smallest p-value on top
+of plot, and largest p-value on bottom of plot).
+
 ```{r plotFoldEnrichment}

### 2019-08-23 README.Rmd
Clarification in README introduction

@@
 enriched pathways from two different gene sets, and finally, the user will identify 
-pathways that are differentially expressed between the two gene sets. In addition to 
-the analysis pipeline, this package also provides functions for data visualizations.
+pathways that are differentially enriched between the two gene sets. In addition to 
+the analysis pipeline, this package also provides a plotting function.
 

### 2019-08-26 README.Rmd
Added code for first version of analysis pipeline flowchart to README

@@
 
-development of methods and software for differential enrichment analysis
+```{r, echo=FALSE, fig.cap='Figure 1. diffEnrich analysis pipeline.', dpi=400}
+suppressMessages(library(diagram))
+
+## open plot
+par(mar = c(0.5, 1, 0.5, 1))
+openplotmat()
+## define number of boxes at each level
+elpos <- coordinates (c(1, 1, 2, 2, 1, 1, 1, 1))
+## generate matrix of from-to arrow coordinates
+fromto <- matrix(ncol = 2, byrow = TRUE,
+                 data = c(1, 2,
+                          2, 3,
+                          2, 4,
+                          3, 5,
+                          4, 6,
+                          5, 7,
+                          6, 7,
+                          7, 8, 
+                          8, 9))
+nr <- nrow(fromto)
+arrpos <- matrix(ncol = 2, nrow = nr)
+
+## Build flow chart
+for (i in 1:nr){
+  arrpos[i, ] <- straightarrow (to = elpos[fromto[i, 2], ],
+                                from = elpos[fromto[i, 1], ],
+                                lwd = 2, arr.pos = 0.5, arr.length = 0.5)
+  textrect(elpos[1,], 0.10, 0.03, lab = "get_kegg()", box.col = "steelblue1",
+              shadow.col = "darkblue", shadow.size = 0.005, cex = 1.0)
+  textround(elpos[2,], 0.10, 0.03,lab = "KEGG REST API", box.col = "orange1",
+            shadow.col = "orange4", shadow.size = 0.005, cex = 1.0)
+  textrect(elpos[3,], 0.10, 0.03,lab = "pathEnrich(List 1)", box.col = "steelblue1",
+            shadow.col = "darkblue", shadow.size = 0.005, cex = 1.0)
+  textrect(elpos[4,], 0.10, 0.03, lab = "pathEnrich(List 2)", box.col = "steelblue1",
+           shadow.col = "darkblue", shadow.size = 0.005, cex = 1.0)
+  textellipse(elpos[5,], 0.10, 0.04, lab = c("pathEnrich","result/object"), box.col = "firebrick1",
+              shadow.col = "firebrick4", shadow.size = 0.005, cex = 1.0)
+  textellipse(elpos[6,], 0.10, 0.04, lab = c("pathEnrich","result/object"),box.col = "firebrick1",
+              shadow.col = "firebrick4", shadow.size = 0.005, cex = 1.0)
+  textrect(elpos[7,], 0.10, 0.03, lab = "diffEnrich()", box.col = "steelblue1",
+           shadow.col = "darkblue", shadow.size = 0.005, cex = 1.0)
+  textellipse(elpos[8,], 0.10, 0.04, lab = c("diffEnrich","result/object"), box.col = "firebrick1",
+              shadow.col = "firebrick4", shadow.size = 0.005, cex = 1.0)
+  textrect(elpos[9,], 0.11, 0.03, lab = "plotFoldEnrichment", box.col = "steelblue1",
+           shadow.col = "darkblue", shadow.size = 0.005, cex = 1.0)
+  text(0.2,0.95, "Figure 1. diffEnrich Analysis pipeline", cex = 0.75)
+}
+```
+
+## Introduction
+
+## Differential Enrichment in practice
 
@@
 
-```{r plotFoldEnrichment}
+```{r plotFoldEnrichment, dpi=400}
 ## Plot fold enrichment

### 2019-08-26 README.Rmd
Added detailed version of analysis pipeline flowchart to README

--- 2019-08-23 README.Rmd
@@
 
-development of methods and software for differential enrichment analysis
+```{r, echo=FALSE, fig.cap='Figure 1. diffEnrich analysis pipeline.', dpi=400}
+suppressMessages(library(diagram))
+
+## open plot
+par(mar = c(0.5, 0.5, 0.5, 0.5))
+openplotmat()
+## define number of boxes at each level
+elpos <- coordinates (c(3, 1, 1, 2, 2, 1, 1, 1))
+## generate matrix of from-to arrow coordinates
+fromto <- matrix(ncol = 2, byrow = TRUE,
+                 data = c(1, 6,
+                          2, 4,
+                          3, 7,
+                          4, 5,
+                          5, 6,
+                          5, 7,
+                          6, 8,
+                          7, 9,
+                          8, 10, 
+                          9, 10,
+                          10, 11,
+                          11, 12))
+nr <- nrow(fromto)
+arrpos <- matrix(ncol = 2, nrow = nr)
+
+## Build flow chart
+for (i in 1:nr){
+  arrpos[i, ] <- straightarrow (to = elpos[fromto[i, 2], ],
+                                from = elpos[fromto[i, 1], ],
+                                lwd = 2, arr.pos = 0.6, arr.length = 0.5)
+  textellipse(elpos[1,], 0.10, 0.04, lab = "geneList1", box.col = "firebrick1",
+              shadow.col = "firebrick4", shadow.size = 0.005, cex = 1.0)
+  textrect(elpos[2,], 0.10, 0.03, lab = "get_kegg()", box.col = "steelblue1",
+              shadow.col = "darkblue", shadow.size = 0.005, cex = 1.0)
+  textellipse(elpos[3,], 0.10, 0.04, lab = "geneList2", box.col = "firebrick1",
+              shadow.col = "firebrick4", shadow.size = 0.005, cex = 1.0)
+  textround(elpos[4,], 0.10, 0.03,lab = "KEGG REST API", box.col = "orange1",
+            shadow.col = "orange4", shadow.size = 0.005, cex = 1.0)
+  textellipse(elpos[5,], 0.15, 0.04, lab = c("cleaned gene-to-pathway", "map"), box.col = "firebrick1",
+              shadow.col = "firebrick4", shadow.size = 0.005, cex = 1.0)
+  textrect(elpos[6,], 0.10, 0.03,lab = "pathEnrich(List 1)", box.col = "steelblue1",
+            shadow.col = "darkblue", shadow.size = 0.005, cex = 1.0)
+  textrect(elpos[7,], 0.10, 0.03, lab = "pathEnrich(List 2)", box.col = "steelblue1",
+           shadow.col = "darkblue", shadow.size = 0.005, cex = 1.0)
+  textellipse(elpos[8,], 0.10, 0.04, lab = c("pathEnrich","result/object"), box.col = "firebrick1",
+              shadow.col = "firebrick4", shadow.size = 0.005, cex = 1.0)
+  textellipse(elpos[9,], 0.10, 0.04, lab = c("pathEnrich","result/object"),box.col = "firebrick1",
+              shadow.col = "firebrick4", shadow.size = 0.005, cex = 1.0)
+  textrect(elpos[10,], 0.10, 0.03, lab = "diffEnrich()", box.col = "steelblue1",
+           shadow.col = "darkblue", shadow.size = 0.005, cex = 1.0)
+  textellipse(elpos[11,], 0.10, 0.04, lab = c("diffEnrich","result/object"), box.col = "firebrick1",
+              shadow.col = "firebrick4", shadow.size = 0.005, cex = 1.0)
+  textrect(elpos[12,], 0.11, 0.03, lab = "plotFoldEnrichment", box.col = "steelblue1",
+           shadow.col = "darkblue", shadow.size = 0.005, cex = 1.0)
+  text(0.15,0.05, "Figure 1. diffEnrich Analysis pipeline", cex = 0.75)
+}
+```
+
+## Introduction
+
+## Differential Enrichment in practice
 

### 2019-08-28 README.Rmd
Added API descrption and diffEnrich in practice to README and vignette

@@
 
+## Introduction
+
 The goal of diffEnrich is simple. Given a list of gene symbols, *diffEnrich*  will
@@
 and Genomes (KEGG) REST API. This package provides a number of functions that are 
-intended to be used in a pipeline (See Figure N). Briefly, the workflow will
+intended to be used in a pipeline (See Figure 1). Briefly, the workflow will
 download and clean species specific ENTREZ gene IDs and map them to their respective
@@
 the analysis pipeline, this package also provides a plotting function.
+
+**The KEGG REST API**
+
+KEGG is a database resource for understanding high-level functions and utilities 
+of the biological system, such as the cell, the organism and the ecosystem, 
+from genomic and molecular-level information [REF](https://www.kegg.jp/kegg/kegg1a.html). KEGG is an integrated database 
+resource consisting of eighteen databases that are clustered into 4 main categories:
+1) systems information (e.g. hierarchies and maps), 2) genomic information 
+(e.g. genes and proteins), 3) chemical information (e.g. biochemical reactions), 
+and 4) health information (e.g. human disease and drugs) [REF](https://www.kegg.jp/kegg/kegg1a.html).
+
+In 2012 KEGG released its first application programming interface (API), and has been
+adding features and functionality ever since. There are benefits to using an API.
+First, API's like KEGG's allow users to perform customized analyses with the most
+up-t-date versions of the data contained in the data base. In addition, accessing the
+KEGG API is very easy using statistical programming tools like R or Python and
+integrating data pulls into user's code makes the program reproducible. To 
+further enforce reproducibilty *diffEnrich* adds a date and KEGG realease tag to
+all data files it generates from accessing the API. For update histories and release 
+notes please go [here](https://www.kegg.jp/kegg/rest/).
 
@@
 
-## Introduction
+**Differential Enrichment in practice**
 
-## Differential Enrichment in practice
+A recent [study](https://onlinelibrary.wiley.com/doi/full/10.1111/acer.13766) explored 
+the hepatic mechanism for removal of toxic lipid aldehydes via conjugation with glutathione (Cheng et al., [2001](https://www.ncbi.nlm.nih.gov/pubmed/11488593?dopt=Abstract);
+Gallagher et al., [2007](https://www.ncbi.nlm.nih.gov/pubmed/17553661?dopt=Abstract)) 
+and the primary enzyme that catalyzes this conjugation, glutathione S‐transferase A4‐4 (GSTA4).
+Specifically, the researchers examin the role of the GSTA4 gene on protein 
+carbonylation and the progression of liver injury in a model consisting of long‐term 
+(116 days) chronic ethanol (EtOH) consumption followed by a single EtOH binge.
 
-Papers from Saba Lab where differentially enrichment has been used:
-1. [Shearn et al 2018 - ACER](https://onlinelibrary.wiley.com/doi/full/10.1111/acer.13766)
+Functional enrichment of carbonylated proteins for KEGG pathways and was tested 
+using a 1‐sided Fisher's exact test in each experimental group 
+(2 genotypes × 3 treatment conditions = 6 experimental groups). Differential 
+enrichment between experimental groups was examined using a 2‐sided Fisher's exact test.
+For KEGG pathway enrichment, proteins were mapped to pathways using their UniProt ID, 
+and databases were downloaded from the KEGG API.
+
+Note: This example uses proteins as the feature of interest, however *diffEnrich*
+currently only supports genes as the feature of interest.
+
+Differential enrichment analysis of carbonylated proteins indicated that compared 
+to chow‐fed animals, the most overrepresented group of proteins that was 
+significantly adducted after long‐term EtOH exposure was ribosomal proteins.
 

### 2019-08-28 README.Rmd
Added appveyor
Added reference subsetction to README
testing appveyor badge status
Updated appvayer badge url
Proof read up to section 4 in README and added travis

@@
 <!-- README.md is generated from README.Rmd. Please edit that file -->
+[![Build status](https://ci.appveyor.com/api/projects/status/xwu8nket2u1pd5ij?svg=true)](https://ci.appveyor.com/project/hsmith9002/diffenrich)
+
+[![Travis build status](https://travis-ci.com/SabaLab/diffEnrich.svg?branch=master)](https://travis-ci.com/SabaLab/diffEnrich)
 
@@
 First, API's like KEGG's allow users to perform customized analyses with the most
-up-t-date versions of the data contained in the data base. In addition, accessing the
+up-to-date versions of the data contained in the data base. In addition, accessing the
 KEGG API is very easy using statistical programming tools like R or Python and
@@
 
-Functional enrichment of carbonylated proteins for KEGG pathways and was tested 
+Functional enrichment of carbonylated proteins for KEGG pathways was tested 
 using a 1‐sided Fisher's exact test in each experimental group 
@@
 ### Step 1: Collect and clean pathways from KEGG API
-First we will use the *get_kegg* function access the KEGG REST API and download the data sets required to perform 
+First we will use the *get_kegg* function to access the KEGG REST API and download the data sets required to perform 
 our downstream analysis. This function takes two arguments. The first, 'species' is required and is the species of 
@@
 create a named list with the three relevant KEGG data sets. The names of this list will describe the the data set.
+For a ddetailed descritpion of list elements use *?get_kegg*.
 
@@
 
-Here are the files:
+Here are examples of the output files:
 
@@
 
-In this step we will use the *pathEnrich* function to identify KEGG pathways that are enriched (i.e. over-represented) based on a list of genes we
-are interested in and based on a list of background genes. This function may not always use the complete list of genes provided 
+In this step we will use the *pathEnrich* function to identify KEGG pathways that are enriched (i.e.
+over-represented) based on two gene lists of interest.
+The ENTREZ gene IDs.
+User gene lists must also be charater vectors and be formatted as ENTREZ gene IDs.
+The *clusterProfiler* offers a nice function (*bitr*) that maps gene symbols and Ensembl IDs to ENTREZ gene IDs, and an example can be seen in their [vignette](https://yulab-smu.github.io/clusterProfiler-book/chapter5.html#supported-organisms).
+
+```{r view_genelists}
+## View sample gene lists from package data
+head(geneLists$list1)
+head(geneLists$list2)
+```
+
+This function may not always use the complete list of genes provided 
 by the user. Specifically, it will only use the genes from the list provided that are also in
 the most current species list pulled from the KEGG REST API using *get_kegg*, or from the older KEGG data loaded by the user from a 
-previous *get_kegg* call. The *pathEnrich* function should be run at least twice, once for the genes of interest and once for the background.
-Each *pathEnrich* call generates a data frame summarizing the results of an enrichment analysis in which a Fisher's
-Exact test is used to identify which KEGG pathways are enriched by the user's list of interesting genes with respect to background
-enrichment, and a p-value is calculated using a hypergeometric distribution (Formula 1). P-values are adjusted for multiple comparisons by controlling the False Discovery Rate (FDR) at 0.05.
+previous *get_kegg* call.
+The *pathEnrich* function should be run at least twice, once for the genes of interest in list 1 and once for the genes of interest in list2.
+Each *pathEnrich* call generates a data frame summarizing the results of an enrichment analysis in which a Fisher's Exact test is used to identify which KEGG pathways are enriched by the user's list of interesting genes with respect to background
+(universe), and a p-value is calculated using a hypergeometric distribution (Formula 1). P-values are adjusted for multiple comparisons by controlling the False Discovery Rate (FDR) at 0.05.
 
@@
 where N is the total number of genes in the background, M is the number of genes within that background that also in the gene set of interest, n is the size of the list of genes of interest and k is the number of genes within that are in the background.
-The background by default is all the genes that have annotation.
+The background (universe) by default is all the genes that have annotation.
 
@@
 
-*pathEnrich* generates a data frame with 9 columns described below.
+*pathEnrich* generates a data frame with 9 columns described below. Details also provided in *pathEnrich* documentation. Use *?pathEnrich*.
 
@@
 
-The diffEnrich function will merge the results from the *pathEnrich* calls generated above. Specifically, the data frame
+The *diffEnrich* function will merge the results from the *pathEnrich* calls generated above. Specifically, the data frame
 'list1_pe' and the data frame 'list2_pe' will be merged using the base *merge* functionality and be merged by the following columns:
@@
 
-The result of the *diffEnrich* call is the merged data frame with the estimated odds ration generated by the Fisher's Exact test 
-and the associated p-value. Users can choose from those supported by *stats::p.adjust*, and the default is the False Discovery Rate (Benjamini and Hochberg).
+The result of the *diffEnrich* call is the merged data frame with the estimated odds ratio generated by the Fisher's Exact test 
+and the associated p-value.
+Users can choose from those supported by *stats::p.adjust*, and the default is the False Discovery Rate (Benjamini and Hochberg, [1995](http://www.jstor.org/stable/2346101)).
 
@@
 ```
+
+## References

### 2019-08-28 README.Rmd
Finished proof reading README

@@
 
-This function generates a grouped bar plot using ggplot2 and the
-ggnewscale package. KEGG pathways are plotted on the y-axis and fold
+*plotFoldEnrichment* generates a grouped bar plot using ggplot2 and the
+*ggnewscale* package. KEGG pathways are plotted on the y-axis and fold
 enrichment is plotted on the x-axis. each KEGG pathway has a bar plotting
 its fold enrichment in list 1 (red) and its fold enrichment in list 2 (blue).
-The transparency of the bars correspond to the adjusted p-value for the
+The transparency of the bars correspond to the adjusted p-value (FDR) for the
 pathway's enrichment in the given list. The p-value presented as text to the
-right of each pair of bars is the adjusted p-value associated with the
+right of each pair of bars is the adjusted p-value (user defined: default is FDR) associated with the
 differential enrichment of the pathway between the two lists, and the pathways
@@
 
-```{r plotFoldEnrichment, dpi=400}
+```{r plotFoldEnrichment, dpi=400, message=FALSE}
 ## Plot fold enrichment

### 2019-08-28 R/plotFoldEnrichment.R
Formatted p-values in plotFoldEnrichment

@@ -115,6 +115,11 @@ plotFoldEnrichment <- function(de_res, pval, N){
   df_ptext <- subset(df_ptext, df_ptext$variable %in% c("diff_enrich_adjusted")) %>%
     dplyr::filter(!duplicated(.data$value))
 
+  ## Format list p-values
+  options(scipen = 999)
+  lpval <- sprintf(as.numeric(summary(bardat$pvals))[c(1,2,3,5,6)], fmt = '%#.5f')
+  lpval <- ifelse(lpval < 0.000001, "< 0.000001", lpval)
+
   ## Generate finale plot
   p <- ggplot(mapping = aes(xmin = .data$xmin, xmax = .data$xmax, ymin = .data$ymin, ymax = .data$ymax)) +
     geom_rect(data = ld[ld$vars == "fold_enrichment_list1", ], aes(fill = .data$pvals)) +
@@ -123,7 +128,7 @@ plotFoldEnrichment <- function(de_res, pval, N){
                         #trans = 'log10',
                         limits = c(min(ld$pvals), 0),
                         breaks = as.numeric(summary(ld$pvals))[c(1,2,3,5,6)],
-                        labels = as.character(formatC(as.numeric(summary(bardat$pvals))[c(1,2,3,5,6)], format = "e", digits = 2)),
+                        labels = lpval,
                         name = "P-values List 1") +
     ggnewscale::new_scale_fill() +
     geom_rect(data = ld[ld$vars == "fold_enrichment_list2", ], aes(fill = .data$pvals)) +
@@ -131,12 +136,12 @@ plotFoldEnrichment <- function(de_res, pval, N){
                         #trans = 'log10',
                         limits = c(min(ld$pvals), 0),
                         breaks = as.numeric(summary(ld$pvals))[c(1,2,3,5,6)],
-                        labels = as.character(formatC(as.numeric(summary(bardat$pvals))[c(1,2,3,5,6)], format = 'e', digits = 2)),
+                        labels = lpval,
                         name = "P-values List 2") +
     scale_x_continuous(breaks = seq_along(unique(ld$descr)),
                        labels = unique(ld$descr)) +
     coord_flip() + theme_bw() +
     geom_text(data=df_ptext,
-              aes(x = 1:N, y = (max(bardat$value) + 0.3), label = round(.data$value, 4)))
+              aes(x = 1:N, y = (max(bardat$value) + 0.4), label = round(.data$value, 5)))
   return(p)
 }

### 2019-08-28 README.Rmd
added Travis badge
Added gene numbers to plot legend and removed devel from travis yaml

@@
 
-[![Travis build status](https://travis-ci.com/SabaLab/diffEnrich.svg?branch=master)](https://travis-ci.com/SabaLab/diffEnrich)
+[![Build Status](https://travis-ci.com/SabaLab/diffEnrich.svg?token=M2jELEzTYYqmxZMY6hpb&branch=master)](https://travis-ci.com/SabaLab/diffEnrich)
 
@@
 are ordered from top to bottom by this p-value (i.e. smallest p-value on top
-of plot, and largest p-value on bottom of plot).
+of plot, and largest p-value on bottom of plot). Finally, the number of genes used
+for analysis from each gene list (recall that this number may not be the same as the number of
+genes in the user's original list) are reported below their repsective p-values
+in the legend.
 

### 2019-08-28 R/plotFoldEnrichment.R

@@ -129,7 +129,7 @@ plotFoldEnrichment <- function(de_res, pval, N){
                         limits = c(min(ld$pvals), 0),
                         breaks = as.numeric(summary(ld$pvals))[c(1,2,3,5,6)],
                         labels = lpval,
-                        name = "P-values List 1") +
+                        name = paste0("P-values List 1", "\n", "gene count: ", de_res[1,6])) +
     ggnewscale::new_scale_fill() +
     geom_rect(data = ld[ld$vars == "fold_enrichment_list2", ], aes(fill = .data$pvals)) +
     scale_fill_gradient(low =  "navy", high = "transparent",
@@ -137,7 +137,7 @@ plotFoldEnrichment <- function(de_res, pval, N){
                         limits = c(min(ld$pvals), 0),
                         breaks = as.numeric(summary(ld$pvals))[c(1,2,3,5,6)],
                         labels = lpval,
-                        name = "P-values List 2") +
+                        name = paste0("P-values List 2", "\n", "gene count: ", de_res[1,12])) +
     scale_x_continuous(breaks = seq_along(unique(ld$descr)),
                        labels = unique(ld$descr)) +
     coord_flip() + theme_bw() +

### 2019-08-28 R/get_kegg.R
Added dontrun call to get_kegg examples

@@ -23,7 +23,9 @@
 #' @importFrom here here
 #'
 #' @examples
+#' \dontrun{
 #' kegg <- get_kegg(species = "rno")
+#' }
 #' \dontrun{
 #' kegg <- get_kegg(species = "mmu", path = "usr/data/out/")
 #' }

### 2019-08-28 README.Rmd
Cleaned up README and vignette
spell/grammer check README and vingette

@@
 integrating data pulls into user's code makes the program reproducible. To 
-further enforce reproducibilty *diffEnrich* adds a date and KEGG realease tag to
+further enforce reproducibilty *diffEnrich* adds a date and KEGG release tag to
 all data files it generates from accessing the API. For update histories and release 
@@
 and the primary enzyme that catalyzes this conjugation, glutathione S‐transferase A4‐4 (GSTA4).
-Specifically, the researchers examin the role of the GSTA4 gene on protein 
+Specifically, the researchers examine the role of the GSTA4 gene on protein 
 carbonylation and the progression of liver injury in a model consisting of long‐term 
@@
 create a named list with the three relevant KEGG data sets. The names of this list will describe the the data set.
-For a ddetailed descritpion of list elements use *?get_kegg*.
+For a detailed description of list elements use *?get_kegg*.
 
@@
 The ENTREZ gene IDs.
-User gene lists must also be charater vectors and be formatted as ENTREZ gene IDs.
+User gene lists must also be character vectors and be formatted as ENTREZ gene IDs.
 The *clusterProfiler* offers a nice function (*bitr*) that maps gene symbols and Ensembl IDs to ENTREZ gene IDs, and an example can be seen in their [vignette](https://yulab-smu.github.io/clusterProfiler-book/chapter5.html#supported-organisms).
@@
 for analysis from each gene list (recall that this number may not be the same as the number of
-genes in the user's original list) are reported below their repsective p-values
+genes in the user's original list) are reported below their respective p-values
 in the legend.
@@
 ```
-
-## References

### 2019-09-13 R/diffEnrich.R
updated fisher test in diffEnrich

@@ -60,7 +60,7 @@ diffEnrich <- function(list1_pe, list2_pe, method = 'BH'){
 
   ## Build diffEnrich Fisher's Exact function
   de <- function(a,b,c,d){
-    y <- stats::fisher.test(matrix(c(a,b,c,d), nrow = 2))
+    y <- stats::fisher.test(matrix(c(a,b,c-a,d-b), nrow = 2))
     est <- y$estimate
     pv <- y$p.value
     out.de <- data.frame(est, pv)

### 2019-09-13 README.Rmd
Updated goal of package
Made changes to intro and kegg api section of README and vignette

@@
 
-The goal of diffEnrich is simple. Given a list of gene symbols, *diffEnrich*  will
+The goal of diffEnrich is to compare functional enrichment between two experimentally-derived groups of genes or proteins. Given a list of gene symbols, *diffEnrich*  will
 perform differential enrichment analysis using the Kyoto Encyclopedia of Genes
 and Genomes (KEGG) REST API. This package provides a number of functions that are 
-intended to be used in a pipeline (See Figure 1). Briefly, the workflow will
+intended to be used in a pipeline (See Figure 1). Briefly, the user provides a KEGG formated species id for either human, mouse or rat, and the package will
 download and clean species specific ENTREZ gene IDs and map them to their respective
-KEGG pathways by accessing KEGG's REST API. This way the user will always have 
-the most up-to-date pathway data from KEGG. Next, the user will identify significantly
+KEGG pathways by accessing KEGG's REST API. KEGG's API is used ot guarantee the most up-to-date pathway data from KEGG. Next, the user will identify significantly
 enriched pathways from two different gene sets, and finally, the user will identify 
@@
 
-KEGG is a database resource for understanding high-level functions and utilities 
-of the biological system, such as the cell, the organism and the ecosystem, 
+KEGG is a database resource for understanding high-level functions 
+of a biological system, such as a cell, an organism and an ecosystem, 
 from genomic and molecular-level information [REF](https://www.kegg.jp/kegg/kegg1a.html). KEGG is an integrated database 
@@
 First, API's like KEGG's allow users to perform customized analyses with the most
-up-to-date versions of the data contained in the data base. In addition, accessing the
+up-to-date versions of the data contained in the database. In addition, accessing the
 KEGG API is very easy using statistical programming tools like R or Python and
@@
 all data files it generates from accessing the API. For update histories and release 
-notes please go [here](https://www.kegg.jp/kegg/rest/).
+notes for the KEGG REST API please go [here](https://www.kegg.jp/kegg/rest/).
 

### 2019-09-13 README.Rmd
Changed geneList bubble colors in flow chart

@@
                                 lwd = 2, arr.pos = 0.6, arr.length = 0.5)
-  textellipse(elpos[1,], 0.10, 0.04, lab = "geneList1", box.col = "firebrick1",
-              shadow.col = "firebrick4", shadow.size = 0.005, cex = 1.0)
+  textellipse(elpos[1,], 0.10, 0.04, lab = "geneList1", box.col = "purple",
+              shadow.col = "purple4", shadow.size = 0.005, cex = 1.0)
   textrect(elpos[2,], 0.10, 0.03, lab = "get_kegg()", box.col = "steelblue1",
               shadow.col = "darkblue", shadow.size = 0.005, cex = 1.0)
-  textellipse(elpos[3,], 0.10, 0.04, lab = "geneList2", box.col = "firebrick1",
-              shadow.col = "firebrick4", shadow.size = 0.005, cex = 1.0)
+  textellipse(elpos[3,], 0.10, 0.04, lab = "geneList2", box.col = "purple",
+              shadow.col = "purple4", shadow.size = 0.005, cex = 1.0)
   textround(elpos[4,], 0.10, 0.03,lab = "KEGG REST API", box.col = "orange1",

### 2019-09-13 README.Rmd
added step numbers and detaile figure legend to Figure 1

@@
   fig.path = "man/figures/README-",
-  out.width = "100%"
+  out.width = "100%",
+  eval.after = 'fig.cap'
 ) 
@@
 
-```{r, echo=FALSE, fig.cap='Figure 1. diffEnrich analysis pipeline.', dpi=400}
+```{r, echo=FALSE, fig.cap=cap, dpi=400}
 suppressMessages(library(diagram))
@@
            shadow.col = "darkblue", shadow.size = 0.005, cex = 1.0)
-  text(0.15,0.05, "Figure 1. diffEnrich Analysis pipeline", cex = 0.75)
+  # text(0.15,0.05, "Figure 1. diffEnrich Analysis pipeline", cex = 0.75)
+  text(0.05,0.87, "Step 1", cex = 1.2, font = 2)
+  text(0.05,0.57, "Step 2", cex = 1.2, font = 2)
+  text(0.05,0.35, "Step 3", cex = 1.2, font = 2)
+  text(0.05,0.10, "Step 4", cex = 1.2, font = 2)
 }
+
+cap <- c("Figure 1.
+diffEnrich Analysis pipeline.
+Functions within the diffEnrich package are represented by blue rectangles.
+The data that must be provided by the user is represented by the pruple ovals.
+Data objects generated by a function in diffEnrich are representaed by red ovals.
+The external call of the get_kegg function to the KEGG REST API is represented in yellow.")
 ```

### 2019-09-13 README.Rmd
Changed section title for motivating example of diff enrich
Edited step 1section in readme and hid library call

@@
 
-**Differential Enrichment in practice**
+**Motivating experimental design for differential enrichment**
 
@@
 First we will use the *get_kegg* function to access the KEGG REST API and download the data sets required to perform 
-our downstream analysis. This function takes two arguments. The first, 'species' is required and is the species of 
-interest. Currently, *diffEnrich* supports three species, and the argument is a character string using the KEGG
+our downstream analysis. This function takes two arguments. The first, 'species' is required. Currently, *diffEnrich* supports three species, and the argument is a character string using the KEGG
 code [REF](https://www.pnas.org/content/suppl/2008/09/11/0806162105.DCSupplemental/ST1_PDF.pdf): 
@@
 and the version of KEGG when the API was accessed. In addition to these flat files, *get_kegg* will also
-create a named list with the three relevant KEGG data sets. The names of this list will describe the the data set.
+create a named list with the three relevant KEGG data sets. The names of this list will describe the data set.
 For a detailed description of list elements use *?get_kegg*.
 
+```{r, results='hide', echo=FALSE}
+suppressMessages(library(diffEnrich))
+```
+
 ```{r get_kegg_exp1}
-suppressMessages(library(diffEnrich))
-
 ## run get_kegg() using rat
@@
 and if it finds even one of the three, it will not re-write any of the data sets. It will still however, let the user know
-it is not writing out new data sets and still generate the named list object. Users can of course generate 'same-day/same-version'
+it is not writing out new data sets and still generate the named list object. Users can generate 'same-day/same-version'
 data sets in different directories if they so choose.

### 2019-09-13 README.Rmd
Added description of get_kegg output in Table 1

@@
 ) 
+options(knitr.table.format = "html")
 ```
@@
 
+```{r get_kegg_descriptions, echo=FALSE}
+# Define column names
+kegg_df_dat <- c("ncbi_to_kegg",
+                 "kegg_to_pathway",
+                 "pathway_to_species")
+
+kegg_df_dscr <- c(paste0("ncbi gene ID",
+                 " <-- mapped to --> ",
+                 "KEGG gene ID"),
+                 paste0("KEGG gene ID",
+                 " <-- mapped to --> ",
+                 "KEGG pathway ID"),
+                 paste0("KEGG pathway ID",
+                 " <-- mapped to --> ",
+                 "KEGG pathway species description"))
+# Make dataframe
+kegg_df <- data.frame("get_kegg list object" = kegg_df_dat, 
+                      "Object description" = kegg_df_dscr)
+```
+
+```{r get_kegg_kable, warning=FALSE, echo=FALSE, message=FALSE}
+if (require(kableExtra)){
+knitr::kable(kegg_df,
+             caption = "Table 1. Description of data sets generated from accessing KEGG REST API", row.names = FALSE) %>%
+  kableExtra::kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = T)
+}
+```
+
 ### Step 2: Perform individual enrichment analysis
@@
 knitr::kable(head(list1_pe),
-             caption = "Table 1. Head of list1_pe data frame generated using pathEnrich", row.names = FALSE) %>%
+             caption = "Table 2. Head of list1_pe data frame generated using pathEnrich", row.names = FALSE) %>%
   kableExtra::kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = T)
@@
 knitr::kable(df1,
-             caption = "Table 2. Description of columns is dataframe generated by pathEnrich", row.names = FALSE) %>%
+             caption = "Table 3. Description of columns is dataframe generated by pathEnrich", row.names = FALSE) %>%
   kableExtra::kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = F)

### 2019-09-13 README.Rmd
testing escape in tables

@@
 ) 
-options(knitr.table.format = "html")
 ```
@@
 if (require(kableExtra)){
-knitr::kable(kegg_df,
+knitr::kable(kegg_df, format = 'html',escape = FALSE,
              caption = "Table 1. Description of data sets generated from accessing KEGG REST API", row.names = FALSE) %>%

### 2019-09-13 README.Rmd
working on escape bug in kables
This escape is a bug that has to do with either R or pandoc version. Works from iMac

@@
 if (require(kableExtra)){
-knitr::kable(kegg_df, format = 'html',escape = FALSE,
+knitr::kable(kegg_df, escape = FALSE,
              caption = "Table 1. Description of data sets generated from accessing KEGG REST API", row.names = FALSE) %>%

### 2019-09-13 README.Rmd
Made dits to step 2 sections of README and vignette

@@
 
-### Step 2: Perform individual enrichment analysis
+### Step 2: Perform enrichment analysis of individual gene sets
 
 In this step we will use the *pathEnrich* function to identify KEGG pathways that are enriched (i.e.
-over-represented) based on two gene lists of interest.
-The ENTREZ gene IDs.
+over-represented) based on a gene list of interest.
 User gene lists must also be character vectors and be formatted as ENTREZ gene IDs.
-The *clusterProfiler* offers a nice function (*bitr*) that maps gene symbols and Ensembl IDs to ENTREZ gene IDs, and an example can be seen in their [vignette](https://yulab-smu.github.io/clusterProfiler-book/chapter5.html#supported-organisms).
+The *clusterProfiler* package offers a nice function (*bitr*) that maps gene symbols and Ensembl IDs to ENTREZ gene IDs, and an example can be seen in their [vignette](https://yulab-smu.github.io/clusterProfiler-book/chapter5.html#supported-organisms).
 
@@
 The *pathEnrich* function should be run at least twice, once for the genes of interest in list 1 and once for the genes of interest in list2.
-Each *pathEnrich* call generates a data frame summarizing the results of an enrichment analysis in which a Fisher's Exact test is used to identify which KEGG pathways are enriched by the user's list of interesting genes with respect to background
-(universe), and a p-value is calculated using a hypergeometric distribution (Formula 1). P-values are adjusted for multiple comparisons by controlling the False Discovery Rate (FDR) at 0.05.
-
-Formula 1:
-
-$$ P = 1-\sum_{i=0}^{k-1} \frac{ \binom{m} {i} \binom{N-M} {n-i} } {\binom{N} {n} } $$
-
-where N is the total number of genes in the background, M is the number of genes within that background that also in the gene set of interest, n is the size of the list of genes of interest and k is the number of genes within that are in the background.
-The background (universe) by default is all the genes that have annotation.
+Each *pathEnrich* call generates a data frame summarizing the results of an enrichment analysis in which a Fisher's Exact test is used to identify which KEGG pathways are enriched within the user's list of interesting genes compared to all genes annotated to a KEGG pathway.
+P-values from the Fisher's Exact test are adjusted for multiple comparisons by controlling the False Discovery Rate (FDR) at 0.05.
 
@@
 knitr::kable(head(list1_pe),
-             caption = "Table 2. Head of list1_pe data frame generated using pathEnrich", row.names = FALSE) %>%
+             caption = "Table 2. First 6 rows of list1_pe data frame generated using pathEnrich", row.names = FALSE) %>%
   kableExtra::kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = T)

### 2019-09-13 README.Rmd
Made edits to Table 3

@@
                  "Number of Genes from gene list in KEGG Database",
-                 "Expected number of genes from list to be in KEGG pathway by chance (i.e., not enriched)",
-                 "P-value for enrichment of list genes related to KEGG pathway",
+                 "Expected number of genes from list to be in KEGG pathway by chance",
+                 "P-value for enrichment within the KEGG pathway for list genes",
                  "False Discovery Rate (Benjamini and Hochberg) to account for multiple testing across KEGG pathways",
-                 "KEGG_PATHWAY_in_list/expected")
+                 "Ratio of number of genes observed from the gene list annotated to the KEGG pathay to the number og genes expected from the gene list annotated to the KEGG pathway if ther was no enrichment (i.e.
+KEGG_PATHWAY_in_list/expected)")
 # Make dataframe

### 2019-09-13 README.Rmd
Made edits to step 3 of README and vignette

@@
 
-The *diffEnrich* function will merge the results from the *pathEnrich* calls generated above. Specifically, the data frame
-'list1_pe' and the data frame 'list2_pe' will be merged using the base *merge* functionality and be merged by the following columns:
-'c("KEGG_PATHWAY_ID", "KEGG_PATHWAY_description", "KEGG_PATHWAY_cnt", "KEGG_DATABASE_cnt")'.
+The *diffEnrich* function will merge two results from the *pathEnrich* calls generated above. Specifically, the data frame
+'list1_pe' and the data frame 'list2_pe' will be merged by the following columns:
+"KEGG_PATHWAY_ID", "KEGG_PATHWAY_description", "KEGG_PATHWAY_cnt", "KEGG_DATABASE_cnt".
 This merged data set will then be used to perform differential enrichment using the same method and p-value calculation as described above.
@@
 for multiple testing adjustment.
+Users can choose from those supported by *stats::p.adjust* for multiple correction options, and the default is the False Discovery Rate (Benjamini and Hochberg, [1995](http://www.jstor.org/stable/2346101)).
 
@@
 
-The result of the *diffEnrich* call is the merged data frame with the estimated odds ratio generated by the Fisher's Exact test 
+The result of the *diffEnrich* call is a data frame with the estimated odds ratio generated by the Fisher's Exact test 
 and the associated p-value.
-Users can choose from those supported by *stats::p.adjust*, and the default is the False Discovery Rate (Benjamini and Hochberg, [1995](http://www.jstor.org/stable/2346101)).
 

### 2019-09-13 README.Rmd
Made edits to step 4 in README and vignette
Added Table 5 describing variables in diffEnrich data frame

@@
                  "False Discovery Rate (Benjamini and Hochberg) to account for multiple testing across KEGG pathways",
-                 "Ratio of number of genes observed from the gene list annotated to the KEGG pathay to the number og genes expected from the gene list annotated to the KEGG pathway if ther was no enrichment (i.e.
+                 "Ratio of number of genes observed from the gene list annotated to the KEGG pathay to the number of genes expected from the gene list annotated to the KEGG pathway if ther was no enrichment (i.e.
 KEGG_PATHWAY_in_list/expected)")
@@
 knitr::kable(head(diff_enrich),
-             caption = "Table 3. Description of columns is data frame generated by diffEnrich", row.names = FALSE) %>%
+             caption = "Table 4.First 6 rows from data frame generated by diffEnrich", row.names = FALSE) %>%
+  kableExtra::kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = F)
+```
+
+```{r diffEnrich_descriptions, echo=FALSE}
+## get vector of column names
+de_col_names <- names(diff_enrich)
+de_col_descr <- c("KEGG Pathway Identifier",
+                  "Description of KEGG Pathway (provided by KEGG)",
+                  "Number of Genes in KEGG Pathway",
+                  "Number of Genes in KEGG Database",
+                  "Number of Genes from gene list 1 in KEGG Pathway",
+                  "Number of Genes from gene list 1 in KEGG Database",
+                  "Expected number of genes from list 1 to be in KEGG pathway by chance",
+                  "P-value for enrichment of list 1 genes related to KEGG pathway",
+                  "False Discovery Rate (Benjamini and Hochberg) of enrich_p_list1 to account for multiple testing across KEGG pathways",
+                  "Ratio of number of genes observed from the gene list 1 annotated to the KEGG pathay to the number of genes expected from the gene list 1 annotated to the KEGG pathway if ther was no enrichment (i.e.
+KEGG_PATHWAY_in_list1/expected_list1)",
+                  "Number of Genes from gene list 2 in KEGG Pathway",
+                  "Number of Genes from gene list 2 in KEGG Database",
+                  "Expected number of genes from list 2 to be in KEGG pathway by chance",
+                  "P-value for enrichment of list 2 genes related to KEGG pathway",
+                  "False Discovery Rate (Benjamini and Hochberg) of enrich_p_list2 to account for multiple testing across KEGG pathways",
+                  "Ratio of number of genes observed from the gene list 2 annotated to the KEGG pathay to the number of genes expected from the gene list 2 annotated to the KEGG pathway if ther was no enrichment (i.e.
+KEGG_PATHWAY_in_list2/expected_list2)",
+                  "Odds of a gene from list 2 being from this KEGG pathway / Odds of a gene from list 1 being from this KEGG pathway",
+                  "P-value for differential enrichment of this KEGG pathway between list 1 and list 2",
+                  "Multiple testing adjustment of diff_enrich_p (default = False Discovery Rate (Benjamini and Hochberg))")
+# Make dataframe
+de_descr_df <- data.frame(de_col_names, de_col_descr)
+colnames(de_descr_df) <- c("Column Names", "Column Description") 
+```
+
+```{r diffEnrich_vars_kable, warning=FALSE, echo=FALSE}
+knitr::kable(de_descr_df,
+             caption = "Table 5. Description of columns is dataframe generated by diffEnrich", row.names = FALSE) %>%
   kableExtra::kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = F)
@@
 *ggnewscale* package. KEGG pathways are plotted on the y-axis and fold
-enrichment is plotted on the x-axis. each KEGG pathway has a bar plotting
+enrichment is plotted on the x-axis. Each KEGG pathway has a bar depicting
 its fold enrichment in list 1 (red) and its fold enrichment in list 2 (blue).

### 2019-09-13 R/plotFoldEnrichment.R
Fixed example in plotfoldenrichment function

@@ -37,7 +37,7 @@
 #'
 #' @examples
 #' \dontrun{
-#' plot <- plotFoldEnrichment(de_res = diffEnrich, pval = 0.05, N = 5)
+#' plot <- plotFoldEnrichment(de_res = diff_enrich, pval = 0.05, N = 5)
 #' }
 plotFoldEnrichment <- function(de_res, pval, N){
   # library(dplyr)

### 2019-09-13 R/plotFoldEnrichment.R
working on sciepen in plot

@@ -116,9 +116,9 @@ plotFoldEnrichment <- function(de_res, pval, N){
     dplyr::filter(!duplicated(.data$value))
 
   ## Format list p-values
-  options(scipen = 999)
-  lpval <- sprintf(as.numeric(summary(bardat$pvals))[c(1,2,3,5,6)], fmt = '%#.5f')
-  lpval <- ifelse(lpval < 0.000001, "< 0.000001", lpval)
+  options(scipen = 0, digits = 2)
+  lpval <-as.numeric(summary(bardat$pvals))[c(1,2,3,5,6)]
+  #lpval <- ifelse(lpval < 0.000001, "< 0.000001", lpval)
 
   ## Generate finale plot
   p <- ggplot(mapping = aes(xmin = .data$xmin, xmax = .data$xmax, ymin = .data$ymin, ymax = .data$ymax)) +

### 2019-09-13 R/plotFoldEnrichment.R
still fixing scipen formatting in plot legend

@@ -118,6 +118,7 @@ plotFoldEnrichment <- function(de_res, pval, N){
   ## Format list p-values
   options(scipen = 0, digits = 2)
   lpval <-as.numeric(summary(bardat$pvals))[c(1,2,3,5,6)]
+  lpval <- as.character(lpval)
   #lpval <- ifelse(lpval < 0.000001, "< 0.000001", lpval)
 
   ## Generate finale plot

### 2019-09-13 R/plotFoldEnrichment.R
fixed scipen formatting in legend

@@ -117,8 +117,7 @@ plotFoldEnrichment <- function(de_res, pval, N){
 
   ## Format list p-values
   options(scipen = 0, digits = 2)
-  lpval <-as.numeric(summary(bardat$pvals))[c(1,2,3,5,6)]
-  lpval <- as.character(lpval)
+  lpval <-formatC(as.numeric(summary(bardat$pvals))[c(1,2,3,5,6)])
   #lpval <- ifelse(lpval < 0.000001, "< 0.000001", lpval)
 
   ## Generate finale plot

### 2019-09-13 R/plotFoldEnrichment.R
Clean/rebuild and reknit README/vignette after fixing formatting issues with plot

@@ -116,13 +116,13 @@ plotFoldEnrichment <- function(de_res, pval, N){
     dplyr::filter(!duplicated(.data$value))
 
   ## Format list p-values
-  options(scipen = 0, digits = 2)
+  options(scipen = 0, digits = 3)
   lpval <-formatC(as.numeric(summary(bardat$pvals))[c(1,2,3,5,6)])
   #lpval <- ifelse(lpval < 0.000001, "< 0.000001", lpval)
 
   ## Generate finale plot
   p <- ggplot(mapping = aes(xmin = .data$xmin, xmax = .data$xmax, ymin = .data$ymin, ymax = .data$ymax)) +
-    geom_rect(data = ld[ld$vars == "fold_enrichment_list1", ], aes(fill = .data$pvals)) +
+    geom_rect(data = ld[ld$vars == "fold_enrichment_list1", ], aes(fill = .data$pvals), color = 'black') +
     ylim(0, max(bardat$value) + 1.0) + xlab("") + ylab("Fold Enrichment") +
     scale_fill_gradient(low = "darkred", high = "transparent",
                         #trans = 'log10',
@@ -131,7 +131,7 @@ plotFoldEnrichment <- function(de_res, pval, N){
                         labels = lpval,
                         name = paste0("P-values List 1", "\n", "gene count: ", de_res[1,6])) +
     ggnewscale::new_scale_fill() +
-    geom_rect(data = ld[ld$vars == "fold_enrichment_list2", ], aes(fill = .data$pvals)) +
+    geom_rect(data = ld[ld$vars == "fold_enrichment_list2", ], aes(fill = .data$pvals), color = 'black') +
     scale_fill_gradient(low =  "navy", high = "transparent",
                         #trans = 'log10',
                         limits = c(min(ld$pvals), 0),
@@ -140,8 +140,9 @@ plotFoldEnrichment <- function(de_res, pval, N){
                         name = paste0("P-values List 2", "\n", "gene count: ", de_res[1,12])) +
     scale_x_continuous(breaks = seq_along(unique(ld$descr)),
                        labels = unique(ld$descr)) +
+    geom_hline(yintercept=1.0, linetype ='dashed') +
     coord_flip() + theme_bw() +
     geom_text(data=df_ptext,
-              aes(x = 1:N, y = (max(bardat$value) + 0.4), label = round(.data$value, 5)))
+              aes(x = 1:N, y = (max(bardat$value) + 0.5), label = round(.data$value, 5)))
   return(p)
 }

### 2019-09-14 README.Rmd
Clean/rebuild and reknit README/vignette after fixing formatting issues with plot
Added details to plot and added figure legends to readme and vignette

@@
 *plotFoldEnrichment* generates a grouped bar plot using ggplot2 and the
-*ggnewscale* package. KEGG pathways are plotted on the y-axis and fold
+*ggnewscale* package.
+There are 3 arguments: 1) *de_res* is the dataframe generated from the *diffEnrich* function, 2) *pval* is the threshold for the adjusted p-value associated with differential enrichment that will filter which KEGG pathways to plot, and 3) after filtering based on *pval* *N* tells the function how many pathways to plot.
+It is important to make a note that the significance of the fold change is associated with the number of genes in the gene list.
+Notice that in this example the pathways in gene list 2 have smaller fold changes (shorter bars) than those in list 1, but that many of them are more significant (darker blue).
+This is because there are more genes in gene list 2 compred to gene list 1.
+
+```{r echo=FALSE}
+cap_plot <- c("Figure 2. Fold enrichment stratified by gene list. KEGG pathways are plotted on the y-axis and fold
 enrichment is plotted on the x-axis. Each KEGG pathway has a bar depicting
 its fold enrichment in list 1 (red) and its fold enrichment in list 2 (blue).
-The transparency of the bars correspond to the adjusted p-value (FDR) for the
+The transparency of the bars correspond to the unadjusted p-value for the
 pathway's enrichment in the given list. The p-value presented as text to the
@@
 are ordered from top to bottom by this p-value (i.e. smallest p-value on top
-of plot, and largest p-value on bottom of plot). Finally, the number of genes used
+of plot, and largest p-value on bottom of plot). The dotted line represents a fold enrichment of 1. Finally, the number of genes used
 for analysis from each gene list (recall that this number may not be the same as the number of
 genes in the user's original list) are reported below their respective p-values
-in the legend.
+in the legend.")
+```
 
-```{r plotFoldEnrichment, dpi=400, message=FALSE}
+```{r plotFoldEnrichment, dpi=400, message=FALSE, fig.cap=cap_plot}
 ## Plot fold enrichment

### 2019-09-16 README.Rmd
Added note on figure legends for README
added arrows to table 1

@@
 the analysis pipeline, this package also provides a plotting function.
+
+**Note on figure legends:** To view figure legends, hover mouse over image.
 
@@
 if (require(kableExtra)){
-knitr::kable(kegg_df, escape = FALSE,
+knitr::kable(kegg_df,
              caption = "Table 1. Description of data sets generated from accessing KEGG REST API", row.names = FALSE) %>%

### 2019-09-16 README.Rmd
Fixed typos in description tables

@@
                  "False Discovery Rate (Benjamini and Hochberg) to account for multiple testing across KEGG pathways",
-                 "Ratio of number of genes observed from the gene list annotated to the KEGG pathay to the number of genes expected from the gene list annotated to the KEGG pathway if ther was no enrichment (i.e.
+                 "Ratio of number of genes observed from the gene list annotated to the KEGG pathway to the number of genes expected from the gene list annotated to the KEGG pathway if there was no enrichment (i.e.
 KEGG_PATHWAY_in_list/expected)")
@@
                   "False Discovery Rate (Benjamini and Hochberg) of enrich_p_list1 to account for multiple testing across KEGG pathways",
-                  "Ratio of number of genes observed from the gene list 1 annotated to the KEGG pathay to the number of genes expected from the gene list 1 annotated to the KEGG pathway if ther was no enrichment (i.e.
+                  "Ratio of number of genes observed from the gene list 1 annotated to the KEGG pathway to the number of genes expected from the gene list 1 annotated to the KEGG pathway if there was no enrichment (i.e.
 KEGG_PATHWAY_in_list1/expected_list1)",
@@
                   "False Discovery Rate (Benjamini and Hochberg) of enrich_p_list2 to account for multiple testing across KEGG pathways",
-                  "Ratio of number of genes observed from the gene list 2 annotated to the KEGG pathay to the number of genes expected from the gene list 2 annotated to the KEGG pathway if ther was no enrichment (i.e.
+                  "Ratio of number of genes observed from the gene list 2 annotated to the KEGG pathway to the number of genes expected from the gene list 2 annotated to the KEGG pathway if there was no enrichment (i.e.
 KEGG_PATHWAY_in_list2/expected_list2)",

### 2019-09-16 README.Rmd
Spell check README and vignette

@@
 and Genomes (KEGG) REST API. This package provides a number of functions that are 
-intended to be used in a pipeline (See Figure 1). Briefly, the user provides a KEGG formated species id for either human, mouse or rat, and the package will
+intended to be used in a pipeline (See Figure 1). Briefly, the user provides a KEGG formatted species id for either human, mouse or rat, and the package will
 download and clean species specific ENTREZ gene IDs and map them to their respective
-KEGG pathways by accessing KEGG's REST API. KEGG's API is used ot guarantee the most up-to-date pathway data from KEGG. Next, the user will identify significantly
+KEGG pathways by accessing KEGG's REST API. KEGG's API is used to guarantee the most up-to-date pathway data from KEGG. Next, the user will identify significantly
 enriched pathways from two different gene sets, and finally, the user will identify 
@@
 Notice that in this example the pathways in gene list 2 have smaller fold changes (shorter bars) than those in list 1, but that many of them are more significant (darker blue).
-This is because there are more genes in gene list 2 compred to gene list 1.
+This is because there are more genes in gene list 2 compared to gene list 1.
 

### 2019-09-18 README.Rmd
Spelling edits and typos in figure legends for README and vignette

@@
 Functions within the diffEnrich package are represented by blue rectangles.
-The data that must be provided by the user is represented by the pruple ovals.
-Data objects generated by a function in diffEnrich are representaed by red ovals.
+The data that must be provided by the user is represented by the purple ovals.
+Data objects generated by a function in diffEnrich are represented by red ovals.
 The external call of the get_kegg function to the KEGG REST API is represented in yellow.")

### 2019-09-23 R/plotFoldEnrichment.R
formatted digit in figure legend

@@ -116,8 +116,8 @@ plotFoldEnrichment <- function(de_res, pval, N){
     dplyr::filter(!duplicated(.data$value))
 
   ## Format list p-values
-  options(scipen = 0, digits = 3)
-  lpval <-formatC(as.numeric(summary(bardat$pvals))[c(1,2,3,5,6)])
+  options(scipen = 0, digits = 1)
+  lpval <-formatC(as.numeric(summary(bardat$pvals))[c(1,2,3,5,6)], digits = 1)
   #lpval <- ifelse(lpval < 0.000001, "< 0.000001", lpval)
 
   ## Generate finale plot

### 2019-09-23 README.Rmd
Changed get_kegg description languahe in README and reknit

@@
 intended to be used in a pipeline (See Figure 1). Briefly, the user provides a KEGG formatted species id for either human, mouse or rat, and the package will
-download and clean species specific ENTREZ gene IDs and map them to their respective
+download and store species specific ENTREZ gene IDs and map them to their respective
 KEGG pathways by accessing KEGG's REST API. KEGG's API is used to guarantee the most up-to-date pathway data from KEGG. Next, the user will identify significantly

### 2019-09-23 R/get_kegg.R
Added details section to get_kegg

@@ -1,5 +1,5 @@
 #' @title get_kegg
-#' @description This function calls an internal helper function that connects to the KEGG API, downloads, and cleans
+#' @description This function calls an internal helper function that connects to the KEGG API, downloads, and stores
 #' ncbi gene ID data, KEGG pathway descriptions, and species specific data.
 #' Currently, this function supports Human, Mouse, and Rat. Files will be
 #' written to the working directory unless otherwise specified by the user.
@@ -7,6 +7,20 @@
 #' @param species character. The species to use in kegg data pull
 #' @param path character. A character string describing the path to write out KEGG
 #' API data sets. If not provided, defaults to current working directory.
+#'
+#' @details the \code{get_kegg} function is used to connect to the KEGG REST API
+#' and download the data sets required to perform downstream analysis.
+#' Currently, this function supports three species, and recognizes the KEGG code
+#' for Homo sapiens (‘hsa’), Mus musculus (‘mmu’), and Rattus norvegicus (‘rno’).
+#' For a given species, three data sets are generated: 1) Because the user must
+#' provide their own gene lists in downstream analysis using ENTREZ gene IDs,
+#' the data set maps NCBI/ENTREZ gene IDs to KEGG gene IDs, 2) a data set that
+#' maps KEGG gene IDs to their respective KEGG pathway IDs, and 3) a data set that
+#' maps KEGG pathway IDs to their respective pathway descriptions.  This function
+#' allows the user save versioned (based on KEGG release) and time-stamped text
+#' files of the three data sets described above. In addition to these flat files,
+#' \code{get_kegg()} will also create a named list with the three relevant KEGG
+#' data sets. The names of this list will describe the data set.
 #'
 #' @return kegg_out: A named list of the data pulled from kegg api when the
 #' function was run. This may be different if the function is run at

### 2019-09-23 README.Rmd
changed 'clean' to 'store' in RERADME and vignette

@@
 
-### Step 1: Collect and clean pathways from KEGG API
+### Step 1: Collect and store pathways from KEGG API
 First we will use the *get_kegg* function to access the KEGG REST API and download the data sets required to perform 

### 2019-09-23 R/get_kegg.R
Added table 1 to get_kegg roxygen header

@@ -22,6 +22,15 @@
 #' \code{get_kegg()} will also create a named list with the three relevant KEGG
 #' data sets. The names of this list will describe the data set.
 #'
+#' \strong{Table 1.} Description of \code{get_kegg} list object
+#' \tabular{ll}{
+#'    \strong{get_kegg_list_object} \tab \strong{Object_description}                              \cr
+#'    ncbi_to_kegg         \tab ncbi gene ID <-- mapped to --> KEGG gene ID                       \cr
+#'    kegg_to_pathway      \tab KEGG gene ID <-- mapped to --> KEGG pathway ID                    \cr
+#'    pathway_to_species   \tab KEGG pathway ID <-- mapped to --> KEGG pathway species description
+#' }
+#'
+#'
 #' @return kegg_out: A named list of the data pulled from kegg api when the
 #' function was run. This may be different if the function is run at
 #' different times. For reproducible results, use text files generated

### 2019-10-02 R/get_kegg.R
removed str_extract from imports

@@ -230,7 +230,6 @@ get_kegg <- function(species, read = FALSE, path = NULL, date, release){
 #'
 #' @importFrom here here
 #' @import utils
-#' @importFrom stringr str_extract
 #'
 
 .data_read <- function(path = path, date = date, release = release){

### 2019-10-02 R/pathEnrich.R
Added method option to pathEnrich. Clean and rebuild.

@@ -10,6 +10,9 @@
 #' generated by \code{get_kegg}, they must first load that data and put it in a named
 #' list that matches the names given in the list generated by \code{get_kegg}.
 #' @param gene_list Vector. Vector of NCBI (ENTREZ) geneIDs.
+#' @param method Character. Character string telling \code{diffEnrich} which method to
+#' use for multiple testing correction. Available methods are thos provided by
+#' \code{\link{p.adjust}}, and the default is "BH", or False Discovery Rate (FDR).
 #'
 #' @return enrich_table: An object of class data.frame that summarizes the results
 #' of the pathway analysis and contains the following variables:
@@ -45,7 +48,7 @@
 #' list2_pe <- pathEnrich(gk_obj = kegg, gene_list = geneLists$list2)
 #' }
 #'
-pathEnrich <- function(gk_obj, gene_list){
+pathEnrich <- function(gk_obj, gene_list, method = 'BH'){
   ## argument check
   if(missing(gk_obj)){stop("Argument missing: gk_obj")}
   if(missing(gene_list)){stop("Argument missing: gene_list. Please provide list of ncbi geneIDs")}
@@ -99,7 +102,7 @@ pathEnrich <- function(gk_obj, gene_list){
 
   ## clean and return
   enrich_table <- enrich_table[order(enrich_table$enrich_p), ]
-  enrich_table$fdr <- stats::p.adjust(enrich_table$enrich_p, method = 'BH')
+  enrich_table$fdr <- stats::p.adjust(enrich_table$enrich_p, method = method)
   enrich_table$V1 <- gsub("path:", "", enrich_table$V1, fixed = TRUE)
 
   ## Calulate fold enrichment column

### 2019-10-02 R/diffEnrich.R
Changed fdr column to more general padj and updated .rd files for pathEnrich and diffEnrich

@@ -28,13 +28,13 @@
 #'   \item{KEGG_DATABASE_in_list1}{Number of Genes from gene list 1 in KEGG Database}
 #'   \item{expected_list1}{Expected number of genes from list 1 to be in KEGG pathway by chance (i.e., not enriched)}
 #'   \item{enrich_p_list1}{P-value for enrichment of list 1 genes related to KEGG pathway}
-#'   \item{fdr_list1}{False Discovery Rate (Benjamini and Hochberg) of enrich_p_list1 to account for multiple testing across KEGG pathways}
+#'   \item{p_adj_list1}{Multiple testing adjustment of enrich_p_list1 (default = False Discovery Rate (Benjamini and Hochberg))}
 #'   \item{fold_enrichment_list1}{KEGG_PATHWAY_in_list1/expected_list1}
 #'   \item{KEGG_PATHWAY_in_list2}{Number of Genes from gene list 2 in KEGG Pathway}
 #'   \item{KEGG_DATABASE_in_list2}{Number of Genes from gene list 2 in KEGG Database}
 #'   \item{expected_list2}{Expected number of genes from list 2 to be in KEGG pathway by chance (i.e., not enriched)}
 #'   \item{enrich_p_list2}{P-value for enrichment of list 2 genes related to KEGG pathway}
-#'   \item{fdr_list2}{False Discovery Rate (Benjamini and Hochberg) of enrich_p_list2 to account for multiple testing across KEGG pathways}
+#'   \item{p_adj_list2}{Multiple testing adjustment of enrich_p_list2 (default = False Discovery Rate (Benjamini and Hochberg))}
 #'   \item{fold_enrichment_list2}{KEGG_PATHWAY_in_list2/expected_list2}
 #'   \item{odd_ratio}{Odds of a gene from list 2 being from this KEGG pathway / Odds of a gene from list 1 being from this KEGG pathway}
 #'   \item{diff_enrich_p}{P-value for differential enrichment of this KEGG pathway between list 1 and list 2}
@@ -73,8 +73,8 @@ diffEnrich <- function(list1_pe, list2_pe, method = 'BH'){
   res$adjusted_p <- stats::p.adjust(res$pv, method = method)
   colnames(res) <- c("KEGG_PATHWAY_ID", "KEGG_PATHWAY_description", "KEGG_PATHWAY_cnt", "KEGG_DATABASE_cnt",
[…]

### 2019-10-02 R/pathEnrich.R

@@ -26,7 +26,7 @@
 #'   \item{KEGG_DATABASE_in_list}{Number of Genes from gene list in KEGG Database}
 #'   \item{expected}{Expected number of genes from list to be in KEGG pathway by chance (i.e., not enriched)}
 #'   \item{enrich_p}{P-value for enrichment of list genes related to KEGG pathway}
-#'   \item{fdr}{False Discovery Rate (Benjamini and Hochberg) to account for multiple testing across KEGG pathways}
+#'   \item{p_adj}{False Discovery Rate (Benjamini and Hochberg) to account for multiple testing across KEGG pathways}
 #'   \item{fold_enrichment}{KEGG_PATHWAY_in_list/expected}
 #' }
 #'
@@ -110,6 +110,6 @@ pathEnrich <- function(gk_obj, gene_list, method = 'BH'){
   ## Add menaingful column names for the C1 and C2
   colnames(enrich_table) <- c("KEGG_PATHWAY_ID", "KEGG_PATHWAY_description", "KEGG_PATHWAY_cnt", "KEGG_PATHWAY_in_list",
                                       "KEGG_DATABASE_cnt", "KEG_DATABASE_in_list", "expected", "enrich_p",
-                                      "fdr", "fold_enrichment")
+                                      "p_adj", "fold_enrichment")
   return(enrich_table)
 }

### 2019-10-02 README.Rmd
Added method option to pathEnrich. Clean and rebuild.
updated documentation in README and vignette

@@
 Each *pathEnrich* call generates a data frame summarizing the results of an enrichment analysis in which a Fisher's Exact test is used to identify which KEGG pathways are enriched within the user's list of interesting genes compared to all genes annotated to a KEGG pathway.
-P-values from the Fisher's Exact test are adjusted for multiple comparisons by controlling the False Discovery Rate (FDR) at 0.05.
+Users can choose from those supported by *stats::p.adjust* for multiple correction options, and the default is the False Discovery Rate (Benjamini and Hochberg, [1995](http://www.jstor.org/stable/2346101)).
 
@@
                  "P-value for enrichment within the KEGG pathway for list genes",
-                 "False Discovery Rate (Benjamini and Hochberg) to account for multiple testing across KEGG pathways",
-                 "Ratio of number of genes observed from the gene list annotated to the KEGG pathway to the number of genes expected from the gene list annotated to the KEGG pathway if there was no enrichment (i.e.
+                 "Multiple testing adjustment of enrich_p (default = False Discovery Rate (Benjamini and Hochberg))",
+                 "Ratio of number of genes observed from the gene list annotated to the KEGG pathway to the number og genes expected from the gene list annotated to the KEGG pathway if there was no enrichment (i.e.
 KEGG_PATHWAY_in_list/expected)")
@@
                   "P-value for enrichment of list 1 genes related to KEGG pathway",
-                  "False Discovery Rate (Benjamini and Hochberg) of enrich_p_list1 to account for multiple testing across KEGG pathways",
+                  "Multiple testing adjustment of enrich_p_list1 (default = False Discovery Rate (Benjamini and Hochberg))",
                   "Ratio of number of genes observed from the gene list 1 annotated to the KEGG pathway to the number of genes expected from the gene list 1 annotated to the KEGG pathway if there was no enrichment (i.e.
@@
                   "P-value for enrichment of list 2 genes related to KEGG pathway",
-                  "False Discovery Rate (Benjamini and Hochberg) of enrich_p_list2 to account for multiple testing across KEGG pathways",
+                  "Multiple testing adjustment of enrich_p_list2 (default = False Discovery Rate (Benjamini and Hochberg))",
                   "Ratio of number of genes observed from the gene list 2 annotated to the KEGG pathway to the number of genes expected from the gene list 2 annotated to the KEGG pathway if there was no enrichment (i.e.

### 2019-10-03 README.Rmd
Updated motivating example section of README and vignette

@@
 
-A recent [study](https://onlinelibrary.wiley.com/doi/full/10.1111/acer.13766) explored 
-the hepatic mechanism for removal of toxic lipid aldehydes via conjugation with glutathione (Cheng et al., [2001](https://www.ncbi.nlm.nih.gov/pubmed/11488593?dopt=Abstract);
-Gallagher et al., [2007](https://www.ncbi.nlm.nih.gov/pubmed/17553661?dopt=Abstract)) 
-and the primary enzyme that catalyzes this conjugation, glutathione S‐transferase A4‐4 (GSTA4).
-Specifically, the researchers examine the role of the GSTA4 gene on protein 
-carbonylation and the progression of liver injury in a model consisting of long‐term 
-(116 days) chronic ethanol (EtOH) consumption followed by a single EtOH binge.
+Often high throughput omics studies include a functional enrichment analysis to glean biological insight from a list of candidate genes, proteins, metabolites, etc.
+Functional enrichment examines whether the number of genes in the list associated with a biological function or particular pathway is more than would be expected by chance.
+As an example, enrichment of a particular pathway among a list of genes that are differentially expressed after an experimental manipulation may indicate that the pathway has been altered by that manipulation.
+This analysis is rather straight forward and many solutions have been offered (e.g., [Haung et al., 2009](https://www.ncbi.nlm.nih.gov/pubmed/19033363); [Kuleshov et al., 2016](https://www.ncbi.nlm.nih.gov/pubmed/27141961); [Liao et al., 2019](https://www.ncbi.nlm.nih.gov/pubmed/31114916); [Subramanian et al., 2005](https://www.pnas.org/content/102/43/15545)) as well as a wide variety of databases used to define these pathways (e.g., [Kanehisa and Goto, 2000](https://www.ncbi.nlm.nih.gov/pubmed/10592173)) and ontologies (e.g., [Ashburn et al., 2000](https://www.ncbi.nlm.nih.gov/pubmed/10802651)).
 
-Functional enrichment of carbonylated proteins for KEGG pathways was tested 
-using a 1‐sided Fisher's exact test in each experimental group 
-(2 genotypes × 3 treatment conditions = 6 experimental groups). Differential 
-enrichment between experimental groups was examined using a 2‐sided Fisher's exact test.
-For KEGG pathway enrichment, proteins were mapped to pathways using their UniProt ID, 
-and databases were downloaded from the KEGG API.
+One key component of a statistically rigorous functional enrichment analysis is the definition of a background data set that can be used to estimate the number of candidate genes that are ‘expected’ to be associated with the pathway by chance, e.g., if 5% of genes in the background data set are associated with a pathway then 5% of candidate gene are expected to be associated with the pathway by chance.
+For many study designs, the background data set is relatively simple to define (e.g., RNA-Seq analyses where the background data set includes genes expressed above background).
 
-Note: This example uses proteins as the feature of interest, however *diffEnrich*
-currently only supports genes as the feature of interest.
+However, for some newer omics technologies, the background data set is hard to define.
+For example, LC-MS analysis can be used to identify carbonylated proteins ( [Peterson et al., 2018](https://www.ncbi.nlm.nih.gov/pubmed/29799837); [Shearn et al., 2019](https://www.ncbi.nlm.nih.gov/pubmed/31377417); [Shearn et al., 2018](https://www.ncbi.nlm.nih.gov/pmc/articles/PMC6028296/)).
+With this study design, carbonylated proteins are isolated using a BH-derivation and then LC-MS is used to identify peptides in this isolated sample.
+The most appropriate background data set would be proteins present in that tissue, but this would require a separate analytical analysis.
+Furthermore, most functional enrichment analyses involve a single gene list.
+However, in protein modification studies, the typical experimental design compares the presence or absence of particular modified proteins between treatment groups.
 
-Differential enrichment analysis of carbonylated proteins indicated that compared 
-to chow‐fed animals, the most overrepresented group of proteins that was 
-significantly adducted after long‐term EtOH exposure was ribosomal proteins.
+When there are two or more gene lists to compare and the background gene list is not clearly defined, as is often the case in protein modification experiments, we propose a differential enrichment analysis.
+In this analysis, we compare the proportion of genes/proteins from one gene list associated with a particular pathway to the proportion of genes/proteins from a second gene list that are associated with that pathway.
+To easily execute this analysis, we have designed an R package that uses the KEGG REST API to obtain the most recent version of the KEGG PATHWAY ( [Kanehisa and Goto, 2000](https://www.ncbi.nlm.nih.gov/pubmed/10592173)) database to initially identify functional enrichment within a gene list using the entire KEGG transcriptome as the background data set and then to identify differentially enriched pathways between two gene lists.
+This R package includes a function to generate a ‘differential enrichment’ graphic.
 
@@
 Each *pathEnrich* call generates a data frame summarizing the results of an enrichment analysis in which a Fisher's Exact test is used to identify which KEGG pathways are enriched within the user's list of interesting genes compared to all genes annotated to a KEGG pathway.
-Users can choose from those supported by *stats::p.adjust* for multiple correction options, and the default is the False Discovery Rate (Benjamini and Hochberg, [1995](http://www.jstor.org/stable/2346101)).
+Users can choose from those supported by *stats::p.adjust* for multiple correction options, and the default is the False Discovery Rate ( [Benjamini and Hochberg, 1995](http://www.jstor.org/stable/2346101)).
 
@@
 for multiple testing adjustment.
-Users can choose from those supported by *stats::p.adjust* for multiple correction options, and the default is the False Discovery Rate (Benjamini and Hochberg, [1995](http://www.jstor.org/stable/2346101)).
+Users can choose from those supported by *stats::p.adjust* for multiple correction options, and the default is the False Discovery Rate ( [Benjamini and Hochberg, 1995](http://www.jstor.org/stable/2346101)).
 
@@
 ```
+
+# References
+
+Ashburner et al. (2000) Gene ontology: tool for the unification of biology. Nat Genet. 25(1):25-29.
+
+Benjamini, Y., and Hochberg, Y. (1995). Controlling the false discovery rate: a practical and powerful approach to multiple testing. Journal of the Royal Statistical Society Series B,.57, 289–300
+
+Huang DW. et al. (2009) Bioinformatics enrichment tools: paths toward the comprehensive functional analysis of large gene lists. Nucleic Acids Res. 37(1):1-13.
+
+Kanehisa, M. and Goto, S. (2000) KEGG: Kyoto Encyclopedia of Genes and Genomes. Nucleic Acids Res. 28, 27-30.
+
+Kuleshov MV. et al. (2016) Enrichr: a comprehensive gene set enrichment analysis web server 2016 update. Nucleic Acids Res. 44(1): 90 – 97.
+
+Liao, Y. et al. (2019) Gene set analysis toolkit with revamped UIs and APIs, Nucleic Acids Res. 47(1):199 – 205.
+
+Petersen DR. et al. (2018) Elevated Nrf-2 responses are insufficient to mitigate protein carbonylation in hepatospecific PTEN deletion mice. PLoS One. 13(5):e0198139.
+
+Shearn CT.
+et al.
+(2019) Cholestatic liver disease results increased production of reactive aldehydes and an atypical periportal hepatic antioxidant response.
+Free Radic Biol Med;143:101-114.
+[Epub ahead ofbprint] PubMed PMID: 31377417.
+
+Shearn CT. et al. (2018) Knockout of the Gsta4 Gene in Male Mice Leads to an
+Altered Pattern of Hepatic Protein Carbonylation and Enhanced Inflammation
+Following Chronic Consumption of an Ethanol Diet. Alcohol Clin Exp Res. 42(7):1192-1205.
+
+Subramanian T. et al. (2005) Gene set enrichment analysis: A knowledge-based approach for interpreting genome-wide expression profiles. PNAS 102, 15545-15550.
+
+Yu G. et al. (2012). clusterProfiler: an R package for comparing biological themes among gene clusters. OMICS. 16(5), 284-287.

### 2019-10-07 R/pathEnrich.R
added summary and print generic methods for pathEnrich which now returns an obkect of class diffEnrich

@@ -14,8 +14,14 @@
 #' use for multiple testing correction. Available methods are thos provided by
 #' \code{\link{p.adjust}}, and the default is "BH", or False Discovery Rate (FDR).
 #'
-#' @return enrich_table: An object of class data.frame that summarizes the results
-#' of the pathway analysis and contains the following variables:
+#' @return A list object of class \code{diffEnrich} that contains 4 items:
+#'
+#' \describe{
+#' \item{species}{The species used in enrichment}
+#' \item{padj}{The method used to correct for multiple testing}
+#' \item{sig_paths}{The KEGG pathways the reached statistical significance after multiple testing correction.}
+#' \item{enrich_table}{A data frame that summarizes the results of the pathway analysis and contains the following variables:}
+#' }
 #'
 #' \describe{
 #'   \item{KEGG_PATHWAY_ID}{KEGG Pathway Identifier}
@@ -33,7 +39,11 @@
 #' @details This function may not always use the complete list of genes provided by the user.
 #' Specifically, it will only use the genes from the list provided that are also in
 #' the most current species list pulled from the KEGG REST API, or from the older data KEGG
-#' loaded by the user.
+#' loaded by the user. S3 generic functions for \code{print} and \code{summary} are
+#' provided. The \code{print} function prints the results table as a \code{tibble}, and the
+#' \code{summary} function returns the number of pathways that reached statistical significance
+#' as well as their descriptions, the number of genes used from the KEGG data base, the KEGG species, and the
+#' method used for multiple testing correction.
 #'
 #' @export
 #' @importFrom  stats fisher.test
@@ -56,6 +66,8 @@ pathEnrich <- function(gk_obj, gene_list, method = 'BH'){
   ## Prepare gene list
   #gene_list <- gene_list[!is.na(gene_list)]
   gene_list <- as.integer(gene_list)
+  # define p.adjust method
+  p.adj <- method
 
   ## prepare kegg data
   ncbi_to_kegg <- gk_obj[["ncbi_to_kegg"]]
@@ -111,16 +123,49 @@ pathEnrich <- function(gk_obj, gene_list, method = 'BH'){
[…]

### 2019-10-07 R/pathEnrich.R
small changes to summary generic for pathEnrich

@@ -153,19 +153,19 @@ print.diffEnrich <- function(x){
 #' @method summary diffEnrich
 #' @export
 summary.diffEnrich <- function(x){
+  ## summary part 1
   l1 <- paste0(
     dim(x$enrich_table)[1], ' KEGG pathways were tested. \n')
   l2 <- paste0("KEGG pathway species: ", x$species, "\n")
   l3 <- paste0(
     x$enrich_table$KEGG_DATABASE_cnt[1], ' genes from gene_list were in the KEGG data pull. \n')
   l4 <- paste0("p-value adjustment method: ", x$padj, "\n")
-  l5 <- paste0(sum(list1_pe$enrich_table$p_adj < 0.05), " pathways reached statistical significance after multiple testing correction. \n")
+  l5 <- paste0(sum(x$enrich_table$p_adj < 0.05), " pathways reached statistical significance after multiple testing correction. \n")
   cat(
     l1, l2, l3, l4, l5, "\n")
-
-  cat("Significant pathways: \n")
+  ## summary part 2
   paths <- paste(x$sig_pathways, collapse = "\n")
-  cat(paths)
+  cat("Significant pathways: \n", paths)
 }
 
 

### 2019-10-07 R/diffEnrich.R
bug fixes in diffEnrich now that pathEnrich results are of class diffEnrich

@@ -112,7 +112,7 @@ diffEnrich <- function(list1_pe, list2_pe, method = 'BH'){
   if(missing(list2_pe)){stop("Argument missing: list2_pe")}
 
   ## Merge results from first enrichment
-  combined_enrich <- merge(list1_pe, list2_pe, by = c("KEGG_PATHWAY_ID", "KEGG_PATHWAY_description", "KEGG_PATHWAY_cnt", "KEGG_DATABASE_cnt"))
+  combined_enrich <- merge(list1_pe$enrich_table, list2_pe$enrich_table, by = c("KEGG_PATHWAY_ID", "KEGG_PATHWAY_description", "KEGG_PATHWAY_cnt", "KEGG_DATABASE_cnt"))
   colnames(combined_enrich) <- gsub(".x", "_list1", colnames(combined_enrich), fixed = TRUE)
   colnames(combined_enrich) <- gsub(".y", "_list2", colnames(combined_enrich), fixed = TRUE)
   colnames(combined_enrich) <- c("KEGG_PATHWAY_ID", "KEGG_PATHWAY_description", "KEGG_PATHWAY_cnt", "KEGG_DATABASE_cnt",

### 2019-10-07 R/pathEnrich.R

@@ -144,7 +144,7 @@ pathEnrich <- function(gk_obj, gene_list, method = 'BH'){
 #' @rdname pathEnrich
 #' @method print diffEnrich
 #' @export
-print.diffEnrich <- function(x){
+print.diffEnrich <- function(x, ...){
   dplyr::as_tibble(x$enrich_table)
 }
 
@@ -152,19 +152,19 @@ print.diffEnrich <- function(x){
 #' @rdname pathEnrich
 #' @method summary diffEnrich
 #' @export
-summary.diffEnrich <- function(x){
+summary.diffEnrich <- function(object, ...){
   ## summary part 1
   l1 <- paste0(
-    dim(x$enrich_table)[1], ' KEGG pathways were tested. \n')
-  l2 <- paste0("KEGG pathway species: ", x$species, "\n")
+    dim(object$enrich_table)[1], ' KEGG pathways were tested. \n')
+  l2 <- paste0("KEGG pathway species: ", object$species, "\n")
   l3 <- paste0(
-    x$enrich_table$KEGG_DATABASE_cnt[1], ' genes from gene_list were in the KEGG data pull. \n')
-  l4 <- paste0("p-value adjustment method: ", x$padj, "\n")
-  l5 <- paste0(sum(x$enrich_table$p_adj < 0.05), " pathways reached statistical significance after multiple testing correction. \n")
+    object$enrich_table$KEGG_DATABASE_cnt[1], ' genes from gene_list were in the KEGG data pull. \n')
+  l4 <- paste0("p-value adjustment method: ", object$padj, "\n")
+  l5 <- paste0(sum(object$enrich_table$p_adj < 0.05), " pathways reached statistical significance after multiple testing correction. \n")
   cat(
     l1, l2, l3, l4, l5, "\n")
   ## summary part 2
-  paths <- paste(x$sig_pathways, collapse = "\n")
+  paths <- paste(object$sig_pathways, collapse = "\n")
   cat("Significant pathways: \n", paths)
 }
 

### 2019-10-07 R/pathEnrich.R
added param descriptions to S3 generics for pathEnrich

@@ -143,16 +143,18 @@ pathEnrich <- function(gk_obj, gene_list, method = 'BH'){
 #' @name print.pathEnrich
 #' @rdname pathEnrich
 #' @method print diffEnrich
+#' @param x object of class \code{diffEnrich}
 #' @export
-print.diffEnrich <- function(x, ...){
+print.diffEnrich <- function(x){
   dplyr::as_tibble(x$enrich_table)
 }
 
 #' @name summary.pathEnrich
 #' @rdname pathEnrich
 #' @method summary diffEnrich
+#' @param object object of class \code{diffEnrich}
 #' @export
-summary.diffEnrich <- function(object, ...){
+summary.diffEnrich <- function(object){
   ## summary part 1
   l1 <- paste0(
     dim(object$enrich_table)[1], ' KEGG pathways were tested. \n')

### 2019-10-07 R/pathEnrich.R
bug fix in S3 documentation

@@ -145,7 +145,7 @@ pathEnrich <- function(gk_obj, gene_list, method = 'BH'){
 #' @method print diffEnrich
 #' @param x object of class \code{diffEnrich}
 #' @export
-print.diffEnrich <- function(x){
+print.diffEnrich <- function(x, ...){
   dplyr::as_tibble(x$enrich_table)
 }
 
@@ -154,7 +154,7 @@ print.diffEnrich <- function(x){
 #' @method summary diffEnrich
 #' @param object object of class \code{diffEnrich}
 #' @export
-summary.diffEnrich <- function(object){
+summary.diffEnrich <- function(object, ...){
   ## summary part 1
   l1 <- paste0(
     dim(object$enrich_table)[1], ' KEGG pathways were tested. \n')

### 2019-10-07 R/pathEnrich.R
added documentation for ... param in S3 generics

@@ -144,6 +144,7 @@ pathEnrich <- function(gk_obj, gene_list, method = 'BH'){
 #' @rdname pathEnrich
 #' @method print diffEnrich
 #' @param x object of class \code{diffEnrich}
+#' @param ... additional arguments to be passed to \code{print}
 #' @export
 print.diffEnrich <- function(x, ...){
   dplyr::as_tibble(x$enrich_table)
@@ -153,6 +154,7 @@ print.diffEnrich <- function(x, ...){
 #' @rdname pathEnrich
 #' @method summary diffEnrich
 #' @param object object of class \code{diffEnrich}
+#' @param ... additional arguments to be passed to \code{summary}
 #' @export
 summary.diffEnrich <- function(object, ...){
   ## summary part 1

### 2019-10-07 R/pathEnrich.R
changed \dots documentation in generics for pathEnrich

@@ -144,7 +144,7 @@ pathEnrich <- function(gk_obj, gene_list, method = 'BH'){
 #' @rdname pathEnrich
 #' @method print diffEnrich
 #' @param x object of class \code{diffEnrich}
-#' @param ... additional arguments to be passed to \code{print}
+#' @param \dots Unused
 #' @export
 print.diffEnrich <- function(x, ...){
   dplyr::as_tibble(x$enrich_table)
@@ -154,7 +154,7 @@ print.diffEnrich <- function(x, ...){
 #' @rdname pathEnrich
 #' @method summary diffEnrich
 #' @param object object of class \code{diffEnrich}
-#' @param ... additional arguments to be passed to \code{summary}
+#' @param \dots Unused
 #' @export
 summary.diffEnrich <- function(object, ...){
   ## summary part 1

### 2019-10-07 README.Rmd
pathEnrich generics bugs are fixed. updated DESCRIPTION, README, and vignette

@@
 ```{r pathEnrich_kable, warning=FALSE, echo=FALSE}
-knitr::kable(head(list1_pe),
+knitr::kable(head(list1_pe$enrich_table),
              caption = "Table 2. First 6 rows of list1_pe data frame generated using pathEnrich", row.names = FALSE) %>%
@@
 # get column names
-names <- colnames(list1_pe)
+names <- colnames(list1_pe$enrich_table)
 # Define column names

### 2019-10-07 R/pathEnrich.R

@@ -140,7 +140,7 @@ pathEnrich <- function(gk_obj, gene_list, method = 'BH'){
   return(res)
 }
 
-#' @name print.pathEnrich
+#' @name print.diffEnrich
 #' @rdname pathEnrich
 #' @method print diffEnrich
 #' @param x object of class \code{diffEnrich}
@@ -150,11 +150,10 @@ print.diffEnrich <- function(x, ...){
   dplyr::as_tibble(x$enrich_table)
 }
 
-#' @name summary.pathEnrich
+#' @name summary.diffEnrich
 #' @rdname pathEnrich
 #' @method summary diffEnrich
 #' @param object object of class \code{diffEnrich}
-#' @param \dots Unused
 #' @export
 summary.diffEnrich <- function(object, ...){
   ## summary part 1

### 2019-10-11 R/pathEnrich.R
changed pathEnrich output to clas pathEnrich

@@ -14,7 +14,7 @@
 #' use for multiple testing correction. Available methods are thos provided by
 #' \code{\link{p.adjust}}, and the default is "BH", or False Discovery Rate (FDR).
 #'
-#' @return A list object of class \code{diffEnrich} that contains 4 items:
+#' @return A list object of class \code{pathEnrich} that contains 4 items:
 #'
 #' \describe{
 #' \item{species}{The species used in enrichment}
@@ -136,26 +136,26 @@ pathEnrich <- function(gk_obj, gene_list, method = 'BH'){
     "enrich_table" = enrich_table,
     "sig_pathways" = sig_paths)
   ## define class attr
-  class(res) <- c("diffEnrich")
+  class(res) <- c("pathEnrich")
   return(res)
 }
 
-#' @name print.diffEnrich
+#' @name print.pathEnrich
 #' @rdname pathEnrich
-#' @method print diffEnrich
-#' @param x object of class \code{diffEnrich}
+#' @method print pathEnrich
+#' @param x object of class \code{pathEnrich}
 #' @param \dots Unused
 #' @export
-print.diffEnrich <- function(x, ...){
+print.pathEnrich <- function(x, ...){
   dplyr::as_tibble(x$enrich_table)
 }
 
-#' @name summary.diffEnrich
+#' @name summary.pathEnrich
 #' @rdname pathEnrich
-#' @method summary diffEnrich
-#' @param object object of class \code{diffEnrich}
+#' @method summary pathEnrich
+#' @param object object of class \code{pathEnrich}
 #' @export
-summary.diffEnrich <- function(object, ...){
+summary.pathEnrich <- function(object, ...){
   ## summary part 1
   l1 <- paste0(
     dim(object$enrich_table)[1], ' KEGG pathways were tested. \n')

### 2019-10-11 R/diffEnrich.R
added print and summary S3 generics for diffEnrich results

@@ -6,7 +6,12 @@
 #' pathways between the terms enriched in the gene-of-interest list and the pathways enriched
 #' in the background. \code{diffEnrich} returns a dataframe containing differentially enriched
 #' pathways with their associated estimated odds ratio, unadjusted p-value, and fdr adjusted
-#' p-value.
+#' p-value. S3 generic functions for \code{print} and \code{summary} are
+#' provided. The \code{print} function prints the results table as a \code{tibble}, and the
+#' \code{summary} function returns the number of pathways that reached statistical significance
+#' as well as their descriptions, the number of genes used from the KEGG data base, the KEGG species,
+#' the number of pathways that were shared (and therefore tested) between the gene lists and the
+#' method used for multiple testing correction.
 #'
 #' @param list1_pe data.frame. Dataframe of enrichment results for genes of interest
 #' generated from \code{\link{pathEnrich}}. See example for \code{\link{pathEnrich}}.
@@ -16,8 +21,15 @@
 #' use for multiple testing correction. Available methods are thos provided by
 #' \code{\link{p.adjust}}, and the default is "BH", or False Discovery Rate (FDR).
 #'
-#' @return res: An object of class data.frame that summarizes the results
-#' of the differential enrichment and contains the following variables:
+#' @return A list object of class \code{diffEnrich} that contains 5 items:
+#'
+#' \describe{
+#' \item{species}{The species used in enrichment}
+#' \item{padj}{The method used to correct for multiple testing for the differential enrichment}
+#' \item{sig_paths}{The KEGG pathways the reached statistical significance after multiple testing correction.}
+#' \item{path_intersect}{the number of pathways that were shared (and therefore tested) between the gene lists.}
+#' \item{de_table}{A data frame that summarizes the results of the differential enrichment analysis and contains the following variables:}
+#' }
 #'
 #' \describe{
 #'   \item{KEGG_PATHWAY_ID}{KEGG Pathway Identifier}
@@ -55,8 +67,13 @@
 #' dif_enrich <- diffEnrich(list1_pe = list1_pe, list2_pe = list2_pe, method = 'none')
 #'
[…]

### 2019-10-11 README.Rmd
updated plot function to handle diffEnrich class object and updated README and vignette

@@
 
-*pathEnrich* generates a data frame with 9 columns described below. Details also provided in *pathEnrich* documentation. Use *?pathEnrich*.
+*pathEnrich* generates a list object that contains a data frame with 9 columns described below.
+Details also provided in *pathEnrich* documentation as well as other metadata from the function call.
+Use *?pathEnrich*.
+S3 generic functions for *print* and *summary* are
+provided. The *print* function prints the results table as a *tibble*, and the
+*summary* function returns the number of pathways that reached statistical significance
+as well as their descriptions, the number of genes used from the KEGG data base, the KEGG species, and the
+method used for multiple testing correction.
+
+```{r}
+summary(list1_pe)
+```
 
@@
 ```{r diffEnrich_kable, warning=FALSE, echo=FALSE}
-knitr::kable(head(diff_enrich),
+knitr::kable(head(diff_enrich$de_table),
              caption = "Table 4.First 6 rows from data frame generated by diffEnrich", row.names = FALSE) %>%
@@
 ## get vector of column names
-de_col_names <- names(diff_enrich)
+de_col_names <- names(diff_enrich$de_table)
 de_col_descr <- c("KEGG Pathway Identifier",
@@
 
-The result of the *diffEnrich* call is a data frame with the estimated odds ratio generated by the Fisher's Exact test 
-and the associated p-value.
+The result of the *diffEnrich* call is a list object that contains a data frame with the estimated odds ratio generated by the Fisher's Exact test 
+and the associated p-value as well as other metadata from the function call. S3 generic functions for *print* and *summary* are
+provided. The *print* function prints the results table as a *tibble*, and the
+*summary* function returns the number of pathways that reached statistical significance
+as well as their descriptions, the number of genes used from the KEGG data base, the KEGG species,
+the number of pathways that were shared (and therefore tested) between the gene lists and the
+method used for multiple testing correction.
+
+```{r}
+summary(diff_enrich)
+```
 

### 2019-10-11 R/plotFoldEnrichment.R

@@ -47,13 +47,14 @@ plotFoldEnrichment <- function(de_res, pval, N){
   if(missing(pval)){stop("Argument missing: pval - please provide a threshold")}
   if(missing(N)){stop("Argument missing: N - if you'd like to plot top pathways, please provide a threshold and make sure N > 0")}
   if(N < 1){stop("Number of top genes (N) must be > 0")}
+  if(class(de_res) != "diffEnrich"){stop("de_res must be an object of class 'diffEnrich'. Please generate this object using the diffEnrich function provided in this package.")}
 
   ###########################################################
   # Prepare and reshape data for plotting using ggplot
   ###########################################################
 
   ## Strip extra columns from de_res and filter based on pval. Then sort by pval.
-  df <- de_res %>%
+  df <- de_res$de_table %>%
     dplyr::select(.data$KEGG_PATHWAY_ID, .data$KEGG_PATHWAY_description,
                   .data$fold_enrichment_list1, .data$fold_enrichment_list2,
                   .data$enrich_p_list1, .data$enrich_p_list2,
@@ -129,7 +130,7 @@ plotFoldEnrichment <- function(de_res, pval, N){
                         limits = c(min(ld$pvals), 0),
                         breaks = as.numeric(summary(ld$pvals))[c(1,2,3,5,6)],
                         labels = lpval,
-                        name = paste0("P-values List 1", "\n", "gene count: ", de_res[1,6])) +
+                        name = paste0("P-values List 1", "\n", "gene count: ", de_res$de_table[1,6])) +
     ggnewscale::new_scale_fill() +
     geom_rect(data = ld[ld$vars == "fold_enrichment_list2", ], aes(fill = .data$pvals), color = 'black') +
     scale_fill_gradient(low =  "navy", high = "transparent",
@@ -137,7 +138,7 @@ plotFoldEnrichment <- function(de_res, pval, N){
                         limits = c(min(ld$pvals), 0),
                         breaks = as.numeric(summary(ld$pvals))[c(1,2,3,5,6)],
                         labels = lpval,
-                        name = paste0("P-values List 2", "\n", "gene count: ", de_res[1,12])) +
+                        name = paste0("P-values List 2", "\n", "gene count: ", de_res$de_table[1,12])) +
     scale_x_continuous(breaks = seq_along(unique(ld$descr)),
                        labels = unique(ld$descr)) +
     geom_hline(yintercept=1.0, linetype ='dashed') +

### 2019-10-15 R/plotFoldEnrichment.R
bug fix for p-val ordering in plotFold Enrichment

@@ -119,7 +119,6 @@ plotFoldEnrichment <- function(de_res, pval, N){
   ## Format list p-values
   options(scipen = 0, digits = 1)
   lpval <-formatC(as.numeric(summary(bardat$pvals))[c(1,2,3,5,6)], digits = 1)
-  #lpval <- ifelse(lpval < 0.000001, "< 0.000001", lpval)
 
   ## Generate finale plot
   p <- ggplot(mapping = aes(xmin = .data$xmin, xmax = .data$xmax, ymin = .data$ymin, ymax = .data$ymax)) +
@@ -144,6 +143,6 @@ plotFoldEnrichment <- function(de_res, pval, N){
     geom_hline(yintercept=1.0, linetype ='dashed') +
     coord_flip() + theme_bw() +
     geom_text(data=df_ptext,
-              aes(x = 1:N, y = (max(bardat$value) + 0.5), label = round(.data$value, 5)))
+              aes(x = 1:N, y = (max(bardat$value) + 0.5), label = sort(round(.data$value, 5), decreasing = TRUE)))
   return(p)
 }

### 2019-10-15 README.Rmd
Started updates to README and vignette per LS comments

@@
 
-The goal of diffEnrich is to compare functional enrichment between two experimentally-derived groups of genes or proteins. Given a list of gene symbols, *diffEnrich*  will
+The goal of diffEnrich is to compare functional enrichment between two experimentally-derived groups of genes or proteins. Given a list of NCBI gene symbols, *diffEnrich*  will
 perform differential enrichment analysis using the Kyoto Encyclopedia of Genes
@@
 download and store species specific ENTREZ gene IDs and map them to their respective
-KEGG pathways by accessing KEGG's REST API. KEGG's API is used to guarantee the most up-to-date pathway data from KEGG. Next, the user will identify significantly
-enriched pathways from two different gene sets, and finally, the user will identify 
+KEGG pathways by accessing the KEGG REST API. The KEGG API is used to guarantee the most up-to-date pathway data from KEGG. Next, the user will identify significantly
+enriched pathways in two different gene sets, and finally, the user will identify 
 pathways that are differentially enriched between the two gene sets. In addition to 
@@
 of a biological system, such as a cell, an organism and an ecosystem, 
-from genomic and molecular-level information [REF](https://www.kegg.jp/kegg/kegg1a.html). KEGG is an integrated database 
+from genomic and molecular-level information [https://www.kegg.jp/kegg/kegg1a.html](https://www.kegg.jp/kegg/kegg1a.html). KEGG is an integrated database 
 resource consisting of eighteen databases that are clustered into 4 main categories:
@@
 (e.g. genes and proteins), 3) chemical information (e.g. biochemical reactions), 
-and 4) health information (e.g. human disease and drugs) [REF](https://www.kegg.jp/kegg/kegg1a.html).
+and 4) health information (e.g. human disease and drugs) [https://www.kegg.jp/kegg/kegg1a.html](https://www.kegg.jp/kegg/kegg1a.html).
 
@@
 KEGG API is very easy using statistical programming tools like R or Python and
-integrating data pulls into user's code makes the program reproducible. To 
+integrating data retrevial into user's code makes the program reproducible. To 
 further enforce reproducibilty *diffEnrich* adds a date and KEGG release tag to
@@
 As an example, enrichment of a particular pathway among a list of genes that are differentially expressed after an experimental manipulation may indicate that the pathway has been altered by that manipulation.
-This analysis is rather straight forward and many solutions have been offered (e.g., [Haung et al., 2009](https://www.ncbi.nlm.nih.gov/pubmed/19033363); [Kuleshov et al., 2016](https://www.ncbi.nlm.nih.gov/pubmed/27141961); [Liao et al., 2019](https://www.ncbi.nlm.nih.gov/pubmed/31114916); [Subramanian et al., 2005](https://www.pnas.org/content/102/43/15545)) as well as a wide variety of databases used to define these pathways (e.g., [Kanehisa and Goto, 2000](https://www.ncbi.nlm.nih.gov/pubmed/10592173)) and ontologies (e.g., [Ashburn et al., 2000](https://www.ncbi.nlm.nih.gov/pubmed/10802651)).
+This analysis is rather straight forward and many solutions have been offered (e.g., [Haung et al., 2009](https://www.ncbi.nlm.nih.gov/pubmed/19033363); [Kuleshov et al., 2016](https://www.ncbi.nlm.nih.gov/pubmed/27141961); [Liao et al., 2019](https://www.ncbi.nlm.nih.gov/pubmed/31114916); [Subramanian et al., 2005](https://www.pnas.org/content/102/43/15545)) as well a wide variety of databases have been used to define these pathways (e.g., [Kanehisa and Goto, 2000](https://www.ncbi.nlm.nih.gov/pubmed/10592173)) and ontologies (e.g., [Ashburn et al., 2000](https://www.ncbi.nlm.nih.gov/pubmed/10802651)).
 
@@
 Furthermore, most functional enrichment analyses involve a single gene list.
-However, in protein modification studies, the typical experimental design compares the presence or absence of particular modified proteins between treatment groups.
+However, in protein modification studies, the typical experimental design compares the presence or absence of particular modified proteins between multiple groups.
 
@@
 our downstream analysis. This function takes two arguments. The first, 'species' is required. Currently, *diffEnrich* supports three species, and the argument is a character string using the KEGG
-code [REF](https://www.pnas.org/content/suppl/2008/09/11/0806162105.DCSupplemental/ST1_PDF.pdf): 
+code [https://www.pnas.org/content/suppl/2008/09/11/0806162105.DCSupplemental/ST1_PDF.pdf](https://www.pnas.org/content/suppl/2008/09/11/0806162105.DCSupplemental/ST1_PDF.pdf): 
 Homo sapiens (human), use 'hsa'; Mus musculus (mouse), use 'mmu', and Rattus norvegicus (rat), use 'rno'. The second,
@@
 and the version of KEGG when the API was accessed. In addition to these flat files, *get_kegg* will also
-create a named list with the three relevant KEGG data sets. The names of this list will describe the data set.
+create a named list in R with the three relevant KEGG data sets. The names of this list will describe the data set.
 For a detailed description of list elements use *?get_kegg*.
@@
 User gene lists must also be character vectors and be formatted as ENTREZ gene IDs.
-The *clusterProfiler* package offers a nice function (*bitr*) that maps gene symbols and Ensembl IDs to ENTREZ gene IDs, and an example can be seen in their [vignette](https://yulab-smu.github.io/clusterProfiler-book/chapter5.html#supported-organisms).
+The *clusterProfiler* package offers a nice function (*bitr*) that maps gene symbols and Ensembl IDs to ENTREZ gene IDs, and an example can be seen in their [vignette](https://yulab-smu.github.io/clusterProfiler-book/chapter14.html#bitr).
 
@@
 previous *get_kegg* call.
-The *pathEnrich* function should be run at least twice, once for the genes of interest in list 1 and once for the genes of interest in list2.
-Each *pathEnrich* call generates a data frame summarizing the results of an enrichment analysis in which a Fisher's Exact test is used to identify which KEGG pathways are enriched within the user's list of interesting genes compared to all genes annotated to a KEGG pathway.
-Users can choose from those supported by *stats::p.adjust* for multiple correction options, and the default is the False Discovery Rate ( [Benjamini and Hochberg, 1995](http://www.jstor.org/stable/2346101)).
+The *pathEnrich* function should be run once for the genes of interest in list 1 and once for the genes of interest in list2.
+Each *pathEnrich* call generates a data frame summarizing the results of an enrichment analysis in which a Fisher's Exact test is used to identify which KEGG pathways are enriched for the user's list of  genes compared to all genes annotated to a KEGG pathway.
+Users can choose a multiple correction option from those supported by *stats::p.adjust*.
+The default is the False Discovery Rate ( [Benjamini and Hochberg, 1995](http://www.jstor.org/stable/2346101)).
 
@@
 
-*pathEnrich* generates a list object that contains a data frame with 9 columns described below.
-Details also provided in *pathEnrich* documentation as well as other metadata from the function call.
-Use *?pathEnrich*.
+*pathEnrich* generates a list object that contains a data frame with 9 columns described below as well as other metadata from the function call.
+Details also provided in *pathEnrich* documentation.
 S3 generic functions for *print* and *summary* are

### 2019-10-15 README.Rmd
Made text changes to readme and vignette. Added figure legends to readme.

@@
 ```
+
+**Figure 1.** diffEnrich Analysis pipeline.
+Functions within the diffEnrich package are represented by blue rectangles.
+The data that must be provided by the user is represented by the purple ovals.
+Data objects generated by a function in diffEnrich are represented by red ovals.
+The external call of the get_kegg function to the KEGG REST API is represented in yellow.
 
@@
                  "P-value for enrichment within the KEGG pathway for list genes",
-                 "Multiple testing adjustment of enrich_p (default = False Discovery Rate (Benjamini and Hochberg))",
-                 "Ratio of number of genes observed from the gene list annotated to the KEGG pathway to the number og genes expected from the gene list annotated to the KEGG pathway if there was no enrichment (i.e.
+                 "Multiple testing adjusted enrichment p-values (default = False Discovery Rate (Benjamini and Hochberg))",
+                 "Ratio of number of genes observed from the gene list annotated to the KEGG pathway to the number of genes expected from the gene list to be annotated to the KEGG pathway if there was no enrichment (i.e.
 KEGG_PATHWAY_in_list/expected)")
@@
 for multiple testing adjustment.
-Users can choose from those supported by *stats::p.adjust* for multiple correction options, and the default is the False Discovery Rate ( [Benjamini and Hochberg, 1995](http://www.jstor.org/stable/2346101)).
+Users can choose from those supported by *stats::p.adjust*.
+The default is the False Discovery Rate ( [Benjamini and Hochberg, 1995](http://www.jstor.org/stable/2346101)).
 
@@
                   "P-value for enrichment of list 1 genes related to KEGG pathway",
-                  "Multiple testing adjustment of enrich_p_list1 (default = False Discovery Rate (Benjamini and Hochberg))",
-                  "Ratio of number of genes observed from the gene list 1 annotated to the KEGG pathway to the number of genes expected from the gene list 1 annotated to the KEGG pathway if there was no enrichment (i.e.
+                  "Multiple testing adjusted enrichment p-values from gene list 1 (default = False Discovery Rate (Benjamini and Hochberg))",
+                  "Ratio of number of genes observed from gene list 1 annotated to the KEGG pathway to the number of genes expected from gene list 1 annotated to the KEGG pathway if there was no enrichment (i.e.
 KEGG_PATHWAY_in_list1/expected_list1)",
@@
                   "P-value for enrichment of list 2 genes related to KEGG pathway",
-                  "Multiple testing adjustment of enrich_p_list2 (default = False Discovery Rate (Benjamini and Hochberg))",
-                  "Ratio of number of genes observed from the gene list 2 annotated to the KEGG pathway to the number of genes expected from the gene list 2 annotated to the KEGG pathway if there was no enrichment (i.e.
+                  "Multiple testing adjusted enrichment p-values from gene list 2 (default = False Discovery Rate (Benjamini and Hochberg))",
+                  "Ratio of number of genes observed from gene list 2 annotated to the KEGG pathway to the number of genes expected from gene list 2 annotated to the KEGG pathway if there was no enrichment (i.e.
 KEGG_PATHWAY_in_list2/expected_list2)",
@@
                   "P-value for differential enrichment of this KEGG pathway between list 1 and list 2",
-                  "Multiple testing adjustment of diff_enrich_p (default = False Discovery Rate (Benjamini and Hochberg))")
+                  "Multiple testing adjusted differential enrichment p-values (default = False Discovery Rate (Benjamini and Hochberg))")
 # Make dataframe
@@
 knitr::kable(de_descr_df,
-             caption = "Table 5. Description of columns is dataframe generated by diffEnrich", row.names = FALSE) %>%
+             caption = "Table 5. Description of columns in dataframe generated by diffEnrich", row.names = FALSE) %>%
   kableExtra::kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = F)
@@
 *summary* function returns the number of pathways that reached statistical significance
-as well as their descriptions, the number of genes used from the KEGG data base, the KEGG species,
-the number of pathways that were shared (and therefore tested) between the gene lists and the
+as well as their descriptions, the number of genes used from the KEGG database, the KEGG species,
+the number of pathways that were shared (i.e. had at least 1 gene from each gene list present in the pathway) between the gene lists and the
 method used for multiple testing correction.
@@
 
+**Figure 2.** Fold enrichment stratified by gene list. KEGG pathways are plotted on the y-axis and fold
+enrichment is plotted on the x-axis. Each KEGG pathway has a bar depicting
+its fold enrichment in list 1 (red) and its fold enrichment in list 2 (blue).
+The transparency of the bars correspond to the unadjusted p-value for the
+pathway's enrichment in the given list. The p-value presented as text to the
+right of each pair of bars is the adjusted p-value (user defined: default is FDR) associated with the
+differential enrichment of the pathway between the two lists, and the pathways
+are ordered from top to bottom by this p-value (i.e. smallest p-value on top
+of plot, and largest p-value on bottom of plot). The dotted line represents a fold enrichment of 1. Finally, the number of genes used
+for analysis from each gene list (recall that this number may not be the same as the number of
+genes in the user's original list) are reported below their respective p-values
+in the legend.
+
 # References

### 2019-10-15 README.Rmd
Added 'cuttoff' and 'N' arguments to pathEnrich and updated documentation .rd, README, and vignette

@@
 previous *get_kegg* call.
+Users can also decide which KEGG pathways should be tested based on how many genes from their gene list are contained in the KEGG pathway.
+Users can set this parameter by changing the 'N' argument.
+The default is N = 2.
 The *pathEnrich* function should be run once for the genes of interest in list 1 and once for the genes of interest in list2.
@@
 Users can choose a multiple correction option from those supported by *stats::p.adjust*.
-The default is the False Discovery Rate ( [Benjamini and Hochberg, 1995](http://www.jstor.org/stable/2346101)).
+The default is the False Discovery Rate ( [Benjamini and Hochberg, 1995](http://www.jstor.org/stable/2346101)), and the default threshold to reach significance is 0.05.
 
@@
 ## List 1
-list1_pe <- pathEnrich(gk_obj = kegg, gene_list = geneLists$list1)
+list1_pe <- pathEnrich(gk_obj = kegg, gene_list = geneLists$list1, cutoff = 0.05, N = 2)
 ## list2
-list2_pe <- pathEnrich(gk_obj = kegg, gene_list = geneLists$list2) 
+list2_pe <- pathEnrich(gk_obj = kegg, gene_list = geneLists$list2, cutoff = 0.05, N = 2) 
 ```

### 2019-10-15 R/pathEnrich.R

@@ -10,16 +10,20 @@
 #' generated by \code{get_kegg}, they must first load that data and put it in a named
 #' list that matches the names given in the list generated by \code{get_kegg}.
 #' @param gene_list Vector. Vector of NCBI (ENTREZ) geneIDs.
+#' @param N Numeric. The number of genes from the gene list that must be present in a KEGG pathway in order for that pathway to be retained and tested.
 #' @param method Character. Character string telling \code{diffEnrich} which method to
-#' use for multiple testing correction. Available methods are thos provided by
+#' use for multiple testing correction. Available methods are those provided by
 #' \code{\link{p.adjust}}, and the default is "BH", or False Discovery Rate (FDR).
+#' @param cutoff Numeric. The p-value threshold to be used as the cutoff when determining statistical significance, and used to filter final results data set.
 #'
-#' @return A list object of class \code{pathEnrich} that contains 4 items:
+#' @return A list object of class \code{pathEnrich} that contains 6 items:
 #'
 #' \describe{
 #' \item{species}{The species used in enrichment}
 #' \item{padj}{The method used to correct for multiple testing}
 #' \item{sig_paths}{The KEGG pathways the reached statistical significance after multiple testing correction.}
+#' \item{cutoff}{The p-value threshold to be used as the cutoff when determining statistical significance, and used to filter final results data set.}
+#' \item{N}{The number of genes from the gene list that must be present in a KEGG pathway in order for that pathway to be retained and tested.}
 #' \item{enrich_table}{A data frame that summarizes the results of the pathway analysis and contains the following variables:}
 #' }
 #'
@@ -41,9 +45,9 @@
 #' the most current species list pulled from the KEGG REST API, or from the older data KEGG
 #' loaded by the user. S3 generic functions for \code{print} and \code{summary} are
 #' provided. The \code{print} function prints the results table as a \code{tibble}, and the
-#' \code{summary} function returns the number of pathways that reached statistical significance
[…]

### 2019-10-17 R/diffEnrich.R
Added cuttoff argument to diffEnrich; updated README and vignette

@@ -20,6 +20,7 @@
 #' @param method character. Character string telling \code{diffEnrich} which method to
 #' use for multiple testing correction. Available methods are thos provided by
 #' \code{\link{p.adjust}}, and the default is "BH", or False Discovery Rate (FDR).
+#' @param cutoff Numeric. The p-value threshold to be used as the cutoff when determining statistical significance, and used to filter list of significant pathways.
 #'
 #' @return A list object of class \code{diffEnrich} that contains 5 items:
 #'
@@ -64,9 +65,9 @@
 #' list2_pe <- pathEnrich(gk_obj = kegg, gene_list = geneLists$list2)
 #'
 #' ## Perform differential enrichment
-#' dif_enrich <- diffEnrich(list1_pe = list1_pe, list2_pe = list2_pe, method = 'none')
+#' dif_enrich <- diffEnrich(list1_pe = list1_pe, list2_pe = list2_pe, method = 'none', cutoff = 0.05)
 #'
-diffEnrich <- function(list1_pe, list2_pe, method = 'BH'){
+diffEnrich <- function(list1_pe, list2_pe, method = 'BH', cutoff = 0.05){
   ## check class
   if(class(list1_pe) != "pathEnrich"){stop("list1_pe must be an object of class 'pathEnrich'. Please generate this object using the pathEnrich function provided in this package.")}
   if(class(list2_pe) != "pathEnrich"){stop("list2_pe must be an object of class 'pathEnrich'. Please generate this object using the pathEnrich function provided in this package.")}
@@ -109,13 +110,14 @@ diffEnrich <- function(list1_pe, list2_pe, method = 'BH'){
 
   ## Extract sicgnificant pathways
   sig_paths <- de_table %>%
-    dplyr::filter(.data$diff_enrich_adjusted < 0.05) %>%
+    dplyr::filter(.data$diff_enrich_adjusted < cutoff) %>%
     pull(.data$KEGG_PATHWAY_description)
   sig_paths <- unlist(lapply(strsplit(sig_paths, split = " - ", fixed = TRUE), function(x) x[1]))
 
   ## build results list
   out <- list("species" = species,
               "padj" = p.adj,
+              "cutoff" = cutoff,
               "path_intersect" = path_intersect,
               "de_table" = de_table,
               "sig_pathways" = sig_paths)
@@ -186,7 +188,7 @@ summary.diffEnrich <- function(object, ...){
   l3 <- paste0(
[…]

### 2019-10-17 R/pathEnrich.R

@@ -14,7 +14,7 @@
 #' @param method Character. Character string telling \code{diffEnrich} which method to
 #' use for multiple testing correction. Available methods are those provided by
 #' \code{\link{p.adjust}}, and the default is "BH", or False Discovery Rate (FDR).
-#' @param cutoff Numeric. The p-value threshold to be used as the cutoff when determining statistical significance, and used to filter final results data set.
+#' @param cutoff Numeric. The p-value threshold to be used as the cutoff when determining statistical significance, and used to filter list of significant pathways.
 #'
 #' @return A list object of class \code{pathEnrich} that contains 6 items:
 #'
@@ -43,7 +43,8 @@
 #' @details This function may not always use the complete list of genes provided by the user.
 #' Specifically, it will only use the genes from the list provided that are also in
 #' the most current species list pulled from the KEGG REST API, or from the older data KEGG
-#' loaded by the user. S3 generic functions for \code{print} and \code{summary} are
+#' loaded by the user. The 'cutoff' only filters the list of pathways provided in the 'sig_paths'
+#' list item. It is not used to filter the 'enrich_table' list object. S3 generic functions for \code{print} and \code{summary} are
 #' provided. The \code{print} function prints the results table as a \code{tibble}, and the
 #' \code{summary} function returns the number of pathways that reached statistical significance,
 #' as well as their descriptions, the number of genes used from the KEGG data base, the KEGG species, and the
@@ -66,7 +67,7 @@ pathEnrich <- function(gk_obj, gene_list, method = 'BH', cutoff = 0.05, N = 2){
   ## argument check
   if(missing(gk_obj)){stop("Argument missing: gk_obj")}
   if(missing(gene_list)){stop("Argument missing: gene_list. Please provide list of ncbi geneIDs")}
-  if(N < 0){stop("N must be a positive number of type float or integer.")}
+  if(N < 0 | is.character(N)){stop("N must be a positive number of type float or integer.")}
 
   ## Prepare gene list
   #gene_list <- gene_list[!is.na(gene_list)]

### 2019-10-17 R/plotFoldEnrichment.R
Added logic response and warning for if N > sig_paths for plotFoldEnrichment.

@@ -64,6 +64,8 @@ plotFoldEnrichment <- function(de_res, pval, N){
     dplyr::filter(.data$diff_enrich_adjusted < pval) %>%
     dplyr::slice(1:N)
 
+  if(dim(df)[1] < N){warning(paste0("'N' is larger than the number of significant pathways based on 'pval'. plotFoldEnrichment will only print the significant pathways. There were ", dim(df)[1]))}
+
   ## Melt data set
   df.melt <-reshape2::melt(df, id.vars = c('KEGG_PATHWAY_ID', 'KEGG_PATHWAY_description'))
 
@@ -120,6 +122,9 @@ plotFoldEnrichment <- function(de_res, pval, N){
   options(scipen = 0, digits = 1)
   lpval <-formatC(as.numeric(summary(bardat$pvals))[c(1,2,3,5,6)], digits = 1)
 
+  ## decide how to print p-val text
+  pt <- ifelse((dim(df)[1] >= N), N, dim(df)[1])
+
   ## Generate finale plot
   p <- ggplot(mapping = aes(xmin = .data$xmin, xmax = .data$xmax, ymin = .data$ymin, ymax = .data$ymax)) +
     geom_rect(data = ld[ld$vars == "fold_enrichment_list1", ], aes(fill = .data$pvals), color = 'black') +
@@ -143,6 +148,6 @@ plotFoldEnrichment <- function(de_res, pval, N){
     geom_hline(yintercept=1.0, linetype ='dashed') +
     coord_flip() + theme_bw() +
     geom_text(data=df_ptext,
-              aes(x = 1:N, y = (max(bardat$value) + 0.5), label = sort(round(.data$value, 5), decreasing = TRUE)))
+              aes(x = 1:pt, y = (max(bardat$value) + 0.5), label = sort(round(.data$value, 5), decreasing = TRUE)))
   return(p)
 }

### 2019-10-17 README.Rmd
Added cuttoff argument to diffEnrich; updated README and vignette
Added arg checks to .data_read, and added get_kegg read = TRUE example to README and vignette

@@
 
-    kegg_to_pathway2019-04-26Release_90.0+_04-26_Apr_19.txt
-    ncbi_to_kegg2019-04-26Release_90.0+_04-26_Apr_19.txt
-    pathway_to_species2019-04-26Release_90.0+_04-26_Apr_19.txt
+    ncbi_to_kegg2019-10-17Release_92.0+_10-17_Oct_19.txt
+    kegg_to_pathway2019-10-17Release_92.0+_10-17_Oct_19.txt
+    pathway_to_species2019-10-17Release_92.0+_10-17_Oct_19.txt
 
@@
 it is not writing out new data sets and still generate the named list object. Users can generate 'same-day/same-version'
-data sets in different directories if they so choose.
+data sets in different directories if they so choose. Additionaly, *get_kegg* can be used to read in saved versions of the txt files
+generated from a previous call (like those displayed above), and generate an R list object that is compatible with downstream
+functions.
 
@@
 ```
+
+```{r get_kegg_read, eval=FALSE}
+## run get_kegg() using rat
+kegg_rno <- get_kegg(read = TRUE, path = "/path/to/files", date = "2019-10-17", release = "92")
+```
+
+Here is an example of the output:
+
+    Reading in the following files:
+    ncbi_to_kegg2019-10-17Release_92.0+_10-17_Oct_19.txt
+    kegg_to_pathway2019-10-17Release_92.0+_10-17_Oct_19.txt
+    pathway_to_species2019-10-17Release_92.0+_10-17_Oct_19.txt
+    File location: ~/Desktop
 
@@
 ## Perform differential enrichment
-diff_enrich <- diffEnrich(list1_pe = list1_pe, list2_pe = list2_pe, method = 'none') 
+diff_enrich <- diffEnrich(list1_pe = list1_pe, list2_pe = list2_pe, method = 'none', cutoff = 0.05) 
 ```

### 2019-10-17 R/get_kegg.R

@@ -235,6 +235,9 @@ get_kegg <- function(species, read = FALSE, path = NULL, date, release){
 .data_read <- function(path = path, date = date, release = release){
   options(stringsAsFactors = F)
   if(is.null(path)){stop("Please provide a file path pointing to files from previous get_kegg() call.")}
+  if(!is.character(date)){stop("'date' must be provided as a character type.")}
+  if(!is.character(release)){stop("'release' must be provided as a character type.")}
+
   # Define user's base file path
   flist <- list.files(path)
   # define regex to match on for files

### 2019-10-21 README.Rmd
Spell check README and vignette

@@
 KEGG API is very easy using statistical programming tools like R or Python and
-integrating data retrevial into user's code makes the program reproducible. To 
+integrating data retrieval into user's code makes the program reproducible. To 
 further enforce reproducibilty *diffEnrich* adds a date and KEGG release tag to
@@
 it is not writing out new data sets and still generate the named list object. Users can generate 'same-day/same-version'
-data sets in different directories if they so choose. Additionaly, *get_kegg* can be used to read in saved versions of the txt files
+data sets in different directories if they so choose. Additionally, *get_kegg* can be used to read in saved versions of the txt files
 generated from a previous call (like those displayed above), and generate an R list object that is compatible with downstream
@@
 Free Radic Biol Med;143:101-114.
-[Epub ahead ofbprint] PubMed PMID: 31377417.
+[Epub ahead of print] PubMed PMID: 31377417.
 

### 2019-10-23 README.Rmd
pathways with 0 genes in both lists are now filtered and a warning is generated.

@@
 The default is the False Discovery Rate ( [Benjamini and Hochberg, 1995](http://www.jstor.org/stable/2346101)).
+KEGG pathways that do not contain any genes from either gene list (i.e., *list1_pe$enrich_table$KEGG_PATHWAY_in_list* for 'rno04530' = 0 AND *list2_pe$enrich_table$KEGG_PATHWAY_in_list* for 'rno04530' = 0) will be removed as these cannot be tested.
+If this is the case a warning will be printed that tells the user how many pathways were removed.
+This can be avoided by setting the 'N' parameter to a value > 0 in the *pathEnrich* calls.
 

### 2019-10-23 R/diffEnrich.R

@@ -1,10 +1,13 @@
 #' diffEnrich
 #' @description This function takes the objects generated from \code{\link{pathEnrich}}.
 #' If performing a dfferential enrichment analysis, the user will have 2 objects. There
-#' will be one for the genes of interest and one for the background (see example for \code{\link{pathEnrich}}).
-#' This function then uses a Fisher's Exact test to iddentify differentially enriched
-#' pathways between the terms enriched in the gene-of-interest list and the pathways enriched
-#' in the background. \code{diffEnrich} returns a dataframe containing differentially enriched
+#' will be one for the genes of interest in gene list 1 and one for the genes of interest in gene list 2 (see example for \code{\link{pathEnrich}}).
+#' This function then uses a Fisher's Exact test to identify differentially enriched
+#' pathways between the terms enriched in the gene-of-interest lists. \code{diffEnrich}
+#' will remove KEGG pathways that do not contain any genes from either gene list as these
+#' cannot be tested, and will print a warning message telling the user how many pathways
+#' were removed.
+#' \code{diffEnrich} returns a dataframe containing differentially enriched
 #' pathways with their associated estimated odds ratio, unadjusted p-value, and fdr adjusted
 #' p-value. S3 generic functions for \code{print} and \code{summary} are
 #' provided. The \code{print} function prints the results table as a \code{tibble}, and the
@@ -13,10 +16,10 @@
 #' the number of pathways that were shared (and therefore tested) between the gene lists and the
 #' method used for multiple testing correction.
 #'
-#' @param list1_pe data.frame. Dataframe of enrichment results for genes of interest
-#' generated from \code{\link{pathEnrich}}. See example for \code{\link{pathEnrich}}.
-#' @param list2_pe data.frame. Dataframe of enrichment results for background genes
-#' generated from \code{\link{pathEnrich}}. See example for \code{\link{pathEnrich}}.
+#' @param list1_pe object of class \code{pathEnrich} generated from \code{\link{pathEnrich}}.
+#' See example for \code{\link{pathEnrich}}.
+#' @param list2_pe object of class \code{pathEnrich} generated from \code{\link{pathEnrich}}.
+#' See example for \code{\link{pathEnrich}}.
[…]

### 2019-10-31 README.Rmd
Made minor text changes to README

@@
 adding features and functionality ever since. There are benefits to using an API.
-First, API's like KEGG's allow users to perform customized analyses with the most
+First, API's, like KEGG's, allow users to perform customized analyses with the most
 up-to-date versions of the data contained in the database. In addition, accessing the
@@
 
-**Figure 1.** diffEnrich Analysis pipeline.
-Functions within the diffEnrich package are represented by blue rectangles.
+**Figure 1.
+diffEnrich Analysis pipeline.** Functions within the diffEnrich package are represented by blue rectangles.
 The data that must be provided by the user is represented by the purple ovals.
@@
 As an example, enrichment of a particular pathway among a list of genes that are differentially expressed after an experimental manipulation may indicate that the pathway has been altered by that manipulation.
-This analysis is rather straight forward and many solutions have been offered (e.g., [Haung et al., 2009](https://www.ncbi.nlm.nih.gov/pubmed/19033363); [Kuleshov et al., 2016](https://www.ncbi.nlm.nih.gov/pubmed/27141961); [Liao et al., 2019](https://www.ncbi.nlm.nih.gov/pubmed/31114916); [Subramanian et al., 2005](https://www.pnas.org/content/102/43/15545)) as well a wide variety of databases have been used to define these pathways (e.g., [Kanehisa and Goto, 2000](https://www.ncbi.nlm.nih.gov/pubmed/10592173)) and ontologies (e.g., [Ashburn et al., 2000](https://www.ncbi.nlm.nih.gov/pubmed/10802651)).
+This analysis is rather straight forward and many solutions have been offered (e.g., [Haung et al., 2009](https://www.ncbi.nlm.nih.gov/pubmed/19033363); [Kuleshov et al., 2016](https://www.ncbi.nlm.nih.gov/pubmed/27141961); [Liao et al., 2019](https://www.ncbi.nlm.nih.gov/pubmed/31114916); [Subramanian et al., 2005](https://www.pnas.org/content/102/43/15545)).
+A wide variety of databases have also been used to define these pathways (e.g., [Kanehisa and Goto, 2000](https://www.ncbi.nlm.nih.gov/pubmed/10592173)) and ontologies (e.g., [Ashburn et al., 2000](https://www.ncbi.nlm.nih.gov/pubmed/10802651)).
 
@@
 code [https://www.pnas.org/content/suppl/2008/09/11/0806162105.DCSupplemental/ST1_PDF.pdf](https://www.pnas.org/content/suppl/2008/09/11/0806162105.DCSupplemental/ST1_PDF.pdf): 
-Homo sapiens (human), use 'hsa'; Mus musculus (mouse), use 'mmu', and Rattus norvegicus (rat), use 'rno'. The second,
+Homo sapiens (human), use 'hsa'; Mus musculus (mouse), use 'mmu'; and Rattus norvegicus (rat), use 'rno'. The second,
 'path' is also passed as a character string, and is the path of the directory in which the user would like to
@@
 over-represented) based on a gene list of interest.
-User gene lists must also be character vectors and be formatted as ENTREZ gene IDs.
+User gene lists must be character vectors and be formatted as ENTREZ gene IDs.
 The *clusterProfiler* package offers a nice function (*bitr*) that maps gene symbols and Ensembl IDs to ENTREZ gene IDs, and an example can be seen in their [vignette](https://yulab-smu.github.io/clusterProfiler-book/chapter14.html#bitr).
@@
                  "P-value for enrichment within the KEGG pathway for list genes",
-                 "Multiple testing adjusted enrichment p-values (default = False Discovery Rate (Benjamini and Hochberg))",
+                 "Multiple testing adjusted enrichment p-values (default = False Discovery Rate (Benjamini and Hochberg, 1995))",
                  "Ratio of number of genes observed from the gene list annotated to the KEGG pathway to the number of genes expected from the gene list to be annotated to the KEGG pathway if there was no enrichment (i.e.
@@
                   "P-value for enrichment of list 1 genes related to KEGG pathway",
-                  "Multiple testing adjusted enrichment p-values from gene list 1 (default = False Discovery Rate (Benjamini and Hochberg))",
+                  "Multiple testing adjusted enrichment p-values from gene list 1 (default = False Discovery Rate (Benjamini and Hochberg, 1995))",
                   "Ratio of number of genes observed from gene list 1 annotated to the KEGG pathway to the number of genes expected from gene list 1 annotated to the KEGG pathway if there was no enrichment (i.e.
@@
                   "P-value for enrichment of list 2 genes related to KEGG pathway",
-                  "Multiple testing adjusted enrichment p-values from gene list 2 (default = False Discovery Rate (Benjamini and Hochberg))",
+                  "Multiple testing adjusted enrichment p-values from gene list 2 (default = False Discovery Rate (Benjamini and Hochberg, 1995))",
                   "Ratio of number of genes observed from gene list 2 annotated to the KEGG pathway to the number of genes expected from gene list 2 annotated to the KEGG pathway if there was no enrichment (i.e.
@@
                   "P-value for differential enrichment of this KEGG pathway between list 1 and list 2",
-                  "Multiple testing adjusted differential enrichment p-values (default = False Discovery Rate (Benjamini and Hochberg))")
+                  "Multiple testing adjusted differential enrichment p-values (default = False Discovery Rate (Benjamini and Hochberg, 1995))")
 # Make dataframe
@@
 as well as their descriptions, the number of genes used from the KEGG database, the KEGG species,
-the number of pathways that were shared (i.e. had at least 1 gene from each gene list present in the pathway) between the gene lists and the
+the number of pathways that were shared (i.e. had at least N gene from each gene list present in the pathway based on what the user chose for N in *pathEnrich*) between the gene lists and the
 method used for multiple testing correction.
@@
 ```{r echo=FALSE}
-cap_plot <- c("Figure 2. Fold enrichment stratified by gene list. KEGG pathways are plotted on the y-axis and fold
+cap_plot <- c("Figure 2. Example of a differential enrichment graphic. KEGG pathways are plotted on the y-axis and fold
 enrichment is plotted on the x-axis. Each KEGG pathway has a bar depicting
@@
 
-**Figure 2.** Fold enrichment stratified by gene list. KEGG pathways are plotted on the y-axis and fold
+**Figure 2.  Example of a differential enrichment graphic.** KEGG pathways are plotted on the y-axis and fold
 enrichment is plotted on the x-axis. Each KEGG pathway has a bar depicting

### 2019-10-31 R/diffEnrich.R
fixed global variable note for .combineEnrich

@@ -158,7 +158,7 @@ diffEnrich <- function(list1_pe, list2_pe, method = 'BH', cutoff = 0.05){
   combined_enrich.tmp <- merge(list1_pe$enrich_table, list2_pe$enrich_table, by = c("KEGG_PATHWAY_ID", "KEGG_PATHWAY_description", "KEGG_PATHWAY_cnt", "KEGG_DATABASE_cnt"))
   ## remove pathways that have 0 genes in both lists
   combined_enrich <- combined_enrich.tmp %>%
-    dplyr::filter(!(KEGG_PATHWAY_in_list.x == 0 & KEGG_PATHWAY_in_list.y == 0))
+    dplyr::filter(!(.data$KEGG_PATHWAY_in_list.x == 0 & .data$KEGG_PATHWAY_in_list.y == 0))
 
   ## get number of pathways that were removed
   paths_removed <- dim(combined_enrich.tmp)[1] - dim(combined_enrich)[1]

### 2019-11-15 jss-article-rnw-diffEnrich/article.Rnw
Added jss template materials

\documentclass[article]{jss}

%% -- LaTeX packages and custom commands ---------------------------------------

%% recommended packages
\usepackage{thumbpdf,lmodern}

%% another package (only for this demo article)
\usepackage{framed}

%% new custom commands
\newcommand{\class}[1]{`\code{#1}'}
\newcommand{\fct}[1]{\code{#1()}}

%% For Sweave-based articles about R packages:
%% need no \usepackage{Sweave}
%% \SweaveOpts{engine=R, eps=FALSE, keep.source = TRUE}
<<preliminaries, echo=FALSE, results='hide'>>=
options(prompt = "R> ", continue = "+  ", width = 70, useFancyQuotes = FALSE)
library("MASS")
@

%% -- Article metainformation (author, title, ...) -----------------------------

%% - \author{} with primary affiliation
%% - Separate authors by \And or \AND (in \author).
%% - \AND starts a new line, \And does not.
\author{Harry A. Smith\\Department of Biostatistics and Informatics
\\Colorado School of Public Health\\Skaggs School of Pharmacy\\and
Pharmaceutical Sciences
   \And Laura Saba\\Skaggs School of Pharmacy\\and
Pharmaceutical Sciences}
%% \Plainauthor{Achim Zeileis, Second Author}

%% - \title{} in title case
%% - \Plaintitle{} without LaTeX markup (if any)
%% - \Shorttitle{} with LaTeX markup (if any), used as running title
\title{diffEnrich: An \proglang{R} Package to Compare Functional Enrichment Between Two 
Experimentally-derived Groups of Genes by Connecting to the KEGG REST API}
\Plaintitle{diffEnrich: An R Package to Compare Functional Enrichment Between Two 
Experimentally-derived Groups of Genes by Connecting to the KEGG REST API}
\Shorttitle{The \proglang{R} Package diffEnrich}

%% - \Abstract{} almost as usual
\Abstract{
  \textbf{Motivation:} To aid in the biological interpretation of a list of candidate 
  genes and proteins generated as part of omics studies, researchers quantitate
  the enrichment of known pathways or biological functions among the genes of
  interest. With the advent of new technologies and new experimental designs, 
  it is often of interest to compare enrichment of a particular pathway between
  two gene lists (i.e., differential enrichment).
  \textbf{Results:} This package provides a number of functions that are intended to be 
  used in a pipeline. Briefly, a function within the package will map 
  species-specific ENTREZ gene IDs to their respective Kyoto Encyclopedia of 
  Genes and Genomes (KEGG) pathways by accessing the KEGG REST API. The KEGG API 
  is used to guarantee the most up-to-date pathway data from KEGG. Next, another 
  function will identify significantly enriched pathways in two gene sets 
  independently. The user can then identify pathways that are differentially 
  enriched between the two gene sets using a third function. This package also 
  provides a plotting function.
  \textbf{Availability and implementation:} diffEnrich is freely available on the 
  Comprehensive R Archive Network (CRAN). Issues and bug reports can be submitted 
  to the GitHub page \url{https://github.com/SabaLab/diffEnrich/issues}. 
  \textbf{Supplementary information:} A step-by-step tutorial is provided on the diffEnrich
  GitHub page \url{https://github.com/SabaLab/diffEnrich}, and example data are 
  included in the package.

}

%% - \Keywords{} with LaTeX markup, at least one required
%% - \Plainkeywords{} without LaTeX markup (if necessary)
%% - Should be comma-separated and in sentence case.
\Keywords{JSS, style guide, comma-separated, not capitalized, \proglang{R}}
\Plainkeywords{JSS, style guide, comma-separated, not capitalized, R}

%% - \Address{} of at least one author
%% - May contain multiple affiliations for each author
%%   (in extra lines, separated by \emph{and}\\).
%% - May contain multiple authors for the same affiliation
%%   (in the same first line, separated by comma).
\Address{
  Achim Zeileis\\
  Journal of Statistical Software\\
  \emph{and}\\
  Department of Statistics\\
  Faculty of Economics and Statistics\\
  Universit\"at Innsbruck\\
  Universit\"atsstr.~15\\
  6020 Innsbruck, Austria\\
  E-mail: \email{Achim.Zeileis@R-project.org}\\
  URL: \url{https://eeecon.uibk.ac.at/~zeileis/}
}

\begin{document}

%% -- Introduction -------------------------------------------------------------

%% - In principle "as usual".
%% - But should typically have some discussion of both _software_ and _methods_.
%% - Use \proglang{}, \pkg{}, and \code{} markup throughout the manuscript.
%% - If such markup is in (sub)section titles, a plain text version has to be
%%   added as well.
%% - All software mentioned should be properly \cite-d.
%% - All abbreviations should be introduced.
%% - Unless the expansions of abbreviations are proper names (like "Journal
%%   of Statistical Software" above) they should be in sentence case (like
%%   "generalized linear models" below).

\section[Introduction: Count data regression in R]{Introduction: Count data regression in \proglang{R}} \label{sec:intro}

\begin{leftbar}
The introduction is in principle ``as usual''. However, it should usually embed
both the implemented \emph{methods} and the \emph{software} into the respective
relevant literature. For the latter both competing and complementary software
should be discussed (within the same software environment and beyond), bringing
out relative (dis)advantages. All software mentioned should be properly
\verb|\cite{}|d. (See also Appendix~\ref{app:bibtex} for more details on
\textsc{Bib}{\TeX}.)

For writing about software JSS requires authors to use the markup
\verb|\proglang{}| (programming languages and large programmable systems),
\verb|\pkg{}| (software packages), \verb|\code{}| (functions, commands,
arguments, etc.). If there is such markup in (sub)section titles (as above), a
plain text version has to be provided in the {\LaTeX} command as well. Below we
also illustrate how abbrevations should be introduced and citation commands can
be employed. See the {\LaTeX} code for more details.
\end{leftbar}

Modeling count variables is a common task in economics and the social sciences.
The classical Poisson regression model for count data is often of limited use in
these disciplines because empirical count data sets typically exhibit
overdispersion and/or an excess number of zeros. The former issue can be
addressed by extending  the plain Poisson regression model in various
directions: e.g., using sandwich covariances or estimating an additional
dispersion parameter (in a so-called quasi-Poisson model). Another more formal
way is to use a negative binomial (NB) regression. All of these models belong to
the family of generalized linear models (GLMs). However, although these models
typically can capture overdispersion rather well, they are in many applications
not sufficient for  modeling excess zeros. Since \cite{Mullahy:1986} there is
increased interest in zero-augmented models that address this issue by a second
model component capturing zero counts. An overview of count data models in
econometrics, including  hurdle and zero-inflated models, is provided in
\cite{Cameron+Trivedi:2013}.

In \proglang{R} \citep{R}, GLMs are provided by the model fitting functions
\fct{glm} in the \pkg{stats} package and \fct{glm.nb} in the \pkg{MASS} package
\citep[][Chapter~7.4]{Venables+Ripley:2002} along with associated methods for
diagnostics and inference. The manuscript that this document is based on
\citep{Zeileis+Kleiber+Jackman:2008} then introduced hurdle and zero-inflated
count models in the functions \fct{hurdle} and \fct{zeroinfl} in the \pkg{pscl}
package \citep{Jackman:2015}. Of course, much more software could be discussed
here, including (but not limited to) generalized additive models for count data
as available in the \proglang{R} packages \pkg{mgcv} \cite{Wood:2006}, 
\pkg{gamlss} \citep{Stasinopoulos+Rigby:2007}, or \pkg{VGAM} \citep{Yee:2009}.

%% -- Manuscript ---------------------------------------------------------------

%% - In principle "as usual" again.
%% - When using equations (e.g., {equation}, {eqnarray}, {align}, etc.
%%   avoid empty lines before and after the equation (which would signal a new
%%   paragraph.
%% - When describing longer chunks of code that are _not_ meant for execution
%%   (e.g., a function synopsis or list of arguments), the environment {Code}
%%   is recommended. Alternatively, a plain {verbatim} can also be used.
%%   (For executed code see the next section.)

\section{Models and software} \label{sec:models}

The basic Poisson regression model for count data is a special case of the GLM
framework \cite{McCullagh+Nelder:1989}. It describes the dependence of a count
response variable $y_i$ ($i = 1, \dots, n$) by assuming a Poisson distribution
$y_i \sim \mathrm{Pois}(\mu_i)$. The dependence of the conditional mean
$\E[y_i \, | \, x_i] = \mu_i$ on the regressors $x_i$ is then specified via a
log link and a linear predictor
%
\begin{equation} \label{eq:mean}
\log(\mu_i) \quad = \quad x_i^\top \beta,
\end{equation}
%
where the regression coefficients $\beta$ are estimated by maximum likelihood
(ML) using the iterative weighted least squares (IWLS) algorithm.

\begin{leftbar}
Note that around the \verb|{equation}| above there should be no spaces (avoided
in the {\LaTeX} code by \verb|%| lines) so that ``normal'' spacing is used and
not a new paragraph started.
\end{leftbar}

\proglang{R} provides a very flexible implementation of the general GLM
framework in the function \fct{glm} \citep{Chambers+Hastie:1992} in the
\pkg{stats} package. Its most important arguments are
\begin{Code}
glm(formula, data, subset, na.action, weights, offset,
  family = gaussian, start = NULL, control = glm.control(...),
  model = TRUE, y = TRUE, x = FALSE, ...)
\end{Code}
where \code{formula} plus \code{data} is the now standard way of specifying
regression relationships in \proglang{R}/\proglang{S} introduced in
\cite{Chambers+Hastie:1992}. The remaining arguments in the first line
(\code{subset}, \code{na.action}, \code{weights}, and \code{offset}) are also
standard  for setting up formula-based regression models in
\proglang{R}/\proglang{S}. The arguments in the second line control aspects
specific to GLMs while the arguments in the last line specify which components
are returned in the fitted model object (of class \class{glm} which inherits
from \class{lm}). For further arguments to \fct{glm} (including alternative
specifications of starting values) see \code{?glm}. For estimating a Poisson
model \code{family = poisson} has to be specified.

\begin{leftbar}
As the synopsis above is a code listing that is not meant to be executed,
one can use either the dedicated \verb|{Code}| environment or a simple
\verb|{verbatim}| environment for this. Again, spaces before and after should be
avoided.

Finally, there might be a reference to a \verb|{table}| such as
Table~\ref{tab:overview}. Usually, these are placed at the top of the page
(\verb|[t!]|), centered (\verb|\centering|), with a caption below the table,
column headers and captions in sentence style, and if possible avoiding vertical
lines.
\end{leftbar}

\begin{table}[t!]
\centering
\begin{tabular}{lllp{7.4cm}}
\hline
Type           & Distribution & Method   & Description \\ \hline
GLM            & Poisson      & ML       & Poisson regression: classical GLM,
                                           estimated by maximum likelihood (ML) \\
               &              & Quasi    & ``Quasi-Poisson regression'':
                                           same mean function, estimated by
                                           quasi-ML (QML) or equivalently
                                           generalized estimating equations (GEE),
                                           inference adjustment via estimated
                                           dispersion parameter \\
               &              & Adjusted & ``Adjusted Poisson regression'':
                                           same mean function, estimated by
                                           QML/GEE, inference adjustment via
                                           sandwich covariances\\
               & NB           & ML       & NB regression: extended GLM,
                                           estimated by ML including additional
                                           shape parameter \\ \hline
Zero-augmented & Poisson      & ML       & Zero-inflated Poisson (ZIP),
                                           hurdle Poisson \\
               & NB           & ML       & Zero-inflated NB (ZINB),
                                           hurdle NB \\ \hline
\end{tabular}
\caption{\label{tab:overview} Overview of various count regression models. The
table is usually placed at the top of the page (\texttt{[t!]}), centered
(\texttt{centering}), has a caption below the table, column headers and captions
are in sentence style, and if possible vertical lines should be avoided.}
\end{table}

%% -- Illustrations ------------------------------------------------------------

%% - Virtually all JSS manuscripts list source code along with the generated
%%   output. The style files provide dedicated environments for this.
%% - In R, the environments {Sinput} and {Soutput} - as produced by Sweave() or
%%   or knitr using the render_sweave() hook - are used (without the need to
%%   load Sweave.sty).
%% - Equivalently, {CodeInput} and {CodeOutput} can be used.
%% - The code input should use "the usual" command prompt in the respective
%%   software system.
%% - For R code, the prompt "R> " should be used with "+  " as the
%%   continuation prompt.
%% - Comments within the code chunks should be avoided - these should be made
%%   within the regular LaTeX text.

\section{Illustrations} \label{sec:illustrations}

For a simple illustration of basic Poisson and NB count regression the
\code{quine} data from the \pkg{MASS} package is used. This provides the number
of \code{Days} that children were absent from school in Australia in a
particular year, along with several covariates that can be employed as regressors.
The data can be loaded by
%
<<data>>=
data("quine", package = "MASS")
@
%
and a basic frequency distribution of the response variable is displayed in
Figure~\ref{fig:quine}.

\begin{leftbar}
For code input and output, the style files provide dedicated environments.
Either the ``agnostic'' \verb|{CodeInput}| and \verb|{CodeOutput}| can be used
or, equivalently, the environments \verb|{Sinput}| and \verb|{Soutput}| as
produced by \fct{Sweave} or \pkg{knitr} when using the \code{render_sweave()}
hook. Please make sure that all code is properly spaced, e.g., using
\code{y = a + b * x} and \emph{not} \code{y=a+b*x}. Moreover, code input should
use ``the usual'' command prompt in the respective software system. For
\proglang{R} code, the prompt \code{"R> "} should be used with \code{"+  "} as
the continuation prompt. Generally, comments within the code chunks should be
avoided -- and made in the regular {\LaTeX} text instead. Finally, empty lines
before and after code input/output should be avoided (see above).
\end{leftbar}

\begin{figure}[t!]
\centering
<<visualization, echo=FALSE, fig=TRUE, height=5.2, width=7>>=
par(mar = c(4, 4, 1, 1))
plot(table(quine$Days), xlab = "Days", ylab = "Frequency", axes = FALSE)
axis(2)
axis(1, at = 0:16 * 5, labels = FALSE)
axis(1, at = 0:8 * 10)
box()
@
\caption{\label{fig:quine} Frequency distribution for number of days absent
from school.}
\end{figure}

As a first model for the \code{quine} data, we fit the basic Poisson regression
model. (Note that JSS prefers when the second line of code is indented by two
spaces.)
%
<<poisson>>=
m_pois <- glm(Days ~ (Eth + Sex + Age + Lrn)^2, data = quine,
  family = poisson)
@
%
To account for potential overdispersion we also consider a negative binomial
GLM.
%
<<negbin>>=
library("MASS")
m_nbin <- glm.nb(Days ~ (Eth + Sex + Age + Lrn)^2, data = quine)
@
%
In a comparison with the BIC the latter model is clearly preferred.
%
<<comparison>>=
BIC(m_pois, m_nbin)
@
%
Hence, the full summary of that model is shown below.
%
<<summary>>=
summary(m_nbin)
@

%% -- Summary/conclusions/discussion -------------------------------------------

\section{Summary and discussion} \label{sec:summary}

\begin{leftbar}
As usual \dots
\end{leftbar}

%% -- Optional special unnumbered sections -------------------------------------

\section*{Computational details}

\begin{leftbar}
If necessary or useful, information about certain computational details
such as version numbers, operating systems, or compilers could be included
in an unnumbered section. Also, auxiliary packages (say, for visualizations,
maps, tables, \dots) that are not cited in the main text can be credited here.
\end{leftbar}

The results in this paper were obtained using
\proglang{R}~\Sexpr{paste(R.Version()[6:7], collapse = ".")} with the
\pkg{MASS}~\Sexpr{packageVersion("MASS")} package. \proglang{R} itself
and all packages used are available from the Comprehensive
\proglang{R} Archive Network (CRAN) at
\url{https://CRAN.R-project.org/}.

\section*{Acknowledgments}

\begin{leftbar}
All acknowledgments (note the AE spelling) should be collected in this
unnumbered section before the references. It may contain the usual information
about funding and feedback from colleagues/reviewers/etc. Furthermore,
information such as relative contributions of the authors may be added here
(if any).
\end{leftbar}

%% -- Bibliography -------------------------------------------------------------
%% - References need to be provided in a .bib BibTeX database.
%% - All references should be made with \cite, \citet, \citep, \citealp etc.
%%   (and never hard-coded). See the FAQ for details.
%% - JSS-specific markup (\proglang, \pkg, \code) should be used in the .bib.
%% - Titles in the .bib should be in title case.
%% - DOIs should be included where available.

\bibliography{refs}

%% -- Appendix (if any) --------------------------------------------------------
%% - After the bibliography with page break.
%% - With proper section titles and _not_ just "Appendix".

\newpage

\begin{appendix}

\section{More technical details} \label{app:technical}

\begin{leftbar}
Appendices can be included after the bibliography (with a page break). Each
section within the appendix should have a proper section title (rather than
just \emph{Appendix}).

For more technical style details, please check out JSS's style FAQ at
\url{https://www.jstatsoft.org/pages/view/style#frequently-asked-questions}
which includes the following topics:
\begin{itemize}
  \item Title vs.\ sentence case.
  \item Graphics formatting.
  \item Naming conventions.
  \item Turning JSS manuscripts into \proglang{R} package vignettes.
  \item Trouble shooting.
  \item Many other potentially helpful details\dots
\end{itemize}
\end{leftbar}

\section[Using BibTeX]{Using \textsc{Bib}{\TeX}} \label{app:bibtex}

\begin{leftbar}
References need to be provided in a \textsc{Bib}{\TeX} file (\code{.bib}). All
references should be made with \verb|\cite|, \verb|\citet|, \verb|\citep|,
\verb|\citealp| etc.\ (and never hard-coded). This commands yield different
formats of author-year citations and allow to include additional details (e.g.,
pages, chapters, \dots) in brackets. In case you are not familiar with these
commands see the JSS style FAQ for details.

Cleaning up \textsc{Bib}{\TeX} files is a somewhat tedious task -- especially
when acquiring the entries automatically from mixed online sources. However,
it is important that informations are complete and presented in a consistent
style to avoid confusions. JSS requires the following format.
\begin{itemize}
  \item JSS-specific markup (\verb|\proglang|, \verb|\pkg|, \verb|\code|) should
    be used in the references.
  \item Titles should be in title case.
  \item Journal titles should not be abbreviated and in title case.
  \item DOIs should be included where available.
  \item Software should be properly cited as well. For \proglang{R} packages
    \code{citation("pkgname")} typically provides a good starting point.
\end{itemize}
\end{leftbar}

\end{appendix}

%% -----------------------------------------------------------------------------

\end{document}

### 2019-11-15 jss-article-rnw-diffEnrich/article.tex
Added jss template materials

--- 2019-11-15 jss-article-rnw-diffEnrich/article.Rnw
@@
-\documentclass[article]{jss}
+\documentclass[article]{jss}\usepackage[]{graphicx}\usepackage[]{color}
+% maxwidth is the original width if it is less than linewidth
+% otherwise use linewidth (to make sure the graphics do not exceed the margin)
+\makeatletter
+\def\maxwidth{ %
+  \ifdim\Gin@nat@width>\linewidth
+    \linewidth
+  \else
+    \Gin@nat@width
+  \fi
+}
+\makeatother
+
+\definecolor{fgcolor}{rgb}{0.345, 0.345, 0.345}
+\newcommand{\hlnum}[1]{\textcolor[rgb]{0.686,0.059,0.569}{#1}}%
+\newcommand{\hlstr}[1]{\textcolor[rgb]{0.192,0.494,0.8}{#1}}%
+\newcommand{\hlcom}[1]{\textcolor[rgb]{0.678,0.584,0.686}{\textit{#1}}}%
+\newcommand{\hlopt}[1]{\textcolor[rgb]{0,0,0}{#1}}%
+\newcommand{\hlstd}[1]{\textcolor[rgb]{0.345,0.345,0.345}{#1}}%
+\newcommand{\hlkwa}[1]{\textcolor[rgb]{0.161,0.373,0.58}{\textbf{#1}}}%
+\newcommand{\hlkwb}[1]{\textcolor[rgb]{0.69,0.353,0.396}{#1}}%
+\newcommand{\hlkwc}[1]{\textcolor[rgb]{0.333,0.667,0.333}{#1}}%
+\newcommand{\hlkwd}[1]{\textcolor[rgb]{0.737,0.353,0.396}{\textbf{#1}}}%
+\let\hlipl\hlkwb
+
+\usepackage{framed}
+\makeatletter
+\newenvironment{kframe}{%
+ \def\at@end@of@kframe{}%
+ \ifinner\ifhmode%
+  \def\at@end@of@kframe{\end{minipage}}%
+  \begin{minipage}{\columnwidth}%
+ \fi\fi%
+ \def\FrameCommand##1{\hskip\@totalleftmargin \hskip-\fboxsep
+ \colorbox{shadecolor}{##1}\hskip-\fboxsep
+     % There is no \\@totalrightmargin, so:
+     \hskip-\linewidth \hskip-\@totalleftmargin \hskip\columnwidth}%
+ \MakeFramed {\advance\hsize-\width
+   \@totalleftmargin\z@ \linewidth\hsize
+   \@setminipage}}%
+ {\par\unskip\endMakeFramed%
+ \at@end@of@kframe}
+\makeatother
+
+\definecolor{shadecolor}{rgb}{.97, .97, .97}
+\definecolor{messagecolor}{rgb}{0, 0, 0}
+\definecolor{warningcolor}{rgb}{1, 0, 1}
+\definecolor{errorcolor}{rgb}{1, 0, 0}
+\newenvironment{knitrout}{}{} % an empty environment to be redefined in TeX
+
+\usepackage{alltt}
 
@@
 %% \SweaveOpts{engine=R, eps=FALSE, keep.source = TRUE}
-<<preliminaries, echo=FALSE, results='hide'>>=
-options(prompt = "R> ", continue = "+  ", width = 70, useFancyQuotes = FALSE)
-library("MASS")
-@
 
@@
 }
-
+\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
 \begin{document}
@@
 %
-<<data>>=
-data("quine", package = "MASS")
-@
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{alltt}
+\hlkwd{data}\hlstd{(}\hlstr{"quine"}\hlstd{,} \hlkwc{package} \hlstd{=} \hlstr{"MASS"}\hlstd{)}
+\end{alltt}
+\end{kframe}
+\end{knitrout}
 %
@@
 \centering
-<<visualization, echo=FALSE, fig=TRUE, height=5.2, width=7>>=
-par(mar = c(4, 4, 1, 1))
-plot(table(quine$Days), xlab = "Days", ylab = "Frequency", axes = FALSE)
-axis(2)
-axis(1, at = 0:16 * 5, labels = FALSE)
-axis(1, at = 0:8 * 10)
-box()
-@
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}
+\includegraphics[width=\maxwidth]{figure/visualization-1}
+
+\end{knitrout}
 \caption{\label{fig:quine} Frequency distribution for number of days absent
@@
 %
-<<poisson>>=
-m_pois <- glm(Days ~ (Eth + Sex + Age + Lrn)^2, data = quine,
-  family = poisson)
-@
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{alltt}
+\hlstd{m_pois} \hlkwb{<-} \hlkwd{glm}\hlstd{(Days} \hlopt{~} \hlstd{(Eth} \hlopt{+} \hlstd{Sex} \hlopt{+} \hlstd{Age} \hlopt{+} \hlstd{Lrn)}\hlopt{^}\hlnum{2}\hlstd{,} \hlkwc{data} \hlstd{= quine,}
+  \hlkwc{family} \hlstd{= poisson)}
+\end{alltt}
+\end{kframe}
+\end{knitrout}
 %
@@
 %
-<<negbin>>=
-library("MASS")
-m_nbin <- glm.nb(Days ~ (Eth + Sex + Age + Lrn)^2, data = quine)
-@
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{alltt}
+\hlkwd{library}\hlstd{(}\hlstr{"MASS"}\hlstd{)}
+\hlstd{m_nbin} \hlkwb{<-} \hlkwd{glm.nb}\hlstd{(Days} \hlopt{~} \hlstd{(Eth} \hlopt{+} \hlstd{Sex} \hlopt{+} \hlstd{Age} \hlopt{+} \hlstd{Lrn)}\hlopt{^}\hlnum{2}\hlstd{,} \hlkwc{data} \hlstd{= quine)}
+\end{alltt}
+\end{kframe}
+\end{knitrout}
 %
@@
 %
-<<comparison>>=
-BIC(m_pois, m_nbin)
-@
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{alltt}
+\hlkwd{BIC}\hlstd{(m_pois, m_nbin)}
+\end{alltt}
+\begin{verbatim}
+##        df      BIC
+## m_pois 18 2046.851
+## m_nbin 19 1157.235
+\end{verbatim}
+\end{kframe}
+\end{knitrout}
 %
@@
 %
-<<summary>>=
-summary(m_nbin)
-@
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{alltt}
+\hlkwd{summary}\hlstd{(m_nbin)}
+\end{alltt}
+\begin{verbatim}
+## 
+## Call:
+## glm.nb(formula = Days ~ (Eth + Sex + Age + Lrn)^2, data = quine, 
+##     init.theta = 1.60364105, link = log)
+## 
+## Deviance Residuals: 
+##     Min       1Q   Median       3Q      Max  
+## -3.0857  -0.8306  -0.2620   0.4282   2.0898  
+## 
+## Coefficients: (1 not defined because of singularities)
+##             Estimate Std. Error z value Pr(>|z|)    
+## (Intercept)  3.00155    0.33709   8.904  < 2e-16 ***
+## EthN        -0.24591    0.39135  -0.628  0.52977    
+## SexM        -0.77181    0.38021  -2.030  0.04236 *  
+## AgeF1       -0.02546    0.41615  -0.061  0.95121    
+## AgeF2       -0.54884    0.54393  -1.009  0.31296    
+## AgeF3       -0.25735    0.40558  -0.635  0.52574    
+## LrnSL        0.38919    0.48421   0.804  0.42153    
+## EthN:SexM    0.36240    0.29430   1.231  0.21818    
+## EthN:AgeF1  -0.70000    0.43646  -1.604  0.10876    
+## EthN:AgeF2  -1.23283    0.42962  -2.870  0.00411 ** 
+## EthN:AgeF3   0.04721    0.44883   0.105  0.91622    
+## EthN:LrnSL   0.06847    0.34040   0.201  0.84059    
+## SexM:AgeF1   0.02257    0.47360   0.048  0.96198    
+## SexM:AgeF2   1.55330    0.51325   3.026  0.00247 ** 
+## SexM:AgeF3   1.25227    0.45539   2.750  0.00596 ** 
+## SexM:LrnSL   0.07187    0.40805   0.176  0.86019    
+## AgeF1:LrnSL -0.43101    0.47948  -0.899  0.36870    
+## AgeF2:LrnSL  0.52074    0.48567   1.072  0.28363    
+## AgeF3:LrnSL       NA         NA      NA       NA    
+## ---
+## Signif. codes:  0 '***' 0.001 '**' 0.01 '*' 0.05 '.' 0.1 ' ' 1
+## 
+## (Dispersion parameter for Negative Binomial(1.6036) family taken to be 1)
+## 
+##     Null deviance: 235.23  on 145  degrees of freedom
+## Residual deviance: 167.53  on 128  degrees of freedom
+## AIC: 1100.5
+## 
+## Number of Fisher Scoring iterations: 1
+## 
+## 
+##               Theta:  1.604 
+##           Std. Err.:  0.214 
+## 
+##  2 x log-likelihood:  -1062.546
+\end{verbatim}
+\end{kframe}
+\end{knitrout}
 
@@
 The results in this paper were obtained using
-\proglang{R}~\Sexpr{paste(R.Version()[6:7], collapse = ".")} with the
-\pkg{MASS}~\Sexpr{packageVersion("MASS")} package. \proglang{R} itself
+\proglang{R}~3.6.1 with the
+\pkg{MASS}~7.3.51.4 package. \proglang{R} itself
 and all packages used are available from the Comprehensive

### 2019-11-15 jss-article-rnw-diffEnrich/article.Rnw
testing appveyor build
added keywords and address to article.Rnw

@@
 %% - Should be comma-separated and in sentence case.
-\Keywords{JSS, style guide, comma-separated, not capitalized, \proglang{R}}
-\Plainkeywords{JSS, style guide, comma-separated, not capitalized, R}
+\Keywords{differential enrichment, KEGG REST API, \proglang{R}}
+\Plainkeywords{differential enrichment, KEGG REST API, R}
 
@@
 \Address{
-  Achim Zeileis\\
-  Journal of Statistical Software\\
-  \emph{and}\\
-  Department of Statistics\\
-  Faculty of Economics and Statistics\\
-  Universit\"at Innsbruck\\
-  Universit\"atsstr.~15\\
-  6020 Innsbruck, Austria\\
-  E-mail: \email{Achim.Zeileis@R-project.org}\\
-  URL: \url{https://eeecon.uibk.ac.at/~zeileis/}
+  Laura Saba\\
+  University of Colorado\\
+  Skaggs School of Pharmacy and Pharmaceutical Sciences\\
+  Mail Stop C238\\
+  12850 E. Montview Blvd. V20-2124\\
+  Aurora, CO 80045\\
+  E-mail: \email{Laura.Saba@cuanschutz.edu}\\
 }

### 2019-11-15 jss-article-rnw-diffEnrich/article.Rnw
Added introduction text. working on bibTex.

@@
 \usepackage{thumbpdf,lmodern}
+\usepackage{textcomp}
 
@@
   Aurora, CO 80045\\
-  E-mail: \email{Laura.Saba@cuanschutz.edu}\\
+  E-mail: \email{Laura.Saba@cuanschutz.edu}
 }
@@
 
-\section[Introduction: Count data regression in R]{Introduction: Count data regression in \proglang{R}} \label{sec:intro}
+\section[Introduction]{Introduction} \label{sec:intro}
 
-\begin{leftbar}
-The introduction is in principle ``as usual''. However, it should usually embed
-both the implemented \emph{methods} and the \emph{software} into the respective
-relevant literature. For the latter both competing and complementary software
-should be discussed (within the same software environment and beyond), bringing
-out relative (dis)advantages. All software mentioned should be properly
-\verb|\cite{}|d. (See also Appendix~\ref{app:bibtex} for more details on
-\textsc{Bib}{\TeX}.)
+Often high throughput omics studies include a functional enrichment analysis to
+glean biological insight from a list of candidate genes, proteins, metabolites,
+etc. Functional enrichment examines whether the number of genes in the list
+associated with a biological function or particular pathway is more than would
+be expected by chance. As an example, enrichment of a particular pathway among
+a list of genes that are differentially expressed after an experimental
+manipulation may indicate that the pathway has been altered by that
+manipulation. This analysis is rather straight forward and many solutions have
+been offered (e.g., \cite{HuangDW:2009}; Kuleshov et al., 2016; Liao et al.,
+2019; Subramanian et al., 2005). A wide variety of databases have also been
+used to define these pathways (e.g., Kanehisa and Goto, 2000) and ontologies
+(e.g., \cite{Ashburner:2000}).
 
-For writing about software JSS requires authors to use the markup
-\verb|\proglang{}| (programming languages and large programmable systems),
-\verb|\pkg{}| (software packages), \verb|\code{}| (functions, commands,
-arguments, etc.). If there is such markup in (sub)section titles (as above), a
-plain text version has to be provided in the {\LaTeX} command as well. Below we
-also illustrate how abbrevations should be introduced and citation commands can
-be employed. See the {\LaTeX} code for more details.
-\end{leftbar}
+One key component of a statistically rigorous functional enrichment analysis is
+the definition of a background data set that can be used to estimate the number
+of candidate genes that are ``expected'' to be associated with the pathway by
+chance, e.g., if 5\% of genes in the background data set are associated with
+a pathway then 5\% of candidate gene are expected to be associated with the
+pathway by chance. For many study designs, the background data set is
+relatively simple to define (e.g., RNA-Seq analyses where the background data
+set includes genes expressed above background).
 
-Modeling count variables is a common task in economics and the social sciences.
-The classical Poisson regression model for count data is often of limited use in
-these disciplines because empirical count data sets typically exhibit
-overdispersion and/or an excess number of zeros. The former issue can be
-addressed by extending  the plain Poisson regression model in various
-directions: e.g., using sandwich covariances or estimating an additional
-dispersion parameter (in a so-called quasi-Poisson model). Another more formal
-way is to use a negative binomial (NB) regression. All of these models belong to
-the family of generalized linear models (GLMs). However, although these models
-typically can capture overdispersion rather well, they are in many applications
-not sufficient for  modeling excess zeros. Since \cite{Mullahy:1986} there is
-increased interest in zero-augmented models that address this issue by a second
-model component capturing zero counts. An overview of count data models in
-econometrics, including  hurdle and zero-inflated models, is provided in
-\cite{Cameron+Trivedi:2013}.
+However, for some newer omics technologies, the background data set is hard to
+define. For example, LC-MS analysis can be used to identify carbonylated
+proteins ( Peterson et al., 2018; Shearn et al., 2019; Shearn et al., 2018).
+With this study design, carbonylated proteins are isolated using a BH-derivation
+and then LC-MS is used to identify peptides in this isolated sample. The most
+appropriate background data set would be proteins present in that tissue, but
+this would require a separate analytical analysis. Furthermore, most functional
+enrichment analyses involve a single gene list. However, in protein modification
+studies, the typical experimental design compares the presence or absence of
+particular modified proteins between multiple groups.
 
-In \proglang{R} \citep{R}, GLMs are provided by the model fitting functions
-\fct{glm} in the \pkg{stats} package and \fct{glm.nb} in the \pkg{MASS} package
-\citep[][Chapter~7.4]{Venables+Ripley:2002} along with associated methods for
-diagnostics and inference. The manuscript that this document is based on
-\citep{Zeileis+Kleiber+Jackman:2008} then introduced hurdle and zero-inflated
-count models in the functions \fct{hurdle} and \fct{zeroinfl} in the \pkg{pscl}
-package \citep{Jackman:2015}. Of course, much more software could be discussed
-here, including (but not limited to) generalized additive models for count data
-as available in the \proglang{R} packages \pkg{mgcv} \cite{Wood:2006},
-\pkg{gamlss} \citep{Stasinopoulos+Rigby:2007}, or \pkg{VGAM} \citep{Yee:2009}.
+When there are two or more gene lists to compare and the background gene list is
+not clearly defined, as is often the case in protein modification experiments,
+we propose a differential enrichment analysis. In this analysis, we compare the
+proportion of genes/proteins from one gene list associated with a particular
+pathway to the proportion of genes/proteins from a second gene list that are
+associated with that pathway. To easily execute this analysis, we have designed
+an R package that uses the KEGG REST API to obtain the most recent version of
+the KEGG PATHWAY ( Kanehisa and Goto, 2000) database to initially identify
+functional enrichment within a gene list using the entire KEGG transcriptome as
+the background data set and then to identify differentially enriched pathways
+between two gene lists. This R package includes a function to generate a
+``differential enrichment'' graphic.
+
+KEGG is a database resource for understanding high-level functions of a
+biological system, such as a cell, an organism and an ecosystem, from genomic
+and molecular-level information \url{https://www.kegg.jp/kegg/kegg1a.html}. KEGG is
+an integrated database resource consisting of eighteen databases that are
+clustered into 4 main categories: 1) systems information (e.g. hierarchies
+and maps), 2) genomic information (e.g. genes and proteins), 3) chemical
+information (e.g. biochemical reactions), and 4) health information (e.g. human
+disease and drugs) \url{https://www.kegg.jp/kegg/kegg1a.html}.
+
+In 2012 KEGG released its first application programming interface (API), and has
+been adding features and functionality ever since. There are benefits to using
+an API. First, API\textquotesingle s, like KEGG\textquotesingle s, allow users to
+perform customized analyses with the most up-to-date versions of the data
+contained in the database. In addition, accessing the KEGG API is very easy
+using statistical programming tools like R or Python and integrating data
+retrieval into user\textquotesingle s code makes the
+program reproducible. To further enforce reproducibilty diffEnrich adds a date
+and KEGG release tag to all data files it generates from accessing the API. For
+update histories and release notes for the KEGG REST API please visit
+\url{https://www.kegg.jp/kegg/rest/}.
 

### 2019-11-15 jss-article-rnw-diffEnrich/article.Rnw
added citations for introduction

@@
 manipulation. This analysis is rather straight forward and many solutions have
-been offered (e.g., \cite{HuangDW:2009}; Kuleshov et al., 2016; Liao et al.,
-2019; Subramanian et al., 2005). A wide variety of databases have also been
-used to define these pathways (e.g., Kanehisa and Goto, 2000) and ontologies
+been offered (e.g., \cite{HuangDW:2009}; \cite{Kuleshov:2016}; \cite{Liao:2019};
+\cite{Subramanian:2005}). A wide variety of databases have also been
+used to define these pathways (e.g., \cite{KEGG:2000}) and ontologies
 (e.g., \cite{Ashburner:2000}).
@@
 define. For example, LC-MS analysis can be used to identify carbonylated
-proteins ( Peterson et al., 2018; Shearn et al., 2019; Shearn et al., 2018).
+proteins ( \cite{Petersen:2018}; \cite{Shearn:2019}; \cite{Shearn:2018}).
 With this study design, carbonylated proteins are isolated using a BH-derivation
@@
 an R package that uses the KEGG REST API to obtain the most recent version of
-the KEGG PATHWAY ( Kanehisa and Goto, 2000) database to initially identify
+the KEGG PATHWAY (\cite{KEGG:2000}) database to initially identify
 functional enrichment within a gene list using the entire KEGG transcriptome as

### 2019-11-15 jss-article-rnw-diffEnrich/article.Rnw
Added flowchart as fig 1 and caption to article.Rnw

@@
 \usepackage{textcomp}
+\usepackage[section]{placeins}
 
@@
 
+\begin{figure}[!ht]
+% \centering
+<<visualization, echo=FALSE, fig=TRUE, height=3.0, width=5>>=
+suppressMessages(library(diagram))
+
+## open plot
+par(mar = c(0.5, 0.5, 0.5, 0.5))
+openplotmat()
+## define number of boxes at each level
+elpos <- coordinates (c(3, 1, 1, 2, 2, 1, 1, 1))
+## generate matrix of from-to arrow coordinates
+fromto <- matrix(ncol = 2, byrow = TRUE,
+                 data = c(1, 6,
+                          2, 4,
+                          3, 7,
+                          4, 5,
+                          5, 6,
+                          5, 7,
+                          6, 8,
+                          7, 9,
+                          8, 10,
+                          9, 10,
+                          10, 11,
+                          11, 12))
+nr <- nrow(fromto)
+arrpos <- matrix(ncol = 2, nrow = nr)
+
+## Build flow chart
+for (i in 1:nr){
+  arrpos[i, ] <- straightarrow (to = elpos[fromto[i, 2], ],
+                                from = elpos[fromto[i, 1], ],
+                                lwd = 2, arr.pos = 0.6, arr.length = 0.5)
+  textellipse(elpos[1,], 0.10, 0.04, lab = "geneList1", box.col = "purple",
+              shadow.col = "purple4", shadow.size = 0.005, cex = 1.0)
+  textrect(elpos[2,], 0.10, 0.03, lab = "get_kegg()", box.col = "steelblue1",
+              shadow.col = "darkblue", shadow.size = 0.005, cex = 1.0)
+  textellipse(elpos[3,], 0.10, 0.04, lab = "geneList2", box.col = "purple",
+              shadow.col = "purple4", shadow.size = 0.005, cex = 1.0)
+  textround(elpos[4,], 0.10, 0.03,lab = "KEGG REST API", box.col = "orange1",
+            shadow.col = "orange4", shadow.size = 0.005, cex = 1.0)
+  textellipse(elpos[5,], 0.15, 0.04, lab = c("cleaned gene-to-pathway", "map"), box.col = "firebrick1",
+              shadow.col = "firebrick4", shadow.size = 0.005, cex = 1.0)
+  textrect(elpos[6,], 0.10, 0.03,lab = "pathEnrich(List 1)", box.col = "steelblue1",
+            shadow.col = "darkblue", shadow.size = 0.005, cex = 1.0)
+  textrect(elpos[7,], 0.10, 0.03, lab = "pathEnrich(List 2)", box.col = "steelblue1",
+           shadow.col = "darkblue", shadow.size = 0.005, cex = 1.0)
+  textellipse(elpos[8,], 0.10, 0.04, lab = c("pathEnrich","result/object"), box.col = "firebrick1",
+              shadow.col = "firebrick4", shadow.size = 0.005, cex = 1.0)
+  textellipse(elpos[9,], 0.10, 0.04, lab = c("pathEnrich","result/object"),box.col = "firebrick1",
+              shadow.col = "firebrick4", shadow.size = 0.005, cex = 1.0)
+  textrect(elpos[10,], 0.10, 0.03, lab = "diffEnrich()", box.col = "steelblue1",
+           shadow.col = "darkblue", shadow.size = 0.005, cex = 1.0)
+  textellipse(elpos[11,], 0.10, 0.04, lab = c("diffEnrich","result/object"), box.col = "firebrick1",
+              shadow.col = "firebrick4", shadow.size = 0.005, cex = 1.0)
+  textrect(elpos[12,], 0.11, 0.03, lab = "plotFoldEnrichment", box.col = "steelblue1",
+           shadow.col = "darkblue", shadow.size = 0.005, cex = 1.0)
+  # text(0.15,0.05, "Figure 1. diffEnrich Analysis pipeline", cex = 0.75)
+  text(0.05,0.87, "Step 1", cex = 1.2, font = 2)
+  text(0.05,0.57, "Step 2", cex = 1.2, font = 2)
+  text(0.05,0.35, "Step 3", cex = 1.2, font = 2)
+  text(0.05,0.10, "Step 4", cex = 1.2, font = 2)
+}
+@
+\caption{\label{fig:flowchart} \textbf{diffEnrich Analysis pipeline.} Functions within
+the diffEnrich package are represented by blue rectangles. The data that must be
+provided by the user is represented by the purple ovals. Data objects generated
+by a function in diffEnrich are represented by red ovals. The external call of
+the \code{get\textunderscore kegg} function to the KEGG REST API is represented in yellow.}
+\end{figure}
+
 In 2012 KEGG released its first application programming interface (API), and has
@@
 
-%% -- Manuscript ---------------------------------------------------------------
+The goal of the diffEnrich package is to compare functional enrichment between
+two experimentally-derived groups of genes or proteins. This package provides
+four functions that are intended to be used in an ordered pipeline
+(Figure~\ref{fig:flowchart}).
 
-%% - In principle "as usual" again.
-%% - When using equations (e.g., {equation}, {eqnarray}, {align}, etc.
-%%   avoid empty lines before and after the equation (which would signal a new
-%%   paragraph.
-%% - When describing longer chunks of code that are _not_ meant for execution
-%%   (e.g., a function synopsis or list of arguments), the environment {Code}
-%%   is recommended. Alternatively, a plain {verbatim} can also be used.
-%%   (For executed code see the next section.)
+%% ------------Section 2: Features----------------------------------------------
 
-\section{Models and software} \label{sec:models}
-
-The basic Poisson regression model for count data is a special case of the GLM
-framework \cite{McCullagh+Nelder:1989}. It describes the dependence of a count
-response variable $y_i$ ($i = 1, \dots, n$) by assuming a Poisson distribution
-$y_i \sim \mathrm{Pois}(\mu_i)$. The dependence of the conditional mean
-$\E[y_i \, | \, x_i] = \mu_i$ on the regressors $x_i$ is then specified via a
-log link and a linear predictor
-%
-\begin{equation} \label{eq:mean}
-\log(\mu_i) \quad = \quad x_i^\top \beta,
-\end{equation}
-%
-where the regression coefficients $\beta$ are estimated by maximum likelihood
-(ML) using the iterative weighted least squares (IWLS) algorithm.
-
-\begin{leftbar}
-Note that around the \verb|{equation}| above there should be no spaces (avoided
-in the {\LaTeX} code by \verb|%| lines) so that ``normal'' spacing is used and
-not a new paragraph started.
-\end{leftbar}
-
-\proglang{R} provides a very flexible implementation of the general GLM
-framework in the function \fct{glm} \citep{Chambers+Hastie:1992} in the
-\pkg{stats} package. Its most important arguments are
-\begin{Code}
-glm(formula, data, subset, na.action, weights, offset,
-  family = gaussian, start = NULL, control = glm.control(...),
-  model = TRUE, y = TRUE, x = FALSE, ...)
-\end{Code}
-where \code{formula} plus \code{data} is the now standard way of specifying
-regression relationships in \proglang{R}/\proglang{S} introduced in
-\cite{Chambers+Hastie:1992}. The remaining arguments in the first line
-(\code{subset}, \code{na.action}, \code{weights}, and \code{offset}) are also
-standard  for setting up formula-based regression models in
-\proglang{R}/\proglang{S}. The arguments in the second line control aspects
-specific to GLMs while the arguments in the last line specify which components
-are returned in the fitted model object (of class \class{glm} which inherits
-from \class{lm}). For further arguments to \fct{glm} (including alternative
-specifications of starting values) see \code{?glm}. For estimating a Poisson
-model \code{family = poisson} has to be specified.
-
-\begin{leftbar}
-As the synopsis above is a code listing that is not meant to be executed,
-one can use either the dedicated \verb|{Code}| environment or a simple
-\verb|{verbatim}| environment for this. Again, spaces before and after should be
-avoided.
-
-Finally, there might be a reference to a \verb|{table}| such as
-Table~\ref{tab:overview}. Usually, these are placed at the top of the page
-(\verb|[t!]|), centered (\verb|\centering|), with a caption below the table,
-column headers and captions in sentence style, and if possible avoiding vertical
-lines.
-\end{leftbar}
-
-\begin{table}[t!]
-\centering
-\begin{tabular}{lllp{7.4cm}}
-\hline
-Type           & Distribution & Method   & Description \\ \hline
-GLM            & Poisson      & ML       & Poisson regression: classical GLM,
-                                           estimated by maximum likelihood (ML) \\
-               &              & Quasi    & ``Quasi-Poisson regression'':
-                                           same mean function, estimated by
-                                           quasi-ML (QML) or equivalently
-                                           generalized estimating equations (GEE),
-                                           inference adjustment via estimated
-                                           dispersion parameter \\
-               &              & Adjusted & ``Adjusted Poisson regression'':
-                                           same mean function, estimated by
-                                           QML/GEE, inference adjustment via
-                                           sandwich covariances\\
-               & NB           & ML       & NB regression: extended GLM,
-                                           estimated by ML including additional
-                                           shape parameter \\ \hline
-Zero-augmented & Poisson      & ML       & Zero-inflated Poisson (ZIP),
-                                           hurdle Poisson \\
-               & NB           & ML       & Zero-inflated NB (ZINB),
-                                           hurdle NB \\ \hline
-\end{tabular}
-\caption{\label{tab:overview} Overview of various count regression models. The
-table is usually placed at the top of the page (\texttt{[t!]}), centered
-(\texttt{centering}), has a caption below the table, column headers and captions
-are in sentence style, and if possible vertical lines should be avoided.}
-\end{table}
-
-%% -- Illustrations ------------------------------------------------------------
-
-%% - Virtually all JSS manuscripts list source code along with the generated
-%%   output. The style files provide dedicated environments for this.
-%% - In R, the environments {Sinput} and {Soutput} - as produced by Sweave() or
-%%   or knitr using the render_sweave() hook - are used (without the need to
-%%   load Sweave.sty).
-%% - Equivalently, {CodeInput} and {CodeOutput} can be used.
-%% - The code input should use "the usual" command prompt in the respective
-%%   software system.
-%% - For R code, the prompt "R> " should be used with "+  " as the
-%%   continuation prompt.
-%% - Comments within the code chunks should be avoided - these should be made
-%%   within the regular LaTeX text.
-
-\section{Illustrations} \label{sec:illustrations}
-
-For a simple illustration of basic Poisson and NB count regression the
-\code{quine} data from the \pkg{MASS} package is used. This provides the number
-of \code{Days} that children were absent from school in Australia in a
-particular year, along with several covariates that can be employed as regressors.
-The data can be loaded by
-%
-<<data>>=
-data("quine", package = "MASS")
-@
-%
-and a basic frequency distribution of the response variable is displayed in
-Figure~\ref{fig:quine}.
-
-\begin{leftbar}
-For code input and output, the style files provide dedicated environments.
-Either the ``agnostic'' \verb|{CodeInput}| and \verb|{CodeOutput}| can be used
-or, equivalently, the environments \verb|{Sinput}| and \verb|{Soutput}| as
-produced by \fct{Sweave} or \pkg{knitr} when using the \code{render_sweave()}
-hook. Please make sure that all code is properly spaced, e.g., using
-\code{y = a + b * x} and \emph{not} \code{y=a+b*x}. Moreover, code input should
-use ``the usual'' command prompt in the respective software system. For
-\proglang{R} code, the prompt \code{"R> "} should be used with \code{"+  "} as
-the continuation prompt. Generally, comments within the code chunks should be
-avoided -- and made in the regular {\LaTeX} text instead. Finally, empty lines
-before and after code input/output should be avoided (see above).
-\end{leftbar}
-
-\begin{figure}[t!]
-\centering
-<<visualization, echo=FALSE, fig=TRUE, height=5.2, width=7>>=
-par(mar = c(4, 4, 1, 1))
-plot(table(quine$Days), xlab = "Days", ylab = "Frequency", axes = FALSE)
-axis(2)
-axis(1, at = 0:16 * 5, labels = FALSE)
-axis(1, at = 0:8 * 10)
-box()
-@
-\caption{\label{fig:quine} Frequency distribution for number of days absent
-from school.}
-\end{figure}
-
-As a first model for the \code{quine} data, we fit the basic Poisson regression
-model. (Note that JSS prefers when the second line of code is indented by two
-spaces.)
-%
-<<poisson>>=
-m_pois <- glm(Days ~ (Eth + Sex + Age + Lrn)^2, data = quine,
-  family = poisson)
-@
-%
-To account for potential overdispersion we also consider a negative binomial
-GLM.
-%
-<<negbin>>=
-library("MASS")
-m_nbin <- glm.nb(Days ~ (Eth + Sex + Age + Lrn)^2, data = quine)
-@
-%
-In a comparison with the BIC the latter model is clearly preferred.
-%
-<<comparison>>=
-BIC(m_pois, m_nbin)
-@
-%
-Hence, the full summary of that model is shown below.
-%
-<<summary>>=
-summary(m_nbin)
-@
-
-%% -- Summary/conclusions/discussion -------------------------------------------
-
-\section{Summary and discussion} \label{sec:summary}
-
-\begin{leftbar}
-As usual \dots
-\end{leftbar}
-
-%% -- Optional special unnumbered sections -------------------------------------
-
-\section*{Computational details}
-
-\begin{leftbar}
-If necessary or useful, information about certain computational details
-such as version numbers, operating systems, or compilers could be included
-in an unnumbered section. Also, auxiliary packages (say, for visualizations,
-maps, tables, \dots) that are not cited in the main text can be credited here.
-\end{leftbar}
-
-The results in this paper were obtained using
-\proglang{R}~\Sexpr{paste(R.Version()[6:7], collapse = ".")} with the
-\pkg{MASS}~\Sexpr{packageVersion("MASS")} package. \proglang{R} itself
-and all packages used are available from the Comprehensive
-\proglang{R} Archive Network (CRAN) at
-\url{https://CRAN.R-project.org/}.
-
-\section*{Acknowledgments}
-
-\begin{leftbar}
-All acknowledgments (note the AE spelling) should be collected in this
-unnumbered section before the references. It may contain the usual information
-about funding and feedback from colleagues/reviewers/etc. Furthermore,
-information such as relative contributions of the authors may be added here
-(if any).
-\end{leftbar}
-
-%% -- Bibliography -------------------------------------------------------------
-%% - References need to be provided in a .bib BibTeX database.
-%% - All references should be made with \cite, \citet, \citep, \citealp etc.
-%%   (and never hard-coded). See the FAQ for details.
-%% - JSS-specific markup (\proglang, \pkg, \code) should be used in the .bib.
-%% - Titles in the .bib should be in title case.
-%% - DOIs should be included where available.
+\section[Features]{Features} \label{sec:feat}
 
 \bibliography{refs}
-
-%% -- Appendix (if any) --------------------------------------------------------
-%% - After the bibliography with page break.
-%% - With proper section titles and _not_ just "Appendix".
-
-\newpage
-
-\begin{appendix}
-
-\section{More technical details} \label{app:technical}
-
-\begin{leftbar}
-Appendices can be included after the bibliography (with a page break). Each
-section within the appendix should have a proper section title (rather than
-just \emph{Appendix}).
-
-For more technical style details, please check out JSS's style FAQ at
-\url{https://www.jstatsoft.org/pages/view/style#frequently-asked-questions}
-which includes the following topics:
-\begin{itemize}
-  \item Title vs.\ sentence case.
-  \item Graphics formatting.
-  \item Naming conventions.
-  \item Turning JSS manuscripts into \proglang{R} package vignettes.
-  \item Trouble shooting.
-  \item Many other potentially helpful details\dots
-\end{itemize}
-\end{leftbar}
-
-\section[Using BibTeX]{Using \textsc{Bib}{\TeX}} \label{app:bibtex}
-
-\begin{leftbar}
-References need to be provided in a \textsc{Bib}{\TeX} file (\code{.bib}). All
-references should be made with \verb|\cite|, \verb|\citet|, \verb|\citep|,
-\verb|\citealp| etc.\ (and never hard-coded). This commands yield different
-formats of author-year citations and allow to include additional details (e.g.,
-pages, chapters, \dots) in brackets. In case you are not familiar with these
-commands see the JSS style FAQ for details.
-
-Cleaning up \textsc{Bib}{\TeX} files is a somewhat tedious task -- especially
-when acquiring the entries automatically from mixed online sources. However,
-it is important that informations are complete and presented in a consistent
-style to avoid confusions. JSS requires the following format.
-\begin{itemize}
-  \item JSS-specific markup (\verb|\proglang|, \verb|\pkg|, \verb|\code|) should
-    be used in the references.
-  \item Titles should be in title case.
-  \item Journal titles should not be abbreviated and in title case.
-  \item DOIs should be included where available.
-  \item Software should be properly cited as well. For \proglang{R} packages
-    \code{citation("pkgname")} typically provides a good starting point.
-\end{itemize}
-\end{leftbar}
-
-\end{appendix}
 

### 2019-11-15 jss-article-rnw-diffEnrich/article.tex
Added flowchart as fig 1 and caption to article.Rnw

@@
 
+\begin{figure}[!ht]
+% \centering
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}
+\includegraphics[width=\maxwidth]{figure/visualization-1}
+
+\end{knitrout}
+\caption{\label{fig:flowchart} \textbf{diffEnrich Analysis pipeline.} Functions within
+the diffEnrich package are represented by blue rectangles. The data that must be
+provided by the user is represented by the purple ovals. Data objects generated
+by a function in diffEnrich are represented by red ovals. The external call of
+the \code{get\textunderscore kegg} function to the KEGG REST API is represented in yellow.}
+\end{figure}
+
 In 2012 KEGG released its first application programming interface (API), and has
@@
 
-%% -- Manuscript ---------------------------------------------------------------
+The goal of the diffEnrich package is to compare functional enrichment between
+two experimentally-derived groups of genes or proteins. This package provides
+four functions that are intended to be used in an ordered pipeline
+(Figure~\ref{fig:flowchart}).
 
-%% - In principle "as usual" again.
-%% - When using equations (e.g., {equation}, {eqnarray}, {align}, etc.
-%%   avoid empty lines before and after the equation (which would signal a new
-%%   paragraph.
-%% - When describing longer chunks of code that are _not_ meant for execution
-%%   (e.g., a function synopsis or list of arguments), the environment {Code}
-%%   is recommended. Alternatively, a plain {verbatim} can also be used.
-%%   (For executed code see the next section.)
+%% ------------Section 2: Features----------------------------------------------
 
-\section{Models and software} \label{sec:models}
-
-The basic Poisson regression model for count data is a special case of the GLM
-framework \cite{McCullagh+Nelder:1989}. It describes the dependence of a count
-response variable $y_i$ ($i = 1, \dots, n$) by assuming a Poisson distribution
-$y_i \sim \mathrm{Pois}(\mu_i)$. The dependence of the conditional mean
-$\E[y_i \, | \, x_i] = \mu_i$ on the regressors $x_i$ is then specified via a
-log link and a linear predictor
-%
-\begin{equation} \label{eq:mean}
-\log(\mu_i) \quad = \quad x_i^\top \beta,
-\end{equation}
-%
-where the regression coefficients $\beta$ are estimated by maximum likelihood
-(ML) using the iterative weighted least squares (IWLS) algorithm.
-
-\begin{leftbar}
-Note that around the \verb|{equation}| above there should be no spaces (avoided
-in the {\LaTeX} code by \verb|%| lines) so that ``normal'' spacing is used and
-not a new paragraph started.
-\end{leftbar}
-
-\proglang{R} provides a very flexible implementation of the general GLM
-framework in the function \fct{glm} \citep{Chambers+Hastie:1992} in the
-\pkg{stats} package. Its most important arguments are
-\begin{Code}
-glm(formula, data, subset, na.action, weights, offset,
-  family = gaussian, start = NULL, control = glm.control(...),
-  model = TRUE, y = TRUE, x = FALSE, ...)
-\end{Code}
-where \code{formula} plus \code{data} is the now standard way of specifying
-regression relationships in \proglang{R}/\proglang{S} introduced in
-\cite{Chambers+Hastie:1992}. The remaining arguments in the first line
-(\code{subset}, \code{na.action}, \code{weights}, and \code{offset}) are also
-standard  for setting up formula-based regression models in
-\proglang{R}/\proglang{S}. The arguments in the second line control aspects
-specific to GLMs while the arguments in the last line specify which components
-are returned in the fitted model object (of class \class{glm} which inherits
-from \class{lm}). For further arguments to \fct{glm} (including alternative
-specifications of starting values) see \code{?glm}. For estimating a Poisson
-model \code{family = poisson} has to be specified.
-
-\begin{leftbar}
-As the synopsis above is a code listing that is not meant to be executed,
-one can use either the dedicated \verb|{Code}| environment or a simple
-\verb|{verbatim}| environment for this. Again, spaces before and after should be
-avoided.
-
-Finally, there might be a reference to a \verb|{table}| such as
-Table~\ref{tab:overview}. Usually, these are placed at the top of the page
-(\verb|[t!]|), centered (\verb|\centering|), with a caption below the table,
-column headers and captions in sentence style, and if possible avoiding vertical
-lines.
-\end{leftbar}
-
-\begin{table}[t!]
-\centering
-\begin{tabular}{lllp{7.4cm}}
-\hline
-Type           & Distribution & Method   & Description \\ \hline
-GLM            & Poisson      & ML       & Poisson regression: classical GLM,
-                                           estimated by maximum likelihood (ML) \\
-               &              & Quasi    & ``Quasi-Poisson regression'':
-                                           same mean function, estimated by
-                                           quasi-ML (QML) or equivalently
-                                           generalized estimating equations (GEE),
-                                           inference adjustment via estimated
-                                           dispersion parameter \\
-               &              & Adjusted & ``Adjusted Poisson regression'':
-                                           same mean function, estimated by
-                                           QML/GEE, inference adjustment via
-                                           sandwich covariances\\
-               & NB           & ML       & NB regression: extended GLM,
-                                           estimated by ML including additional
-                                           shape parameter \\ \hline
-Zero-augmented & Poisson      & ML       & Zero-inflated Poisson (ZIP),
-                                           hurdle Poisson \\
-               & NB           & ML       & Zero-inflated NB (ZINB),
-                                           hurdle NB \\ \hline
-\end{tabular}
-\caption{\label{tab:overview} Overview of various count regression models. The
-table is usually placed at the top of the page (\texttt{[t!]}), centered
-(\texttt{centering}), has a caption below the table, column headers and captions
-are in sentence style, and if possible vertical lines should be avoided.}
-\end{table}
-
-%% -- Illustrations ------------------------------------------------------------
-
-%% - Virtually all JSS manuscripts list source code along with the generated
-%%   output. The style files provide dedicated environments for this.
-%% - In R, the environments {Sinput} and {Soutput} - as produced by Sweave() or
-%%   or knitr using the render_sweave() hook - are used (without the need to
-%%   load Sweave.sty).
-%% - Equivalently, {CodeInput} and {CodeOutput} can be used.
-%% - The code input should use "the usual" command prompt in the respective
-%%   software system.
-%% - For R code, the prompt "R> " should be used with "+  " as the
-%%   continuation prompt.
-%% - Comments within the code chunks should be avoided - these should be made
-%%   within the regular LaTeX text.
-
-\section{Illustrations} \label{sec:illustrations}
-
-For a simple illustration of basic Poisson and NB count regression the
-\code{quine} data from the \pkg{MASS} package is used. This provides the number
-of \code{Days} that children were absent from school in Australia in a
-particular year, along with several covariates that can be employed as regressors.
-The data can be loaded by
-%
-\begin{knitrout}
-\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
-\begin{alltt}
-\hlkwd{data}\hlstd{(}\hlstr{"quine"}\hlstd{,} \hlkwc{package} \hlstd{=} \hlstr{"MASS"}\hlstd{)}
-\end{alltt}
-\end{kframe}
-\end{knitrout}
-%
-and a basic frequency distribution of the response variable is displayed in
-Figure~\ref{fig:quine}.
-
-\begin{leftbar}
-For code input and output, the style files provide dedicated environments.
-Either the ``agnostic'' \verb|{CodeInput}| and \verb|{CodeOutput}| can be used
-or, equivalently, the environments \verb|{Sinput}| and \verb|{Soutput}| as
-produced by \fct{Sweave} or \pkg{knitr} when using the \code{render_sweave()}
-hook. Please make sure that all code is properly spaced, e.g., using
-\code{y = a + b * x} and \emph{not} \code{y=a+b*x}. Moreover, code input should
-use ``the usual'' command prompt in the respective software system. For
-\proglang{R} code, the prompt \code{"R> "} should be used with \code{"+  "} as
-the continuation prompt. Generally, comments within the code chunks should be
-avoided -- and made in the regular {\LaTeX} text instead. Finally, empty lines
-before and after code input/output should be avoided (see above).
-\end{leftbar}
-
-\begin{figure}[t!]
-\centering
-\begin{knitrout}
-\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}
-\includegraphics[width=\maxwidth]{figure/visualization-1}
-
-\end{knitrout}
-\caption{\label{fig:quine} Frequency distribution for number of days absent
-from school.}
-\end{figure}
-
-As a first model for the \code{quine} data, we fit the basic Poisson regression
-model. (Note that JSS prefers when the second line of code is indented by two
-spaces.)
-%
-\begin{knitrout}
-\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
-\begin{alltt}
-\hlstd{m_pois} \hlkwb{<-} \hlkwd{glm}\hlstd{(Days} \hlopt{~} \hlstd{(Eth} \hlopt{+} \hlstd{Sex} \hlopt{+} \hlstd{Age} \hlopt{+} \hlstd{Lrn)}\hlopt{^}\hlnum{2}\hlstd{,} \hlkwc{data} \hlstd{= quine,}
-  \hlkwc{family} \hlstd{= poisson)}
-\end{alltt}
-\end{kframe}
-\end{knitrout}
-%
-To account for potential overdispersion we also consider a negative binomial
-GLM.
-%
-\begin{knitrout}
-\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
-\begin{alltt}
-\hlkwd{library}\hlstd{(}\hlstr{"MASS"}\hlstd{)}
-\hlstd{m_nbin} \hlkwb{<-} \hlkwd{glm.nb}\hlstd{(Days} \hlopt{~} \hlstd{(Eth} \hlopt{+} \hlstd{Sex} \hlopt{+} \hlstd{Age} \hlopt{+} \hlstd{Lrn)}\hlopt{^}\hlnum{2}\hlstd{,} \hlkwc{data} \hlstd{= quine)}
-\end{alltt}
-\end{kframe}
-\end{knitrout}
-%
-In a comparison with the BIC the latter model is clearly preferred.
-%
-\begin{knitrout}
-\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
-\begin{alltt}
-\hlkwd{BIC}\hlstd{(m_pois, m_nbin)}
-\end{alltt}
-\begin{verbatim}
-##        df      BIC
-## m_pois 18 2046.851
-## m_nbin 19 1157.235
-\end{verbatim}
-\end{kframe}
-\end{knitrout}
-%
-Hence, the full summary of that model is shown below.
-%
-\begin{knitrout}
-\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
-\begin{alltt}
-\hlkwd{summary}\hlstd{(m_nbin)}
-\end{alltt}
-\begin{verbatim}
-## 
-## Call:
-## glm.nb(formula = Days ~ (Eth + Sex + Age + Lrn)^2, data = quine, 
-##     init.theta = 1.60364105, link = log)
-## 
-## Deviance Residuals: 
-##     Min       1Q   Median       3Q      Max  
-## -3.0857  -0.8306  -0.2620   0.4282   2.0898  
-## 
-## Coefficients: (1 not defined because of singularities)
-##             Estimate Std. Error z value Pr(>|z|)    
-## (Intercept)  3.00155    0.33709   8.904  < 2e-16 ***
-## EthN        -0.24591    0.39135  -0.628  0.52977    
-## SexM        -0.77181    0.38021  -2.030  0.04236 *  
-## AgeF1       -0.02546    0.41615  -0.061  0.95121    
-## AgeF2       -0.54884    0.54393  -1.009  0.31296    
-## AgeF3       -0.25735    0.40558  -0.635  0.52574    
-## LrnSL        0.38919    0.48421   0.804  0.42153    
-## EthN:SexM    0.36240    0.29430   1.231  0.21818    
-## EthN:AgeF1  -0.70000    0.43646  -1.604  0.10876    
-## EthN:AgeF2  -1.23283    0.42962  -2.870  0.00411 ** 
-## EthN:AgeF3   0.04721    0.44883   0.105  0.91622    
-## EthN:LrnSL   0.06847    0.34040   0.201  0.84059    
-## SexM:AgeF1   0.02257    0.47360   0.048  0.96198    
-## SexM:AgeF2   1.55330    0.51325   3.026  0.00247 ** 
-## SexM:AgeF3   1.25227    0.45539   2.750  0.00596 ** 
-## SexM:LrnSL   0.07187    0.40805   0.176  0.86019    
-## AgeF1:LrnSL -0.43101    0.47948  -0.899  0.36870    
-## AgeF2:LrnSL  0.52074    0.48567   1.072  0.28363    
-## AgeF3:LrnSL       NA         NA      NA       NA    
-## ---
-## Signif. codes:  0 '***' 0.001 '**' 0.01 '*' 0.05 '.' 0.1 ' ' 1
-## 
-## (Dispersion parameter for Negative Binomial(1.6036) family taken to be 1)
-## 
-##     Null deviance: 235.23  on 145  degrees of freedom
-## Residual deviance: 167.53  on 128  degrees of freedom
-## AIC: 1100.5
-## 
-## Number of Fisher Scoring iterations: 1
-## 
-## 
-##               Theta:  1.604 
-##           Std. Err.:  0.214 
-## 
-##  2 x log-likelihood:  -1062.546
-\end{verbatim}
-\end{kframe}
-\end{knitrout}
-
-%% -- Summary/conclusions/discussion -------------------------------------------
-
-\section{Summary and discussion} \label{sec:summary}
-
-\begin{leftbar}
-As usual \dots
-\end{leftbar}
-
-%% -- Optional special unnumbered sections -------------------------------------
-
-\section*{Computational details}
-
-\begin{leftbar}
-If necessary or useful, information about certain computational details
-such as version numbers, operating systems, or compilers could be included
-in an unnumbered section. Also, auxiliary packages (say, for visualizations,
-maps, tables, \dots) that are not cited in the main text can be credited here.
-\end{leftbar}
-
-The results in this paper were obtained using
-\proglang{R}~3.6.1 with the
-\pkg{MASS}~7.3.51.4 package. \proglang{R} itself
-and all packages used are available from the Comprehensive
-\proglang{R} Archive Network (CRAN) at
-\url{https://CRAN.R-project.org/}.
-
-\section*{Acknowledgments}
-
-\begin{leftbar}
-All acknowledgments (note the AE spelling) should be collected in this
-unnumbered section before the references. It may contain the usual information
-about funding and feedback from colleagues/reviewers/etc. Furthermore,
-information such as relative contributions of the authors may be added here
-(if any).
-\end{leftbar}
-
-%% -- Bibliography -------------------------------------------------------------
-%% - References need to be provided in a .bib BibTeX database.
-%% - All references should be made with \cite, \citet, \citep, \citealp etc.
-%%   (and never hard-coded). See the FAQ for details.
-%% - JSS-specific markup (\proglang, \pkg, \code) should be used in the .bib.
-%% - Titles in the .bib should be in title case.
-%% - DOIs should be included where available.
+\section[Features]{Features} \label{sec:feat}
 
 \bibliography{refs}
-
-%% -- Appendix (if any) --------------------------------------------------------
-%% - After the bibliography with page break.
-%% - With proper section titles and _not_ just "Appendix".
-
-\newpage
-
-\begin{appendix}
-
-\section{More technical details} \label{app:technical}
-
-\begin{leftbar}
-Appendices can be included after the bibliography (with a page break). Each
-section within the appendix should have a proper section title (rather than
-just \emph{Appendix}).
-
-For more technical style details, please check out JSS's style FAQ at
-\url{https://www.jstatsoft.org/pages/view/style#frequently-asked-questions}
-which includes the following topics:
-\begin{itemize}
-  \item Title vs.\ sentence case.
-  \item Graphics formatting.
-  \item Naming conventions.
-  \item Turning JSS manuscripts into \proglang{R} package vignettes.
-  \item Trouble shooting.
-  \item Many other potentially helpful details\dots
-\end{itemize}
-\end{leftbar}
-
-\section[Using BibTeX]{Using \textsc{Bib}{\TeX}} \label{app:bibtex}
-
-\begin{leftbar}
-References need to be provided in a \textsc{Bib}{\TeX} file (\code{.bib}). All
-references should be made with \verb|\cite|, \verb|\citet|, \verb|\citep|,
-\verb|\citealp| etc.\ (and never hard-coded). This commands yield different
-formats of author-year citations and allow to include additional details (e.g.,
-pages, chapters, \dots) in brackets. In case you are not familiar with these
-commands see the JSS style FAQ for details.
-
-Cleaning up \textsc{Bib}{\TeX} files is a somewhat tedious task -- especially
-when acquiring the entries automatically from mixed online sources. However,
-it is important that informations are complete and presented in a consistent
-style to avoid confusions. JSS requires the following format.
-\begin{itemize}
-  \item JSS-specific markup (\verb|\proglang|, \verb|\pkg|, \verb|\code|) should
-    be used in the references.
-  \item Titles should be in title case.
-  \item Journal titles should not be abbreviated and in title case.
-  \item DOIs should be included where available.
-  \item Software should be properly cited as well. For \proglang{R} packages
-    \code{citation("pkgname")} typically provides a good starting point.
-\end{itemize}
-\end{leftbar}
-
-\end{appendix}
 

### 2019-11-15 jss-article-rnw-diffEnrich/article.Rnw
Started subsection 2.1 in article.Rnw

@@
 
+%% ------------Section 2: Features----------------------------------------------
+
+\section[Features]{Features} \label{sec:feat}
+
 The goal of the diffEnrich package is to compare functional enrichment between
@@
 
-%% ------------Section 2: Features----------------------------------------------
+\medskip
 
-\section[Features]{Features} \label{sec:feat}
+You can install the released version of diffEnrich from CRAN with:
+
+<<installation, echo=TRUE, eval=FALSE>>=
+install.packages("diffEnrich")
+@
+
+\subsection{\code{get\textunderscore kegg}: Download and prepare pathways from KEGG API}
+
+First, the \code{get\textunderscore kegg} function is used to connect to the
+KEGG REST API and download the data sets required to perform downstream
+analysis. Currently, this function supports three species: Homo sapiens, Mus
+musculus, and Rattus norvegicus. For a given species, three data sets are
+generated: 1) ncbi\textunderscore to\textunderscore kegg: this data set maps NCBI/ENTREZ gene IDs to KEGG gene
+IDs, 2) kegg\textunderscore to\textunderscore pathway: this data set maps KEGG gene IDs to their respective
+KEGG pathway IDs, and 3) pathway\textunderscore to\textunderscore species: this data set maps KEGG pathway IDs
+to their respective pathway descriptions. This function allows the user to save
+versioned (based on KEGG release) and time-stamped text files of the three data
+sets described above. In addition to these flat files, \code{get\textunderscore kegg} will also create
+a named list in R with the three data sets. \code{get\textunderscore kegg} can also read in the saved
+text files from a past run and format them for downstream analyses. This
+function typically completes in a few seconds, but it is important to note that
+the finishing time is dependent on the time it takes to connect to the KEGG API.
 

### 2019-11-15 jss-article-rnw-diffEnrich/article.tex
Started subsection 2.1 in article.Rnw

@@
 
-%% ------------Section 2: Features----------------------------------------------
+\medskip
 
-\section[Features]{Features} \label{sec:feat}
+You can install the released version of diffEnrich from CRAN with:
+
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{alltt}
+\hlkwd{install.packages}\hlstd{(}\hlstr{"diffEnrich"}\hlstd{)}
+\end{alltt}
+\end{kframe}
+\end{knitrout}
+
+\subsection{\code{get\textunderscore kegg}: Download and prepare pathways from KEGG API}
+
+First, the \code{get\textunderscore kegg} function is used to connect to the
+KEGG REST API and download the data sets required to perform downstream
+analysis. Currently, this function supports three species: Homo sapiens, Mus
+musculus, and Rattus norvegicus. For a given species, three data sets are
+generated: 1) ncbi\textunderscore to\textunderscore kegg: this data set maps NCBI/ENTREZ gene IDs to KEGG gene
+IDs, 2) kegg\textunderscore to\textunderscore pathway: this data set maps KEGG gene IDs to their respective
+KEGG pathway IDs, and 3) pathway\textunderscore to\textunderscore species: this data set maps KEGG pathway IDs
+to their respective pathway descriptions. This function allows the user to save
+versioned (based on KEGG release) and time-stamped text files of the three data
+sets described above. In addition to these flat files, \code{get\textunderscore kegg} will also create
+a named list in R with the three data sets. \code{get\textunderscore kegg} can also read in the saved
+text files from a past run and format them for downstream analyses. This
+function typically completes in a few seconds, but it is important to note that
+the finishing time is dependent on the time it takes to connect to the KEGG API.
 

### 2019-11-18 README.Rmd
Fixed format error in README

@@
 The default is the False Discovery Rate ( [Benjamini and Hochberg, 1995](http://www.jstor.org/stable/2346101)).
-KEGG pathways that do not contain any genes from either gene list (i.e., *list1_pe$enrich_table$KEGG_PATHWAY_in_list* for 'rno04530' = 0 AND *list2_pe$enrich_table$KEGG_PATHWAY_in_list* for 'rno04530' = 0) will be removed as these cannot be tested.
+KEGG pathways that do not contain any genes from either gene list (i.e., list1_pe\$enrich_table\$KEGG_PATHWAY_in_list for 'rno04530' = 0 AND list2_pe\$enrich_table\$KEGG_PATHWAY_in_list for 'rno04530' = 0) will be removed as these cannot be tested.
 If this is the case a warning will be printed that tells the user how many pathways were removed.

### 2019-11-18 jss-article-rnw-diffEnrich/article.Rnw
Started Step 1 instruction in Rnw

@@
 
+The \code{get_kegg} function accesses the KEGG REST API and downloads the data
+sets required to perform downstream analysis. This function takes two arguments.
+The first, 'species' is required. Currently, diffEnrich supports three species,
+and the argument is a character string using the KEGG code: Homo sapiens
+(human), use 'hsa'; Mus musculus (mouse), use 'mmu'; and Rattus norvegicus
+(rat), use 'rno'. The second, 'path' is also passed as a character string, and
+is the path of the directory in which the user would like to write the data sets
+downloaded from the KEGG REST API. If the user does not provide a path, the data
+sets will be automatically written to the current working directory using the
+\code{here::here()} (\cite{kirill:2017}) functionality. These data sets will be
+tab delimited files with a name describing the data, and for reproducibility,
+the date they were generated and the version of KEGG when the API was accessed.
+In addition to these flat files, \code{get_kegg} will also create a named list
+in R with the three relevant KEGG data sets. The names of this list will
+describe the data set, and are described in Table 1.
+
 \bibliography{refs}

### 2019-11-20 jss-article-rnw-diffEnrich/article.Rnw
Added table 1 to .Rnw

@@
 
-The \code{get_kegg} function accesses the KEGG REST API and downloads the data
+The \code{get_\textunderscore kegg} function accesses the KEGG REST API and downloads the data
 sets required to perform downstream analysis. This function takes two arguments.
@@
 the date they were generated and the version of KEGG when the API was accessed.
-In addition to these flat files, \code{get_kegg} will also create a named list
+In addition to these flat files, \code{get\textunderscore kegg} will also create a named list
 in R with the three relevant KEGG data sets. The names of this list will
 describe the data set, and are described in Table 1.
+
+\begin{table}[ht!]
+\centering
+\begin{tabular}{lllp{7.4cm}}
+\hline
+\code{get\textunderscore kegg} list object & Description \\ \hline
+ncbi\textunderscore to\textunderscore kegg                & ncbi gene ID <-- mapped to --> KEGG gene ID \\
+kegg\textunderscore to\textunderscore pathway             & KEGG gene ID <-- mapped to --> KEGG pathway ID \\
+pathway\textunderscore to\textunderscore species          & KEGG pathway ID <-- mapped to --> KEGG pathway species description \\ \hline
+\end{tabular}
+\caption{\label{tab:description} Description of the data sets retrieved by \code{get\textunderscore kegg}'s
+connection to the KEGG REST API.}
+\end{table}
 

### 2019-11-20 jss-article-rnw-diffEnrich/article.Rnw
Added get_kegg code examples

@@
 KEGG pathway IDs, and 3) pathway\textunderscore to\textunderscore species: this data set maps KEGG pathway IDs
-to their respective pathway descriptions. This function allows the user to save
-versioned (based on KEGG release) and time-stamped text files of the three data
-sets described above. In addition to these flat files, \code{get\textunderscore kegg} will also create
-a named list in R with the three data sets. \code{get\textunderscore kegg} can also read in the saved
-text files from a past run and format them for downstream analyses. This
-function typically completes in a few seconds, but it is important to note that
-the finishing time is dependent on the time it takes to connect to the KEGG API.
+to their respective pathway descriptions. This function typically completes in a
+few seconds, but it is important to note that the finishing time is dependent on
+the time it takes to connect to the KEGG API.
 
@@
 describe the data set, and are described in Table 1.
+
+<<get_kegg1, echo = TRUE>>=
+## Load package
+suppressMessages(library(diffEnrich))
+## run get_kegg() using rat
+kegg_rno <- get_kegg('rno')
+@
+
+\textbf{Note:} Because it is assumed that a user might want to use the data sets
+generated by \code{get\textunderscore kegg}, it is careful not to overwrite data
+sets with exact names. \code{get\textunderscore kegg} checks the path provided
+for data sets generated 'same-day/same-version', and if it finds even one of the
+three, it will not re-write any of the data sets. It will still however, let the
+user know it is not writing out new data sets and still generate the named list
+object. Users can generate 'same-day/same-version' data sets in different
+directories if they so choose.
+
+<<get_kegg2, echo = TRUE>>=
+## run get_kegg() using rat
+kegg_rno <- get_kegg('rno')
+@
+
+Additionally, \code{get\textunderscore kegg} can be used to read in saved
+versions of the txt files generated from a previous call, and generate an R list
+object that is compatible with downstream functions.
+
+<<<get_kegg_read, echo = TRUE>>=
+## run get_kegg() using rat
+date <- as.character(Sys.Date())
+kegg_rno <- get_kegg(read = TRUE,
+                     path = here::here(),
+                     date = date,
+                     release = "92")
+@
 

### 2019-11-20 jss-article-rnw-diffEnrich/article.tex
Added get_kegg code examples

@@
 describe the data set, and are described in Table 1.
+
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{alltt}
+\hlcom{## Load package}
+\hlkwd{suppressMessages}\hlstd{(}\hlkwd{library}\hlstd{(diffEnrich))}
+\hlcom{## run get_kegg() using rat}
+\hlstd{kegg_rno} \hlkwb{<-} \hlkwd{get_kegg}\hlstd{(}\hlstr{'rno'}\hlstd{)}
+\end{alltt}
+
+{\ttfamily\noindent\itshape\color{messagecolor}{\#\# 3 data sets will be written as tab delimited text files}}
+
+{\ttfamily\noindent\itshape\color{messagecolor}{\#\# File location: /Users/smiharry/Documents/packages/diffEnrich}}
+
+{\ttfamily\noindent\itshape\color{messagecolor}{\#\# Kegg Release: Release\_92.0+\_11-20\_Nov\_19}}\end{kframe}
+\end{knitrout}
+
+\textbf{Note:} Because it is assumed that a user might want to use the data sets
+generated by \code{get\textunderscore kegg}, it is careful not to overwrite data
+sets with exact names. \code{get\textunderscore kegg} checks the path provided
+for data sets generated 'same-day/same-version', and if it finds even one of the
+three, it will not re-write any of the data sets. It will still however, let the
+user know it is not writing out new data sets and still generate the named list
+object. Users can generate 'same-day/same-version' data sets in different
+directories if they so choose.
+
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{alltt}
+\hlcom{## run get_kegg() using rat}
+\hlstd{kegg_rno} \hlkwb{<-} \hlkwd{get_kegg}\hlstd{(}\hlstr{'rno'}\hlstd{)}
+\end{alltt}
+
+{\ttfamily\noindent\itshape\color{messagecolor}{\#\# These files already exist in your working directory. New files will not be generated.}}
+
+{\ttfamily\noindent\itshape\color{messagecolor}{\#\# Kegg Release: Release\_92.0+\_11-20\_Nov\_19}}\end{kframe}
+\end{knitrout}
+
+Additionally, \code{get\textunderscore kegg} can be used to read in saved
+versions of the txt files generated from a previous call, and generate an R list
+object that is compatible with downstream functions.
+
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{alltt}
+\hlcom{## run get_kegg() using rat}
+\hlstd{date} \hlkwb{<-} \hlkwd{as.character}\hlstd{(}\hlkwd{Sys.Date}\hlstd{())}
+\hlstd{kegg_rno} \hlkwb{<-} \hlkwd{get_kegg}\hlstd{(}\hlkwc{read} \hlstd{=} \hlnum{TRUE}\hlstd{,}
+                     \hlkwc{path} \hlstd{= here}\hlopt{::}\hlkwd{here}\hlstd{(),}
+                     \hlkwc{date} \hlstd{= date,}
+                     \hlkwc{release} \hlstd{=} \hlstr{"92"}\hlstd{)}
+\end{alltt}
+
+{\ttfamily\noindent\itshape\color{messagecolor}{\#\# Reading in the following files:}}
+
+{\ttfamily\noindent\itshape\color{messagecolor}{\#\# ncbi\_to\_kegg2019-11-20Release\_92.0+\_11-20\_Nov\_19.txt}}
+
+{\ttfamily\noindent\itshape\color{messagecolor}{\#\# kegg\_to\_pathway2019-11-20Release\_92.0+\_11-20\_Nov\_19.txt}}
+
+{\ttfamily\noindent\itshape\color{messagecolor}{\#\# pathway\_to\_species2019-11-20Release\_92.0+\_11-20\_Nov\_19.txt}}
+
+{\ttfamily\noindent\itshape\color{messagecolor}{\#\# File location: /Users/smiharry/Documents/packages/diffEnrich}}\end{kframe}
+\end{knitrout}
 

### 2019-11-20 jss-article-rnw-diffEnrich/article.Rnw
Added text for sec 2.2 (pathEnrich) in .Rnw

@@
 
+\subsection{\code{pathEnrich}: Perform enrichment analysis of individual gene sets.}
+
+In this step, the \code{pathEnrich} function is used to identify KEGG pathways
+that are enriched (i.e. over-represented) based on a gene list of interest
+provided by the user. User gene lists must be ENTREZ gene IDs. If a user only
+has gene symbols, the \code{clusterProfiler} package (3.9) (Yu:2012) offers a
+function (\code{bitr}) that maps gene symbols and Ensembl IDs to ENTREZ gene
+IDs. An example of this function's use can be found in their vignette
+(\url{https://yulab-smu.github.io/clusterProfiler-book/chapter14.html#bitr}).
+
+<<view_genelists, echo = TRUE>>=
+## View sample gene lists from package data
+head(geneLists$list1)
+head(geneLists$list2)
+@
+
+The pathEnrich function will only use the genes from the list provided that are
+also in the KEGG database. The \code{pathEnrich} function should be run at least
+twice, once for the genes of interest in list 1 and once for the genes of
+interest in list 2. Each \code{pathEnrich} call generates a data frame
+summarizing the results of enrichment analyses in which a Fisher’s Exact test is
+used to identify which KEGG pathways are enriched within the user’s list of
+genes compared to all genes annotated to a KEGG pathway. Users can limit which
+pathways are tested by requiring that they contain a minimum number of genes
+from the list, and this can be set by changing the 'N' arguement. The default is
+that a KEGG pathway must contain at least 2 genes (N = 2) from the user’s list
+to be tested.
+
+<<<pathEnrich, echo = FALSE>>=
+# run pathEnrich
+## List 1
+list1_pe <- pathEnrich(gk_obj = kegg, gene_list = geneLists$list1, cutoff = 0.05, N = 2)
+## list2
+list2_pe <- pathEnrich(gk_obj = kegg, gene_list = geneLists$list2, cutoff = 0.05, N = 2)
+@
+
+By default, p-values from the Fisher’s Exact test are adjusted for multiple
+comparisons with a False Discovery Rate (FDR) (Benjamini:1995), however users
+have the option of choosing any type of multiple testing correction supported
+by \code{p.adjust}. In addition to the unadjusted p-value and FDR, pathEnrich
+will calculate for each KEGG pathway, its fold enrichment defined as the ratio
+of number of genes observed from the gene list annotated to the KEGG pathway to
+the expected number of genes from the gene list to be annotated to the KEGG
+pathway by chance. An example of the first 5 results generated by \code{pathEnrich}
+are in Table 2. For a detailed description of the variables in this table see Table 3.
+
+S3 generic functions for \code{print} and \code{summary} are provided.
+The \code{print} function prints the results table as a \code{tibble}, and the
+\code{summary} function returns the number of pathways that reached statistical
+significance as well as their descriptions, the number of genes used from the
+KEGG data base, the KEGG species, and the method used for multiple testing
+correction.
+
+<<path_summary, echo = TRUE>>=
+summary(list1_pe)
+@
+
 \bibliography{refs}

### 2019-11-20 jss-article-rnw-diffEnrich/article.tex
Added text for sec 2.2 (pathEnrich) in .Rnw

@@
 
-{\ttfamily\noindent\itshape\color{messagecolor}{\#\# 3 data sets will be written as tab delimited text files}}
-
-{\ttfamily\noindent\itshape\color{messagecolor}{\#\# File location: /Users/smiharry/Documents/packages/diffEnrich}}
+{\ttfamily\noindent\itshape\color{messagecolor}{\#\# These files already exist in your working directory. New files will not be generated.}}
 
@@
 
+\subsection{\code{pathEnrich}: Perform enrichment analysis of individual gene sets.}
+
+In this step, the \code{pathEnrich} function is used to identify KEGG pathways
+that are enriched (i.e. over-represented) based on a gene list of interest
+provided by the user. User gene lists must be ENTREZ gene IDs. If a user only
+has gene symbols, the \code{clusterProfiler} package (3.9) (Yu:2012) offers a
+function (\code{bitr}) that maps gene symbols and Ensembl IDs to ENTREZ gene
+IDs. An example of this function's use can be found in their vignette
+(\url{https://yulab-smu.github.io/clusterProfiler-book/chapter14.html#bitr}).
+
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{alltt}
+\hlcom{## View sample gene lists from package data}
+\hlkwd{head}\hlstd{(geneLists}\hlopt{$}\hlstd{list1)}
+\end{alltt}
+\begin{verbatim}
+## [1] "361692"    "293654"    "293655"    "500974"    "100361529"
+## [6] "171434"
+\end{verbatim}
+\begin{alltt}
+\hlkwd{head}\hlstd{(geneLists}\hlopt{$}\hlstd{list2)}
+\end{alltt}
+\begin{verbatim}
+## [1] "315547" "315548" "315549" "315550" "50938"  "58856"
+\end{verbatim}
+\end{kframe}
+\end{knitrout}
+
+The pathEnrich function will only use the genes from the list provided that are
+also in the KEGG database. The \code{pathEnrich} function should be run at least
+twice, once for the genes of interest in list 1 and once for the genes of
+interest in list 2. Each \code{pathEnrich} call generates a data frame
+summarizing the results of enrichment analyses in which a Fisher’s Exact test is
+used to identify which KEGG pathways are enriched within the user’s list of
+genes compared to all genes annotated to a KEGG pathway. Users can limit which
+pathways are tested by requiring that they contain a minimum number of genes
+from the list, and this can be set by changing the 'N' arguement. The default is
+that a KEGG pathway must contain at least 2 genes (N = 2) from the user’s list
+to be tested.
+
+By default, p-values from the Fisher’s Exact test are adjusted for multiple
+comparisons with a False Discovery Rate (FDR) (Benjamini:1995), however users
+have the option of choosing any type of multiple testing correction supported
+by \code{p.adjust}. In addition to the unadjusted p-value and FDR, pathEnrich
+will calculate for each KEGG pathway, its fold enrichment defined as the ratio
+of number of genes observed from the gene list annotated to the KEGG pathway to
+the expected number of genes from the gene list to be annotated to the KEGG
+pathway by chance. An example of the first 5 results generated by \code{pathEnrich}
+are in Table 2. For a detailed description of the variables in this table see Table 3.
+
+S3 generic functions for \code{print} and \code{summary} are provided.
+The \code{print} function prints the results table as a \code{tibble}, and the
+\code{summary} function returns the number of pathways that reached statistical
+significance as well as their descriptions, the number of genes used from the
+KEGG data base, the KEGG species, and the method used for multiple testing
+correction.
+
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{alltt}
+\hlkwd{summary}\hlstd{(list1_pe)}
+\end{alltt}
+\begin{verbatim}
+## 219 KEGG pathways were tested. 
+##  Only KEGG pathways that contained at least 2 genes from gene_list were tested. 
+##  KEGG pathway species: Rattus norvegicus (rat)
+##  8856 genes from gene_list were in the KEGG data pull. 
+##  p-value adjustment method: BH
+##  36 pathways reached statistical significance after multiple testing correction at a cutoff of 0.05. 
+##  
+## Significant pathways: 
+##  Tight junction
+## Yersinia infection
+## Colorectal cancer
+## Choline metabolism in cancer
+## Endometrial cancer
+## Endocytosis
+## Neurotrophin signaling pathway
+## Thermogenesis
+## Oocyte meiosis
+## VEGF signaling pathway
+## Thyroid hormone signaling pathway
+## Hippo signaling pathway
+## T cell receptor signaling pathway
+## Apoptosis
+## Hepatocellular carcinoma
+## MAPK signaling pathway
+## Focal adhesion
+## Salmonella infection
+## Non-alcoholic fatty liver disease (NAFLD)
+## ErbB signaling pathway
+## Sphingolipid signaling pathway
+## Pancreatic cancer
+## Progesterone-mediated oocyte maturation
+## Alzheimer disease
+## Endocrine resistance
+## Adrenergic signaling in cardiomyocytes
+## IL-17 signaling pathway
+## Chronic myeloid leukemia
+## Dopaminergic synapse
+## Prostate cancer
+## EGFR tyrosine kinase inhibitor resistance
+## Hepatitis C
+## Ras signaling pathway
+## Acute myeloid leukemia
+## Insulin signaling pathway
+## Fc epsilon RI signaling pathway
+\end{verbatim}
+\end{kframe}
+\end{knitrout}
+
 \bibliography{refs}

### 2019-11-20 R/plotFoldEnrichment.R
Bug fix in plotFold Enrichment. Should now retain proper ordering when plotting more than 6 pathways

@@ -72,24 +72,26 @@ plotFoldEnrichment <- function(de_res, pval, N){
   ## Clean up melted data frame
   df.ss <- df.melt %>%
     dplyr::filter(.data$variable %in% c("fold_enrichment_list1", "fold_enrichment_list2",
-                         "enrich_p_list1", "enrich_p_list2",
-                         "diff_enrich_adjusted"))
+                                        "enrich_p_list1", "enrich_p_list2",
+                                        "diff_enrich_adjusted"))
 
   ## get vector of pvals
   pvals <- subset(df.ss, df.ss$variable %in% c("enrich_p_list1", "enrich_p_list2"))
 
   ## Generate data set to be used for plotting
-  bardat <- subset(df.ss, df.ss$variable %in% c("fold_enrichment_list1", "fold_enrichment_list2")) %>%
+  bardat.tmp <- subset(df.ss, df.ss$variable %in% c("fold_enrichment_list1", "fold_enrichment_list2")) %>%
     dplyr::mutate(alpha = log10(pvals$value),
-           pvals = pvals$value) %>%
+                  pvals = pvals$value) %>%
     dplyr::arrange(.data$pvals)
+
+  bardat <- merge(bardat.tmp, df, by = "KEGG_PATHWAY_ID")
 
   ###########################################################
   # Generate plot
   ###########################################################
-    # library(ggnewscale)
+  # library(ggnewscale)
   # First, we'll make a plot and save it as a variable
-    g <- ggplot(bardat, aes(x=stats::reorder(.data$KEGG_PATHWAY_description, -.data$pvals), y=.data$value)) +
+  g <- ggplot(bardat, aes(x=stats::reorder(.data$KEGG_PATHWAY_description.x, -.data$diff_enrich_adjusted), y=.data$value)) +
     geom_bar(stat="identity", aes(col=.data$variable, group=.data$variable, fill=.data$pvals), position="dodge") +
     ylim(0, max(bardat$value) + 0.6) + xlab("") +
     coord_flip() +
@@ -110,7 +112,7 @@ plotFoldEnrichment <- function(de_res, pval, N){
 
   # Supplement with original data
   ld$pvals <- log10(bardat$pvals[matches])
-  ld$descr <- bardat$KEGG_PATHWAY_description[matches]
+  ld$descr <- bardat$KEGG_PATHWAY_description.x[matches]
   ld$vars <- bardat$variable[matches]
 
   ## Merge ld with df.ss

### 2019-11-20 R/plotFoldEnrichment.R
debugging travis build issue

@@ -85,13 +85,14 @@ plotFoldEnrichment <- function(de_res, pval, N){
     dplyr::arrange(.data$pvals)
 
   bardat <- merge(bardat.tmp, df, by = "KEGG_PATHWAY_ID")
+  colnames(bardat)[2] <- "KEGG_PATHWAY_description"
 
   ###########################################################
   # Generate plot
   ###########################################################
   # library(ggnewscale)
   # First, we'll make a plot and save it as a variable
-  g <- ggplot(bardat, aes(x=stats::reorder(.data$KEGG_PATHWAY_description.x, -.data$diff_enrich_adjusted), y=.data$value)) +
+  g <- ggplot(bardat, aes(x=stats::reorder(.data$KEGG_PATHWAY_description, -.data$diff_enrich_adjusted), y=.data$value)) +
     geom_bar(stat="identity", aes(col=.data$variable, group=.data$variable, fill=.data$pvals), position="dodge") +
     ylim(0, max(bardat$value) + 0.6) + xlab("") +
     coord_flip() +
@@ -112,7 +113,7 @@ plotFoldEnrichment <- function(de_res, pval, N){
 
   # Supplement with original data
   ld$pvals <- log10(bardat$pvals[matches])
-  ld$descr <- bardat$KEGG_PATHWAY_description.x[matches]
+  ld$descr <- bardat$KEGG_PATHWAY_description[matches]
   ld$vars <- bardat$variable[matches]
 
   ## Merge ld with df.ss

### 2019-11-21 R/plotFoldEnrichment.R
dbtb

@@ -84,8 +84,7 @@ plotFoldEnrichment <- function(de_res, pval, N){
                   pvals = pvals$value) %>%
     dplyr::arrange(.data$pvals)
 
-  bardat <- merge(bardat.tmp, df, by = "KEGG_PATHWAY_ID")
-  colnames(bardat)[2] <- "KEGG_PATHWAY_description"
+  bardat <- merge(bardat.tmp, df[, c(1,8)], by = "KEGG_PATHWAY_ID")
 
   ###########################################################
   # Generate plot

### 2019-11-21 R/plotFoldEnrichment.R
revert back to old version

@@ -79,19 +79,19 @@ plotFoldEnrichment <- function(de_res, pval, N){
   pvals <- subset(df.ss, df.ss$variable %in% c("enrich_p_list1", "enrich_p_list2"))
 
   ## Generate data set to be used for plotting
-  bardat.tmp <- subset(df.ss, df.ss$variable %in% c("fold_enrichment_list1", "fold_enrichment_list2")) %>%
+  bardat <- subset(df.ss, df.ss$variable %in% c("fold_enrichment_list1", "fold_enrichment_list2")) %>%
     dplyr::mutate(alpha = log10(pvals$value),
                   pvals = pvals$value) %>%
     dplyr::arrange(.data$pvals)
 
-  bardat <- merge(bardat.tmp, df[, c(1,8)], by = "KEGG_PATHWAY_ID")
+  # bardat <- merge(bardat.tmp, df[, c(1,8)], by = "KEGG_PATHWAY_ID")
 
   ###########################################################
   # Generate plot
   ###########################################################
   # library(ggnewscale)
   # First, we'll make a plot and save it as a variable
-  g <- ggplot(bardat, aes(x=stats::reorder(.data$KEGG_PATHWAY_description, -.data$diff_enrich_adjusted), y=.data$value)) +
+  g <- ggplot(bardat, aes(x=stats::reorder(.data$KEGG_PATHWAY_description, -.data$pvals), y=.data$value)) +
     geom_bar(stat="identity", aes(col=.data$variable, group=.data$variable, fill=.data$pvals), position="dodge") +
     ylim(0, max(bardat$value) + 0.6) + xlab("") +
     coord_flip() +

### 2019-11-21 R/plotFoldEnrichment.R
dbtb

@@ -79,19 +79,19 @@ plotFoldEnrichment <- function(de_res, pval, N){
   pvals <- subset(df.ss, df.ss$variable %in% c("enrich_p_list1", "enrich_p_list2"))
 
   ## Generate data set to be used for plotting
-  bardat <- subset(df.ss, df.ss$variable %in% c("fold_enrichment_list1", "fold_enrichment_list2")) %>%
+  bardat.tmp <- subset(df.ss, df.ss$variable %in% c("fold_enrichment_list1", "fold_enrichment_list2")) %>%
     dplyr::mutate(alpha = log10(pvals$value),
                   pvals = pvals$value) %>%
     dplyr::arrange(.data$pvals)
 
-  # bardat <- merge(bardat.tmp, df[, c(1,8)], by = "KEGG_PATHWAY_ID")
+  bardat <- merge(bardat.tmp, df[, c(1,8)], by = "KEGG_PATHWAY_ID")
 
   ###########################################################
   # Generate plot
   ###########################################################
   # library(ggnewscale)
   # First, we'll make a plot and save it as a variable
-  g <- ggplot(bardat, aes(x=stats::reorder(.data$KEGG_PATHWAY_description, -.data$pvals), y=.data$value)) +
+  g <- ggplot(bardat, aes(x=stats::reorder(.data$KEGG_PATHWAY_description, -.data$diff_enrich_adjusted), y=.data$value)) +
     geom_bar(stat="identity", aes(col=.data$variable, group=.data$variable, fill=.data$pvals), position="dodge") +
     ylim(0, max(bardat$value) + 0.6) + xlab("") +
     coord_flip() +

### 2019-11-22 jss-article-rnw-diffEnrich/article.tex
updated version number in NEWS

@@
 
-{\ttfamily\noindent\itshape\color{messagecolor}{\#\# These files already exist in your working directory. New files will not be generated.}}
+{\ttfamily\noindent\itshape\color{messagecolor}{\#\# 3 data sets will be written as tab delimited text files}}
 
-{\ttfamily\noindent\itshape\color{messagecolor}{\#\# Kegg Release: Release\_92.0+\_11-20\_Nov\_19}}\end{kframe}
+{\ttfamily\noindent\itshape\color{messagecolor}{\#\# File location: /Users/harry/Documents/Saba\_Lab/diffEnrich}}
+
+{\ttfamily\noindent\itshape\color{messagecolor}{\#\# Kegg Release: Release\_92.0+\_11-21\_Nov\_19}}\end{kframe}
 \end{knitrout}
@@
 
-{\ttfamily\noindent\itshape\color{messagecolor}{\#\# Kegg Release: Release\_92.0+\_11-20\_Nov\_19}}\end{kframe}
+{\ttfamily\noindent\itshape\color{messagecolor}{\#\# Kegg Release: Release\_92.0+\_11-21\_Nov\_19}}\end{kframe}
 \end{knitrout}
@@
 
-{\ttfamily\noindent\itshape\color{messagecolor}{\#\# ncbi\_to\_kegg2019-11-20Release\_92.0+\_11-20\_Nov\_19.txt}}
+{\ttfamily\noindent\itshape\color{messagecolor}{\#\# ncbi\_to\_kegg2019-11-21Release\_92.0+\_11-21\_Nov\_19.txt}}
 
-{\ttfamily\noindent\itshape\color{messagecolor}{\#\# kegg\_to\_pathway2019-11-20Release\_92.0+\_11-20\_Nov\_19.txt}}
+{\ttfamily\noindent\itshape\color{messagecolor}{\#\# kegg\_to\_pathway2019-11-21Release\_92.0+\_11-21\_Nov\_19.txt}}
 
-{\ttfamily\noindent\itshape\color{messagecolor}{\#\# pathway\_to\_species2019-11-20Release\_92.0+\_11-20\_Nov\_19.txt}}
+{\ttfamily\noindent\itshape\color{messagecolor}{\#\# pathway\_to\_species2019-11-21Release\_92.0+\_11-21\_Nov\_19.txt}}
 
-{\ttfamily\noindent\itshape\color{messagecolor}{\#\# File location: /Users/smiharry/Documents/packages/diffEnrich}}\end{kframe}
+{\ttfamily\noindent\itshape\color{messagecolor}{\#\# File location: /Users/harry/Documents/Saba\_Lab/diffEnrich}}\end{kframe}
 \end{knitrout}

### 2019-11-22 jss-article-rnw-diffEnrich/article.Rnw
Formatting table 1 and added sup table 1 to .Rnw

@@
 \usepackage[section]{placeins}
+\usepackage{tabularx}
+\usepackage{graphicx}
+\usepackage{adjustbox}
+\newcommand\mc[1]{\multicolumn{1}{c}{#1}}
 
@@
   \centering
-\begin{tabular}{lllp{7.4cm}}
+  \caption{\label{tab:description} Description of the data sets retrieved by \code{get\textunderscore kegg}'s
+connection to the KEGG REST API.}
+  \begin{adjustbox}{width=\textwidth}
+  \begin{tabular}{rr}
 \hline
@@
 \end{tabular}
-\caption{\label{tab:description} Description of the data sets retrieved by \code{get\textunderscore kegg}'s
-connection to the KEGG REST API.}
+\end{adjustbox}
 \end{table}
@@
 the expected number of genes from the gene list to be annotated to the KEGG
-pathway by chance. An example of the first 5 results generated by \code{pathEnrich}
-are in Table 2. For a detailed description of the variables in this table see Table 3.
+pathway by chance. An example of the first 6 results generated by \code{pathEnrich}
+are in Supplementary Table 1. For a detailed description of the variables in this table see Table 3.
+
+<<list1_head, echo = FALSE, eval = FALSE>>=
+library(xtable)
+tb2 <- head(list1_pe$enrich_table)
+xtable(tb2, caption = "Table 2. First 6 results generated by pathEnrich.")
+@
 
@@
 
+%% -----Supplementary Information------------------------------------
+\section[Supplementary Information]{Supplementary Information} \label{sec:supinfo}
+
+\begin{table}[h]
+  \centering
+  \caption{Supplementary Table 1. First 6 results generated by pathEnrich.}
+  \tiny
+  \begin{adjustbox}{width=\textwidth}
+  \begin{tabular}{rllrrrrrrrr}
+  \hline
+ & KEGG\_PATHWAY\_ID & KEGG\_PATHWAY\_description & KEGG\_PATHWAY\_cnt & KEGG\_PATHWAY\_in\_list & KEGG\_DATABASE\_cnt & KEG\_DATABASE\_in\_list & expected & enrich\_p & p\_adj & fold\_enrichment \\
+  \hline
+95 & rno04530 & Tight junction - Rattus norvegicus (rat) & 170 & 19.00 & 8856 & 295 & 5.66 & 0.00 & 0.00 & 3.36 \\
+  172 & rno05135 & Yersinia infection - Rattus norvegicus (rat) & 128 & 16.00 & 8856 & 295 & 4.26 & 0.00 & 0.00 & 3.75 \\
+  194 & rno05210 & Colorectal cancer - Rattus norvegicus (rat) &  88 & 12.00 & 8856 & 295 & 2.93 & 0.00 & 0.00 & 4.09 \\
+  212 & rno05231 & Choline metabolism in cancer - Rattus norvegicus (rat) &  99 & 12.00 & 8856 & 295 & 3.30 & 0.00 & 0.00 & 3.64 \\
+  197 & rno05213 & Endometrial cancer - Rattus norvegicus (rat) &  58 & 9.00 & 8856 & 295 & 1.93 & 0.00 & 0.00 & 4.66 \\
+  66 & rno04144 & Endocytosis - Rattus norvegicus (rat) & 275 & 22.00 & 8856 & 295 & 9.16 & 0.00 & 0.00 & 2.40 \\
+   \hline
+  \end{tabular}
+\end{adjustbox}
+\end{table}
+
 \bibliography{refs}
 
-%% -----------------------------------------------------------------------------
+%% ----------------------------------------------------------------------
 

### 2019-11-22 jss-article-rnw-diffEnrich/article.tex
Formatting table 1 and added sup table 1 to .Rnw

@@
 
-{\ttfamily\noindent\itshape\color{messagecolor}{\#\# 3 data sets will be written as tab delimited text files}}
-
-{\ttfamily\noindent\itshape\color{messagecolor}{\#\# File location: /Users/harry/Documents/Saba\_Lab/diffEnrich}}
+{\ttfamily\noindent\itshape\color{messagecolor}{\#\# These files already exist in your working directory. New files will not be generated.}}
 
@@
 the expected number of genes from the gene list to be annotated to the KEGG
-pathway by chance. An example of the first 5 results generated by \code{pathEnrich}
-are in Table 2. For a detailed description of the variables in this table see Table 3.
+pathway by chance. An example of the first 6 results generated by \code{pathEnrich}
+are in Supplementary Table 1. For a detailed description of the variables in this table see Table 3.
 

### 2019-11-22 jss-article-rnw-diffEnrich/article.Rnw
Added sup table 2

@@
 \usepackage{adjustbox}
+\usepackage{makecell}
 \newcommand\mc[1]{\multicolumn{1}{c}{#1}}
@@
 pathway by chance. An example of the first 6 results generated by \code{pathEnrich}
-are in Supplementary Table 1. For a detailed description of the variables in this table see Table 3.
-
-<<list1_head, echo = FALSE, eval = FALSE>>=
-library(xtable)
-tb2 <- head(list1_pe$enrich_table)
-xtable(tb2, caption = "Table 2. First 6 results generated by pathEnrich.")
-@
+are in Supplementary Table 1. For a detailed description of the variables in this table see Supplementary Table 2.
 
@@
 
-\begin{table}[h]
+<<list1_head, echo = FALSE, eval = FALSE>>=
+library(xtable)
+suptb1 <- head(list1_pe$enrich_table)
+xtable(suptb1)
+@
+
+\begin{table}[ht]
   \centering
@@
 
+<<pathEnrich_description, echo=FALSE, eval = FALSE>>=
+# get column names
+names <- colnames(list1_pe$enrich_table)
+# Define column names
+description <- c("KEGG Pathway Identifier",
+                 "Description of KEGG Pathway (provided by KEGG)",
+                 "Number of Genes in KEGG Pathway",
+                 "Number of Genes from gene list in KEGG Pathway",
+                 "Number of Genes in KEGG Database",
+                 "Number of Genes from gene list in KEGG Database",
+                 "Expected number of genes from list to be in KEGG pathway by chance",
+                 "P-value for enrichment within the KEGG pathway for list genes",
+                 "Multiple testing adjusted enrichment p-values (default = False Discovery Rate (Benjamini and Hochberg, 1995))",
+                 "Ratio of number of genes observed from the gene list annotated to the KEGG pathway to the number of genes expected from the gene list to be annotated to the KEGG pathway if there was no enrichment (i.e.
+KEGG_PATHWAY_in_list/expected)")
+# Make dataframe
+suptb2 <- data.frame(names, description)
+colnames(suptb2) <- c("Column Names", "Column Description")
+
+## Generate xtable
+xtable(suptb2)
+@
+
+\begin{table}[h]
+\centering
+  \caption{Supplementary Table 2. Description of columns is dataframe generated by pathEnrich.}
+  \begin{adjustbox}{width=\textwidth}
+  \begin{tabular}{rll}
+  \hline
+& Column Names & Column Description \\
+  \hline
+& KEGG\_PATHWAY\_ID & \makecell{KEGG Pathway Identifier \\} \\
+& & \\
+& KEGG\_PATHWAY\_description & \makecell{Description of KEGG Pathway \\ (provided by KEGG) \\ } \\
+  & & \\
+& KEGG\_PATHWAY\_cnt & \makecell{Number of Genes in KEGG Pathway} \\
+& & \\
+& KEGG\_PATHWAY\_in\_list & \makecell{Number of Genes from  \\ gene list in KEGG Pathway} \\
+& & \\
+& KEGG\_DATABASE\_cnt & \makecell{Number of Genes in KEGG Database} \\
+& & \\
+& KEG\_DATABASE\_in\_list & \makecell{Number of Genes from \\ gene list in KEGG Database} \\
+& & \\
+& expected & \makecell{Expected number of genes \\ from list to be in \\ KEGG pathway by chance} \\
+& & \\
+& enrich\_p & \makecell{P-value for enrichment within \\ the KEGG pathway for list genes} \\
+& & \\
+& p\_adj & \makecell{Multiple testing adjusted enrichment \\ p-values (default = False Discovery Rate \\ (Benjamini and Hochberg, 1995))} \\
+& & \\
+& fold\_enrichment & \makecell{Ratio of number of genes observed  \\ from the gene list annotated to the \\ KEGG pathway to the number of genes  \\ expected from the gene list to  \\ be annotated to the KEGG pathway if  \\ there was no enrichment \\ (i.e.
+KEGG\_PATHWAY\_in\_list/expected)} \\
+   \hline
+  \end{tabular}
+\end{adjustbox}
+\end{table}
+
+%% ----------------------------------------------------------------------
 \bibliography{refs}

### 2019-11-22 jss-article-rnw-diffEnrich/article.tex
Added sup table 2

@@
 
-{\ttfamily\noindent\itshape\color{messagecolor}{\#\# These files already exist in your working directory. New files will not be generated.}}
+{\ttfamily\noindent\itshape\color{messagecolor}{\#\# 3 data sets will be written as tab delimited text files}}
 
-{\ttfamily\noindent\itshape\color{messagecolor}{\#\# Kegg Release: Release\_92.0+\_11-21\_Nov\_19}}\end{kframe}
+{\ttfamily\noindent\itshape\color{messagecolor}{\#\# File location: /Users/harry/Documents/Saba\_Lab/diffEnrich}}
+
+{\ttfamily\noindent\itshape\color{messagecolor}{\#\# Kegg Release: Release\_92.0+\_11-22\_Nov\_19}}\end{kframe}
 \end{knitrout}
@@
 
-{\ttfamily\noindent\itshape\color{messagecolor}{\#\# Kegg Release: Release\_92.0+\_11-21\_Nov\_19}}\end{kframe}
+{\ttfamily\noindent\itshape\color{messagecolor}{\#\# Kegg Release: Release\_92.0+\_11-22\_Nov\_19}}\end{kframe}
 \end{knitrout}
@@
 
-{\ttfamily\noindent\itshape\color{messagecolor}{\#\# ncbi\_to\_kegg2019-11-21Release\_92.0+\_11-21\_Nov\_19.txt}}
+{\ttfamily\noindent\itshape\color{messagecolor}{\#\# ncbi\_to\_kegg2019-11-21Release\_92.0+\_11-22\_Nov\_19.txt}}
 
-{\ttfamily\noindent\itshape\color{messagecolor}{\#\# kegg\_to\_pathway2019-11-21Release\_92.0+\_11-21\_Nov\_19.txt}}
+{\ttfamily\noindent\itshape\color{messagecolor}{\#\# kegg\_to\_pathway2019-11-21Release\_92.0+\_11-22\_Nov\_19.txt}}
 
-{\ttfamily\noindent\itshape\color{messagecolor}{\#\# pathway\_to\_species2019-11-21Release\_92.0+\_11-21\_Nov\_19.txt}}
+{\ttfamily\noindent\itshape\color{messagecolor}{\#\# pathway\_to\_species2019-11-21Release\_92.0+\_11-22\_Nov\_19.txt}}
 
@@
 pathway by chance. An example of the first 6 results generated by \code{pathEnrich}
-are in Supplementary Table 1. For a detailed description of the variables in this table see Table 3.
+are in Supplementary Table 1. For a detailed description of the variables in this table see Supplementary Table 2.
 
@@
 
-\begin{table}[h]
+\begin{table}[ht]
   \centering
@@
 
+\begin{table}[h]
+\centering
+  \caption{Supplementary Table 2. Description of columns is dataframe generated by pathEnrich.}
+  \begin{adjustbox}{width=\textwidth}
+  \begin{tabular}{rll}
+  \hline
+& Column Names & Column Description \\
+  \hline
+& KEGG\_PATHWAY\_ID & \makecell{KEGG Pathway Identifier \\} \\
+& & \\
+& KEGG\_PATHWAY\_description & \makecell{Description of KEGG Pathway \\ (provided by KEGG) \\ } \\
+  & & \\
+& KEGG\_PATHWAY\_cnt & \makecell{Number of Genes in KEGG Pathway} \\
+& & \\
+& KEGG\_PATHWAY\_in\_list & \makecell{Number of Genes from  \\ gene list in KEGG Pathway} \\
+& & \\
+& KEGG\_DATABASE\_cnt & \makecell{Number of Genes in KEGG Database} \\
+& & \\
+& KEG\_DATABASE\_in\_list & \makecell{Number of Genes from \\ gene list in KEGG Database} \\
+& & \\
+& expected & \makecell{Expected number of genes \\ from list to be in \\ KEGG pathway by chance} \\
+& & \\
+& enrich\_p & \makecell{P-value for enrichment within \\ the KEGG pathway for list genes} \\
+& & \\
+& p\_adj & \makecell{Multiple testing adjusted enrichment \\ p-values (default = False Discovery Rate \\ (Benjamini and Hochberg, 1995))} \\
+& & \\
+& fold\_enrichment & \makecell{Ratio of number of genes observed  \\ from the gene list annotated to the \\ KEGG pathway to the number of genes  \\ expected from the gene list to  \\ be annotated to the KEGG pathway if  \\ there was no enrichment \\ (i.e.
+KEGG\_PATHWAY\_in\_list/expected)} \\
+   \hline
+  \end{tabular}
+\end{adjustbox}
+\end{table}
+
+%% ----------------------------------------------------------------------
 \bibliography{refs}

### 2019-11-22 jss-article-rnw-diffEnrich/article.Rnw
Formatted table 2 in .Rnw

@@
 \usepackage{makecell}
+\usepackage{float}
+\usepackage{longtable}
 \newcommand\mc[1]{\multicolumn{1}{c}{#1}}
@@
 pathway by chance. An example of the first 6 results generated by \code{pathEnrich}
-are in Supplementary Table 1. For a detailed description of the variables in this table see Supplementary Table 2.
-
-S3 generic functions for \code{print} and \code{summary} are provided.
-The \code{print} function prints the results table as a \code{tibble}, and the
-\code{summary} function returns the number of pathways that reached statistical
-significance as well as their descriptions, the number of genes used from the
-KEGG data base, the KEGG species, and the method used for multiple testing
-correction.
-
-<<path_summary, echo = TRUE>>=
-summary(list1_pe)
-@
-
-%% -----Supplementary Information------------------------------------
-\section[Supplementary Information]{Supplementary Information} \label{sec:supinfo}
-
-<<list1_head, echo = FALSE, eval = FALSE>>=
-library(xtable)
-suptb1 <- head(list1_pe$enrich_table)
-xtable(suptb1)
-@
-
-\begin{table}[ht]
-  \centering
-  \caption{Supplementary Table 1. First 6 results generated by pathEnrich.}
-  \tiny
-  \begin{adjustbox}{width=\textwidth}
-  \begin{tabular}{rllrrrrrrrr}
-  \hline
- & KEGG\_PATHWAY\_ID & KEGG\_PATHWAY\_description & KEGG\_PATHWAY\_cnt & KEGG\_PATHWAY\_in\_list & KEGG\_DATABASE\_cnt & KEG\_DATABASE\_in\_list & expected & enrich\_p & p\_adj & fold\_enrichment \\
-  \hline
-95 & rno04530 & Tight junction - Rattus norvegicus (rat) & 170 & 19.00 & 8856 & 295 & 5.66 & 0.00 & 0.00 & 3.36 \\
-  172 & rno05135 & Yersinia infection - Rattus norvegicus (rat) & 128 & 16.00 & 8856 & 295 & 4.26 & 0.00 & 0.00 & 3.75 \\
-  194 & rno05210 & Colorectal cancer - Rattus norvegicus (rat) &  88 & 12.00 & 8856 & 295 & 2.93 & 0.00 & 0.00 & 4.09 \\
-  212 & rno05231 & Choline metabolism in cancer - Rattus norvegicus (rat) &  99 & 12.00 & 8856 & 295 & 3.30 & 0.00 & 0.00 & 3.64 \\
-  197 & rno05213 & Endometrial cancer - Rattus norvegicus (rat) &  58 & 9.00 & 8856 & 295 & 1.93 & 0.00 & 0.00 & 4.66 \\
-  66 & rno04144 & Endocytosis - Rattus norvegicus (rat) & 275 & 22.00 & 8856 & 295 & 9.16 & 0.00 & 0.00 & 2.40 \\
-   \hline
-  \end{tabular}
-\end{adjustbox}
-\end{table}
+are in Supplementary Table 1. For a detailed description of the variables in this table see Table 2.
 
@@
 
-\begin{table}[h]
+%\begin{table}[hp!]
+%\centering
+\begin{longtable}{cc}
+  \caption{Description of columns is dataframe generated by pathEnrich.}
+  % \begin{adjustbox}{width=\textwidth}
+  % \begin{tabular}{rll}
+  \hline
+Column Names & Column Description \\
+  \hline
+  \endhead
+  \hline
+  \endfoot
+KEGG\_PATHWAY\_ID & \makecell{KEGG Pathway Identifier \\} \\
+& & \\
+KEGG\_PATHWAY\_description & \makecell{Description of KEGG Pathway \\ (provided by KEGG) \\ } \\
+  & & \\
+KEGG\_PATHWAY\_cnt & \makecell{Number of Genes in KEGG Pathway} \\
+& & \\
+KEGG\_PATHWAY\_in\_list & \makecell{Number of Genes from  \\ gene list in KEGG Pathway} \\
+& & \\
+KEGG\_DATABASE\_cnt & \makecell{Number of Genes in KEGG Database} \\
+& & \\
+KEGG\_DATABASE\_in\_list & \makecell{Number of Genes from \\ gene list in KEGG Database} \\
+& & \\
+expected & \makecell{Expected number of genes \\ from list to be in \\ KEGG pathway by chance} \\
+& & \\
+enrich\_p & \makecell{P-value for enrichment within \\ the KEGG pathway for list genes} \\
+& & \\
+p\_adj & \makecell{Multiple testing adjusted enrichment \\ p-values (default = False Discovery Rate \\ (Benjamini and Hochberg, 1995))} \\
+& & \\
+fold\_enrichment & \makecell{Ratio of number of genes observed  \\ from the gene list annotated to the \\ KEGG pathway to the number of genes  \\ expected from the gene list to  \\ be annotated to the KEGG pathway if  \\ there was no enrichment \\ (i.e.
+KEGG\_PATHWAY\_in\_list/expected)} \\
+  % \end{tabular}
+% \end{adjustbox}
+\end{longtable}
+%\end{table}
+
+S3 generic functions for \code{print} and \code{summary} are provided.
+The \code{print} function prints the results table as a \code{tibble}, and the
+\code{summary} function returns the number of pathways that reached statistical
+significance as well as their descriptions, the number of genes used from the
+KEGG data base, the KEGG species, and the method used for multiple testing
+correction. See supplementary information for examples.
+
+\subsection{\code{diffEnrich}: Identify differentially enriched KEGG pathways.}
+
+The \code{diffEnrich} function merges results from the \code{pathEnrich} function
+generated in section 2.2. This merged data set is then used to perform differential
+enrichment using a Fisher’s exact test as described in 2.2. The resulting odds
+ratio is defined as the odds of a gene from list 2 belonging to a given KEGG
+pathway divided by the odds of a gene from list 1 belonging to a given KEGG
+pathway. Users have the same options for multiple testing methods that are
+provided in the \code{pathEnrich} function. KEGG pathways that do not contain
+any genes from either gene list (e.g. ‘rno04530’ contains 0 genes from list 1
+and 0 genes from list 2) are removed from the analysis. If this is the case a
+warning will be printed that tells the user how many pathways were removed.
+This can be avoided by setting the 'N' parameter to a value > 0 in the
+\code{pathEnrich calls}. This diffEnrich function generates a table that
+contains the results from the analyses performed in section 2.2 for each gene
+list as well as odds ratios and their associated unadjusted and adjusted
+p-values for each KEGG pathway. For a detailed description of the results
+generated by \code{diffEnrich} see Table 3, and an example of the first 6
+results generated by \code{diffEnrich} are in Supplementary Table 2.
+
+%% -----Supplementary Information------------------------------------
+\section[Supplementary Information]{Supplementary Information} \label{sec:supinfo}
+
+<<list1_head, echo = FALSE, eval = FALSE>>=
+library(xtable)
+suptb1 <- head(list1_pe$enrich_table)
+xtable(suptb1)
+@
+
+\begin{table}[H]
   \centering
-  \caption{Supplementary Table 2. Description of columns is dataframe generated by pathEnrich.}
+  \caption{Supplementary Table 1. First 6 results generated by pathEnrich.}
+  \tiny
   \begin{adjustbox}{width=\textwidth}
-  \begin{tabular}{rll}
+  \begin{tabular}{rllrrrrrrrr}
   \hline
-& Column Names & Column Description \\
+ & KEGG\_PATHWAY\_ID & KEGG\_PATHWAY\_description & KEGG\_PATHWAY\_cnt & KEGG\_PATHWAY\_in\_list & KEGG\_DATABASE\_cnt & KEG\_DATABASE\_in\_list & expected & enrich\_p & p\_adj & fold\_enrichment \\
   \hline
-& KEGG\_PATHWAY\_ID & \makecell{KEGG Pathway Identifier \\} \\
-& & \\
-& KEGG\_PATHWAY\_description & \makecell{Description of KEGG Pathway \\ (provided by KEGG) \\ } \\
-  & & \\
-& KEGG\_PATHWAY\_cnt & \makecell{Number of Genes in KEGG Pathway} \\
-& & \\
-& KEGG\_PATHWAY\_in\_list & \makecell{Number of Genes from  \\ gene list in KEGG Pathway} \\
-& & \\
-& KEGG\_DATABASE\_cnt & \makecell{Number of Genes in KEGG Database} \\
-& & \\
-& KEG\_DATABASE\_in\_list & \makecell{Number of Genes from \\ gene list in KEGG Database} \\
-& & \\
-& expected & \makecell{Expected number of genes \\ from list to be in \\ KEGG pathway by chance} \\
-& & \\
-& enrich\_p & \makecell{P-value for enrichment within \\ the KEGG pathway for list genes} \\
-& & \\
-& p\_adj & \makecell{Multiple testing adjusted enrichment \\ p-values (default = False Discovery Rate \\ (Benjamini and Hochberg, 1995))} \\
-& & \\
-& fold\_enrichment & \makecell{Ratio of number of genes observed  \\ from the gene list annotated to the \\ KEGG pathway to the number of genes  \\ expected from the gene list to  \\ be annotated to the KEGG pathway if  \\ there was no enrichment \\ (i.e.
-KEGG\_PATHWAY\_in\_list/expected)} \\
+95 & rno04530 & Tight junction - Rattus norvegicus (rat) & 170 & 19.00 & 8856 & 295 & 5.66 & 0.00 & 0.00 & 3.36 \\
+  172 & rno05135 & Yersinia infection - Rattus norvegicus (rat) & 128 & 16.00 & 8856 & 295 & 4.26 & 0.00 & 0.00 & 3.75 \\
+  194 & rno05210 & Colorectal cancer - Rattus norvegicus (rat) &  88 & 12.00 & 8856 & 295 & 2.93 & 0.00 & 0.00 & 4.09 \\
+  212 & rno05231 & Choline metabolism in cancer - Rattus norvegicus (rat) &  99 & 12.00 & 8856 & 295 & 3.30 & 0.00 & 0.00 & 3.64 \\
+  197 & rno05213 & Endometrial cancer - Rattus norvegicus (rat) &  58 & 9.00 & 8856 & 295 & 1.93 & 0.00 & 0.00 & 4.66 \\
+  66 & rno04144 & Endocytosis - Rattus norvegicus (rat) & 275 & 22.00 & 8856 & 295 & 9.16 & 0.00 & 0.00 & 2.40 \\
    \hline
@@
 \end{table}
+
+Example of S3 functions
+<<path_summary, echo = TRUE>>=
+summary(list1_pe)
+@
 

### 2019-11-22 jss-article-rnw-diffEnrich/article.tex
Formatted table 2 in .Rnw

--- 2019-11-22 jss-article-rnw-diffEnrich/article.tex
@@
 \usepackage{adjustbox}
+\usepackage{makecell}
+\usepackage{float}
+\usepackage{longtable}
 \newcommand\mc[1]{\multicolumn{1}{c}{#1}}
@@
 
-{\ttfamily\noindent\itshape\color{messagecolor}{\#\# ncbi\_to\_kegg2019-11-21Release\_92.0+\_11-21\_Nov\_19.txt}}
+{\ttfamily\noindent\itshape\color{messagecolor}{\#\# ncbi\_to\_kegg2019-11-22Release\_92.0+\_11-22\_Nov\_19.txt}}
 
-{\ttfamily\noindent\itshape\color{messagecolor}{\#\# kegg\_to\_pathway2019-11-21Release\_92.0+\_11-21\_Nov\_19.txt}}
+{\ttfamily\noindent\itshape\color{messagecolor}{\#\# kegg\_to\_pathway2019-11-22Release\_92.0+\_11-22\_Nov\_19.txt}}
 
-{\ttfamily\noindent\itshape\color{messagecolor}{\#\# pathway\_to\_species2019-11-21Release\_92.0+\_11-21\_Nov\_19.txt}}
+{\ttfamily\noindent\itshape\color{messagecolor}{\#\# pathway\_to\_species2019-11-22Release\_92.0+\_11-22\_Nov\_19.txt}}
 
@@
 pathway by chance. An example of the first 6 results generated by \code{pathEnrich}
-are in Supplementary Table 1. For a detailed description of the variables in this table see Table 3.
+are in Supplementary Table 1. For a detailed description of the variables in this table see Table 2.
+
+%\begin{table}[hp!]
+%\centering
+\begin{longtable}{cc}
+  \caption{Description of columns is dataframe generated by pathEnrich.}
+  % \begin{adjustbox}{width=\textwidth}
+  % \begin{tabular}{rll}
+  \hline
+Column Names & Column Description \\
+  \hline
+  \endhead
+  \hline
+  \endfoot
+KEGG\_PATHWAY\_ID & \makecell{KEGG Pathway Identifier \\} \\
+& & \\
+KEGG\_PATHWAY\_description & \makecell{Description of KEGG Pathway \\ (provided by KEGG) \\ } \\
+  & & \\
+KEGG\_PATHWAY\_cnt & \makecell{Number of Genes in KEGG Pathway} \\
+& & \\
+KEGG\_PATHWAY\_in\_list & \makecell{Number of Genes from  \\ gene list in KEGG Pathway} \\
+& & \\
+KEGG\_DATABASE\_cnt & \makecell{Number of Genes in KEGG Database} \\
+& & \\
+KEGG\_DATABASE\_in\_list & \makecell{Number of Genes from \\ gene list in KEGG Database} \\
+& & \\
+expected & \makecell{Expected number of genes \\ from list to be in \\ KEGG pathway by chance} \\
+& & \\
+enrich\_p & \makecell{P-value for enrichment within \\ the KEGG pathway for list genes} \\
+& & \\
+p\_adj & \makecell{Multiple testing adjusted enrichment \\ p-values (default = False Discovery Rate \\ (Benjamini and Hochberg, 1995))} \\
+& & \\
+fold\_enrichment & \makecell{Ratio of number of genes observed  \\ from the gene list annotated to the \\ KEGG pathway to the number of genes  \\ expected from the gene list to  \\ be annotated to the KEGG pathway if  \\ there was no enrichment \\ (i.e.
+KEGG\_PATHWAY\_in\_list/expected)} \\
+  % \end{tabular}
+% \end{adjustbox}
+\end{longtable}
+%\end{table}
 
@@
 KEGG data base, the KEGG species, and the method used for multiple testing
-correction.
+correction. See supplementary information for examples.
 
+\subsection{\code{diffEnrich}: Identify differentially enriched KEGG pathways.}
+
+The \code{diffEnrich} function merges results from the \code{pathEnrich} function
+generated in section 2.2. This merged data set is then used to perform differential
+enrichment using a Fisher’s exact test as described in 2.2. The resulting odds
+ratio is defined as the odds of a gene from list 2 belonging to a given KEGG
+pathway divided by the odds of a gene from list 1 belonging to a given KEGG
+pathway. Users have the same options for multiple testing methods that are
+provided in the \code{pathEnrich} function. KEGG pathways that do not contain
+any genes from either gene list (e.g. ‘rno04530’ contains 0 genes from list 1
+and 0 genes from list 2) are removed from the analysis. If this is the case a
+warning will be printed that tells the user how many pathways were removed.
+This can be avoided by setting the 'N' parameter to a value > 0 in the
+\code{pathEnrich calls}. This diffEnrich function generates a table that
+contains the results from the analyses performed in section 2.2 for each gene
+list as well as odds ratios and their associated unadjusted and adjusted
+p-values for each KEGG pathway. For a detailed description of the results
+generated by \code{diffEnrich} see Table 3, and an example of the first 6
+results generated by \code{diffEnrich} are in Supplementary Table 2.
+
+%% -----Supplementary Information------------------------------------
+\section[Supplementary Information]{Supplementary Information} \label{sec:supinfo}
+
+\begin{table}[H]
+  \centering
+  \caption{Supplementary Table 1. First 6 results generated by pathEnrich.}
+  \tiny
+  \begin{adjustbox}{width=\textwidth}
+  \begin{tabular}{rllrrrrrrrr}
+  \hline
+ & KEGG\_PATHWAY\_ID & KEGG\_PATHWAY\_description & KEGG\_PATHWAY\_cnt & KEGG\_PATHWAY\_in\_list & KEGG\_DATABASE\_cnt & KEG\_DATABASE\_in\_list & expected & enrich\_p & p\_adj & fold\_enrichment \\
+  \hline
+95 & rno04530 & Tight junction - Rattus norvegicus (rat) & 170 & 19.00 & 8856 & 295 & 5.66 & 0.00 & 0.00 & 3.36 \\
+  172 & rno05135 & Yersinia infection - Rattus norvegicus (rat) & 128 & 16.00 & 8856 & 295 & 4.26 & 0.00 & 0.00 & 3.75 \\
+  194 & rno05210 & Colorectal cancer - Rattus norvegicus (rat) &  88 & 12.00 & 8856 & 295 & 2.93 & 0.00 & 0.00 & 4.09 \\
+  212 & rno05231 & Choline metabolism in cancer - Rattus norvegicus (rat) &  99 & 12.00 & 8856 & 295 & 3.30 & 0.00 & 0.00 & 3.64 \\
+  197 & rno05213 & Endometrial cancer - Rattus norvegicus (rat) &  58 & 9.00 & 8856 & 295 & 1.93 & 0.00 & 0.00 & 4.66 \\
+  66 & rno04144 & Endocytosis - Rattus norvegicus (rat) & 275 & 22.00 & 8856 & 295 & 9.16 & 0.00 & 0.00 & 2.40 \\
+   \hline
+  \end{tabular}
+\end{adjustbox}
+\end{table}
+
+Example of S3 functions
 \begin{knitrout}
@@
 
-%% -----Supplementary Information------------------------------------
-\section[Supplementary Information]{Supplementary Information} \label{sec:supinfo}
-
-\begin{table}[h]
-  \centering
-  \caption{Supplementary Table 1. First 6 results generated by pathEnrich.}
-  \tiny
-  \begin{adjustbox}{width=\textwidth}
-  \begin{tabular}{rllrrrrrrrr}
-  \hline
- & KEGG\_PATHWAY\_ID & KEGG\_PATHWAY\_description & KEGG\_PATHWAY\_cnt & KEGG\_PATHWAY\_in\_list & KEGG\_DATABASE\_cnt & KEG\_DATABASE\_in\_list & expected & enrich\_p & p\_adj & fold\_enrichment \\
-  \hline
-95 & rno04530 & Tight junction - Rattus norvegicus (rat) & 170 & 19.00 & 8856 & 295 & 5.66 & 0.00 & 0.00 & 3.36 \\
-  172 & rno05135 & Yersinia infection - Rattus norvegicus (rat) & 128 & 16.00 & 8856 & 295 & 4.26 & 0.00 & 0.00 & 3.75 \\
-  194 & rno05210 & Colorectal cancer - Rattus norvegicus (rat) &  88 & 12.00 & 8856 & 295 & 2.93 & 0.00 & 0.00 & 4.09 \\
-  212 & rno05231 & Choline metabolism in cancer - Rattus norvegicus (rat) &  99 & 12.00 & 8856 & 295 & 3.30 & 0.00 & 0.00 & 3.64 \\
-  197 & rno05213 & Endometrial cancer - Rattus norvegicus (rat) &  58 & 9.00 & 8856 & 295 & 1.93 & 0.00 & 0.00 & 4.66 \\
-  66 & rno04144 & Endocytosis - Rattus norvegicus (rat) & 275 & 22.00 & 8856 & 295 & 9.16 & 0.00 & 0.00 & 2.40 \\
-   \hline
-  \end{tabular}
-\end{adjustbox}
-\end{table}
-
+%% ----------------------------------------------------------------------
 \bibliography{refs}

### 2019-11-22 jss-article-rnw-diffEnrich/article.Rnw
Added text for sec 2.3 in .Rnw

@@
 
+<<diffEnrich, echo=TRUE>>=
+## Perform differential enrichment
+diff_enrich <- diffEnrich(list1_pe = list1_pe,
+                          list2_pe = list2_pe,
+                          method = 'none',
+                          cutoff = 0.05)
+@
+
+The result of the \code{diffEnrich} call is a list object that
+contains a data frame with the estimated odds ratio generated
+by the Fisher's Exact test and the associated p-value. S3 generic
+functions for \code{print} and \code{summary} are provided. The
+\code{print} function prints the results table as a \code{tibble},
+and the \code{summary} function returns the number of pathways
+that reached statistical significance as well as their descriptions,
+the number of genes used from the KEGG database, the KEGG species,
+the number of pathways that were shared (i.e. had at least N gene
+from each gene list present in the pathway based on what the user
+chose for N in \code{pathEnrich}) between the gene lists and the
+method used for multiple testing correction.
+
 %% -----Supplementary Information------------------------------------

### 2019-11-22 jss-article-rnw-diffEnrich/article.tex
Added text for sec 2.3 in .Rnw

@@
 
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{alltt}
+\hlcom{## Perform differential enrichment}
+\hlstd{diff_enrich} \hlkwb{<-} \hlkwd{diffEnrich}\hlstd{(}\hlkwc{list1_pe} \hlstd{= list1_pe,}
+                          \hlkwc{list2_pe} \hlstd{= list2_pe,}
+                          \hlkwc{method} \hlstd{=} \hlstr{'none'}\hlstd{,}
+                          \hlkwc{cutoff} \hlstd{=} \hlnum{0.05}\hlstd{)}
+\end{alltt}
+\end{kframe}
+\end{knitrout}
+
+The result of the \code{diffEnrich} call is a list object that
+contains a data frame with the estimated odds ratio generated
+by the Fisher's Exact test and the associated p-value. S3 generic
+functions for \code{print} and \code{summary} are provided. The
+\code{print} function prints the results table as a \code{tibble},
+and the \code{summary} function returns the number of pathways
+that reached statistical significance as well as their descriptions,
+the number of genes used from the KEGG database, the KEGG species,
+the number of pathways that were shared (i.e. had at least N gene
+from each gene list present in the pathway based on what the user
+chose for N in \code{pathEnrich}) between the gene lists and the
+method used for multiple testing correction.
+
 %% -----Supplementary Information------------------------------------

### 2019-11-22 jss-article-rnw-diffEnrich/article.Rnw
Added table 3 to .Rnw

@@
 <<pathEnrich_description, echo=FALSE, eval = FALSE>>=
+library(xtable)
 # get column names
@@
 \begin{longtable}{cc}
-  \caption{Description of columns is dataframe generated by pathEnrich.}
+  \caption{Description of columns in dataframe generated by pathEnrich.}
   % \begin{adjustbox}{width=\textwidth}
@@
 
+<<diffEnrich_descriptions, echo=FALSE, eval=FALSE>>=
+## get vector of column names
+de_col_names <- names(diff_enrich$de_table)
+de_col_descr <- c("KEGG Pathway Identifier",
+                  "Description of KEGG Pathway (provided by KEGG)",
+                  "Number of Genes in KEGG Pathway",
+                  "Number of Genes in KEGG Database",
+                  "Number of Genes from gene list 1 in KEGG Pathway",
+                  "Number of Genes from gene list 1 in KEGG Database",
+                  "Expected number of genes from list 1 to be in KEGG pathway by chance",
+                  "P-value for enrichment of list 1 genes related to KEGG pathway",
+                  "Multiple testing adjusted enrichment p-values from gene list 1 (default = False Discovery Rate (Benjamini and Hochberg, 1995))",
+                  "Ratio of number of genes observed from gene list 1 annotated to the KEGG pathway to the number of genes expected from gene list 1 annotated to the KEGG pathway if there was no enrichment (i.e.
+KEGG_PATHWAY_in_list1/expected_list1)",
+                  "Number of Genes from gene list 2 in KEGG Pathway",
+                  "Number of Genes from gene list 2 in KEGG Database",
+                  "Expected number of genes from list 2 to be in KEGG pathway by chance",
+                  "P-value for enrichment of list 2 genes related to KEGG pathway",
+                  "Multiple testing adjusted enrichment p-values from gene list 2 (default = False Discovery Rate (Benjamini and Hochberg, 1995))",
+                  "Ratio of number of genes observed from gene list 2 annotated to the KEGG pathway to the number of genes expected from gene list 2 annotated to the KEGG pathway if there was no enrichment (i.e.
+KEGG_PATHWAY_in_list2/expected_list2)",
+                  "Odds of a gene from list 2 being from this KEGG pathway / Odds of a gene from list 1 being from this KEGG pathway",
+                  "P-value for differential enrichment of this KEGG pathway between list 1 and list 2",
+                  "Multiple testing adjusted differential enrichment p-values (default = False Discovery Rate (Benjamini and Hochberg, 1995))")
+# Make dataframe
+tbl3 <- data.frame(de_col_names, de_col_descr)
+colnames(tbl3) <- c("Column Names", "Column Description")
+xtable(tbl3)
+@
+
+%\begin{table}[ht]
+%\centering
+%\begin{tabular}{rll}
+\begin{longtable}{cc}
+  \caption{Description of columns in dataframe generated by diffEnrich}
+  \hline
+ & Column Names & Column Description \\
+  \hline
+  \endhead
+  \hline
+  \endfoot
+KEGG\_PATHWAY\_ID & \makecell{KEGG Pathway Identifier} \\
+& & \\
+KEGG\_PATHWAY\_description & \makecell{Description of KEGG Pathway \\ (provided by KEGG)} \\
+& & \\
+KEGG\_PATHWAY\_cnt & \makecell{Number of Genes in KEGG Pathway} \\
+& & \\
+KEGG\_DATABASE\_cnt & \makecell{Number of Genes in KEGG Database} \\
+& & \\
+KEGG\_PATHWAY\_in\_list1 & \makecell{Number of Genes from  \\ gene list 1 in KEGG Pathway} \\
+& & \\
+KEGG\_DATABASE\_in\_list1 & \makecell{Number of Genes from \\ gene list 1 in KEGG Database} \\
+& & \\
+expected\_list1 & \makecell{Expected number of genes \\ from list 1 to be in \\ KEGG pathway by chance} \\
+& & \\
+enrich\_p\_list1 & \makecell{P-value for enrichment of \\ list 1 genes related to \\ KEGG pathway} \\
+& & \\
+p\_adj\_list1 & \makecell{Multiple testing adjusted enrichment \\ p-values from gene list 1 \\ (default = False Discovery Rate \\ (Benjamini and Hochberg, 1995))} \\
+& & \\
+fold\_enrichment\_list1 & \makecell{Ratio of number of genes \\ observed from gene list 1 \\ annotated to the KEGG pathway \\ to the number of genes expected \\ from gene list 1 annotated to \\ the KEGG pathway if there was no \\ enrichment (i.e.
+KEGG\_PATHWAY\_in\_list1/expected\_list1)} \\
+& & \\
+KEGG\_PATHWAY\_in\_list2 & \makecell{Number of Genes from gene \\ list 2 in KEGG Pathway} \\
+& & \\
+KEGG\_DATABASE\_in\_list2 & \makecell{Number of Genes from gene \\ list 2 in KEGG Database} \\
+& & \\
+expected\_list2 & \makecell{Expected number of genes \\ from list 2 to be in \\ KEGG pathway by chance} \\
+& & \\
+enrich\_p\_list2 & \makecell{P-value for enrichment of \\ list 2 genes related to \\ KEGG pathway} \\
+& & \\
+p\_adj\_list2 & \makecell{Multiple testing adjusted enrichment \\ p-values from gene list 2 \\ (default = False Discovery Rate \\ (Benjamini and Hochberg, 1995))} \\
+& & \\
+fold\_enrichment\_list2 & \makecell{Ratio of number of genes observed \\ from gene list 2 \\ annotated to the KEGG pathway \\ to the number of genes expected \\ from gene list 2 annotated to \\ the KEGG pathway if there was no \\ enrichment (i.e.
+KEGG\_PATHWAY\_in\_list2/expected\_list2)} \\
+& & \\
+odd\_ratio & \makecell{Odds of a gene from list 2 being \\ from this KEGG pathway / Odds of a gene from list 1\\  being from this KEGG pathway} \\
+& & \\
+diff\_enrich\_p & \makecell{P-value for differential enrichment \\ of this KEGG pathway between \\ list 1 and list 2} \\
+& & \\
+diff\_enrich\_adjusted & \makecell{Multiple testing adjusted differential \\ enrichment p-values \\ (default = False Discovery Rate \\ (Benjamini and Hochberg, 1995))} \\
+   \hline
+%\end{tabular}
+\end{longtable}
+
 %% -----Supplementary Information------------------------------------
@@
 <<list1_head, echo = FALSE, eval = FALSE>>=
-library(xtable)
 suptb1 <- head(list1_pe$enrich_table)

### 2019-11-22 jss-article-rnw-diffEnrich/article.tex
Added table 3 to .Rnw

@@
 
+%\begin{table}[ht]
+%\centering
+%\begin{tabular}{rll}
+\begin{longtable}{cc}
+  \caption{Description of columns in dataframe generated by diffEnrich}
+  \hline
+ & Column Names & Column Description \\
+  \hline
+  \endhead
+  \hline
+  \endfoot
+KEGG\_PATHWAY\_ID & \makecell{KEGG Pathway Identifier} \\
+& & \\
+KEGG\_PATHWAY\_description & \makecell{Description of KEGG Pathway \\ (provided by KEGG)} \\
+& & \\
+KEGG\_PATHWAY\_cnt & \makecell{Number of Genes in KEGG Pathway} \\
+& & \\
+KEGG\_DATABASE\_cnt & \makecell{Number of Genes in KEGG Database} \\
+& & \\
+KEGG\_PATHWAY\_in\_list1 & \makecell{Number of Genes from  \\ gene list 1 in KEGG Pathway} \\
+& & \\
+KEGG\_DATABASE\_in\_list1 & \makecell{Number of Genes from \\ gene list 1 in KEGG Database} \\
+& & \\
+expected\_list1 & \makecell{Expected number of genes \\ from list 1 to be in \\ KEGG pathway by chance} \\
+& & \\
+enrich\_p\_list1 & \makecell{P-value for enrichment of \\ list 1 genes related to \\ KEGG pathway} \\
+& & \\
+p\_adj\_list1 & \makecell{Multiple testing adjusted enrichment \\ p-values from gene list 1 \\ (default = False Discovery Rate \\ (Benjamini and Hochberg, 1995))} \\
+& & \\
+fold\_enrichment\_list1 & \makecell{Ratio of number of genes \\ observed from gene list 1 \\ annotated to the KEGG pathway \\ to the number of genes expected \\ from gene list 1 annotated to \\ the KEGG pathway if there was no \\ enrichment (i.e.
+KEGG\_PATHWAY\_in\_list1/expected\_list1)} \\
+& & \\
+KEGG\_PATHWAY\_in\_list2 & \makecell{Number of Genes from gene \\ list 2 in KEGG Pathway} \\
+& & \\
+KEGG\_DATABASE\_in\_list2 & \makecell{Number of Genes from gene \\ list 2 in KEGG Database} \\
+& & \\
+expected\_list2 & \makecell{Expected number of genes \\ from list 2 to be in \\ KEGG pathway by chance} \\
+& & \\
+enrich\_p\_list2 & \makecell{P-value for enrichment of \\ list 2 genes related to \\ KEGG pathway} \\
+& & \\
+p\_adj\_list2 & \makecell{Multiple testing adjusted enrichment \\ p-values from gene list 2 \\ (default = False Discovery Rate \\ (Benjamini and Hochberg, 1995))} \\
+& & \\
+fold\_enrichment\_list2 & \makecell{Ratio of number of genes observed \\ from gene list 2 \\ annotated to the KEGG pathway \\ to the number of genes expected \\ from gene list 2 annotated to \\ the KEGG pathway if there was no \\ enrichment (i.e.
+KEGG\_PATHWAY\_in\_list2/expected\_list2)} \\
+& & \\
+odd\_ratio & \makecell{Odds of a gene from list 2 being \\ from this KEGG pathway / Odds of a gene from list 1\\  being from this KEGG pathway} \\
+& & \\
+diff\_enrich\_p & \makecell{P-value for differential enrichment \\ of this KEGG pathway between \\ list 1 and list 2} \\
+& & \\
+diff\_enrich\_adjusted & \makecell{Multiple testing adjusted differential \\ enrichment p-values \\ (default = False Discovery Rate \\ (Benjamini and Hochberg, 1995))} \\
+   \hline
+%\end{tabular}
+\end{longtable}
+
 %% -----Supplementary Information------------------------------------

### 2019-11-23 jss-article-rnw-diffEnrich/article.Rnw
First draft jss article

@@
 
+\subsection{\code{plotFoldEnrichment}}
+
+\code{plotFoldEnrichment} generates a grouped bar plot using
+\code{ggplot2} (\cite{Hadley:2106}) and the \code{ggnewscale}
+package (\cite{Elio:2019}). There are 3 arguments: 1)
+\code{de_res} is the dataframe generated from the
+\code{diffEnrich} function, 2) \code{pval} is the threshold
+for the adjusted p-value associated with differential
+enrichment that will filter which KEGG pathways to plot, and
+3) after filtering based on \code{pval}, \code{N} tells the
+function how many pathways to plot. It is important to make
+a note that the significance of the fold change is associated
+with the number of genes in the gene list. Notice that in this
+example the pathways in gene list 2 have smaller fold changes
+(shorter bars) than those in list 1, but that many of them are
+more significant (darker blue). This is because there are
+more genes in gene list 2 compared to gene list 1.
+
+\begin{figure}[H]
+% \centering
+<<visualization_2, echo=TRUE, message=FALSE, out.height='4in', out.width='5.0in'>>=
+## Plot fold enrichment
+plotFoldEnrichment(de_res = diff_enrich, pval = 0.05, N = 5)
+@
+\caption{\label{fig:barchart} \textbf{Example of a differential enrichment graphic.}
+KEGG pathways are plotted on the y-axis and fold enrichment is
+plotted on the x-axis. Each KEGG pathway has a bar depicting its
+fold enrichment in list 1 (red) and its fold enrichment in list 2
+(blue). The transparency of the bars correspond to the unadjusted
+p-value for the pathway's enrichment in the given list. The p-value
+presented as text to the right of each pair of bars is the adjusted
+p-value (user defined: default is FDR) associated with the
+differential enrichment of the pathway between the two lists, and
+the pathways are ordered from top to bottom by this p-value (i.e.
+smallest p-value on top of plot, and largest p-value on bottom of
+plot). The dotted line represents no enrichment (i.e a fold enrichment of 1).
+The number of genes used for analysis from each gene list (recall
+that this number may not be the same as the number of genes in
+the user’s original list) are reported below their respective
+p-values in the legend. In this example, all five pathways are
+differentially enriched with more enrichment in List 1 than in
+List 2.}
+\end{figure}
+
 %% -----Supplementary Information------------------------------------
@@
   \caption{Supplementary Table 1. First 6 results generated by pathEnrich.}
-  \tiny
-  \begin{adjustbox}{width=\textwidth}
+  \begin{adjustbox}{width=\textwidth, angle=90}
+  \fontsize{90}{125}\selectfont
   \begin{tabular}{rllrrrrrrrr}
@@
   66 & rno04144 & Endocytosis - Rattus norvegicus (rat) & 275 & 22.00 & 8856 & 295 & 9.16 & 0.00 & 0.00 & 2.40 \\
+   \hline
+  \end{tabular}
+\end{adjustbox}
+\end{table}
+
+<<diffEnrich_head, echo = FALSE, eval = FALSE>>=
+suptb2 <- head(diff_enrich$de_table)
+xtable(suptb2)
+@
+
+\begin{table}[H]
+\centering
+  \caption{Supplementary Table 2. First 6 results generated by diffEnrich}
+  \begin{adjustbox}{width=\textwidth, angle=90}
+  \fontsize{50}{80}\selectfont
+\begin{tabular}{rllrrrrrrrrrrrrrrrrr}
+  \hline
+ & KEGG\_PATHWAY\_ID & KEGG\_PATHWAY\_description & KEGG\_PATHWAY\_cnt & KEGG\_DATABASE\_cnt & KEGG\_PATHWAY\_in\_list1 & KEGG\_DATABASE\_in\_list1 & expected\_list1 & enrich\_p\_list1 & p\_adj\_list1 & fold\_enrichment\_list1 & KEGG\_PATHWAY\_in\_list2 & KEGG\_DATABASE\_in\_list2 & expected\_list2 & enrich\_p\_list2 & p\_adj\_list2 & fold\_enrichment\_list2 & odd\_ratio & diff\_enrich\_p & diff\_enrich\_adjusted \\
+  \hline
+rno04530 & rno04530 & Tight junction - Rattus norvegicus (rat) & 170 & 8856 & 19.00 & 295 & 5.66 & 0.00 & 0.00 & 3.36 & 131.00 & 5308 & 101.89 & 0.00 & 0.00 & 1.29 & 0.37 & 0.00 & 0.00 \\
+  rno05135 & rno05135 & Yersinia infection - Rattus norvegicus (rat) & 128 & 8856 & 16.00 & 295 & 4.26 & 0.00 & 0.00 & 3.75 & 105.00 & 5308 & 76.72 & 0.00 & 0.00 & 1.37 & 0.35 & 0.00 & 0.00 \\
+  rno05210 & rno05210 & Colorectal cancer - Rattus norvegicus (rat) &  88 & 8856 & 12.00 & 295 & 2.93 & 0.00 & 0.00 & 4.09 & 81.00 & 5308 & 52.74 & 0.00 & 0.00 & 1.54 & 0.37 & 0.00 & 0.00 \\
+  rno05213 & rno05213 & Endometrial cancer - Rattus norvegicus (rat) &  58 & 8856 & 9.00 & 295 & 1.93 & 0.00 & 0.00 & 4.66 & 55.00 & 5308 & 34.76 & 0.00 & 0.00 & 1.58 & 0.33 & 0.01 & 0.01 \\
+  rno04660 & rno04660 & T cell receptor signaling pathway - Rattus norvegicus (rat) & 106 & 8856 & 11.00 & 295 & 3.53 & 0.00 & 0.01 & 3.12 & 79.00 & 5308 & 63.53 & 0.00 & 0.00 & 1.24 & 0.39 & 0.01 & 0.01 \\
+  rno04657 & rno04657 & IL-17 signaling pathway - Rattus norvegicus (rat) &  95 & 8856 & 9.00 & 295 & 3.16 & 0.00 & 0.03 & 2.84 & 59.00 & 5308 & 56.94 & 0.37 & 0.46 & 1.04 & 0.36 & 0.01 & 0.01 \\
    \hline

### 2019-11-23 jss-article-rnw-diffEnrich/article.tex
First draft jss article

@@
 
-{\ttfamily\noindent\itshape\color{messagecolor}{\#\# Kegg Release: Release\_92.0+\_11-22\_Nov\_19}}\end{kframe}
+{\ttfamily\noindent\itshape\color{messagecolor}{\#\# Kegg Release: Release\_92.0+\_11-23\_Nov\_19}}\end{kframe}
 \end{knitrout}
@@
 
-{\ttfamily\noindent\itshape\color{messagecolor}{\#\# ncbi\_to\_kegg2019-11-22Release\_92.0+\_11-22\_Nov\_19.txt}}
+{\ttfamily\noindent\itshape\color{messagecolor}{\#\# ncbi\_to\_kegg2019-11-22Release\_92.0+\_11-23\_Nov\_19.txt}}
 
-{\ttfamily\noindent\itshape\color{messagecolor}{\#\# kegg\_to\_pathway2019-11-22Release\_92.0+\_11-22\_Nov\_19.txt}}
+{\ttfamily\noindent\itshape\color{messagecolor}{\#\# kegg\_to\_pathway2019-11-22Release\_92.0+\_11-23\_Nov\_19.txt}}
 
-{\ttfamily\noindent\itshape\color{messagecolor}{\#\# pathway\_to\_species2019-11-22Release\_92.0+\_11-22\_Nov\_19.txt}}
+{\ttfamily\noindent\itshape\color{messagecolor}{\#\# pathway\_to\_species2019-11-22Release\_92.0+\_11-23\_Nov\_19.txt}}
 
@@
 
+\subsection{\code{plotFoldEnrichment}}
+
+\code{plotFoldEnrichment} generates a grouped bar plot using
+\code{ggplot2} (\cite{Hadley:2106}) and the \code{ggnewscale}
+package (\cite{Elio:2019}). There are 3 arguments: 1)
+\code{de_res} is the dataframe generated from the
+\code{diffEnrich} function, 2) \code{pval} is the threshold
+for the adjusted p-value associated with differential
+enrichment that will filter which KEGG pathways to plot, and
+3) after filtering based on \code{pval}, \code{N} tells the
+function how many pathways to plot. It is important to make
+a note that the significance of the fold change is associated
+with the number of genes in the gene list. Notice that in this
+example the pathways in gene list 2 have smaller fold changes
+(shorter bars) than those in list 1, but that many of them are
+more significant (darker blue). This is because there are
+more genes in gene list 2 compared to gene list 1.
+
+\begin{figure}[H]
+% \centering
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{alltt}
+\hlcom{## Plot fold enrichment}
+\hlkwd{plotFoldEnrichment}\hlstd{(}\hlkwc{de_res} \hlstd{= diff_enrich,} \hlkwc{pval} \hlstd{=} \hlnum{0.05}\hlstd{,} \hlkwc{N} \hlstd{=} \hlnum{5}\hlstd{)}
+\end{alltt}
+\end{kframe}
+\includegraphics[width=5.0in,height=4in]{figure/visualization_2-1}
+
+\end{knitrout}
+\caption{\label{fig:barchart} \textbf{Example of a differential enrichment graphic.}
+KEGG pathways are plotted on the y-axis and fold enrichment is
+plotted on the x-axis. Each KEGG pathway has a bar depicting its
+fold enrichment in list 1 (red) and its fold enrichment in list 2
+(blue). The transparency of the bars correspond to the unadjusted
+p-value for the pathway's enrichment in the given list. The p-value
+presented as text to the right of each pair of bars is the adjusted
+p-value (user defined: default is FDR) associated with the
+differential enrichment of the pathway between the two lists, and
+the pathways are ordered from top to bottom by this p-value (i.e.
+smallest p-value on top of plot, and largest p-value on bottom of
+plot). The dotted line represents no enrichment (i.e a fold enrichment of 1).
+The number of genes used for analysis from each gene list (recall
+that this number may not be the same as the number of genes in
+the user’s original list) are reported below their respective
+p-values in the legend. In this example, all five pathways are
+differentially enriched with more enrichment in List 1 than in
+List 2.}
+\end{figure}
+
 %% -----Supplementary Information------------------------------------
@@
   66 & rno04144 & Endocytosis - Rattus norvegicus (rat) & 275 & 22.00 & 8856 & 295 & 9.16 & 0.00 & 0.00 & 2.40 \\
+   \hline
+  \end{tabular}
+\end{adjustbox}
+\end{table}
+
+\begin{table}[H]
+\centering
+  \caption{Supplementary Table 2. First 6 results generated by diffEnrich}
+  \begin{adjustbox}{width=\textwidth, angle=90}
+  \fontsize{50}{80}\selectfont
+\begin{tabular}{rllrrrrrrrrrrrrrrrrr}
+  \hline
+ & KEGG\_PATHWAY\_ID & KEGG\_PATHWAY\_description & KEGG\_PATHWAY\_cnt & KEGG\_DATABASE\_cnt & KEGG\_PATHWAY\_in\_list1 & KEGG\_DATABASE\_in\_list1 & expected\_list1 & enrich\_p\_list1 & p\_adj\_list1 & fold\_enrichment\_list1 & KEGG\_PATHWAY\_in\_list2 & KEGG\_DATABASE\_in\_list2 & expected\_list2 & enrich\_p\_list2 & p\_adj\_list2 & fold\_enrichment\_list2 & odd\_ratio & diff\_enrich\_p & diff\_enrich\_adjusted \\
+  \hline
+rno04530 & rno04530 & Tight junction - Rattus norvegicus (rat) & 170 & 8856 & 19.00 & 295 & 5.66 & 0.00 & 0.00 & 3.36 & 131.00 & 5308 & 101.89 & 0.00 & 0.00 & 1.29 & 0.37 & 0.00 & 0.00 \\
+  rno05135 & rno05135 & Yersinia infection - Rattus norvegicus (rat) & 128 & 8856 & 16.00 & 295 & 4.26 & 0.00 & 0.00 & 3.75 & 105.00 & 5308 & 76.72 & 0.00 & 0.00 & 1.37 & 0.35 & 0.00 & 0.00 \\
+  rno05210 & rno05210 & Colorectal cancer - Rattus norvegicus (rat) &  88 & 8856 & 12.00 & 295 & 2.93 & 0.00 & 0.00 & 4.09 & 81.00 & 5308 & 52.74 & 0.00 & 0.00 & 1.54 & 0.37 & 0.00 & 0.00 \\
+  rno05213 & rno05213 & Endometrial cancer - Rattus norvegicus (rat) &  58 & 8856 & 9.00 & 295 & 1.93 & 0.00 & 0.00 & 4.66 & 55.00 & 5308 & 34.76 & 0.00 & 0.00 & 1.58 & 0.33 & 0.01 & 0.01 \\
+  rno04660 & rno04660 & T cell receptor signaling pathway - Rattus norvegicus (rat) & 106 & 8856 & 11.00 & 295 & 3.53 & 0.00 & 0.01 & 3.12 & 79.00 & 5308 & 63.53 & 0.00 & 0.00 & 1.24 & 0.39 & 0.01 & 0.01 \\
+  rno04657 & rno04657 & IL-17 signaling pathway - Rattus norvegicus (rat) &  95 & 8856 & 9.00 & 295 & 3.16 & 0.00 & 0.03 & 2.84 & 59.00 & 5308 & 56.94 & 0.37 & 0.46 & 1.04 & 0.36 & 0.01 & 0.01 \\
    \hline

### 2019-12-03 README.Rmd
reknit README with new plot
Format fixes after RStudio update

@@
 ```
-# diffEnrich
+#diffEnrich
 

### 2019-12-05 jss-article-rnw-diffEnrich/article.Rnw
incorp supp info into text body and remove supp section

@@
 pathway by chance. An example of the first 6 results generated by \code{pathEnrich}
-are in Supplementary Table 1. For a detailed description of the variables in this table see Table 2.
+are in displayed below. For a detailed description of the variables in this table see Table 2.
+
+<<pathEnrich_example, echo=TRUE>>=
+head(list1_pe$enrich_table)
+@
 
@@
 KEGG data base, the KEGG species, and the method used for multiple testing
-correction. See supplementary information for examples.
+correction.
+
+<<path_summary, echo = TRUE>>=
+summary(list1_pe)
+@
 
@@
 generated by \code{diffEnrich} see Table 3, and an example of the first 6
-results generated by \code{diffEnrich} are in Supplementary Table 2.
+results generated by \code{diffEnrich} are displayed below.
 
@@
                           cutoff = 0.05)
+@
+
+<<diffEnrich_example, echo=TRUE>>=
+head(diff_enrich$de_table)
 @
@@
 
+<<diff_summary, echo = TRUE>>=
+summary(diff_enrich)
+@
+
 \subsection{\code{plotFoldEnrichment}}
@@
 %% -----Supplementary Information------------------------------------
-\section[Supplementary Information]{Supplementary Information} \label{sec:supinfo}
+% \section[Supplementary Information]{Supplementary Information} \label{sec:supinfo}
 
@@
 
-\begin{table}[H]
-  \centering
-  \caption{Supplementary Table 1. First 6 results generated by pathEnrich.}
-  \begin{adjustbox}{width=\textwidth, angle=90}
-  \fontsize{90}{125}\selectfont
-  \begin{tabular}{rllrrrrrrrr}
-  \hline
- & KEGG\_PATHWAY\_ID & KEGG\_PATHWAY\_description & KEGG\_PATHWAY\_cnt & KEGG\_PATHWAY\_in\_list & KEGG\_DATABASE\_cnt & KEG\_DATABASE\_in\_list & expected & enrich\_p & p\_adj & fold\_enrichment \\
-  \hline
-95 & rno04530 & Tight junction - Rattus norvegicus (rat) & 170 & 19.00 & 8856 & 295 & 5.66 & 0.00 & 0.00 & 3.36 \\
-  172 & rno05135 & Yersinia infection - Rattus norvegicus (rat) & 128 & 16.00 & 8856 & 295 & 4.26 & 0.00 & 0.00 & 3.75 \\
-  194 & rno05210 & Colorectal cancer - Rattus norvegicus (rat) &  88 & 12.00 & 8856 & 295 & 2.93 & 0.00 & 0.00 & 4.09 \\
-  212 & rno05231 & Choline metabolism in cancer - Rattus norvegicus (rat) &  99 & 12.00 & 8856 & 295 & 3.30 & 0.00 & 0.00 & 3.64 \\
-  197 & rno05213 & Endometrial cancer - Rattus norvegicus (rat) &  58 & 9.00 & 8856 & 295 & 1.93 & 0.00 & 0.00 & 4.66 \\
-  66 & rno04144 & Endocytosis - Rattus norvegicus (rat) & 275 & 22.00 & 8856 & 295 & 9.16 & 0.00 & 0.00 & 2.40 \\
-   \hline
-  \end{tabular}
-\end{adjustbox}
-\end{table}
+% \begin{table}[H]
+%   \centering
+%   \caption{Supplementary Table 1. First 6 results generated by pathEnrich.}
+%   \begin{adjustbox}{width=\textwidth, angle=90}
+%   \fontsize{90}{125}\selectfont
+%   \begin{tabular}{rllrrrrrrrr}
+%   \hline
+%  & KEGG\_PATHWAY\_ID & KEGG\_PATHWAY\_description & KEGG\_PATHWAY\_cnt & KEGG\_PATHWAY\_in\_list & KEGG\_DATABASE\_cnt & KEG\_DATABASE\_in\_list & expected & enrich\_p & p\_adj & fold\_enrichment \\
+%   \hline
+% 95 & rno04530 & Tight junction - Rattus norvegicus (rat) & 170 & 19.00 & 8856 & 295 & 5.66 & 0.00 & 0.00 & 3.36 \\
+%   172 & rno05135 & Yersinia infection - Rattus norvegicus (rat) & 128 & 16.00 & 8856 & 295 & 4.26 & 0.00 & 0.00 & 3.75 \\
+%   194 & rno05210 & Colorectal cancer - Rattus norvegicus (rat) &  88 & 12.00 & 8856 & 295 & 2.93 & 0.00 & 0.00 & 4.09 \\
+%   212 & rno05231 & Choline metabolism in cancer - Rattus norvegicus (rat) &  99 & 12.00 & 8856 & 295 & 3.30 & 0.00 & 0.00 & 3.64 \\
+%   197 & rno05213 & Endometrial cancer - Rattus norvegicus (rat) &  58 & 9.00 & 8856 & 295 & 1.93 & 0.00 & 0.00 & 4.66 \\
+%   66 & rno04144 & Endocytosis - Rattus norvegicus (rat) & 275 & 22.00 & 8856 & 295 & 9.16 & 0.00 & 0.00 & 2.40 \\
+%    \hline
+%   \end{tabular}
+% \end{adjustbox}
+% \end{table}
 
@@
 
-\begin{table}[H]
-\centering
-  \caption{Supplementary Table 2. First 6 results generated by diffEnrich}
-  \begin{adjustbox}{width=\textwidth, angle=90}
-  \fontsize{50}{80}\selectfont
-\begin{tabular}{rllrrrrrrrrrrrrrrrrr}
-  \hline
- & KEGG\_PATHWAY\_ID & KEGG\_PATHWAY\_description & KEGG\_PATHWAY\_cnt & KEGG\_DATABASE\_cnt & KEGG\_PATHWAY\_in\_list1 & KEGG\_DATABASE\_in\_list1 & expected\_list1 & enrich\_p\_list1 & p\_adj\_list1 & fold\_enrichment\_list1 & KEGG\_PATHWAY\_in\_list2 & KEGG\_DATABASE\_in\_list2 & expected\_list2 & enrich\_p\_list2 & p\_adj\_list2 & fold\_enrichment\_list2 & odd\_ratio & diff\_enrich\_p & diff\_enrich\_adjusted \\
-  \hline
-rno04530 & rno04530 & Tight junction - Rattus norvegicus (rat) & 170 & 8856 & 19.00 & 295 & 5.66 & 0.00 & 0.00 & 3.36 & 131.00 & 5308 & 101.89 & 0.00 & 0.00 & 1.29 & 0.37 & 0.00 & 0.00 \\
-  rno05135 & rno05135 & Yersinia infection - Rattus norvegicus (rat) & 128 & 8856 & 16.00 & 295 & 4.26 & 0.00 & 0.00 & 3.75 & 105.00 & 5308 & 76.72 & 0.00 & 0.00 & 1.37 & 0.35 & 0.00 & 0.00 \\
-  rno05210 & rno05210 & Colorectal cancer - Rattus norvegicus (rat) &  88 & 8856 & 12.00 & 295 & 2.93 & 0.00 & 0.00 & 4.09 & 81.00 & 5308 & 52.74 & 0.00 & 0.00 & 1.54 & 0.37 & 0.00 & 0.00 \\
-  rno05213 & rno05213 & Endometrial cancer - Rattus norvegicus (rat) &  58 & 8856 & 9.00 & 295 & 1.93 & 0.00 & 0.00 & 4.66 & 55.00 & 5308 & 34.76 & 0.00 & 0.00 & 1.58 & 0.33 & 0.01 & 0.01 \\
-  rno04660 & rno04660 & T cell receptor signaling pathway - Rattus norvegicus (rat) & 106 & 8856 & 11.00 & 295 & 3.53 & 0.00 & 0.01 & 3.12 & 79.00 & 5308 & 63.53 & 0.00 & 0.00 & 1.24 & 0.39 & 0.01 & 0.01 \\
-  rno04657 & rno04657 & IL-17 signaling pathway - Rattus norvegicus (rat) &  95 & 8856 & 9.00 & 295 & 3.16 & 0.00 & 0.03 & 2.84 & 59.00 & 5308 & 56.94 & 0.37 & 0.46 & 1.04 & 0.36 & 0.01 & 0.01 \\
-   \hline
-\end{tabular}
-\end{adjustbox}
-\end{table}
-
-Example of S3 functions
-<<path_summary, echo = TRUE>>=
-summary(list1_pe)
-@
+% \begin{table}[H]
+% \centering
+%   \caption{Supplementary Table 2. First 6 results generated by diffEnrich}
+%   \begin{adjustbox}{width=\textwidth, angle=90}
+%   \fontsize{50}{80}\selectfont
+% \begin{tabular}{rllrrrrrrrrrrrrrrrrr}
+%   \hline
+%  & KEGG\_PATHWAY\_ID & KEGG\_PATHWAY\_description & KEGG\_PATHWAY\_cnt & KEGG\_DATABASE\_cnt & KEGG\_PATHWAY\_in\_list1 & KEGG\_DATABASE\_in\_list1 & expected\_list1 & enrich\_p\_list1 & p\_adj\_list1 & fold\_enrichment\_list1 & KEGG\_PATHWAY\_in\_list2 & KEGG\_DATABASE\_in\_list2 & expected\_list2 & enrich\_p\_list2 & p\_adj\_list2 & fold\_enrichment\_list2 & odd\_ratio & diff\_enrich\_p & diff\_enrich\_adjusted \\
+%   \hline
+% rno04530 & rno04530 & Tight junction - Rattus norvegicus (rat) & 170 & 8856 & 19.00 & 295 & 5.66 & 0.00 & 0.00 & 3.36 & 131.00 & 5308 & 101.89 & 0.00 & 0.00 & 1.29 & 0.37 & 0.00 & 0.00 \\
+%   rno05135 & rno05135 & Yersinia infection - Rattus norvegicus (rat) & 128 & 8856 & 16.00 & 295 & 4.26 & 0.00 & 0.00 & 3.75 & 105.00 & 5308 & 76.72 & 0.00 & 0.00 & 1.37 & 0.35 & 0.00 & 0.00 \\
+%   rno05210 & rno05210 & Colorectal cancer - Rattus norvegicus (rat) &  88 & 8856 & 12.00 & 295 & 2.93 & 0.00 & 0.00 & 4.09 & 81.00 & 5308 & 52.74 & 0.00 & 0.00 & 1.54 & 0.37 & 0.00 & 0.00 \\
+%   rno05213 & rno05213 & Endometrial cancer - Rattus norvegicus (rat) &  58 & 8856 & 9.00 & 295 & 1.93 & 0.00 & 0.00 & 4.66 & 55.00 & 5308 & 34.76 & 0.00 & 0.00 & 1.58 & 0.33 & 0.01 & 0.01 \\
+%   rno04660 & rno04660 & T cell receptor signaling pathway - Rattus norvegicus (rat) & 106 & 8856 & 11.00 & 295 & 3.53 & 0.00 & 0.01 & 3.12 & 79.00 & 5308 & 63.53 & 0.00 & 0.00 & 1.24 & 0.39 & 0.01 & 0.01 \\
+%   rno04657 & rno04657 & IL-17 signaling pathway - Rattus norvegicus (rat) &  95 & 8856 & 9.00 & 295 & 3.16 & 0.00 & 0.03 & 2.84 & 59.00 & 5308 & 56.94 & 0.37 & 0.46 & 1.04 & 0.36 & 0.01 & 0.01 \\
+%    \hline
+% \end{tabular}
+% \end{adjustbox}
+% \end{table}
 

### 2019-12-05 jss-article-rnw-diffEnrich/article.tex
incorp supp info into text body and remove supp section

@@
 
-{\ttfamily\noindent\itshape\color{messagecolor}{\#\# Kegg Release: Release\_92.0+\_11-23\_Nov\_19}}\end{kframe}
+{\ttfamily\noindent\itshape\color{messagecolor}{\#\# Kegg Release: Release\_92.0+\_12-05\_Dec\_19}}\end{kframe}
 \end{knitrout}
@@
 
-{\ttfamily\noindent\itshape\color{messagecolor}{\#\# ncbi\_to\_kegg2019-11-22Release\_92.0+\_11-23\_Nov\_19.txt}}
+{\ttfamily\noindent\itshape\color{messagecolor}{\#\# ncbi\_to\_kegg2019-12-05Release\_92.0+\_12-05\_Dec\_19.txt}}
 
-{\ttfamily\noindent\itshape\color{messagecolor}{\#\# kegg\_to\_pathway2019-11-22Release\_92.0+\_11-23\_Nov\_19.txt}}
+{\ttfamily\noindent\itshape\color{messagecolor}{\#\# kegg\_to\_pathway2019-12-05Release\_92.0+\_12-05\_Dec\_19.txt}}
 
-{\ttfamily\noindent\itshape\color{messagecolor}{\#\# pathway\_to\_species2019-11-22Release\_92.0+\_11-23\_Nov\_19.txt}}
+{\ttfamily\noindent\itshape\color{messagecolor}{\#\# pathway\_to\_species2019-12-05Release\_92.0+\_12-05\_Dec\_19.txt}}
 
@@
 pathway by chance. An example of the first 6 results generated by \code{pathEnrich}
-are in Supplementary Table 1. For a detailed description of the variables in this table see Table 2.
+are in displayed below. For a detailed description of the variables in this table see Table 2.
+
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{alltt}
+\hlkwd{head}\hlstd{(list1_pe}\hlopt{$}\hlstd{enrich_table)}
+\end{alltt}
+\begin{verbatim}
+##     KEGG_PATHWAY_ID
+## 95         rno04530
+## 172        rno05135
+## 194        rno05210
+## 212        rno05231
+## 197        rno05213
+## 66         rno04144
+##                                   KEGG_PATHWAY_description
+## 95                Tight junction - Rattus norvegicus (rat)
+## 172           Yersinia infection - Rattus norvegicus (rat)
+## 194            Colorectal cancer - Rattus norvegicus (rat)
+## 212 Choline metabolism in cancer - Rattus norvegicus (rat)
+## 197           Endometrial cancer - Rattus norvegicus (rat)
+## 66                   Endocytosis - Rattus norvegicus (rat)
+##     KEGG_PATHWAY_cnt KEGG_PATHWAY_in_list KEGG_DATABASE_cnt
+## 95               170                   19              8856
+## 172              128                   16              8856
+## 194               88                   12              8856
+## 212               99                   12              8856
+## 197               58                    9              8856
+## 66               275                   22              8856
+##     KEG_DATABASE_in_list expected     enrich_p        p_adj
+## 95                   295 5.662827 3.485551e-06 0.0005387284
+## 172                  295 4.263776 4.919894e-06 0.0005387284
+## 194                  295 2.931346 3.179635e-05 0.0023211334
+## 212                  295 3.297764 1.032192e-04 0.0044701134
+## 197                  295 1.932023 1.132216e-04 0.0044701134
+## 66                   295 9.160456 1.224689e-04 0.0044701134
+##     fold_enrichment
+## 95         3.355214
+## 172        3.752542
+## 194        4.093683
+## 212        3.638829
+## 197        4.658328
+## 66         2.401627
+\end{verbatim}
+\end{kframe}
+\end{knitrout}
 
@@
 KEGG data base, the KEGG species, and the method used for multiple testing
-correction. See supplementary information for examples.
+correction.
+
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{alltt}
+\hlkwd{summary}\hlstd{(list1_pe)}
+\end{alltt}
+\begin{verbatim}
+## 219 KEGG pathways were tested. 
+##  Only KEGG pathways that contained at least 2 genes from gene_list were tested. 
+##  KEGG pathway species: Rattus norvegicus (rat)
+##  8856 genes from gene_list were in the KEGG data pull. 
+##  p-value adjustment method: BH
+##  36 pathways reached statistical significance after multiple testing correction at a cutoff of 0.05. 
+##  
+## Significant pathways: 
+##  Tight junction
+## Yersinia infection
+## Colorectal cancer
+## Choline metabolism in cancer
+## Endometrial cancer
+## Endocytosis
+## Neurotrophin signaling pathway
+## Thermogenesis
+## Oocyte meiosis
+## VEGF signaling pathway
+## Thyroid hormone signaling pathway
+## Hippo signaling pathway
+## T cell receptor signaling pathway
+## Apoptosis
+## Hepatocellular carcinoma
+## MAPK signaling pathway
+## Focal adhesion
+## Salmonella infection
+## Non-alcoholic fatty liver disease (NAFLD)
+## ErbB signaling pathway
+## Sphingolipid signaling pathway
+## Pancreatic cancer
+## Progesterone-mediated oocyte maturation
+## Alzheimer disease
+## Endocrine resistance
+## Adrenergic signaling in cardiomyocytes
+## IL-17 signaling pathway
+## Chronic myeloid leukemia
+## Dopaminergic synapse
+## Prostate cancer
+## EGFR tyrosine kinase inhibitor resistance
+## Hepatitis C
+## Ras signaling pathway
+## Acute myeloid leukemia
+## Insulin signaling pathway
+## Fc epsilon RI signaling pathway
+\end{verbatim}
+\end{kframe}
+\end{knitrout}
 
@@
 \end{alltt}
+\end{kframe}
+\end{knitrout}
+
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{alltt}
+\hlkwd{head}\hlstd{(diff_enrich}\hlopt{$}\hlstd{de_table)}
+\end{alltt}
+\begin{verbatim}
+##          KEGG_PATHWAY_ID
+## rno04530        rno04530
+## rno05135        rno05135
+## rno05210        rno05210
+## rno05213        rno05213
+## rno04660        rno04660
+## rno04657        rno04657
+##                                             KEGG_PATHWAY_description
+## rno04530                    Tight junction - Rattus norvegicus (rat)
+## rno05135                Yersinia infection - Rattus norvegicus (rat)
+## rno05210                 Colorectal cancer - Rattus norvegicus (rat)
+## rno05213                Endometrial cancer - Rattus norvegicus (rat)
+## rno04660 T cell receptor signaling pathway - Rattus norvegicus (rat)
+## rno04657           IL-17 signaling pathway - Rattus norvegicus (rat)
+##          KEGG_PATHWAY_cnt KEGG_DATABASE_cnt KEGG_PATHWAY_in_list1
+## rno04530              170              8856                    19
+## rno05135              128              8856                    16
+## rno05210               88              8856                    12
+## rno05213               58              8856                     9
+## rno04660              106              8856                    11
+## rno04657               95              8856                     9
+##          KEGG_DATABASE_in_list1 expected_list1 enrich_p_list1
+## rno04530                    295       5.662827   3.485551e-06
+## rno05135                    295       4.263776   4.919894e-06
+## rno05210                    295       2.931346   3.179635e-05
+## rno05213                    295       1.932023   1.132216e-04
+## rno04660                    295       3.530939   7.743416e-04
+## rno04657                    295       3.164521   4.270915e-03
+##           p_adj_list1 fold_enrichment_list1 KEGG_PATHWAY_in_list2
+## rno04530 0.0005387284              3.355214                   131
+## rno05135 0.0005387284              3.752542                   105
+## rno05210 0.0023211334              4.093683                    81
+## rno05213 0.0044701134              4.658328                    55
+## rno04660 0.0129121321              3.115318                    79
+## rno04657 0.0346418644              2.844032                    59
+##          KEGG_DATABASE_in_list2 expected_list2 enrich_p_list2
+## rno04530                   5308      101.89250   1.459992e-06
+## rno05135                   5308       76.71906   5.537739e-08
+## rno05210                   5308       52.74435   9.806872e-12
+## rno05213                   5308       34.76332   1.160182e-09
+## rno04660                   5308       63.53297   1.107477e-03
+## rno04657                   5308       56.93993   3.737223e-01
+##           p_adj_list2 fold_enrichment_list2 odd_ratio diff_enrich_p
+## rno04530 5.568342e-06              1.285669 0.3676651  0.0002935876
+## rno05135 2.594826e-07              1.368630 0.3520039  0.0005434729
+## rno05210 1.237175e-10              1.535709 0.3655602  0.0032572306
+## rno05213 8.849757e-09              1.582127 0.3328275  0.0058774642
+## rno04660 2.256227e-03              1.243449 0.3901694  0.0072047984
+## rno04657 4.591045e-01              1.036180 0.3572935  0.0087537607
+##          diff_enrich_adjusted
+## rno04530         0.0002935876
+## rno05135         0.0005434729
+## rno05210         0.0032572306
+## rno05213         0.0058774642
+## rno04660         0.0072047984
+## rno04657         0.0087537607
+\end{verbatim}
 \end{kframe}
@@
 
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{alltt}
+\hlkwd{summary}\hlstd{(diff_enrich)}
+\end{alltt}
+\begin{verbatim}
+## 219 KEGG pathways were shared between gene lists and were tested. 
+##  KEGG pathway species: Rattus norvegicus (rat)
+##  8856 genes from gene_list were in the KEGG data pull. 
+##  p-value adjustment method: none
+##  19 pathways reached statistical significance after multiple testing correction at a cutoff of 0.05. 
+##  
+## Significant pathways: 
+##  Tight junction
+## Yersinia infection
+## Colorectal cancer
+## Endometrial cancer
+## T cell receptor signaling pathway
+## IL-17 signaling pathway
+## Salmonella infection
+## Choline metabolism in cancer
+## Endocytosis
+## VEGF signaling pathway
+## Oocyte meiosis
+## Thermogenesis
+## Hippo signaling pathway
+## Neurotrophin signaling pathway
+## Apoptosis
+## Hepatocellular carcinoma
+## Fc epsilon RI signaling pathway
+## Thyroid hormone signaling pathway
+## Alzheimer disease
+\end{verbatim}
+\end{kframe}
+\end{knitrout}
+
 \subsection{\code{plotFoldEnrichment}}
@@
 %% -----Supplementary Information------------------------------------
-\section[Supplementary Information]{Supplementary Information} \label{sec:supinfo}
+% \section[Supplementary Information]{Supplementary Information} \label{sec:supinfo}
 
-\begin{table}[H]
-  \centering
-  \caption{Supplementary Table 1. First 6 results generated by pathEnrich.}
-  \begin{adjustbox}{width=\textwidth, angle=90}
-  \fontsize{90}{125}\selectfont
-  \begin{tabular}{rllrrrrrrrr}
-  \hline
- & KEGG\_PATHWAY\_ID & KEGG\_PATHWAY\_description & KEGG\_PATHWAY\_cnt & KEGG\_PATHWAY\_in\_list & KEGG\_DATABASE\_cnt & KEG\_DATABASE\_in\_list & expected & enrich\_p & p\_adj & fold\_enrichment \\
-  \hline
-95 & rno04530 & Tight junction - Rattus norvegicus (rat) & 170 & 19.00 & 8856 & 295 & 5.66 & 0.00 & 0.00 & 3.36 \\
-  172 & rno05135 & Yersinia infection - Rattus norvegicus (rat) & 128 & 16.00 & 8856 & 295 & 4.26 & 0.00 & 0.00 & 3.75 \\
-  194 & rno05210 & Colorectal cancer - Rattus norvegicus (rat) &  88 & 12.00 & 8856 & 295 & 2.93 & 0.00 & 0.00 & 4.09 \\
-  212 & rno05231 & Choline metabolism in cancer - Rattus norvegicus (rat) &  99 & 12.00 & 8856 & 295 & 3.30 & 0.00 & 0.00 & 3.64 \\
-  197 & rno05213 & Endometrial cancer - Rattus norvegicus (rat) &  58 & 9.00 & 8856 & 295 & 1.93 & 0.00 & 0.00 & 4.66 \\
-  66 & rno04144 & Endocytosis - Rattus norvegicus (rat) & 275 & 22.00 & 8856 & 295 & 9.16 & 0.00 & 0.00 & 2.40 \\
-   \hline
-  \end{tabular}
-\end{adjustbox}
-\end{table}
+% \begin{table}[H]
+%   \centering
+%   \caption{Supplementary Table 1. First 6 results generated by pathEnrich.}
+%   \begin{adjustbox}{width=\textwidth, angle=90}
+%   \fontsize{90}{125}\selectfont
+%   \begin{tabular}{rllrrrrrrrr}
+%   \hline
+%  & KEGG\_PATHWAY\_ID & KEGG\_PATHWAY\_description & KEGG\_PATHWAY\_cnt & KEGG\_PATHWAY\_in\_list & KEGG\_DATABASE\_cnt & KEG\_DATABASE\_in\_list & expected & enrich\_p & p\_adj & fold\_enrichment \\
+%   \hline
+% 95 & rno04530 & Tight junction - Rattus norvegicus (rat) & 170 & 19.00 & 8856 & 295 & 5.66 & 0.00 & 0.00 & 3.36 \\
+%   172 & rno05135 & Yersinia infection - Rattus norvegicus (rat) & 128 & 16.00 & 8856 & 295 & 4.26 & 0.00 & 0.00 & 3.75 \\
+%   194 & rno05210 & Colorectal cancer - Rattus norvegicus (rat) &  88 & 12.00 & 8856 & 295 & 2.93 & 0.00 & 0.00 & 4.09 \\
+%   212 & rno05231 & Choline metabolism in cancer - Rattus norvegicus (rat) &  99 & 12.00 & 8856 & 295 & 3.30 & 0.00 & 0.00 & 3.64 \\
+%   197 & rno05213 & Endometrial cancer - Rattus norvegicus (rat) &  58 & 9.00 & 8856 & 295 & 1.93 & 0.00 & 0.00 & 4.66 \\
+%   66 & rno04144 & Endocytosis - Rattus norvegicus (rat) & 275 & 22.00 & 8856 & 295 & 9.16 & 0.00 & 0.00 & 2.40 \\
+%    \hline
+%   \end{tabular}
+% \end{adjustbox}
+% \end{table}
 
-\begin{table}[H]
-\centering
-  \caption{Supplementary Table 2. First 6 results generated by diffEnrich}
-  \begin{adjustbox}{width=\textwidth, angle=90}
-  \fontsize{50}{80}\selectfont
-\begin{tabular}{rllrrrrrrrrrrrrrrrrr}
-  \hline
- & KEGG\_PATHWAY\_ID & KEGG\_PATHWAY\_description & KEGG\_PATHWAY\_cnt & KEGG\_DATABASE\_cnt & KEGG\_PATHWAY\_in\_list1 & KEGG\_DATABASE\_in\_list1 & expected\_list1 & enrich\_p\_list1 & p\_adj\_list1 & fold\_enrichment\_list1 & KEGG\_PATHWAY\_in\_list2 & KEGG\_DATABASE\_in\_list2 & expected\_list2 & enrich\_p\_list2 & p\_adj\_list2 & fold\_enrichment\_list2 & odd\_ratio & diff\_enrich\_p & diff\_enrich\_adjusted \\
-  \hline
-rno04530 & rno04530 & Tight junction - Rattus norvegicus (rat) & 170 & 8856 & 19.00 & 295 & 5.66 & 0.00 & 0.00 & 3.36 & 131.00 & 5308 & 101.89 & 0.00 & 0.00 & 1.29 & 0.37 & 0.00 & 0.00 \\
-  rno05135 & rno05135 & Yersinia infection - Rattus norvegicus (rat) & 128 & 8856 & 16.00 & 295 & 4.26 & 0.00 & 0.00 & 3.75 & 105.00 & 5308 & 76.72 & 0.00 & 0.00 & 1.37 & 0.35 & 0.00 & 0.00 \\
-  rno05210 & rno05210 & Colorectal cancer - Rattus norvegicus (rat) &  88 & 8856 & 12.00 & 295 & 2.93 & 0.00 & 0.00 & 4.09 & 81.00 & 5308 & 52.74 & 0.00 & 0.00 & 1.54 & 0.37 & 0.00 & 0.00 \\
-  rno05213 & rno05213 & Endometrial cancer - Rattus norvegicus (rat) &  58 & 8856 & 9.00 & 295 & 1.93 & 0.00 & 0.00 & 4.66 & 55.00 & 5308 & 34.76 & 0.00 & 0.00 & 1.58 & 0.33 & 0.01 & 0.01 \\
-  rno04660 & rno04660 & T cell receptor signaling pathway - Rattus norvegicus (rat) & 106 & 8856 & 11.00 & 295 & 3.53 & 0.00 & 0.01 & 3.12 & 79.00 & 5308 & 63.53 & 0.00 & 0.00 & 1.24 & 0.39 & 0.01 & 0.01 \\
-  rno04657 & rno04657 & IL-17 signaling pathway - Rattus norvegicus (rat) &  95 & 8856 & 9.00 & 295 & 3.16 & 0.00 & 0.03 & 2.84 & 59.00 & 5308 & 56.94 & 0.37 & 0.46 & 1.04 & 0.36 & 0.01 & 0.01 \\
-   \hline
-\end{tabular}
-\end{adjustbox}
-\end{table}
-
-Example of S3 functions
-\begin{knitrout}
-\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
-\begin{alltt}
-\hlkwd{summary}\hlstd{(list1_pe)}
-\end{alltt}
-\begin{verbatim}
-## 219 KEGG pathways were tested. 
-##  Only KEGG pathways that contained at least 2 genes from gene_list were tested. 
-##  KEGG pathway species: Rattus norvegicus (rat)
-##  8856 genes from gene_list were in the KEGG data pull. 
-##  p-value adjustment method: BH
-##  36 pathways reached statistical significance after multiple testing correction at a cutoff of 0.05. 
-##  
-## Significant pathways: 
-##  Tight junction
-## Yersinia infection
-## Colorectal cancer
-## Choline metabolism in cancer
-## Endometrial cancer
-## Endocytosis
-## Neurotrophin signaling pathway
-## Thermogenesis
-## Oocyte meiosis
-## VEGF signaling pathway
-## Thyroid hormone signaling pathway
-## Hippo signaling pathway
-## T cell receptor signaling pathway
-## Apoptosis
-## Hepatocellular carcinoma
-## MAPK signaling pathway
-## Focal adhesion
-## Salmonella infection
-## Non-alcoholic fatty liver disease (NAFLD)
-## ErbB signaling pathway
-## Sphingolipid signaling pathway
-## Pancreatic cancer
-## Progesterone-mediated oocyte maturation
-## Alzheimer disease
-## Endocrine resistance
-## Adrenergic signaling in cardiomyocytes
-## IL-17 signaling pathway
-## Chronic myeloid leukemia
-## Dopaminergic synapse
-## Prostate cancer
-## EGFR tyrosine kinase inhibitor resistance
-## Hepatitis C
-## Ras signaling pathway
-## Acute myeloid leukemia
-## Insulin signaling pathway
-## Fc epsilon RI signaling pathway
-\end{verbatim}
-\end{kframe}
-\end{knitrout}
+% \begin{table}[H]
+% \centering
+%   \caption{Supplementary Table 2. First 6 results generated by diffEnrich}
+%   \begin{adjustbox}{width=\textwidth, angle=90}
+%   \fontsize{50}{80}\selectfont
+% \begin{tabular}{rllrrrrrrrrrrrrrrrrr}
+%   \hline
+%  & KEGG\_PATHWAY\_ID & KEGG\_PATHWAY\_description & KEGG\_PATHWAY\_cnt & KEGG\_DATABASE\_cnt & KEGG\_PATHWAY\_in\_list1 & KEGG\_DATABASE\_in\_list1 & expected\_list1 & enrich\_p\_list1 & p\_adj\_list1 & fold\_enrichment\_list1 & KEGG\_PATHWAY\_in\_list2 & KEGG\_DATABASE\_in\_list2 & expected\_list2 & enrich\_p\_list2 & p\_adj\_list2 & fold\_enrichment\_list2 & odd\_ratio & diff\_enrich\_p & diff\_enrich\_adjusted \\
+%   \hline
+% rno04530 & rno04530 & Tight junction - Rattus norvegicus (rat) & 170 & 8856 & 19.00 & 295 & 5.66 & 0.00 & 0.00 & 3.36 & 131.00 & 5308 & 101.89 & 0.00 & 0.00 & 1.29 & 0.37 & 0.00 & 0.00 \\
+%   rno05135 & rno05135 & Yersinia infection - Rattus norvegicus (rat) & 128 & 8856 & 16.00 & 295 & 4.26 & 0.00 & 0.00 & 3.75 & 105.00 & 5308 & 76.72 & 0.00 & 0.00 & 1.37 & 0.35 & 0.00 & 0.00 \\
+%   rno05210 & rno05210 & Colorectal cancer - Rattus norvegicus (rat) &  88 & 8856 & 12.00 & 295 & 2.93 & 0.00 & 0.00 & 4.09 & 81.00 & 5308 & 52.74 & 0.00 & 0.00 & 1.54 & 0.37 & 0.00 & 0.00 \\
+%   rno05213 & rno05213 & Endometrial cancer - Rattus norvegicus (rat) &  58 & 8856 & 9.00 & 295 & 1.93 & 0.00 & 0.00 & 4.66 & 55.00 & 5308 & 34.76 & 0.00 & 0.00 & 1.58 & 0.33 & 0.01 & 0.01 \\
+%   rno04660 & rno04660 & T cell receptor signaling pathway - Rattus norvegicus (rat) & 106 & 8856 & 11.00 & 295 & 3.53 & 0.00 & 0.01 & 3.12 & 79.00 & 5308 & 63.53 & 0.00 & 0.00 & 1.24 & 0.39 & 0.01 & 0.01 \\
+%   rno04657 & rno04657 & IL-17 signaling pathway - Rattus norvegicus (rat) &  95 & 8856 & 9.00 & 295 & 3.16 & 0.00 & 0.03 & 2.84 & 59.00 & 5308 & 56.94 & 0.37 & 0.46 & 1.04 & 0.36 & 0.01 & 0.01 \\
+%    \hline
+% \end{tabular}
+% \end{adjustbox}
+% \end{table}
 

### 2019-12-05 jss-article-rnw-diffEnrich/article.Rnw
V2 jss draft push

@@
 \usepackage{longtable}
+\usepackage{listings}
 \newcommand\mc[1]{\multicolumn{1}{c}{#1}}
@@
 options(prompt = "R> ", continue = "+  ", width = 70, useFancyQuotes = FALSE)
-library("MASS")
+
+ options(width=70)
+
+  listing <- function(x, options) {
+    paste("\\begin{lstlisting}[basicstyle=\\ttfamily,breaklines=true]\n",
+      x, "\\end{lstlisting}\n", sep = "")
+  }
+  knit_hooks$set(source=listing, output=listing)
+
 @

### 2019-12-05 jss-article-rnw-diffEnrich/article.tex
V2 jss draft push

@@
 \definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
-\begin{alltt}
-\hlkwd{install.packages}\hlstd{(}\hlstr{"diffEnrich"}\hlstd{)}
-\end{alltt}
+\begin{lstlisting}[basicstyle=\ttfamily,breaklines=true]
+install.packages("diffEnrich")\end{lstlisting}
 \end{kframe}
@@
 \definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
-\begin{alltt}
-\hlcom{## Load package}
-\hlkwd{suppressMessages}\hlstd{(}\hlkwd{library}\hlstd{(diffEnrich))}
-\hlcom{## run get_kegg() using rat}
-\hlstd{kegg_rno} \hlkwb{<-} \hlkwd{get_kegg}\hlstd{(}\hlstr{'rno'}\hlstd{)}
-\end{alltt}
+\begin{lstlisting}[basicstyle=\ttfamily,breaklines=true]
+## Load package\end{lstlisting}
+\begin{lstlisting}[basicstyle=\ttfamily,breaklines=true]
+suppressMessages(library(diffEnrich))\end{lstlisting}
+\begin{lstlisting}[basicstyle=\ttfamily,breaklines=true]
+## run get_kegg() using rat\end{lstlisting}
+\begin{lstlisting}[basicstyle=\ttfamily,breaklines=true]
+kegg_rno <- get_kegg('rno')\end{lstlisting}
 
@@
 \definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
-\begin{alltt}
-\hlcom{## run get_kegg() using rat}
-\hlstd{kegg_rno} \hlkwb{<-} \hlkwd{get_kegg}\hlstd{(}\hlstr{'rno'}\hlstd{)}
-\end{alltt}
+\begin{lstlisting}[basicstyle=\ttfamily,breaklines=true]
+## run get_kegg() using rat\end{lstlisting}
+\begin{lstlisting}[basicstyle=\ttfamily,breaklines=true]
+kegg_rno <- get_kegg('rno')\end{lstlisting}
 
@@
 \definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
-\begin{alltt}
-\hlcom{## run get_kegg() using rat}
-\hlstd{date} \hlkwb{<-} \hlkwd{as.character}\hlstd{(}\hlkwd{Sys.Date}\hlstd{())}
-\hlstd{kegg_rno} \hlkwb{<-} \hlkwd{get_kegg}\hlstd{(}\hlkwc{read} \hlstd{=} \hlnum{TRUE}\hlstd{,}
-                     \hlkwc{path} \hlstd{= here}\hlopt{::}\hlkwd{here}\hlstd{(),}
-                     \hlkwc{date} \hlstd{= date,}
-                     \hlkwc{release} \hlstd{=} \hlstr{"92"}\hlstd{)}
-\end{alltt}
+\begin{lstlisting}[basicstyle=\ttfamily,breaklines=true]
+## run get_kegg() using rat\end{lstlisting}
+\begin{lstlisting}[basicstyle=\ttfamily,breaklines=true]
+date <- as.character(Sys.Date())\end{lstlisting}
+\begin{lstlisting}[basicstyle=\ttfamily,breaklines=true]
+kegg_rno <- get_kegg(read = TRUE,
+                     path = here::here(),
+                     date = date,
+                     release = "92")\end{lstlisting}
 
@@
 \definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
-\begin{alltt}
-\hlcom{## View sample gene lists from package data}
-\hlkwd{head}\hlstd{(geneLists}\hlopt{$}\hlstd{list1)}
-\end{alltt}
-\begin{verbatim}
+\begin{lstlisting}[basicstyle=\ttfamily,breaklines=true]
+## View sample gene lists from package data\end{lstlisting}
+\begin{lstlisting}[basicstyle=\ttfamily,breaklines=true]
+head(geneLists$list1)\end{lstlisting}
+\begin{lstlisting}[basicstyle=\ttfamily,breaklines=true]
 ## [1] "361692"    "293654"    "293655"    "500974"    "100361529"
 ## [6] "171434"
-\end{verbatim}
-\begin{alltt}
-\hlkwd{head}\hlstd{(geneLists}\hlopt{$}\hlstd{list2)}
-\end{alltt}
-\begin{verbatim}
+\end{lstlisting}
+\begin{lstlisting}[basicstyle=\ttfamily,breaklines=true]
+head(geneLists$list2)\end{lstlisting}
+\begin{lstlisting}[basicstyle=\ttfamily,breaklines=true]
 ## [1] "315547" "315548" "315549" "315550" "50938"  "58856"
-\end{verbatim}
+\end{lstlisting}
 \end{kframe}
@@
 \definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
-\begin{alltt}
-\hlkwd{head}\hlstd{(list1_pe}\hlopt{$}\hlstd{enrich_table)}
-\end{alltt}
-\begin{verbatim}
+\begin{lstlisting}[basicstyle=\ttfamily,breaklines=true]
+head(list1_pe$enrich_table)\end{lstlisting}
+\begin{lstlisting}[basicstyle=\ttfamily,breaklines=true]
 ##     KEGG_PATHWAY_ID
@@
 ## 66         2.401627
-\end{verbatim}
+\end{lstlisting}
 \end{kframe}
@@
 \definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
-\begin{alltt}
-\hlkwd{summary}\hlstd{(list1_pe)}
-\end{alltt}
-\begin{verbatim}
+\begin{lstlisting}[basicstyle=\ttfamily,breaklines=true]
+summary(list1_pe)\end{lstlisting}
+\begin{lstlisting}[basicstyle=\ttfamily,breaklines=true]
 ## 219 KEGG pathways were tested. 
@@
 ## Fc epsilon RI signaling pathway
-\end{verbatim}
+\end{lstlisting}
 \end{kframe}
@@
 \definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
-\begin{alltt}
-\hlcom{## Perform differential enrichment}
-\hlstd{diff_enrich} \hlkwb{<-} \hlkwd{diffEnrich}\hlstd{(}\hlkwc{list1_pe} \hlstd{= list1_pe,}
-                          \hlkwc{list2_pe} \hlstd{= list2_pe,}
-                          \hlkwc{method} \hlstd{=} \hlstr{'none'}\hlstd{,}
-                          \hlkwc{cutoff} \hlstd{=} \hlnum{0.05}\hlstd{)}
-\end{alltt}
+\begin{lstlisting}[basicstyle=\ttfamily,breaklines=true]
+## Perform differential enrichment\end{lstlisting}
+\begin{lstlisting}[basicstyle=\ttfamily,breaklines=true]
+diff_enrich <- diffEnrich(list1_pe = list1_pe,
+                          list2_pe = list2_pe,
+                          method = 'none',
+                          cutoff = 0.05)\end{lstlisting}
 \end{kframe}
@@
 \definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
-\begin{alltt}
-\hlkwd{head}\hlstd{(diff_enrich}\hlopt{$}\hlstd{de_table)}
-\end{alltt}
-\begin{verbatim}
+\begin{lstlisting}[basicstyle=\ttfamily,breaklines=true]
+head(diff_enrich$de_table)\end{lstlisting}
+\begin{lstlisting}[basicstyle=\ttfamily,breaklines=true]
 ##          KEGG_PATHWAY_ID
@@
 ## rno04657         0.0087537607
-\end{verbatim}
+\end{lstlisting}
 \end{kframe}
@@
 \definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
-\begin{alltt}
-\hlkwd{summary}\hlstd{(diff_enrich)}
-\end{alltt}
-\begin{verbatim}
+\begin{lstlisting}[basicstyle=\ttfamily,breaklines=true]
+summary(diff_enrich)\end{lstlisting}
+\begin{lstlisting}[basicstyle=\ttfamily,breaklines=true]
 ## 219 KEGG pathways were shared between gene lists and were tested. 
@@
 ## Alzheimer disease
-\end{verbatim}
+\end{lstlisting}
 \end{kframe}
@@
 \definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
-\begin{alltt}
-\hlcom{## Plot fold enrichment}
-\hlkwd{plotFoldEnrichment}\hlstd{(}\hlkwc{de_res} \hlstd{= diff_enrich,} \hlkwc{pval} \hlstd{=} \hlnum{0.05}\hlstd{,} \hlkwc{N} \hlstd{=} \hlnum{5}\hlstd{)}
-\end{alltt}
+\begin{lstlisting}[basicstyle=\ttfamily,breaklines=true]
+## Plot fold enrichment\end{lstlisting}
+\begin{lstlisting}[basicstyle=\ttfamily,breaklines=true]
+plotFoldEnrichment(de_res = diff_enrich, pval = 0.05, N = 5)\end{lstlisting}
 \end{kframe}

### 2020-04-06 jss-article-rnw-diffEnrich/article.Rnw
Added parameter description for get_kegg function in app note

@@
 
+\begin{table}[]
+  \centering
+  \caption{\label{tab:description} Description of parameter usage for \code{get\textunderscore kegg}.}
+  \begin{adjustbox}{width=\textwidth}
+  \begin{tabular}{lll}
+  \hline
+\textbf{Argument} & \textbf{Description}                                                                                                                                                                                                                                     & \textbf{Example}            \\
+\hline
+species           & \begin{tabular}[c]{@{}l@{}}character.
+The species to use in kegg  \\ data pull\end{tabular}                                                                                                                                                                                                         & species = 'rno'             \\
+read              & \begin{tabular}[c]{@{}l@{}}logical.
+Should \code{get\textunderscore kegg} read in files from previous call.
+If TRUE, all 3 files \\ generated by \code{get\textunderscore kegg} must be in the same directory and the user must \\ provide a file path that points to that directory.\end{tabular} & read = TRUE                 \\
+path              & \begin{tabular}[c]{@{}l@{}}character.
+A character string describing the path \\ to write out KEGG API data sets.
+If not provided, \\ defaults to current working directory.\end{tabular}                                                                 & path = "path/to/directory/" \\
+date              & \begin{tabular}[c]{@{}l@{}}character.
+A character string describing the date that \\ was used to time stamp files from previous call.
+\\ Must be formatted like YYYY-MM-DD.\end{tabular}                                                                 & date = "2020-04-10"          \\
+release           & \begin{tabular}[c]{@{}l@{}}character.
+A character string describing the KEGG\\  release that was used to time stamp \\ files from previous call (e.g.
+"90", "92", "94")\end{tabular}                                                                     & release = "92"
+\end{tabular}
+\end{adjustbox}
+\end{table}
+
 \textbf{Note:} Because it is assumed that a user might want to use the data sets
@@
                      date = date,
-                     release = "92")
+                     release = "94")
 @

### 2020-04-06 jss-article-rnw-diffEnrich/article.tex
Added parameter description for get_kegg function in app note

@@
 
-{\ttfamily\noindent\itshape\color{messagecolor}{\#\# Kegg Release: Release\_92.0+\_12-05\_Dec\_19}}\end{kframe}
+{\ttfamily\noindent\itshape\color{messagecolor}{\#\# Kegg Release: Release\_94.0+\_04-06\_Apr\_20}}\end{kframe}
 \end{knitrout}
+
+\begin{table}[]
+  \centering
+  \caption{\label{tab:description} Description of parameter usage for \code{get\textunderscore kegg}.}
+  \begin{adjustbox}{width=\textwidth}
+  \begin{tabular}{lll}
+  \hline
+\textbf{Argument} & \textbf{Description}                                                                                                                                                                                                                                     & \textbf{Example}            \\
+\hline
+species           & \begin{tabular}[c]{@{}l@{}}character.
+The species to use in kegg  \\ data pull\end{tabular}                                                                                                                                                                                                         & species = 'rno'             \\
+read              & \begin{tabular}[c]{@{}l@{}}logical.
+Should \code{get\textunderscore kegg} read in files from previous call.
+If TRUE, all 3 files \\ generated by \code{get\textunderscore kegg} must be in the same directory and the user must \\ provide a file path that points to that directory.\end{tabular} & read = TRUE                 \\
+path              & \begin{tabular}[c]{@{}l@{}}character.
+A character string describing the path \\ to write out KEGG API data sets.
+If not provided, \\ defaults to current working directory.\end{tabular}                                                                 & path = "path/to/directory/" \\
+date              & \begin{tabular}[c]{@{}l@{}}character.
+A character string describing the date that \\ was used to time stamp files from previous call.
+\\ Must be formatted like YYYY-MM-DD.\end{tabular}                                                                 & date = "2020-04-10"          \\
+release           & \begin{tabular}[c]{@{}l@{}}character.
+A character string describing the KEGG\\  release that was used to time stamp \\ files from previous call (e.g.
+"90", "92", "94")\end{tabular}                                                                     & release = "92"
+\end{tabular}
+\end{adjustbox}
+\end{table}
 
@@
 
-{\ttfamily\noindent\itshape\color{messagecolor}{\#\# Kegg Release: Release\_92.0+\_12-05\_Dec\_19}}\end{kframe}
+{\ttfamily\noindent\itshape\color{messagecolor}{\#\# Kegg Release: Release\_94.0+\_04-06\_Apr\_20}}\end{kframe}
 \end{knitrout}
@@
                      date = date,
-                     release = "92")\end{lstlisting}
+                     release = "94")\end{lstlisting}
 
@@
 
-{\ttfamily\noindent\itshape\color{messagecolor}{\#\# ncbi\_to\_kegg2019-12-05Release\_92.0+\_12-05\_Dec\_19.txt}}
+{\ttfamily\noindent\itshape\color{messagecolor}{\#\# ncbi\_to\_kegg2020-04-06Release\_94.0+\_04-06\_Apr\_20.txt}}
 
-{\ttfamily\noindent\itshape\color{messagecolor}{\#\# kegg\_to\_pathway2019-12-05Release\_92.0+\_12-05\_Dec\_19.txt}}
+{\ttfamily\noindent\itshape\color{messagecolor}{\#\# kegg\_to\_pathway2020-04-06Release\_94.0+\_04-06\_Apr\_20.txt}}
 
-{\ttfamily\noindent\itshape\color{messagecolor}{\#\# pathway\_to\_species2019-12-05Release\_92.0+\_12-05\_Dec\_19.txt}}
+{\ttfamily\noindent\itshape\color{messagecolor}{\#\# pathway\_to\_species2020-04-06Release\_94.0+\_04-06\_Apr\_20.txt}}
 

### 2020-04-06 jss-article-rnw-diffEnrich/article.Rnw
Addded parameter descriptions for pathEnrich to app note

@@
 
+\begin{table}[]
+\centering
+  \caption{\label{tab:description} Description of parameter usage for \code{pathEnrich}.}
+  \begin{adjustbox}{width=\textwidth}
+\begin{tabular}{lll}
+\hline
+\textbf{Argument} & \textbf{Description}                                                                                                                                                                                                                                                                                                                                                                                                            & \textbf{Example}      \\
+\hline
+gk\_object        & \begin{tabular}[c]{@{}l@{}}list.
+Object genrated from get\_kegg, or a list containing the output \\ generated from a past get\_kegg call.
+Names of the list must match \\ those defined in get\_kegg.
+If the user wishes to use an older version\\  of data generated by get\_kegg, they must first load that data and\\  put it in a named list that matches the names given in the list \\ generated by get\_kegg.\end{tabular} & gk\_obj = kegg        \\
+gene\_list        & \begin{tabular}[c]{@{}l@{}}Vector.
+Vector of NCBI (ENTREZ)\\  geneIDs.\end{tabular}                                                                                                                                                                                                                                                                                                                                               & gene\_list = yourList \\
+method            & \begin{tabular}[c]{@{}l@{}}Character.
+Character string telling diffEnrich which method to \\ use for multiple testing correction.
+Available methods are \\ those provided by p.adjust, and the default is "BH", or \\ False Discovery Rate (FDR).\end{tabular}                                                                                                                                                                    & method = "BH"         \\
+cutoff            & \begin{tabular}[c]{@{}l@{}}Numeric.
+The p-value threshold to be used as the cutoff when\\  determining statistical significance, and used to filter list\\  of significant pathways.\end{tabular}                                                                                                                                                                                                                                 & cutoff = 0.05         \\
+N                 & \begin{tabular}[c]{@{}l@{}}Numeric.
+The number of genes from the gene list that must\\  be present in a KEGG pathway in order for that pathway \\ to be retained and tested.\end{tabular}                                                                                                                                                                                                                                         & N = 2
+\end{tabular}
+\end{adjustbox}
+\end{table}
+
 <<<pathEnrich, echo = FALSE>>=

### 2020-04-06 jss-article-rnw-diffEnrich/article.tex
Addded parameter descriptions for pathEnrich to app note

@@
 to be tested.
+
+\begin{table}[]
+\centering
+  \caption{\label{tab:description} Description of parameter usage for \code{pathEnrich}.}
+  \begin{adjustbox}{width=\textwidth}
+\begin{tabular}{lll}
+\hline
+\textbf{Argument} & \textbf{Description}                                                                                                                                                                                                                                                                                                                                                                                                            & \textbf{Example}      \\
+\hline
+gk\_object        & \begin{tabular}[c]{@{}l@{}}list.
+Object genrated from get\_kegg, or a list containing the output \\ generated from a past get\_kegg call.
+Names of the list must match \\ those defined in get\_kegg.
+If the user wishes to use an older version\\  of data generated by get\_kegg, they must first load that data and\\  put it in a named list that matches the names given in the list \\ generated by get\_kegg.\end{tabular} & gk\_obj = kegg        \\
+gene\_list        & \begin{tabular}[c]{@{}l@{}}Vector.
+Vector of NCBI (ENTREZ)\\  geneIDs.\end{tabular}                                                                                                                                                                                                                                                                                                                                               & gene\_list = yourList \\
+method            & \begin{tabular}[c]{@{}l@{}}Character.
+Character string telling diffEnrich which method to \\ use for multiple testing correction.
+Available methods are \\ those provided by p.adjust, and the default is "BH", or \\ False Discovery Rate (FDR).\end{tabular}                                                                                                                                                                    & method = "BH"         \\
+cutoff            & \begin{tabular}[c]{@{}l@{}}Numeric.
+The p-value threshold to be used as the cutoff when\\  determining statistical significance, and used to filter list\\  of significant pathways.\end{tabular}                                                                                                                                                                                                                                 & cutoff = 0.05         \\
+N                 & \begin{tabular}[c]{@{}l@{}}Numeric.
+The number of genes from the gene list that must\\  be present in a KEGG pathway in order for that pathway \\ to be retained and tested.\end{tabular}                                                                                                                                                                                                                                         & N = 2
+\end{tabular}
+\end{adjustbox}
+\end{table}
 

### 2020-04-06 jss-article-rnw-diffEnrich/article.Rnw
Updated R to 3.6.3 and recompiled article

@@
 In addition to these flat files, \code{get\textunderscore kegg} will also create a named list
-in R with the three relevant KEGG data sets. The names of this list will
-describe the data set, and are described in Table 1.
+in R with the three relevant KEGG data sets. A description of the paramters in the
+\code{get\textunderscore kegg} function can be found in Table 1, and the names of the list of
+data sets generated by the function that will describe the data set are described in Table 2.
 

### 2022-06-23 R/get_kegg.R
updated api pathe to use https instead of http

@@ -109,7 +109,7 @@ get_kegg <- function(species, read = FALSE, path = NULL, date, release){
   if(missing(species) | !(species %in% c('hsa','mmu','rno'))){stop("Must choose one of the 3 species options: human: use 'hsa', mouse: use 'mmu', rat: use rno")}
 
   # Define base api path and define list of operations/arguments
-  api_base <- "http://rest.kegg.jp/"
+  api_base <- "https://rest.kegg.jp/"
   op <- list("info"="info", "list"="list", "find"="find", "get"="get",
              "conv"="conv", "link"="link", "ddi"="ddi")
   db <- list("pathway"="pathway")

### 2022-06-24 README.Rmd
Debuggig osx build errors
updated urls that moved in READMEs and vingette

@@
 
-[![Build Status](https://travis-ci.com/SabaLab/diffEnrich.svg?token=M2jELEzTYYqmxZMY6hpb&branch=master)](https://travis-ci.com/SabaLab/diffEnrich)
+[![Build Status](https://travis-ci.com/SabaLab/diffEnrich.svg?token=M2jELEzTYYqmxZMY6hpb&branch=master)](https://www.travis-ci.com/SabaLab/diffEnrich)
 
@@
 As an example, enrichment of a particular pathway among a list of genes that are differentially expressed after an experimental manipulation may indicate that the pathway has been altered by that manipulation.
-This analysis is rather straight forward and many solutions have been offered (e.g., [Haung et al., 2009](https://www.ncbi.nlm.nih.gov/pubmed/19033363); [Kuleshov et al., 2016](https://www.ncbi.nlm.nih.gov/pubmed/27141961); [Liao et al., 2019](https://www.ncbi.nlm.nih.gov/pubmed/31114916); [Subramanian et al., 2005](https://www.pnas.org/content/102/43/15545)).
-A wide variety of databases have also been used to define these pathways (e.g., [Kanehisa and Goto, 2000](https://www.ncbi.nlm.nih.gov/pubmed/10592173)) and ontologies (e.g., [Ashburn et al., 2000](https://www.ncbi.nlm.nih.gov/pubmed/10802651)).
+This analysis is rather straight forward and many solutions have been offered (e.g., [Haung et al., 2009](https://pubmed.ncbi.nlm.nih.gov/19033363/); [Kuleshov et al., 2016](https://pubmed.ncbi.nlm.nih.gov/27141961/); [Liao et al., 2019](https://pubmed.ncbi.nlm.nih.gov/31114916/); [Subramanian et al., 2005](https://www.pnas.org/doi/10.1073/pnas.0506580102)).
+A wide variety of databases have also been used to define these pathways (e.g., [Kanehisa and Goto, 2000](https://pubmed.ncbi.nlm.nih.gov/10592173/)) and ontologies (e.g., [Ashburn et al., 2000](https://pubmed.ncbi.nlm.nih.gov/10802651/)).
 
@@
 However, for some newer omics technologies, the background data set is hard to define.
-For example, LC-MS analysis can be used to identify carbonylated proteins ( [Peterson et al., 2018](https://www.ncbi.nlm.nih.gov/pubmed/29799837); [Shearn et al., 2019](https://www.ncbi.nlm.nih.gov/pubmed/31377417); [Shearn et al., 2018](https://www.ncbi.nlm.nih.gov/pmc/articles/PMC6028296/)).
+For example, LC-MS analysis can be used to identify carbonylated proteins ( [Peterson et al., 2018](https://pubmed.ncbi.nlm.nih.gov/29799837/); [Shearn et al., 2019](https://pubmed.ncbi.nlm.nih.gov/31377417/); [Shearn et al., 2018](https://www.ncbi.nlm.nih.gov/pmc/articles/PMC6028296/)).
 With this study design, carbonylated proteins are isolated using a BH-derivation and then LC-MS is used to identify peptides in this isolated sample.
@@
 To easily execute this analysis, we have designed an R package that uses the KEGG REST API to obtain the most recent version of the KEGG PATHWAY ( [Kanehisa and Goto, 2000](https://www.ncbi.nlm.nih.gov/pubmed/10592173)) database to initially identify functional enrichment within a gene list using the entire KEGG transcriptome as the background data set and then to identify differentially enriched pathways between two gene lists.
-This R package includes a function to generate a ‘differential enrichment’ graphic.
+This R package includes a function to generate a "differential enrichment" graphic.
 
@@
 our downstream analysis. This function takes two arguments. The first, 'species' is required. Currently, *diffEnrich* supports three species, and the argument is a character string using the KEGG
-code [https://www.pnas.org/content/suppl/2008/09/11/0806162105.DCSupplemental/ST1_PDF.pdf](https://www.pnas.org/content/suppl/2008/09/11/0806162105.DCSupplemental/ST1_PDF.pdf): 
+code [https://www.pnas.org/doi/10.1073/pnas.0806162105](https://www.pnas.org/doi/10.1073/pnas.0806162105): 
 Homo sapiens (human), use 'hsa'; Mus musculus (mouse), use 'mmu'; and Rattus norvegicus (rat), use 'rno'. The second,
@@
 User gene lists must be character vectors and be formatted as ENTREZ gene IDs.
-The *clusterProfiler* package offers a nice function (*bitr*) that maps gene symbols and Ensembl IDs to ENTREZ gene IDs, and an example can be seen in their [vignette](https://yulab-smu.github.io/clusterProfiler-book/chapter14.html#bitr).
+The *clusterProfiler* package offers a nice function (*bitr*) that maps gene symbols and Ensembl IDs to ENTREZ gene IDs, and an example can be seen in their [vignette](https://yulab-smu.top/biomedical-knowledge-mining-book/useful-utilities.html).
 

### 2022-06-24 README.Rmd
Updated travis build link in README

@@
 <!-- README.md is generated from README.Rmd. Please edit that file -->
-[![Build status](https://ci.appveyor.com/api/projects/status/xwu8nket2u1pd5ij?svg=true)](https://ci.appveyor.com/project/hsmith9002/diffenrich)
+[![Build Status](https://app.travis-ci.com/SabaLab/diffEnrich.svg?branch=master)](https://app.travis-ci.com/SabaLab/diffEnrich)
 

### 2022-06-27 README.Rmd
updated broken URL and removed travis build link

@@
 <!-- README.md is generated from README.Rmd. Please edit that file -->
-[![Build Status](https://app.travis-ci.com/SabaLab/diffEnrich.svg?branch=master)](https://app.travis-ci.com/SabaLab/diffEnrich)
 
@@
 In this analysis, we compare the proportion of genes/proteins from one gene list associated with a particular pathway to the proportion of genes/proteins from a second gene list that are associated with that pathway.
-To easily execute this analysis, we have designed an R package that uses the KEGG REST API to obtain the most recent version of the KEGG PATHWAY ( [Kanehisa and Goto, 2000](https://www.ncbi.nlm.nih.gov/pubmed/10592173)) database to initially identify functional enrichment within a gene list using the entire KEGG transcriptome as the background data set and then to identify differentially enriched pathways between two gene lists.
+To easily execute this analysis, we have designed an R package that uses the KEGG REST API to obtain the most recent version of the KEGG PATHWAY ( [Kanehisa and Goto, 2000](https://pubmed.ncbi.nlm.nih.gov/10592173/)) database to initially identify functional enrichment within a gene list using the entire KEGG transcriptome as the background data set and then to identify differentially enriched pathways between two gene lists.
 This R package includes a function to generate a "differential enrichment" graphic.
